### 2018-02-25 Report/lab-meeting.Rmd
Adds first lab meeting report

---
title: "lab-meeting"
author: "Lukas 'hejtmy' Hejtmanek"
date: "24 February 2018"
output: 
  revealjs::revealjs_presentation:
    theme: night
    transition: none
    self_contained: true
    css: slides.css
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
sop_all <- read.table("../sop.csv", sep=";", header = T, stringsAsFactors = F)
walk_all <- read.table("../walk.csv", sep=";", header = T, stringsAsFactors = F)
```

```r

```

# What

---

Transfer of spatial knowledge across differet

# Methods

## Overview

Building has in total XX doors

6 doors are ascribed to fictional workers: Emily, Michael, Daniel, Jesicca, Jacob and Sarah 
 - Taken from the list of most used names in CA from 2000

4 pointing locations

XX Image of their placement

## Design

Full experiemntal design with 4 possible conditions

- vr - vr
- real - vr
- vr - real
- real - real

## Procedure
Participants fill questionnaires

They do the virtualiser training regardless of their condition (even in real-to real)

Experiment has 2 phases, each has two parts: walking/pointing
- Free roaming from door to door - all 6 ofices in 3 blocks (18 trials)
- Pointing from 2 points to all 6 offices; exactly one office is always visible

Drawing blank map

## Settings

6 possible settings

Example of settings

Question: Should we run two same settings? To directly compare performance in the same sequence of goals and points?

# Participants

## Descriptives

## Number crunching
So far we:

- scheduled 56 timeslots
- administered 45 participants
- have `r length(unique(walk_all$id))` usable subjects
- wasted 5 working days

There are some issues with high dropout rate, so far about a 1/3

---

Dropout by gender

---

Dropout by ethnicity

asdasdas

---

Dropout by condition

asdasadas

# Analysis

## Normalisation
Min normalising

# Results

## The bad
Unfortunatelly, out of `r as.numeric(walk_all %>% filter(type == "restimote") %>% count())` trials in the restimote we have `r as.numeric(walk_all %>% filter(type == "restimote") %>% filter(is.na(distance)) %>% count())` not fully executed.

## Distance

Distance absolute
```{r, echo = F,warning=FALSE}
ggplot(walk_all, aes(x = condition, y = distance, fill=factor(phase))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```

---

Distance normalised
```{r, echo = F, warning=FALSE}
ggplot(walk_all, aes(x = condition, y = min_norm_distance, fill=factor(phase))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```

---

```{r, echo = T}
walk_all %>% 
  group_by(condition) %>% 
  summarise(p.value = t.test(distance ~ phase)$p.value,
            phase1 = t.test(distance ~ phase)$estimate[[1]],
            phase2 = t.test(distance ~ phase)$estimate[[2]])
```

## Time

### 2018-02-25 Scripts/analysis.R

@@ -1,4 +1,4 @@
-library(dplyr)
+library(ggplot2)
 sop_all <- read.table("sop.csv", sep=";", header = T, stringsAsFactors = F)
 walk_all <- read.table("walk.csv", sep=";", header = T, stringsAsFactors = F)
 
@@ -6,9 +6,6 @@ t.test(abs(sop_all$error)~sop_all$phase)
 t.test(abs(sop_all$time)~sop_all$phase)
 
 t.test(abs(sop_all$error)~sop_all$type)
-
-ggplot(sop_all, aes(error, phase))
-
 
 aov_point_condition <- aov(abs(error)~condition*phase, data = sop_all)
 summary(aov_point_condition)
@@ -33,6 +30,10 @@ ggplot(walk_all, aes(x = type, y = time, fill=factor(phase))) +
   stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
 
 ggplot(walk_all, aes(x = condition, y = distance, fill=factor(phase))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
+
+ggplot(walk_all, aes(x = condition, y = min_norm_distance, fill=factor(phase))) +
   stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
   stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
 

### 2018-02-25 Report/lab-meeting.Rmd
Updates presentation

@@
 
+XXX example image
+
 ## Distance
@@
 
-```{r, echo = T}
+```{r, echo = F,results='hide'}
 walk_all %>% 
@@
 ```
+```{r, echo = T}
+aov_distance_condition <- aov(min_norm_distance~condition*phase, 
+                              data = walk_all)
+summary(aov_distance_condition)
+```
+```{r, echo = T}
+aov_distance_condition_2 <- aov(min_norm_distance~condition, 
+                              data = walk_all[walk_all$phase == 2,])
+summary(aov_distance_condition_2)
+```
+## Time
 
-## Time
+Time absolute
+```{r, echo = F,warning=FALSE}
+ggplot(walk_all, aes(x = condition, y = time, fill=factor(phase))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
+```
+
+---
+
+Time normalised
+```{r, echo = F,warning=FALSE}
+ggplot(walk_all, aes(x = condition, y = min_norm_time, fill=factor(phase))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
+```
+
+---
+
+```{r, echo = F,results='hide'}
+walk_all %>% 
+  group_by(condition) %>% 
+  summarise(p.value = t.test(time ~ phase)$p.value,
+            phase1 = t.test(time ~ phase)$estimate[[1]],
+            phase2 = t.test(time ~ phase)$estimate[[2]])
+```
+```{r, echo = T}
+aov_time_condition <- aov(min_norm_time~condition*phase, 
+                              data = walk_all)
+summary(aov_time_condition)
+```
+```{r, echo = T}
+aov_time_condition_2 <- aov(min_norm_time~condition, 
+                              data = walk_all[walk_all$phase == 2,])
+summary(aov_time_condition_2)
+```
+
+## Pointing accuracy
+
+```{r, echo = F,warning=FALSE}
+ggplot(sop_all, aes(x = condition, y = abs(error), fill=factor(phase))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
+```
+
+---
+
+```{r, echo = T}
+aov_point_condition <- aov(abs(error)~condition*phase, 
+                           data = sop_all)
+summary(aov_point_condition)
+```
+
+```{r, echo = T}
+aov_point_condition_2 <- aov(abs(error)~condition, 
+                             data = sop_all[sop_all$phase == 2, ])
+summary(aov_point_condition_2)
+```

### 2018-02-25 Scripts/analysis.R

@@ -10,12 +10,11 @@ t.test(abs(sop_all$error)~sop_all$type)
 aov_point_condition <- aov(abs(error)~condition*phase, data = sop_all)
 summary(aov_point_condition)
 
-aov_distance_condition <- aov(distance~condition*phase, data = walk_all)
+aov_distance_condition <- aov(min_norm_distance~condition*phase, data = walk_all)
 summary(aov_distance_condition)
 
 aov_time_condition <- aov(time~condition*phase, data = walk_all)
 summary(aov_time_condition)
-
 
 ggplot(sop_all, aes(x = condition, y = abs(error), fill=factor(phase))) +
   stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 

### 2018-02-25 Scripts/main-preprocess.R

@@ -21,21 +21,37 @@ conditions$condition <- gsub('restimote', 'real', conditions$condition)
 sop_all <- right_join(sop_all, conditions, by = "id")
 walk_all <-  right_join(walk_all, conditions, by = "id")
 
-#normalisations
+#Distance normalisations
 distance_offices_summary <- walk_all %>% 
   group_by(type, start, goal) %>% 
-  summarise(mean = mean(distance, na.rm = T), 
-           sd = sd(distance, na.rm = T), 
-           min = min(distance, na.rm = T), 
-           max = max(distance, na.rm = T)) %>%
+  summarise(mean.dist = mean(distance, na.rm = T), 
+           sd.dist = sd(distance, na.rm = T), 
+           min.dist = min(distance, na.rm = T), 
+           max.dist = max(distance, na.rm = T)) %>%
   arrange(start, goal)
-  
+ 
+ 
 wide_distance_offices_summary <- distance_offices_summary %>% melt(id.vars=c("type", "start", "goal")) %>% dcast(start + goal ~ type + variable)
-office_distance_type_diff_min <- distance_offices_summary %>% group_by(start, goal) %>% summarise(min_diff = diff(min))
+office_distance_type_diff_min <- distance_offices_summary %>% group_by(start, goal) %>% summarise(min_diff = diff(min.dist))
 qplot(data = office_distance_type_diff_min, x = 1:nrow(office_distance_type_diff_min), y = min_diff, geom = "line")
 
 walk_all <- right_join(walk_all, distance_offices_summary[, c(1:3, 6)], by = c("type", "start", "goal"))
-walk_all$min_norm_distance <- walk_all$distance/walk_all$min
+walk_all$min_norm_distance <- walk_all$distance/walk_all$min.dist
+
+# TIme normalisation
+time_offices_summary <- walk_all %>% 
+  filter(time > 10) %>%
+  group_by(type, start, goal) %>% 
+  summarise(mean.time = mean(time, na.rm = T), 
+            sd.time = sd(time, na.rm = T), 
+            min.time = min(time, na.rm = T), 
+            max.time = max(time, na.rm = T)) %>%
+  arrange(start, goal)
+wide_time_offices_summary <- time_offices_summary %>% melt(id.vars=c("type", "start", "goal")) %>% dcast(start + goal ~ type + variable)
[…]

### 2018-02-25 Report/lab-meeting.Rmd
Adds todo section

@@
 
-## Normalisation
-Min normalising
+## Preprocessing
+
+So far no outliers have been removed
+Paths haven't been smoothed
+because of VR-real world differences, distances and times have been min-normalised
 
@@
 ```
+
+# TODO
+
+---
+
+Get about 15-16 more participants. Not sure if 80 is necessary given current results.
+Smooth paths and try to clean it enough to get less estimote dropouts
+Remove outliers
+Work with "wrong-door" stats
+
+## Questions
+
+Look are initial 3 trials performance after environment shift vs no shift
+
+Look at progressive learning curve from block to block, rather than phase summaries
+
+Add in the EEG

### 2018-02-26 Report/lab-meeting.Rmd
Finalised lab report

--- 2018-02-25 Report/lab-meeting.Rmd
@@
 library(dplyr)
+library(restimoter)
+library(navr)
+library(brainvr.R)
+source_folder <- function(path){
+  sapply(list.files(pattern = "[.]R$", path = path, full.names = TRUE), source);
+}
+source_folder("../TwoWorlds/")
+
 sop_all <- read.table("../sop.csv", sep=";", header = T, stringsAsFactors = F)
 walk_all <- read.table("../walk.csv", sep=";", header = T, stringsAsFactors = F)
+load("../multi.data")
+load("../settings.data")
+df_participants <- settings$participants
+df_participants$condition <- paste0(df_participants$First.phase, "-", df_participants$Second.phase)
 ```
-
-```r
-
-```
-
-# What
-
----
-
-Transfer of spatial knowledge across differet
 
@@
 
-Building has in total XX doors
+Free roaming phase while searching fo 6 offices
 
-6 doors are ascribed to fictional workers: Emily, Michael, Daniel, Jesicca, Jacob and Sarah 
+The CNS has 30 eligible doors in total, 6 doors are ascribed to fictional workers: Emily, Michael, Daniel, Jesicca, Jacob and Sarah 
  - Taken from the list of most used names in CA from 2000
@@
 
-XX Image of their placement
+![](images/layout.png){width=500px}
 
@@
 ## Procedure
-Participants fill questionnaires
 
-They do the virtualiser training regardless of their condition (even in real-to real)
+Questionnaire
+
+Virtualiser training; regardless of their condition (even in real-to real)
 
 Experiment has 2 phases, each has two parts: walking/pointing
+
 - Free roaming from door to door - all 6 ofices in 3 blocks (18 trials)
@@
 
-Drawing blank map
+Drawing building layout
 
@@
 
-Example of settings
+Example: Sarah - Jacob - Emily - Michael - Daniel - Jesicca
 
-Question: Should we run two same settings? To directly compare performance in the same sequence of goals and points?
+Pointing fromn Viewpoint 1 and 3
+
+Never pointing from the same viewpoints in both cases
 
 # Participants
-
-## Descriptives
 
@@
 
+Dropout by condition
+
+```{r, echo = F,warning=FALSE}
+ggplot(df_participants, aes(x = condition, fill = finished)) + geom_bar(position = "dodge")
+```
+
+---
+Average time spent in the device per condition
+
+```{r, echo = F}
+ggplot(walk_all, aes(x = type, y = 18*(time), fill = factor(phase))) + geom_bar( stat = "summary", fun.y = "mean", position = "dodge")
+```
+
+---
+
 Dropout by gender
+```{r, echo = F,warning=FALSE}
+ggplot(df_participants, aes(x = condition, fill = finished)) + 
+  geom_bar(position = "dodge") + facet_grid(.~sex)
+```
 
@@
 
-asdasdas
+```{r, echo = F,warning=FALSE}
+ggplot(df_participants, aes(x = condition, fill = finished)) + 
+  geom_bar(position = "dodge") + facet_grid(.~ethnicity)
+```
+
+# Preprocessing
 
@@
 
-Dropout by condition
+- So far no outliers have been removed
+- Paths haven't been smoothed
+- Because of VR-real world differences, distances and times have been *min-normalised*
 
-asdasadas
+Min-normalisation:
 
-# Analysis
-
-## Normalisation
-Min normalising
-
-# Results
+- Dividing path traveled in a particular task by the shortest path traveled in that particular task and device
+- Esentially gives you how many times the participant traveled more than necessary (2x, 1.5x)
 
@@
 
-XXX example image
+```{r, echo=F, warning=FALSE, message=F}
+plot_walk_trial(ls$tw11$phase2, 8)
+```
+
+## Path examples
+
+Unity
+
+```{r, echo=F, warning=FALSE, message=F}
+plot_walk_trial(ls$tw11$phase1, 8)
+```
+
+---
+
+Real
+
+```{r, echo=F, warning=FALSE, message=F}
+plot_walk_trial(ls$tw29$phase1, 3)
+```
+
+## Pointing examples
+
+Unity
+
+```{r, echo=F, warning=FALSE, message=F}
+plot_sop_point(ls$tw18$phase2, 4)
+```
+
+---
+
+Real
+
+```{r, echo=F, warning=FALSE, message=F}
+plot_sop_point(ls$tw29$phase1, 3)
+```
+
+# Results
 
@@
 
-```{r, echo = F,results='hide'}
+```{r, echo = F, results='hide'}
 walk_all %>% 
@@
 ```
+
 ```{r, echo = T}
@@
 ```
+
 ## Time
@@
 ```
+
+# What now?
+
+## TODO
+
+Get about 15-16 more participants
+
+Smooth paths and try to clean it enough to get less estimote dropouts
+
+Remove outliers
+
+Work with "wrong-door" stats
+
+## Questions
+
+Should we run two same settings? To directly compare performance in the same sequence of goals and points?
+
+Look are initial 3 trials performance after environment shift vs no shift
+
+Look at progressive learning curve from block to block, rather than phase summaries
+
+Add in scalp EEG

### 2018-02-26 TwoWorlds/twoworld-getters-twunity.R

@@ -85,7 +85,7 @@ get_goal_name.brainvr <- function(obj, iGoal){
 }
 get_all_goal_positions.brainvr <- function(obj, include_SOP = FALSE){
   if(include_SOP){i <- 1:10}else{i <- 1:6}
-  ls_goals <- setNames(split(obj$data$experiment_log$positions$GoalPositions[i, c(1,3)], i), i)
+  ls_goals <- setNames(split(obj$data$experiment_log$positions$GoalPositions[i, c(1,3)], i), obj$data$experiment_log$settings$GoalNames[i])
   return(ls_goals)
 }
 get_point_start.sop <- function(obj, trialId){

### 2018-02-28 TwoWorlds/helpers-results.R
CHanges creation of results data frames

@@ -1,6 +1,6 @@
-create_walk_df <- function(obj){
+create_walk_df <- function(participant_id){
   N_TRIALS <- 18
-  df <- data.frame(id = rep(obj$participant_id, N_TRIALS), 
+  df <- data.frame(id = rep(participant_id, N_TRIALS), 
                    time = rep(NA, N_TRIALS), 
                    distance = rep(NA, N_TRIALS),
                    optimal_distance = rep(NA, N_TRIALS),
@@ -10,9 +10,9 @@ create_walk_df <- function(obj){
   return(df)
 }
 
-create_sop_df <- function(obj){
+create_sop_df <- function(participant_id){
   N_POINTING <- 12
-  df <- data.frame(id = rep(obj$participant_id, N_POINTING), 
+  df <- data.frame(id = rep(participant_id, N_POINTING), 
                    time = rep(NA, N_POINTING), 
                    error = rep(NA, N_POINTING),
                    start = rep(NA, N_POINTING),

### 2018-02-28 TwoWorlds/twoworld-results.R

@@ -13,7 +13,7 @@ walk_trial_results <- function(obj, trialId){
 }
 ## UNITY ----
 walk_results.twunity <- function(obj){
-  df_results <- create_walk_df(obj$walk)
+  df_results <- create_walk_df(obj$walk$participant_id)
   return(walk_results.general(obj, df_results))
 }
 walk_trial_results.twunity <- function(obj, trialId){
@@ -26,7 +26,7 @@ walk_trial_results.twunity <- function(obj, trialId){
   return(ls)
 }
 sop_results.twunity <- function(obj){
-  df_results <- create_sop_df(obj$sop)
+  df_results <- create_sop_df(obj$sop$participant_id)
   return(sop_results.general(obj, df_results))
 }
 sop_trial_results.twunity <- function(obj, trialId){
@@ -45,7 +45,7 @@ sop_trial_results.twunity <- function(obj, trialId){
 }
 ## RESTIMOTE ----
 walk_results.restimote <- function(obj){
-  df_results <- create_walk_df(obj)
+  df_results <- create_walk_df(obj$participant_id)
   return(walk_results.general(obj, df_results))
 }
 walk_trial_results.restimote <- function(obj, trialId){
@@ -59,7 +59,7 @@ walk_trial_results.restimote <- function(obj, trialId){
   return(ls)
 }
 sop_results.restimote <- function(obj){
-  df_results <- create_sop_df(obj)
+  df_results <- create_sop_df(obj$participant_id)
   return(sop_results.general(obj, df_results))
 }
 sop_trial_results.restimote <- function(obj, trialId){

### 2018-02-28 TwoWorlds/helpers-results.R
Simplifies results code

@@ -18,3 +18,12 @@ create_sop_df <- function(participant_id){
                    start = rep(NA, N_POINTING),
                    goal = rep(NA, N_POINTING))
 }
+
+list_to_row <- function(df, ls, iRow){
+  df_cols <- colnames(df)
+  ls_names <- names(ls)
+  for(name in ls_names){
+    if(name %in% df_cols) df[iRow, name] <- ls[[name]]
+  }
+  return(df)
+}

### 2018-02-28 TwoWorlds/twoworld-results.R

@@ -78,20 +78,14 @@ sop_trial_results.restimote <- function(obj, trialId){
 sop_results.general <- function (obj, df_results){
   for (trialId in 1:12){
     ls <- sop_trial_results(obj, trialId)
-    df_results[trialId, "time"] <- ls$time
-    df_results[trialId, "error"] <- ls$error
-    df_results[trialId, "start"] <- ls$start
-    df_results[trialId, "goal"] <- ls$goal
+    df_results <- list_to_row(df_results, ls, trialId)
   }
   return(df_results)
 }
 walk_results.general <- function(obj, df_results){
   for (trialId in 1:18){
     ls <- walk_trial_results(obj, trialId)
-    df_results[trialId, "time"] <- ls$time
-    df_results[trialId, "distance"] <- ls$distance
-    df_results[trialId, "start"] <- ls$start
-    df_results[trialId, "goal"] <- ls$goal
+    df_results <- list_to_row(df_results, ls, trialId)
   }
   df_results$distance[df_results$distance == 0] <- NA
   return(df_results)

### 2018-02-28 TwoWorlds/twoworld-results.R
Adds restimote errors

@@ -56,6 +56,7 @@ walk_trial_results.restimote <- function(obj, trialId){
   ls$distance <- sum(log_true$distance)
   ls$start <- get_trial_start_name.restimote(obj, trialId)
   ls$goal <- get_trial_goal_name.restimote(obj, trialId)
+  ls$errors <- get_trial_n_actions(obj, trialId, "Wrong door")
   return(ls)
 }
 sop_results.restimote <- function(obj){

### 2018-02-28 TwoWorlds/twoworld-getters-generics.R
Renames the functions to better fit

@@ -67,3 +67,6 @@ get_all_goal_positions <- function(obj, include_SOP){
 get_trial_point <- function(obj, trialId){
   UseMethod("get_trial_point")
 }
+get_trial_errors <- function(obj, trialId){
+  UseMethod("get_trial_errors")
+}

### 2018-02-28 TwoWorlds/twoworld-getters-twunity.R

@@ -35,6 +35,10 @@ get_trial_start_goal.twunity <- function(obj, trialId){
   ls <- list(start = start, goal = goal)
   return(ls)
 }
+get_trial_errors <- function(obj, trialId){
+  return(NA)
+}
+## SOP ----
 get_point_start_id.twunity <- function(obj, trialId){
   iStart <- obj$sop$data$experiment_log$settings$GoalOrder[trialId]
   return(iStart + 1)

### 2018-02-28 TwoWorlds/twoworld-results.R

@@ -23,6 +23,7 @@ walk_trial_results.twunity <- function(obj, trialId){
   ls$distance <- diff(range(log$cumulative_distance))
   ls$start <- get_trial_start_name.twunity(obj, trialId)
   ls$goal <- get_trial_goal_name.twunity(obj, trialId)
+  ls$errors <- get_trial_errors.twunity(obj, trialId)
   return(ls)
 }
 sop_results.twunity <- function(obj){
@@ -56,7 +57,7 @@ walk_trial_results.restimote <- function(obj, trialId){
   ls$distance <- sum(log_true$distance)
   ls$start <- get_trial_start_name.restimote(obj, trialId)
   ls$goal <- get_trial_goal_name.restimote(obj, trialId)
-  ls$errors <- get_trial_n_actions(obj, trialId, "Wrong door")
+  ls$errors <- get_trial_errors.restimote(obj, trialId)
   return(ls)
 }
 sop_results.restimote <- function(obj){

### 2018-02-28 TwoWorlds/helpers-getters.R
Adds calculation of twunity wrong doors

@@ -4,3 +4,9 @@ get_rotations <- function(df){
   return(as.data.frame(df[,..ids]))
   return(as.data.frame(df[,ids]))
 }
+
+get_all_doors <- function(settings){
+  i <- 1:nrow(settings$door_positions)
+  ls_doors <- setNames(split(settings$door_positions[i, 2:3], i), settings$door_positions[i, 1])
+  return(ls_doors)
+}

### 2018-02-28 TwoWorlds/twoworld-getters-twunity.R

@@ -35,8 +35,20 @@ get_trial_start_goal.twunity <- function(obj, trialId){
   ls <- list(start = start, goal = goal)
   return(ls)
 }
-get_trial_errors <- function(obj, trialId){
-  return(NA)
+get_trial_errors.twunity <- function(obj, trialId){
+  #gets_log
+  DIAMETER <- 0.52
+  ALL_DOORS_POSITIONS <- settings$door_positions #a bit problmatic, fethces stuff from the global env
+  n_errors <- -2 # we will get one correct hit at the goal start and end
+  df_log <- get_trial_log(obj$walk, trialId)
+  df_positions <- df_log[,c('Position.x', 'Position.z')]
+  for(i in 1:nrow(ALL_DOORS_POSITIONS)){
+    position <- c(ALL_DOORS_POSITIONS$Estimote.x[i], ALL_DOORS_POSITIONS$Estimote.y[i])
+    diffs <- sweep(df_positions, 2, position)
+    distances <- apply(diffs, 1, function(x){sqrt(sum(x ^ 2))})
+    if(any(distances < DIAMETER)) n_errors <- n_errors + 1
+  }
+  return(n_errors)
 }
 ## SOP ----
 get_point_start_id.twunity <- function(obj, trialId){

### 2018-02-28 TwoWorlds/helpers-getters.R
Updates getting all doors

@@ -7,6 +7,6 @@ get_rotations <- function(df){
 
 get_all_doors <- function(settings){
   i <- 1:nrow(settings$door_positions)
-  ls_doors <- setNames(split(settings$door_positions[i, 2:3], i), settings$door_positions[i, 1])
+  ls_doors <- setNames(split(settings$door_positions[i, c(2,3)], i), as.character(settings$door_positions[i, 1][[1]]))
   return(ls_doors)
 }

### 2018-02-28 TwoWorlds/twoworld-getters-twunity.R
Better wrong doors for unity

@@ -38,8 +38,9 @@ get_trial_start_goal.twunity <- function(obj, trialId){
 get_trial_errors.twunity <- function(obj, trialId){
   #gets_log
   DIAMETER <- 0.52
+  n_errors <- ifelse(trialId == 1, -1, -2) #we always get to the start and end door
   ALL_DOORS_POSITIONS <- settings$door_positions #a bit problmatic, fethces stuff from the global env
-  n_errors <- -2 # we will get one correct hit at the goal start and end
+   # we will get one correct hit at the goal start and end
   df_log <- get_trial_log(obj$walk, trialId)
   df_positions <- df_log[,c('Position.x', 'Position.z')]
   for(i in 1:nrow(ALL_DOORS_POSITIONS)){
@@ -48,6 +49,7 @@ get_trial_errors.twunity <- function(obj, trialId){
     distances <- apply(diffs, 1, function(x){sqrt(sum(x ^ 2))})
     if(any(distances < DIAMETER)) n_errors <- n_errors + 1
   }
+  if(n_errors < 0) n_errors <- 0
   return(n_errors)
 }
 ## SOP ----

### 2018-03-01 Scripts/main-preprocess.R
Moved files around

@@ -29,8 +29,7 @@ distance_offices_summary <- walk_all %>%
            min.dist = min(distance, na.rm = T), 
            max.dist = max(distance, na.rm = T)) %>%
   arrange(start, goal)
- 
- 
+
 wide_distance_offices_summary <- distance_offices_summary %>% melt(id.vars=c("type", "start", "goal")) %>% dcast(start + goal ~ type + variable)
 office_distance_type_diff_min <- distance_offices_summary %>% group_by(start, goal) %>% summarise(min_diff = diff(min.dist))
 qplot(data = office_distance_type_diff_min, x = 1:nrow(office_distance_type_diff_min), y = min_diff, geom = "line")
@@ -47,6 +46,7 @@ time_offices_summary <- walk_all %>%
             min.time = min(time, na.rm = T), 
             max.time = max(time, na.rm = T)) %>%
   arrange(start, goal)
+
 wide_time_offices_summary <- time_offices_summary %>% melt(id.vars=c("type", "start", "goal")) %>% dcast(start + goal ~ type + variable)
 office_time_type_diff_min <- time_offices_summary %>% group_by(start, goal) %>% summarise(min_diff = diff(min.time))
 qplot(data = office_time_type_diff_min, x = 1:nrow(office_time_type_diff_min), y = min_diff, geom = "line")

### 2018-03-01 TwoWorlds/helpers-results.R

@@ -1,29 +1,0 @@
-create_walk_df <- function(participant_id){
-  N_TRIALS <- 18
-  df <- data.frame(id = rep(participant_id, N_TRIALS), 
-                   time = rep(NA, N_TRIALS), 
-                   distance = rep(NA, N_TRIALS),
-                   optimal_distance = rep(NA, N_TRIALS),
-                   errors = rep(NA, N_TRIALS),
-                   start = rep(NA, N_TRIALS),
-                   goal = rep(NA, N_TRIALS))
-  return(df)
-}
-
-create_sop_df <- function(participant_id){
-  N_POINTING <- 12
-  df <- data.frame(id = rep(participant_id, N_POINTING), 
-                   time = rep(NA, N_POINTING), 
-                   error = rep(NA, N_POINTING),
-                   start = rep(NA, N_POINTING),
-                   goal = rep(NA, N_POINTING))
-}
-
-list_to_row <- function(df, ls, iRow){
-  df_cols <- colnames(df)
-  ls_names <- names(ls)
-  for(name in ls_names){
-    if(name %in% df_cols) df[iRow, name] <- ls[[name]]
-  }
-  return(df)
-}

### 2018-03-01 TwoWorlds/twoworld-results-restimote.R

@@ -1,0 +1,32 @@
+## RESTIMOTE ----
+walk_results.restimote <- function(obj){
+  df_results <- create_walk_df(obj$participant_id)
+  return(walk_results.general(obj, df_results))
+}
+walk_trial_results.restimote <- function(obj, trialId){
+  log <- get_trial_log(obj, trialId)
+  log_true <- true_trial_log(obj, trialId, 30, radius = 3)
+  ls <- list()
+  ls$time <- diff(range(log$Time)) 
+  ls$distance <- sum(log_true$distance)
+  ls$start <- get_trial_start_name.restimote(obj, trialId)
+  ls$goal <- get_trial_goal_name.restimote(obj, trialId)
+  ls$errors <- get_trial_errors.restimote(obj, trialId)
+  return(ls)
+}
+sop_results.restimote <- function(obj){
+  df_results <- create_sop_df(obj$participant_id)
+  return(sop_results.general(obj, df_results))
+}
+sop_trial_results.restimote <- function(obj, trialId){
+  ls <- list()
+  times <- get_trial_point_times.restimote(obj, trialId)
+  ls_pos <- get_point_start_goal.restimote(obj, trialId)
+  ls$pointed_angle <- get_trial_point.restimote(obj, trialId)
+  ls$correct_angle <- navr::angle_from_positions(unlist(ls_pos$start), unlist(ls_pos$goal))
+  ls$error <- navr::angle_to_180(ls$correct_angle - ls$pointed_angle)
+  ls$time <- times$end - times$start
+  ls$start <- get_point_start_name.restimote(obj, trialId)
+  ls$goal <- get_point_goal_name.restimote(obj, trialId)
+  return(ls)
+}

### 2018-03-01 TwoWorlds/twoworld-results.R

@@ -1,4 +1,4 @@
-# GENERIC ----
+# GENERICS ----
 sop_results <- function(obj){
   UseMethod("sop_results")
 }
@@ -10,71 +10,6 @@ walk_results <- function(obj){
 }
 walk_trial_results <- function(obj, trialId){
   UseMethod("walk_trial_results")
-}
-## UNITY ----
-walk_results.twunity <- function(obj){
-  df_results <- create_walk_df(obj$walk$participant_id)
-  return(walk_results.general(obj, df_results))
-}
-walk_trial_results.twunity <- function(obj, trialId){
-  ls <- list()
-  log <- get_trial_log(obj$walk, trialId)
-  ls$time <- diff(range(log$Time))
-  ls$distance <- diff(range(log$cumulative_distance))
-  ls$start <- get_trial_start_name.twunity(obj, trialId)
-  ls$goal <- get_trial_goal_name.twunity(obj, trialId)
-  ls$errors <- get_trial_errors.twunity(obj, trialId)
-  return(ls)
-}
-sop_results.twunity <- function(obj){
-  df_results <- create_sop_df(obj$sop$participant_id)
-  return(sop_results.general(obj, df_results))
-}
-sop_trial_results.twunity <- function(obj, trialId){
-  UNITY_SHIFT <- -90 # WEIRD thing that the controller actually was transformed by this amout, so we need to input it back
-  ls_pos <- get_point_start_goal(obj, trialId)
-  point_line <- get_trial_point(obj, trialId)
-  ls <- as.list(get_rotations(point_line))
-  ls$pointed_angle <- navr::angle_to_360(point_line$Rotation.Controller.x) + UNITY_SHIFT
-  ls$correct_angle <- navr::angle_from_positions(ls_pos$start, ls_pos$goal)
-  ls$error <- navr::angle_to_180(ls$correct_angle - ls$pointed_angle)
-  times <- get_trial_times(obj$sop, trialId)
-  ls$time <- times$finish - times$start
-  ls$start <- get_point_start_name(obj, trialId)
-  ls$goal <- get_point_goal_name(obj, trialId)
-  return(ls)
-}
-## RESTIMOTE ----
-walk_results.restimote <- function(obj){
-  df_results <- create_walk_df(obj$participant_id)
-  return(walk_results.general(obj, df_results))
-}
-walk_trial_results.restimote <- function(obj, trialId){
-  log <- get_trial_log(obj, trialId)
[…]

### 2018-03-01 TwoWorlds/twoworld-results.R
Adds trial id and block id columns to results tables

@@ -27,10 +27,11 @@ walk_results.general <- function(obj, df_results){
   df_results$distance[df_results$distance == 0] <- NA
   return(df_results)
 }
-
 create_walk_df <- function(participant_id){
   N_TRIALS <- 18
-  df <- data.frame(id = rep(participant_id, N_TRIALS), 
+  df <- data.frame(id = rep(participant_id, N_TRIALS),
+                   trial_id = rep(NA, N_TRIALS),
+                   block_id = rep(NA, N_TRIALS),
                    time = rep(NA, N_TRIALS), 
                    distance = rep(NA, N_TRIALS),
                    optimal_distance = rep(NA, N_TRIALS),
@@ -39,10 +40,11 @@ create_walk_df <- function(participant_id){
                    goal = rep(NA, N_TRIALS))
   return(df)
 }
-
 create_sop_df <- function(participant_id){
   N_POINTING <- 12
   df <- data.frame(id = rep(participant_id, N_POINTING), 
+                   trial_id = rep(NA, N_TRIALS),
+                   block_id = rep(NA, N_TRIALS),
                    time = rep(NA, N_POINTING), 
                    error = rep(NA, N_POINTING),
                    start = rep(NA, N_POINTING),

### 2018-03-01 TwoWorlds/twoworld-results.R
closes #17

@@ -16,6 +16,7 @@ sop_results.general <- function (obj, df_results){
   for (trialId in 1:12){
     ls <- sop_trial_results(obj, trialId)
     df_results <- list_to_row(df_results, ls, trialId)
+    df_results <- add_trial_block_id(df_results, trialId)
   }
   return(df_results)
 }
@@ -23,6 +24,7 @@ walk_results.general <- function(obj, df_results){
   for (trialId in 1:18){
     ls <- walk_trial_results(obj, trialId)
     df_results <- list_to_row(df_results, ls, trialId)
+    df_results <- add_trial_block_id(df_results, trialId)
   }
   df_results$distance[df_results$distance == 0] <- NA
   return(df_results)
@@ -43,14 +45,13 @@ create_walk_df <- function(participant_id){
 create_sop_df <- function(participant_id){
   N_POINTING <- 12
   df <- data.frame(id = rep(participant_id, N_POINTING), 
-                   trial_id = rep(NA, N_TRIALS),
-                   block_id = rep(NA, N_TRIALS),
+                   trial_id = rep(NA, N_POINTING),
+                   block_id = rep(NA, N_POINTING),
                    time = rep(NA, N_POINTING), 
                    error = rep(NA, N_POINTING),
                    start = rep(NA, N_POINTING),
                    goal = rep(NA, N_POINTING))
 }
-
 list_to_row <- function(df, ls, iRow){
   df_cols <- colnames(df)
   ls_names <- names(ls)
@@ -58,4 +59,9 @@ list_to_row <- function(df, ls, iRow){
     if(name %in% df_cols) df[iRow, name] <- ls[[name]]
   }
   return(df)
+}
+add_trial_block_id <- function(df_results, trialId){
+  df_results$trial_id[trialId] <- trialId
+  df_results$block_id[trialId] <- floor(trialId/6.01) + 1 #hack that works for up to 100 blocks
+  return(df_results)
 }

### 2018-03-01 Scripts/main-preprocess.R
Adds exp_complete id and block id

@@ -15,6 +15,14 @@ ls <- load_all(settings, dir)
 sop_all <- multi_sop_results(ls)
 walk_all <- multi_walk_results(ls)
 
+### Total trial ----
+walk_all$exp_trial_id <- walk_all$trial_id + (walk_all$phase-1)*18
+walk_all$exp_block_id <- walk_all$block_id + (walk_all$phase-1)*3
+
+sop_all$exp_trial_id <- sop_all$trial_id + (sop_all$phase-1)*18
+sop_all$exp_block_id <- sop_all$block_id + (sop_all$phase-1)*3
+
+### Conditions ----
 conditions <- sop_all %>% select(id, type, phase) %>% unique() %>% group_by(id) %>% summarise(condition = paste(type, collapse = "-"))
 conditions$condition <- gsub('twunity', 'vr', conditions$condition)
 conditions$condition <- gsub('restimote', 'real', conditions$condition)

### 2018-03-01 TwoWorlds/twowold-results-twunity.R
closes #15

@@ -6,7 +6,7 @@ walk_results.twunity <- function(obj){
 walk_trial_results.twunity <- function(obj, trialId){
   ls <- list()
   log <- get_trial_log(obj$walk, trialId)
-  ls$time <- diff(range(log$Time))
+  ls$time <- get_trial_duration(obj, trialId, withoutPauses = T)
   ls$distance <- diff(range(log$cumulative_distance))
   ls$start <- get_trial_start_name.twunity(obj, trialId)
   ls$goal <- get_trial_goal_name.twunity(obj, trialId)

### 2018-03-01 TwoWorlds/twoworld-results-restimote.R

@@ -7,7 +7,7 @@ walk_trial_results.restimote <- function(obj, trialId){
   log <- get_trial_log(obj, trialId)
   log_true <- true_trial_log(obj, trialId, 30, radius = 3)
   ls <- list()
-  ls$time <- diff(range(log$Time)) 
+  ls$time <- get_trial_duration(obj, trialId, withoutPauses = T)
   ls$distance <- sum(log_true$distance)
   ls$start <- get_trial_start_name.restimote(obj, trialId)
   ls$goal <- get_trial_goal_name.restimote(obj, trialId)

### 2018-05-04 Report/progress-report.Rmd
Adds progress report

---
title: "Progress report"
author: "Lukáš 'hejtmy' Hejtmánek"
date: "3 May 2018"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(restimoter)
library(navr)
library(brainvr.R)
source_folder <- function(path){
  sapply(list.files(pattern = "[.]R$", path = path, full.names = TRUE), source);
}
source_folder("../TwoWorlds/")

sop_all <- read.table("../sop.csv", sep=";", header = T, stringsAsFactors = F)
walk_all <- read.table("../walk.csv", sep=";", header = T, stringsAsFactors = F)
load("../multi.data")
load("../settings.data")
df_participants <- settings$participants
df_participants$condition <- paste0(df_participants$First.phase, "-", df_participants$Second.phase)
```

# Results

## Distance

Distance absolute - in Unity units and in 
```{r, echo = F,warning=FALSE}
ggplot(walk_all, aes(x = condition, y = distance, fill=factor(phase))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```

Distance normalised - path traveled (in Unity/real) divided by the shortest path for the particular path.

1 = traveled the exactly as needed, 2 = traveled twice more than needed

```{r, echo = F, warning=FALSE}
ggplot(walk_all, aes(x = condition, y = min_norm_distance, fill=factor(phase))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```

T tests comparing the normalised distance in the first and second phase.

```{r, echo = F}
walk_all %>% 
  group_by(condition) %>% 
  summarise(p.value = t.test(min_norm_distance ~ phase)$p.value,
            phase1 = t.test(min_norm_distance ~ phase)$estimate[[1]],
            phase2 = t.test(min_norm_distance ~ phase)$estimate[[2]])
```

There seems to be little to no effect of the condition on the improvement, acording to a simple anova. Effect of condition is apparent, but the improvement seems across the board, even in the Vr-vr condition, where the reasults are the same as in the real-real in the second phase.

```{r, echo = T}
aov_distance_condition <- aov(min_norm_distance~condition*phase, 
                              data = walk_all)
summary(aov_distance_condition)
```

## Time

Time absolute in seconds

```{r, echo = F,warning=FALSE}
ggplot(walk_all, aes(x = condition, y = time, fill=factor(phase))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```

Time normalised  - time traveled (in Unity/real) divided by the shortest time for the particular path (Jacob's office to Emily's)

1 = traveled the minimum time(exactly as needed), 2 = traveled twice longer than needed, etc
```{r, echo = F,warning=FALSE}
ggplot(walk_all, aes(x = condition, y = min_norm_time, fill=factor(phase))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```

T tests comparing the normalised time in the first and second phase.

```{r, echo = F}
walk_all %>% 
  group_by(condition) %>% 
  summarise(p.value = t.test(min_norm_time ~ phase)$p.value,
            phase1 = t.test(min_norm_time ~ phase)$estimate[[1]],
            phase2 = t.test(min_norm_time ~ phase)$estimate[[2]])
```

Unlike the distance, there seems to be an effect between time and phase - it is probaby due to the vast spread and longer times we can see in the virtual reality, and increase of time needed int the real-vr condition.

```{r, echo = T}
aov_time_condition <- aov(min_norm_time~condition*phase, 
                              data = walk_all)
summary(aov_time_condition)
```

## Pointing accuracy

Pointing accuracy - error in degrees.

It looks like general improvement across the board,m with bigger mistakes in vr-real condition. POssibly caused by outliers, I am looking into that. Or it might be genuinely that the environemtn change is more confusing in that case. I am waiting on the desktop data to see if there are any changes.

```{r, echo = F,warning=FALSE}
ggplot(sop_all, aes(x = condition, y = abs(error), fill=factor(phase))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```

There is an effect, probably caused by the real-real and vr-real conditions. but the second phase perfomance seems better again across the board, even in the vr-vr condition, getting to the level of performance of the real - real

```{r, echo = T}
aov_point_condition <- aov(abs(error)~condition*phase, 
                           data = sop_all)
summary(aov_point_condition)
```

## Number of errors
Number of wrong doors chosen

```{r, echo = F,warning=FALSE}
ggplot(walk_all, aes(x = condition, y = errors, fill=factor(phase))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```

# Per block stats

Condition changes in the 4th block (1-3 1st condtition, 4-6th second condition.) In case of low transfer, we should expect an upward peak in the 4th block

Normalised distance to the goal per block. There are slight (non significant, but still) bumps in the 4th block in vr-real and real-vr conditions.
```{r, echo = F,warning=FALSE}
ggplot(walk_all, aes(x = condition, y = min_norm_distance, fill=factor(exp_block_id))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```

Normalised time per block. There is an evident bump in real-vr condition, probably caused by the "novelty" of the treadmill. Participants trained, but now have been 45 minutes away, so there might be some "getting used to pahse". Despite longer times their number of errors is still lower than during the 3rd trial.

```{r, echo = F,warning=FALSE}
ggplot(walk_all, aes(x = condition, y = min_norm_time, fill=factor(exp_block_id))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```

Number of errors per block. It looks like it is decreasing nicely.

```{r, echo = F,warning=FALSE}
ggplot(walk_all, aes(x = condition, y = errors, fill=factor(exp_block_id))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")

### 2018-05-04 Scripts/main-preprocess.R

@@ -11,6 +11,7 @@ dir <- "D:/OneDrive/Vyzkum/Davis/Transfer/Data/"
 dir <- "C:/Users/hejtm/OneDrive/Vyzkum/Davis/Transfer/Data/"
 
 settings <- load_google_sheets()
+save(settings, file = "settings.data")
 ls <- load_all(settings, dir)
 save(ls, file = "multi.data")
 
@@ -41,10 +42,6 @@ distance_offices_summary <- walk_all %>%
            max.dist = max(distance, na.rm = T)) %>%
   arrange(start, goal)
 
-wide_distance_offices_summary <- distance_offices_summary %>% melt(id.vars=c("type", "start", "goal")) %>% dcast(start + goal ~ type + variable)
-office_distance_type_diff_min <- distance_offices_summary %>% group_by(start, goal) %>% summarise(min_diff = diff(min.dist))
-qplot(data = office_distance_type_diff_min, x = 1:nrow(office_distance_type_diff_min), y = min_diff, geom = "line")
-
 walk_all <- right_join(walk_all, distance_offices_summary[, c(1:3, 6)], by = c("type", "start", "goal"))
 walk_all$min_norm_distance <- walk_all$distance/walk_all$min.dist
 
@@ -58,9 +55,6 @@ time_offices_summary <- walk_all %>%
             max.time = max(time, na.rm = T)) %>%
   arrange(start, goal)
 
-wide_time_offices_summary <- time_offices_summary %>% melt(id.vars=c("type", "start", "goal")) %>% dcast(start + goal ~ type + variable)
-office_time_type_diff_min <- time_offices_summary %>% group_by(start, goal) %>% summarise(min_diff = diff(min.time))
-qplot(data = office_time_type_diff_min, x = 1:nrow(office_time_type_diff_min), y = min_diff, geom = "line")
 walk_all <- right_join(walk_all, time_offices_summary[, c(1:3, 6)], by = c("type", "start", "goal"))
 walk_all$min_norm_time <- walk_all$time/walk_all$min.time
 

### 2018-05-04 TwoWorlds/twoworld-loading.R

@@ -1,6 +1,6 @@
 load_all <- function(settings, dir, select = NULL){
   ls <- list()
-  finished <- settings$versions[settings$versions$finished == "yes" & (settings$versions$is_ok == "yes" | is.na(settings$versions$is_ok)), "Code"][[1]]
+  finished <- settings$versions[settings$versions$finished == "yes" & (settings$versions$is_ok == "yes" & !is.na(settings$versions$is_ok)), "Code"][[1]]
   if(!is.null(select)) finished <- finished[select] 
   for(i in 1:length(finished)){
     code <- finished[i]

### 2018-05-25 Scripts/analysis.R
some participant check changes

@@ -1,4 +1,5 @@
 library(ggplot2)
+source('Scripts/analysis-helpers.R')
 sop_all <- read.table("sop.csv", sep=";", header = T, stringsAsFactors = F)
 walk_all <- read.table("walk.csv", sep=";", header = T, stringsAsFactors = F)
 
@@ -20,35 +21,18 @@ ggplot(sop_all, aes(x = condition, y = abs(error), fill=factor(phase))) +
   stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
   stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
 
-ggplot(walk_all, aes(x = type, y = distance, fill=factor(phase))) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
-  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
+make_graph(sop_all[sop_all$time < 12, ], "condition", "time", "phase")
 
-ggplot(walk_all, aes(x = type, y = time, fill=factor(phase))) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
-  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
+make_graph(sop_all[sop_all$time < 12, ], "condition", "time", "exp_block_id")
+ggplot(sop_all, aes(time))+ geom_histogram()
+ggplot(sop_all[sop_all$time < 12, ], aes(time))+ geom_histogram()
 
-ggplot(walk_all, aes(x = condition, y = distance, fill=factor(phase))) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
-  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
-
-ggplot(walk_all, aes(x = condition, y = min_norm_distance, fill=factor(phase))) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
-  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
-
-ggplot(walk_all, aes(x = condition, y = time, fill=factor(phase))) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
[…]

### 2018-05-26 TwoWorlds/twoworld-loading.R
Adds loadig of unity to ve as well

@@ -34,10 +34,10 @@ load_participant <- function(code, settings, dir){
   line <- get_participant_line(code, settings)
   if(is.null(line)) return(NULL)
   
-  if(line$First.phase == "vr") ls$phase1 <- load_unity(exp_folder, line$WalkTimestamp1, line$SOPTimestamp1)
+  if(line$First.phase %in% c("vr", "ve")) ls$phase1 <- load_unity(exp_folder, line$WalkTimestamp1, line$SOPTimestamp1)
   if(line$First.phase == "real") ls$phase1 <- load_restimote(exp_folder, code, line$WalkTimestamp1, 1, settings)
   
-  if(line$Second.phase == "vr") ls$phase2 <- load_unity(exp_folder, line$WalkTimestamp2, line$SOPTimestamp2)
+  if(line$Second.phase %in% c("vr", "ve")) ls$phase2 <- load_unity(exp_folder, line$WalkTimestamp2, line$SOPTimestamp2)
   if(line$Second.phase == "real") ls$phase2 <- load_restimote(exp_folder, code, line$WalkTimestamp2, 2, settings)
   return(ls)
 }

### 2018-05-26 TwoWorlds/twoworld-loading.R
Adds sop for ve

@@ -59,6 +59,8 @@ load_unity <- function(dir, walk_timestamp, sop_timestamp){
   
   walk <- preprocess_unity_log(walk, dir)
   sop <- preprocess_unity_log(sop, dir)
+  
+  #messes up Y rotation axis - looking up and down
   walk <- transform_unity_coordinates(walk)
   sop <- transform_unity_coordinates(sop)
   
@@ -66,6 +68,7 @@ load_unity <- function(dir, walk_timestamp, sop_timestamp){
   sop$map_limits <- BUILDING_LIMITS
   
   ls <- list(walk = walk, sop = sop)
+  
   class(ls) <-  append(class(ls), "twunity")
   return(ls)
 }

### 2018-05-26 TwoWorlds/twoworld-results-twunity.R

@@ -18,15 +18,23 @@ sop_results.twunity <- function(obj){
   return(sop_results.general(obj, df_results))
 }
 sop_trial_results.twunity <- function(obj, trialId){
-  UNITY_SHIFT <- -90 # WEIRD thing that the controller actually was transformed by this amout, so we need to input it back
   ls_pos <- get_point_start_goal(obj, trialId)
   point_line <- get_trial_point(obj, trialId)
   ls <- as.list(get_rotations(point_line))
-  ls$pointed_angle <- navr::angle_to_360(point_line$Rotation.Controller.x) + UNITY_SHIFT
+  ls$pointed_angle <- ifelse("Rotation.Controller.x" %in% colnames(point_line), get_pointed_angle.vr(point_line), get_pointed_angle.ve(point_line))
   ls$correct_angle <- navr::angle_from_positions(ls_pos$start, ls_pos$goal)
   ls$error <- navr::angle_to_180(ls$correct_angle - ls$pointed_angle)
   ls$time <- brainvr.R::get_trial_duration(obj$sop, trialId, without_pauses = F)
   ls$start <- get_point_start_name(obj, trialId)
   ls$goal <- get_point_goal_name(obj, trialId)
   return(ls)
+}
+get_pointed_angle.vr <- function(point_line){
+  UNITY_SHIFT <- -90 # WEIRD thing that the controller actually was transformed by this amout, so we need to input it back
+  pointed_angle <- navr::angle_to_360(point_line$Rotation.Controller.x) + UNITY_SHIFT
+  return(pointed_angle)
+  
+}
+get_pointed_angle.ve <- function(point_line){
+  navr::angle_to_360(point_line$Rotation.X)
 }

### 2018-05-26 TwoWorlds/twoworld-loading.R
Adds vee categories and separate analysis

@@ -1,7 +1,11 @@
-load_all <- function(settings, dir, select = NULL){
+load_all <- function(settings, dir, select = NULL, only_ok = TRUE){
   ls <- list()
-  finished <- settings$versions[settings$versions$finished == "yes" & (settings$versions$is_ok == "yes" & !is.na(settings$versions$is_ok)), "Code"][[1]]
-  if(!is.null(select)) finished <- finished[select] 
+  if(only_ok){
+    finished <- settings$versions[settings$versions$finished == "yes" & (settings$versions$is_ok == "yes" & !is.na(settings$versions$is_ok)), "Code"][[1]]
+  } else {
+    finished <- settings$versions[settings$versions$finished == "yes", "Code"][[1]]
+  }
+  if(!is.null(select)) finished <- finished[select]
   for(i in 1:length(finished)){
     code <- finished[i]
     print(paste0("Loading participant ", i,  " with code ", code))

### 2018-05-27 Scripts/analysis.R
Adds new analyses of ve data

@@ -1,7 +1,16 @@
 library(ggplot2)
+library(dplyr)
 source('Scripts/analysis-helpers.R')
 sop_all <- read.table("sop.csv", sep=";", header = T, stringsAsFactors = F)
 walk_all <- read.table("walk.csv", sep=";", header = T, stringsAsFactors = F)
+
+# Descriptives ----
+df_missing <- walk_all %>% 
+  group_by(id, phase) %>% 
+  summarise(condition = first(condition), n_na = sum(is.na(distance))) %>%
+  arrange(n_na, id)
+
+# Analysis ----
 
 t.test(abs(sop_all$error)~sop_all$phase)
 t.test(abs(sop_all$time)~sop_all$phase)
@@ -17,22 +26,48 @@ summary(aov_distance_condition)
 aov_time_condition <- aov(time~condition*phase, data = walk_all)
 summary(aov_time_condition)
 
+# Plots ----
+
+## Pointing ----
 ggplot(sop_all, aes(x = condition, y = abs(error), fill=factor(phase))) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
   stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
 
-make_graph(sop_all[sop_all$time < 12, ], "condition", "time", "phase")
+ggplot(sop_all, aes(x = phase, y = abs(error), color = factor(condition))) +
+       stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
+       stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line")
 
+ggplot(sop_all, aes(x = condition, y = abs(error), fill=factor(exp_block_id))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95), geom="bar")
+
+make_graph(sop_all[sop_all$time < 12,], "condition", "time", "phase")
+ 
 make_graph(sop_all[sop_all$time < 12, ], "condition", "time", "exp_block_id")
-ggplot(sop_all, aes(time))+ geom_histogram()
-ggplot(sop_all[sop_all$time < 12, ], aes(time))+ geom_histogram()
+
+## Walking ----
[…]

### 2018-05-27 Scripts/participant-check.R

@@ -10,14 +10,16 @@ dir <- "D:/OneDrive/Vyzkum/Davis/Transfer/Data/"
 
 settings <- load_google_sheets()
 
-code <- "tw125"
+code <- "tw103"
 ls <- load_participant(code, settings, dir)
 sop_results(ls$phase1)
 sop_results(ls$phase2)
 walk_results(ls$phase1) #takes long time and is correct if sop is correct
 walk_results(ls$phase2)
 
-plot_walk_trial(ls$phase2, 9)
+plot_walk_trial(ls$phase1, 5)
+
+plot_all(ls$phase2, 1:12, plot_sop_point)
 
 plot_all(ls$phase2, 1:18, plot_walk_trial)
 

### 2018-05-27 Report/arne-conference.Rmd
Adds report statistics

---
title: "Conference output"
author: "Lukáš 'hejtmy' Hejtmánek"
date: "3 May 2018"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
source('../Scripts/analysis-helpers.R')
sop_all <- read.table("../sop.csv", sep=";", header = T, stringsAsFactors = F)
walk_all <- read.table("../walk.csv", sep=";", header = T, stringsAsFactors = F)
```

```{r}
walk_all %>% group_by(condition) %>% summarise(N = n_distinct(id))
```

## Walking ----

## Per phase separation
```{r warning=F, echo=F}
make_graph(walk_all, "condition", "distance", "phase") + ylab("Average absolute trial distance") + xlab("condition") + scale_fill_discrete(name = "Phase")
make_graph(walk_all, "condition", "min_norm_distance", "phase") + ylab("Average normalised trial distance") + xlab("condition") + scale_fill_discrete(name = "Phase")
make_graph(walk_all, "condition", "time", "phase") + ylab("Average absolute trial duration") + xlab("condition") + scale_fill_discrete(name = "Phase")
make_graph(walk_all, "condition", "min_norm_time", "phase")+ ylab("Average normalised trial duration") + xlab("condition") + scale_fill_discrete(name = "Phase")
make_graph(walk_all, "condition", "errors", "phase")+ ylab("Average number of errors") + xlab("condition") + scale_fill_discrete(name = "Phase")
```

## PER BLOCK
```{r warning=F, echo=F}
make_graph(walk_all, "condition", "min_norm_distance", "exp_block_id") + ylab("Average normalised trial distance") + xlab("condition") + scale_fill_discrete(name = "Block")
make_graph(walk_all, "condition", "min_norm_time", "exp_block_id")+ ylab("Average normalised trial duration") + xlab("condition") + scale_fill_discrete(name = "Block")
make_graph(walk_all, "condition", "errors", "exp_block_id")+ ylab("Average number of errors") + xlab("condition") + scale_fill_discrete(name = "Block")

ggplot(walk_all, aes(x = exp_block_id, y = min_norm_distance, color = factor(condition))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
  ylab("Average normalised trial distance") + xlab("Block") + scale_fill_discrete("condition")

ggplot(walk_all, aes(x = exp_block_id, y = errors, color = factor(condition))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
  ylab("Number of mistakes in office decision") + xlab("Block") + scale_fill_discrete("condition")

```

## Pointing
```{r warning=F, echo=F}
ggplot(sop_all[sop_all$time < 12, ], aes(x = condition, y = abs(error), fill=factor(phase))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
  ylab("Average pointing error in degrees") + xlab("Phase") + scale_fill_discrete("Condition")

ggplot(sop_all[sop_all$time < 12, ], aes(x = phase, y = abs(error), color = factor(condition))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
  ylab("Average pointing error in degrees") + xlab("Phase") + scale_fill_discrete("Condition")

ggplot(sop_all[sop_all$time < 12, ], aes(x = condition, y = time, fill=factor(phase))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
  ylab("Average pointing time") + xlab("Phase") + scale_fill_discrete("Condition")

ggplot(sop_all[sop_all$time < 12, ], aes(x = phase, y = time, color = factor(condition))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
  ylab("Average pointing time") + xlab("Phase") + scale_fill_discrete("Condition")

```

### 2018-05-27 Scripts/analysis.R

@@ -42,7 +42,6 @@ ggplot(sop_all, aes(x = condition, y = abs(error), fill=factor(exp_block_id))) +
   stat_summary(fun.y=mean,position=position_dodge(width=0.95), geom="bar")
 
 make_graph(sop_all[sop_all$time < 12,], "condition", "time", "phase")
- 
 make_graph(sop_all[sop_all$time < 12, ], "condition", "time", "exp_block_id")
 
 ## Walking ----

### 2018-05-28 Report/arne-conference.Rmd
Updates report

@@
 
+#Number of participants
 ```{r}
@@
 make_graph(walk_all, "condition", "min_norm_distance", "exp_block_id") + ylab("Average normalised trial distance") + xlab("condition") + scale_fill_discrete(name = "Block")
-make_graph(walk_all, "condition", "min_norm_time", "exp_block_id")+ ylab("Average normalised trial duration") + xlab("condition") + scale_fill_discrete(name = "Block")
+
+make_graph(walk_all, "condition", "min_norm_time", "exp_block_id") + ylab("Average normalised trial duration") + xlab("condition") + scale_fill_discrete(name = "Block")
+
 make_graph(walk_all, "condition", "errors", "exp_block_id")+ ylab("Average number of errors") + xlab("condition") + scale_fill_discrete(name = "Block")
@@
 
-ggplot(sop_all[sop_all$time < 12, ], aes(x = phase, y = abs(error), color = factor(condition))) +
+ggplot(sop_all[sop_all$time < 12, ], aes(x = phase, y = abs(error), color = factor(type))) +
   stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
@@
 ```
+
+# Block measures
+
+```{r, echo = F}
+condition_all_sums <- walk_all %>% 
+  group_by(exp_block_id, id, condition) %>% 
+  summarise(sum.distance = sum(min_norm_distance, na.rm = T), 
+            sum.errors = sum(errors, na.rm = T),
+            mean.distance = mean(min_norm_distance, na.rm = T),
+            mean.error = mean(errors, na.rm = T))
+
+block_measure <- function(df, value){
+  df_wide <- dcast(df, condition + id ~ exp_block_id, value.var = value)
+  colnames(df_wide) <- c("condition","id", "block1","block2","block3","block4","block5","block6")
+  df_wide %>% 
+    group_by(condition) %>%
+    summarise(block1.mean  = mean(block1, na.rm = T),
+              block3.mean = mean(block3, na.rm = T),
+              block4.mean = mean(block4, na.rm = T),
+              block43diff = sum(block4-block3, na.rm = T)/sum(block4+block3, na.rm = T),
+              block41diff = sum(block4-block1, na.rm = T)/sum(block4+block1, na.rm = T))
+}
+```
+Block 43 diff is as we discussed (block4 - block3)/(block4 + block3) and block41 is the forth and first.
+
+There were several ways how I could calculate it. Eventually, the block 4 data are sums of errors per person and block and average path is the average normalised path for a block per person.
+
+MEasure of overall sum of errors. Mean columns define mean sum of errors for all participants in that block
+```{r, warnings = F}
+df <- block_measure(condition_all_sums, "sum.errors")
+df
+ggplot(melt(df[,1:4]), aes(x = condition, y = value, fill = variable)) + geom_bar(stat = 'identity', position="dodge") + ylab("Mean number of mistakes")
+
+ggplot(melt(df[,c(1,5:6)]), aes(x = condition, y = value, fill = variable)) + geom_bar(stat = 'identity', position="dodge") + ylab("Block improvement")
+```
+
+Measure of mean distance. First three columns define mean normalised distance in each block.
+```{r, warnings = F}
+df <- block_measure(condition_all_sums, "mean.distance")
+df
+ggplot(melt(df[,1:4]), aes(x = condition, y = value, fill = variable)) + geom_bar(stat = 'identity', position="dodge") + ylab("Mean number of mistakes")
+
+ggplot(melt(df[,c(1,5:6)]), aes(x = condition, y = value, fill = variable)) + geom_bar(stat = 'identity', position="dodge") + ylab("Block improvement")
+```

### 2018-05-28 Scripts/analysis.R

@@ -1,4 +1,5 @@
 library(ggplot2)
+library(reshape2)
 library(dplyr)
 source('Scripts/analysis-helpers.R')
 sop_all <- read.table("sop.csv", sep=";", header = T, stringsAsFactors = F)
@@ -26,6 +27,57 @@ summary(aov_distance_condition)
 aov_time_condition <- aov(time~condition*phase, data = walk_all)
 summary(aov_time_condition)
 
+
+## Arne ----
+#' for each dependent measure, (block 4 - block 3) / (block 4+block3) 
+#' so we can compare transfer for ve vs vr vs reality (ie all 5 conditions)
+#' 
+
+## Per office pair ---- 
+condition_offices_sums <- walk_all %>% 
+  group_by(exp_block_id, start, goal, condition) %>% 
+  summarise(sum.distance = sum(min_norm_distance, na.rm = T), 
+            sum.errors = sum(errors, na.rm = T))
+
+### Distances
+condition_distance_sums_wide <- dcast(condition_offices_sums, condition + goal + start ~ exp_block_id, value.var = "sum.distance")
+colnames(condition_distance_sums_wide) <- c("condition","goal","start","block1","block2","block3","block4","block5","block6")
+
+distance_block_measures <- condition_distance_sums_wide %>% 
+  group_by(condition) %>%
+  summarise(block1.mean  = mean(block1, na.rm = T), block3.mean = mean(block3, na.rm = T), block4.mean = mean(block4, na.rm = T),
+            block43diff = sum(block4-block3, na.rm = T)/sum(block4+block3, na.rm = T),
+            block41diff = sum(block4-block1, na.rm = T)/sum(block4+block1, na.rm = T))
+
+### Errors
+condition_errors_sums_wide <- dcast(condition_offices_sums, condition + goal + start ~ exp_block_id, value.var = "sum.errors")
+colnames(condition_errors_sums_wide) <- c("condition","goal","start","block1","block2","block3","block4","block5","block6")
+
+errors_block_measures <- condition_errors_sums_wide %>% 
+  group_by(condition) %>%
+  summarise(block1.mean  = mean(block1, na.rm = T),  block3.mean = mean(block3, na.rm = T), block4.mean = mean(block4, na.rm = T),
+            block43diff = sum(block4-block3, na.rm = T)/sum(block4+block3, na.rm = T),
+            block41diff = sum(block4-block1, na.rm = T)/sum(block4+block1, na.rm = T))
+
[…]

### 2018-05-28 Report/progress-report.Rmd

@@
   stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
+```

### 2018-05-28 Report/arne-conference.Rmd
Fixes ylab

@@
 library(dplyr)
+library(reshape2)
 source('../Scripts/analysis-helpers.R')
@@
 MEasure of overall sum of errors. Mean columns define mean sum of errors for all participants in that block
-```{r, warnings = F}
+```{r, echo=F, warnings = F}
 df <- block_measure(condition_all_sums, "sum.errors")
@@
 Measure of mean distance. First three columns define mean normalised distance in each block.
-```{r, warnings = F}
+```{r, echo=F, warnings = F}
 df <- block_measure(condition_all_sums, "mean.distance")
 df
-ggplot(melt(df[,1:4]), aes(x = condition, y = value, fill = variable)) + geom_bar(stat = 'identity', position="dodge") + ylab("Mean number of mistakes")
+ggplot(melt(df[,1:4]), aes(x = condition, y = value, fill = variable)) + geom_bar(stat = 'identity', position="dodge") + ylab("Mean normalised distance")
 

### 2018-05-29 Report/arne-conference.Rmd
Updates conference graphs

@@
 walk_all <- read.table("../walk.csv", sep=";", header = T, stringsAsFactors = F)
+conference_walk_all <- walk_all %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
+conference_sop_all <- sop_all %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
 ```
@@
 ```{r warning=F, echo=F}
-make_graph(walk_all, "condition", "distance", "phase") + ylab("Average absolute trial distance") + xlab("condition") + scale_fill_discrete(name = "Phase")
-make_graph(walk_all, "condition", "min_norm_distance", "phase") + ylab("Average normalised trial distance") + xlab("condition") + scale_fill_discrete(name = "Phase")
-make_graph(walk_all, "condition", "time", "phase") + ylab("Average absolute trial duration") + xlab("condition") + scale_fill_discrete(name = "Phase")
-make_graph(walk_all, "condition", "min_norm_time", "phase")+ ylab("Average normalised trial duration") + xlab("condition") + scale_fill_discrete(name = "Phase")
-make_graph(walk_all, "condition", "errors", "phase")+ ylab("Average number of errors") + xlab("condition") + scale_fill_discrete(name = "Phase")
+make_graph(conference_walk_all, "condition", "distance", "phase") + ylab("Average absolute trial distance") + xlab("condition") + scale_fill_discrete(name = "Phase")
+make_graph(conference_walk_all, "condition", "min_norm_distance", "phase") + ylab("Average normalised trial distance") + xlab("condition") + scale_fill_discrete(name = "Phase")
+make_graph(conference_walk_all, "condition", "time", "phase") + ylab("Average absolute trial duration") + xlab("condition") + scale_fill_discrete(name = "Phase")
+make_graph(conference_walk_all, "condition", "min_norm_time", "phase")+ ylab("Average normalised trial duration") + xlab("condition") + scale_fill_discrete(name = "Phase")
+make_graph(conference_walk_all, "condition", "errors", "phase")+ ylab("Average number of errors") + xlab("condition") + scale_fill_discrete(name = "Phase")
 ```
@@
 ```{r warning=F, echo=F}
-make_graph(walk_all, "condition", "min_norm_distance", "exp_block_id") + ylab("Average normalised trial distance") + xlab("condition") + scale_fill_discrete(name = "Block")
+make_graph(conference_walk_all, "condition", "min_norm_distance", "exp_block_id") + ylab("Average normalised trial distance") + xlab("condition") + scale_fill_discrete(name = "Block")
 
-make_graph(walk_all, "condition", "min_norm_time", "exp_block_id") + ylab("Average normalised trial duration") + xlab("condition") + scale_fill_discrete(name = "Block")
+make_graph(conference_walk_all, "condition", "min_norm_time", "exp_block_id") + ylab("Average normalised trial duration") + xlab("condition") + scale_fill_discrete(name = "Block")
 
-make_graph(walk_all, "condition", "errors", "exp_block_id")+ ylab("Average number of errors") + xlab("condition") + scale_fill_discrete(name = "Block")
+make_graph(conference_walk_all, "condition", "errors", "exp_block_id")+ ylab("Average number of errors") + xlab("condition") + scale_fill_discrete(name = "Block")
 
-ggplot(walk_all, aes(x = exp_block_id, y = min_norm_distance, color = factor(condition))) +
+ggplot(conference_walk_all, aes(x = exp_block_id, y = min_norm_distance, color = factor(condition))) +
   stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
   stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
-  ylab("Average normalised trial distance") + xlab("Block") + scale_fill_discrete("condition")
+  ylab("Average normalised trial distance") + xlab("Block") + scale_fill_discrete("Condition")
 
-ggplot(walk_all, aes(x = exp_block_id, y = errors, color = factor(condition))) +
+ggplot(conference_walk_all, aes(x = exp_block_id, y = errors, color = factor(condition))) +
   stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
   stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
-  ylab("Number of mistakes in office decision") + xlab("Block") + scale_fill_discrete("condition")
+  ylab("Number of mistakes in office decision") + xlab("Block") + scale_fill_discrete("Condition")
 
@@
 ```{r warning=F, echo=F}
-ggplot(sop_all[sop_all$time < 12, ], aes(x = condition, y = abs(error), fill=factor(phase))) +
+ggplot(conference_sop_all, aes(x = condition, y = abs(error), fill=factor(phase))) +
   stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
   stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
-  ylab("Average pointing error in degrees") + xlab("Phase") + scale_fill_discrete("Condition")
+  ylab("Average pointing error in degrees") + xlab("Condition") + scale_fill_discrete("Phase")
 
-ggplot(sop_all[sop_all$time < 12, ], aes(x = phase, y = abs(error), color = factor(type))) +
+ggplot(conference_sop_all, aes(x = phase, y = abs(error), color = factor(condition))) +
   stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
@@
 
-ggplot(sop_all[sop_all$time < 12, ], aes(x = condition, y = time, fill=factor(phase))) +
+ggplot(conference_sop_all[conference_sop_all$time < 12, ], aes(x = condition, y = time, fill=factor(phase))) +
   stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
   stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
-  ylab("Average pointing time") + xlab("Phase") + scale_fill_discrete("Condition")
+  ylab("Average pointing time") + xlab("Condition") + scale_fill_discrete("Phase")
 
-ggplot(sop_all[sop_all$time < 12, ], aes(x = phase, y = time, color = factor(condition))) +
+ggplot(conference_sop_all[conference_sop_all$time < 12,],aes(x = phase, y = time, color = factor(condition))) +
   stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
@@
 ```{r, echo = F}
-condition_all_sums <- walk_all %>% 
+condition_all_sums <- conference_walk_all %>% 
   group_by(exp_block_id, id, condition) %>% 
@@
 
-MEasure of overall sum of errors. Mean columns define mean sum of errors for all participants in that block
+Measure of overall sum of errors. Mean columns define mean sum of errors for all participants in that block
 ```{r, echo=F, warnings = F}
@@
 df
-ggplot(melt(df[,1:4]), aes(x = condition, y = value, fill = variable)) + geom_bar(stat = 'identity', position="dodge") + ylab("Mean number of mistakes")
+ggplot(melt(df[,1:4]), aes(x = condition, y = value, fill = variable)) + geom_bar(stat = 'identity', position="dodge") + ylab("Mean normalised distance")
+ggplot(melt(df[,c(1,5)]), aes(x = condition, y = value)) + geom_bar(stat = 'identity', position="dodge") + ylab("Mean sum of errors improvement from block 3 to 4")
 
-ggplot(melt(df[,c(1,5:6)]), aes(x = condition, y = value, fill = variable)) + geom_bar(stat = 'identity', position="dodge") + ylab("Block improvement")
+ggplot(melt(df[,c(1,6)]), aes(x = condition, y = value)) + geom_bar(stat = 'identity', position="dodge") + ylab("Mean sum of errors improvement from block 1 to 4")
 ```
@@
 
-ggplot(melt(df[,c(1,5:6)]), aes(x = condition, y = value, fill = variable)) + geom_bar(stat = 'identity', position="dodge") + ylab("Block improvement")
+ggplot(melt(df[,c(1,5)]), aes(x = condition, y = value)) + geom_bar(stat = 'identity', position="dodge") + ylab("Mean normalised distance improvement from block 3 to 4")
+
+ggplot(melt(df[,c(1,6)]), aes(x = condition, y = value)) + geom_bar(stat = 'identity', position="dodge") + ylab("Mean normalised distance improvement from block 1 to 4")
 ```

### 2018-05-30 Report/arne-conference.Rmd
Adds stats

@@
 library(reshape2)
+library(broom)
+
 source('../Scripts/analysis-helpers.R')
@@
               block4.mean = mean(block4, na.rm = T),
-              block43diff = sum(block4-block3, na.rm = T)/sum(block4+block3, na.rm = T),
-              block41diff = sum(block4-block1, na.rm = T)/sum(block4+block1, na.rm = T))
+              block34.diff = sum(block3-block4, na.rm = T)/sum(block4+block3, na.rm = T),
+              block34.se  = sqrt(var((block3-block4)/(block4+block3), na.rm = T)/sum(!is.na(block3-block4))),
+              block14.diff = sum(block1-block4, na.rm = T)/sum(block4+block1, na.rm = T),
+              block14.se = sqrt(var((block1-block4)/(block1+block4), na.rm = T)/sum(!is.na(block1-block4)))
+  )
 }
 ```
-Block 43 diff is as we discussed (block4 - block3)/(block4 + block3) and block41 is the forth and first.
+Block 43 diff is now flipped to (block4 - block3)/(block4 + block3) and block41 is the forth and first.
 
@@
 df
-ggplot(melt(df[,1:4]), aes(x = condition, y = value, fill = variable)) + geom_bar(stat = 'identity', position="dodge") + ylab("Mean normalised distance")
-ggplot(melt(df[,c(1,5)]), aes(x = condition, y = value)) + geom_bar(stat = 'identity', position="dodge") + ylab("Mean sum of errors improvement from block 3 to 4")
+ggplot(melt(df[,1:4]), aes(x = condition, y = value, fill = variable)) + 
+  geom_bar(stat = 'identity', position="dodge") + ylab("Mean normalised distance")
 
-ggplot(melt(df[,c(1,6)]), aes(x = condition, y = value)) + geom_bar(stat = 'identity', position="dodge") + ylab("Mean sum of errors improvement from block 1 to 4")
+ggplot(df, aes(x = condition, y = block34.diff, fill = condition)) + 
+  geom_bar(stat = 'identity', position="dodge") + 
+  geom_errorbar(aes(ymin=block34.diff-block34.se, ymax=block34.diff+block34.se), 
+                width=.2, position=position_dodge(.9)) 
+  ylab("Mean sum of errors improvement from block 3 to 4")
+
+ggplot(df, aes(x = condition, y = block14.diff, fill = condition)) + 
+  geom_bar(stat = 'identity', position="dodge") + 
+  geom_errorbar(aes(ymin=block14.diff-block14.se, ymax=block14.diff+block14.se), 
+                width=.2, position=position_dodge(.9)) +
+  ylab("Mean sum of errors improvement from block 1 to 4")
 ```
@@
 
-ggplot(melt(df[,c(1,5)]), aes(x = condition, y = value)) + geom_bar(stat = 'identity', position="dodge") + ylab("Mean normalised distance improvement from block 3 to 4")
+ggplot(df, aes(x = condition, y = block34.diff, fill = condition)) + 
+  geom_bar(stat = 'identity', position="dodge") + 
+  geom_errorbar(aes(ymin=block34.diff-block34.se, ymax=block34.diff+block34.se), 
+                width=.2, position=position_dodge(.9)) +
+  ylab("Mean normalised distance improvement from block 3 to 4")
 
-ggplot(melt(df[,c(1,6)]), aes(x = condition, y = value)) + geom_bar(stat = 'identity', position="dodge") + ylab("Mean normalised distance improvement from block 1 to 4")
+ggplot(df, aes(x = condition, y = block14.diff, fill = condition)) + 
+  geom_bar(stat = 'identity', position="dodge") + 
+  geom_errorbar(aes(ymin=block14.diff-block14.se, ymax=block14.diff+block14.se), 
+                width=.2, position=position_dodge(.9)) +
+  ylab("Mean normalised distance improvement from block 1 to 4")
 ```
+
+# Stats
+
+## Block T tests
+
+```{r, echo =F}
+block_t_test <- function(df, block1, block2, value){
+  df <- df %>% filter(exp_block_id %in% c(block1,block2))
+  mean.na <- function(data){return(mean(data, na.rm = T))}
+  df_wide <- dcast(df, condition + id ~ exp_block_id, value.var = value, fun.aggregate = mean.na)
+  colnames(df_wide) <- c("condition","id", "block1","block2")
+  df_wide %>% group_by(condition) %>% do(tidy(t.test(.$block1, .$block2, paired = T)))
+}
+
+```
+
+### Block 3-4
+Paired t tests between min normalised distance difference in blocks 3 and 4.
+```{r}
+block_t_test(conference_walk_all, 4, 3,"min_norm_distance")
+```
+
+And for number of errors
+```{r}
+block_t_test(conference_walk_all, 4, 3,"errors")
+```
+
+### Block 1 and 4
+Paired t tests between min normalised distance difference in blocks 1 and 4.
+```{r}
+block_t_test(conference_walk_all, 4, 1,"min_norm_distance")
+```
+And for number of errors
+```{r}
+block_t_test(conference_walk_all, 4, 1,"errors")
+```
+
+### Block 1 and 6
+Paired t tests between min normalised distance difference in blocks 1 and 6.
+```{r}
+block_t_test(conference_walk_all, 6, 1,"min_norm_distance")
+```
+And for number of errors
+```{r}
+block_t_test(conference_walk_all, 6, 1,"errors")
+```
+
+## Anova
+This is simple mixed anova (for those three conditions) I know how to do in R.
+Not sure how to do contrasts and within add subject error variability, nor how to run post hocs.
+I expect Mike will be able to do that in his models for the paper.
+
+```{r}
+aov_block_distance <- aov(min_norm_distance~exp_block_id*condition, conference_walk_all)
+summary(aov_block_distance)
+```

### 2018-05-31 Report/arne-conference.Rmd
More arnes comments

@@
               block4.mean = mean(block4, na.rm = T),
-              block34.diff = sum(block3-block4, na.rm = T)/sum(block4+block3, na.rm = T),
+              block34.diff = mean((block3-block4)/(block4+block3), na.rm = T),
               block34.se  = sqrt(var((block3-block4)/(block4+block3), na.rm = T)/sum(!is.na(block3-block4))),
-              block14.diff = sum(block1-block4, na.rm = T)/sum(block4+block1, na.rm = T),
+              block14.diff = mean((block1-block4)/(block1+block4), na.rm = T),
               block14.se = sqrt(var((block1-block4)/(block1+block4), na.rm = T)/sum(!is.na(block1-block4)))
@@
   geom_errorbar(aes(ymin=block34.diff-block34.se, ymax=block34.diff+block34.se), 
-                width=.2, position=position_dodge(.9)) 
+                width=.2, position=position_dodge(.9)) +
   ylab("Mean sum of errors improvement from block 3 to 4")
@@
 ```
+
+## Independent comparisons
+
+```{r, echo=F}
+df_wide <- dcast(condition_all_sums, condition + id ~ exp_block_id, value.var = "mean.distance")
+colnames(df_wide) <- c("condition","id", "block1","block2","block3","block4","block5","block6")
+df_wide$diff14 <- (df_wide$block1-df_wide$block4)/(df_wide$block1+df_wide$block3)
+```
+
+1 to 4 improvement comparison between VE-real and real-real
+```{r}
+t.test(df_wide$diff14[df_wide$condition=="ve-real"],df_wide$diff14[df_wide$condition=="real-real"])
+
+```
+
+1 to 4 improvement comparison between VR-real and real-real
+```{r}
+t.test(df_wide$diff14[df_wide$condition=="vr-real"],df_wide$diff14[df_wide$condition=="real-real"])  
+```
+
+1 to 4 improvement comparison between VR-real and VE-real
+```{r}
+t.test(df_wide$diff14[df_wide$condition=="vr-real"],df_wide$diff14[df_wide$condition=="ve-real"])
+```

### 2018-06-10 Report/arne-conference.Rmd
updates to smoothing process

@@
 ---
-title: "Conference output"
+title: "Transfer report"
 author: "Lukáš 'hejtmy' Hejtmánek"
-date: "3 May 2018"
+date: "4 June 2018"
 output: html_document
@@
 make_graph(conference_walk_all, "condition", "min_norm_distance", "exp_block_id") + ylab("Average normalised trial distance") + xlab("condition") + scale_fill_discrete(name = "Block")
-
 make_graph(conference_walk_all, "condition", "min_norm_time", "exp_block_id") + ylab("Average normalised trial duration") + xlab("condition") + scale_fill_discrete(name = "Block")
-
 make_graph(conference_walk_all, "condition", "errors", "exp_block_id")+ ylab("Average number of errors") + xlab("condition") + scale_fill_discrete(name = "Block")
-
 ggplot(conference_walk_all, aes(x = exp_block_id, y = min_norm_distance, color = factor(condition))) +
@@
   ylab("Average normalised trial distance") + xlab("Block") + scale_fill_discrete("Condition")
-
 ggplot(conference_walk_all, aes(x = exp_block_id, y = errors, color = factor(condition))) +
@@
 t.test(df_wide$diff14[df_wide$condition=="ve-real"],df_wide$diff14[df_wide$condition=="real-real"])
-
 ```

### 2018-06-10 Report/progress-report.Rmd
updates to smoothing process

@@
 walk_all <- read.table("../walk.csv", sep=";", header = T, stringsAsFactors = F)
-load("../multi.data")
+#load("../multi.data")
 load("../settings.data")
@@
 # Per block stats
+Condition changes in the 4th block (1-3 1st condtition, 4-6th second condition.) In case of low transfer, we should expect an upward peak in the 4th block
 
-Condition changes in the 4th block (1-3 1st condtition, 4-6th second condition.) In case of low transfer, we should expect an upward peak in the 4th block
+Absolute distance to the goal per block.
+```{r, echo = F,warning=FALSE}
+ggplot(walk_all, aes(x = condition, y = distance, fill=factor(exp_block_id))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
+```
 

### 2018-06-10 Report/smoothness-report.Rmd
updates to smoothing process

---
title: "Progress report"
author: "Lukáš 'hejtmy' Hejtmánek"
date: "3 May 2018"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(restimoter)
library(navr)
library(brainvr.R)
source_folder <- function(path){
  sapply(list.files(pattern = "[.]R$", path = path, full.names = TRUE), source);
}
source_folder("../TwoWorlds/")

sop_all <- read.table("../sop.csv", sep=";", header = T, stringsAsFactors = F)
walk_all <- read.table("../walk.csv", sep=";", header = T, stringsAsFactors = F)
walk_smoothed <- read.table("../walk_smoothed.csv", sep=";", header = T, stringsAsFactors = F)
```

Distance absolute - in Unity units and in meteds

```{r, echo = F,warning=FALSE}
ggplot(walk_all, aes(x = condition, y = distance, fill=factor(phase))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```

Distance smoothed absolute

```{r, echo = F,warning=FALSE}
ggplot(walk_smoothed, aes(x = condition, y = distance, fill=factor(phase))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```

Distance normalised

```{r, echo = F, warning=FALSE}
ggplot(walk_all, aes(x = condition, y = min_norm_distance, fill=factor(phase))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```

Normalised smoothed distance

```{r, echo = F, warning=FALSE}
ggplot(walk_smoothed, aes(x = condition, y = min_norm_distance, fill=factor(phase))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```

# Per block stats
Condition changes in the 4th block (1-3 1st condtition, 4-6th second condition.) In case of low transfer, we should expect an upward peak in the 4th block

```{r, echo = F,warning=FALSE}
ggplot(walk_all, aes(x = condition, y = distance, fill=factor(exp_block_id))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") + ggtitle("Absolute distance")
```

```{r, echo = F,warning=FALSE}
ggplot(walk_smoothed, aes(x = condition, y = distance, fill=factor(exp_block_id))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") + ggtitle("Absolute distance - smoothed")
```

```{r, echo = F,warning=FALSE}
ggplot(walk_all, aes(x = condition, y = min_norm_distance, fill=factor(exp_block_id))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")+ ggtitle("Normalised distance to the goal per block")
```

```{r, echo = F,warning=FALSE}
ggplot(walk_smoothed, aes(x = condition, y = min_norm_distance, fill=factor(exp_block_id))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")+ ggtitle("Normalised distance to the goal per block - smoothed")
```

```{r, echo = F,warning=FALSE}
ggplot(walk_all, aes(x = condition, y = errors, fill=factor(exp_block_id))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")+ ggtitle("Number of errors per block.")
```

```{r, echo = F,warning=FALSE}
ggplot(walk_smoothed, aes(x = condition, y = errors, fill=factor(exp_block_id))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")+ ggtitle("Number of errors per block - smoothed.")
```

### 2018-06-10 Scripts/analysis.R

@@ -2,8 +2,8 @@ library(ggplot2)
 library(reshape2)
 library(dplyr)
 source('Scripts/analysis-helpers.R')
-sop_all <- read.table("sop.csv", sep=";", header = T, stringsAsFactors = F)
-walk_all <- read.table("walk.csv", sep=";", header = T, stringsAsFactors = F)
+sop_all <- read.table("sop_new.csv", sep=";", header = T, stringsAsFactors = F)
+walk_all <- read.table("walk_new.csv", sep=";", header = T, stringsAsFactors = F)
 
 # Descriptives ----
 df_missing <- walk_all %>% 
@@ -107,7 +107,10 @@ make_graph(walk_all, "condition", "min_norm_time", "phase")
 make_graph(walk_all, "condition", "errors", "phase")
 
 ### PER BLOCK ----
+
+make_graph(walk_all, "condition", "mean_norm_distance", "exp_block_id")
 make_graph(walk_all, "condition", "min_norm_distance", "exp_block_id")
+make_graph(walk_all, "condition", "norm_distance", "exp_block_id")
 make_graph(walk_all, "condition", "min_norm_time", "exp_block_id")
 make_graph(walk_all, "condition", "errors", "exp_block_id")
 

### 2018-06-10 Scripts/main-preprocess.R

@@ -13,11 +13,15 @@ dir <- "C:/Users/hejtm/OneDrive/Vyzkum/Davis/Transfer/Data/"
 settings <- load_google_sheets()
 save(settings, file = "settings.data")
 ls <- load_all(settings, dir, only_ok = T)
-save(ls, file = "multi.data")
-#load(file = "multi.data")
+save(ls, file = "multi_smoothed_2.data")
+#load(file = "multi_smoothed.data")
 
 sop_all <- multi_sop_results(ls)
 walk_all <- multi_walk_results(ls)
+
+##renames all jesiccas
+walk_all <- walk_all %>% mutate(goal = replace(goal, goal=="Jesicca's office", "Jessica's office"),
+                                start = replace(start, start=="Jesicca's office", "Jessica's office"))
 
 ### Total trial ----
 walk_all$exp_trial_id <- walk_all$trial_id + (walk_all$phase-1)*18
@@ -52,7 +56,6 @@ walk_all <- right_join(walk_all, distance_offices_summary, by = c("type", "start
 walk_all$min_norm_distance <- walk_all$distance/walk_all$min.dist
 walk_all$norm_distance <- (walk_all$distance-walk_all$mean.dist)/walk_all$sd.dist
 
-
 # TIme normalisation
 time_offices_summary <- walk_all %>% 
   filter(time > 10) %>%
@@ -68,10 +71,10 @@ walk_all$min_norm_time <- walk_all$time/walk_all$min.time
 walk_all$norm_time <- (walk_all$time-walk_all$mean.time)/walk_all$sd.time
 
 drop.cols <- c('sd.dist', 'sd.time', 'mean.dist', "mean.time", 'min.dist', 'min.time', 
-               'max.dist', 'max.time', 'optimal_distance', 'na_values', 'na_values_phase')
+               'max.dist', 'max.time', 'optimal_distance')
 walk_all_new <- walk_all %>% select(-one_of(drop.cols))
 
 sop_all$abs_error <- abs(sop_all$error)
 
-write.table(sop_all, file = "sop.csv", sep=";")
-write.table(walk_all_new, file = "walk.csv", sep=";")
+write.table(sop_all, file = "sop_smoothed.csv", sep=";")
+write.table(walk_all_new, file = "walk_smoothed.csv", sep=";")

### 2018-06-10 Scripts/participant-check.R

@@ -10,7 +10,7 @@ dir <- "D:/OneDrive/Vyzkum/Davis/Transfer/Data/"
 
 settings <- load_google_sheets()
 
-code <- "tw103"
+code <- "tw42"
 ls <- load_participant(code, settings, dir)
 sop_results(ls$phase1)
 sop_results(ls$phase2)
@@ -19,9 +19,8 @@ walk_results(ls$phase2)
 
 plot_walk_trial(ls$phase1, 5)
 
-plot_all(ls$phase2, 1:12, plot_sop_point)
-
-plot_all(ls$phase2, 1:18, plot_walk_trial)
+plot_all(ls$phase1, 1:12, plot_sop_point)
+plot_all(ls$phase1, 1:18, plot_walk_trial)
 
 diff(get_action_times.restimote(ls$phase1, "New trial"))
 ls$phase2$log[ls$phase2$log$Action == "pointed","Time"]

### 2018-06-10 Scripts/smoothin-test.R

@@ -1,0 +1,18 @@
+library(brainvr.R)
+library(restimoter)
+library(ggplot2)
+library(dplyr)
+library(reshape2)
+library(googlesheets)
+source("Scripts/loading.R")
+
+load(file = "multi.data")
+
+tw <- ls$tw42
+real <- tw$phase1
+vr <- tw$phase2
+
+get_trial_distance(real, 1)
+plot_trial_path(real, 1)
+get_trial_distance(vr$walk, 1)
+get_trial_log(vr$walk, 1)

### 2018-06-10 TwoWorlds/twoworld-getters-twunity.R

@@ -37,7 +37,7 @@ get_trial_start_goal.twunity <- function(obj, trialId){
 }
 get_trial_errors.twunity <- function(obj, trialId){
   #gets_log
-  DIAMETER <- 0.52
+  DIAMETER <- 1
   n_errors <- ifelse(trialId == 1, -1, -2) #we always get to the start and end door
   ALL_DOORS_POSITIONS <- settings$door_positions #a bit problmatic, fethces stuff from the global env
    # we will get one correct hit at the goal start and end

### 2018-06-10 TwoWorlds/twoworld-loading.R

@@ -64,12 +64,15 @@ load_unity <- function(dir, walk_timestamp, sop_timestamp){
   walk <- preprocess_unity_log(walk, dir)
   sop <- preprocess_unity_log(sop, dir)
   
-  #messes up Y rotation axis - looking up and down
+  # messes up Y rotation axis - looking up and down
   walk <- transform_unity_coordinates(walk)
   sop <- transform_unity_coordinates(sop)
   
   walk$map_limits <- BUILDING_LIMITS
   sop$map_limits <- BUILDING_LIMITS
+  
+  # smooths walking
+  walk <- smooth_unity_log(walk)
   
   ls <- list(walk = walk, sop = sop)
   

### 2018-06-10 TwoWorlds/twoworld-preparing.R

@@ -12,3 +12,8 @@ transform_unity_coordinates <- function(obj){
   obj <- resize_layout(obj, 1/4)
   return(obj)
 }
+
+smooth_unity_log <- function(obj){
+  obj <- brainvr.R::smooth_positions.brainvr(obj, "median", points = 201)
+  return(obj)
+}

### 2018-06-10 TwoWorlds/twoworld-getters-twunity.R
Fixes get trial start id

@@ -1,6 +1,6 @@
 ### twunity ----
 get_trial_start_id.twunity <- function(obj, trialId){
-  if(trialId == 1) return(NULL)
+  if(trialId <= 1) return(NULL)
   return(get_trial_goal_id.twunity(obj, trialId - 1))
 }
 get_trial_start.twunity <- function(obj, trialId){

### 2018-06-10 TwoWorlds/twoworld-optimal-path.R
Preparing for computation of optimal path

@@ -1,0 +1,33 @@
+calculate_optimal_path <- function(office1, office2){
+  
+}
+
+tw <- ls$tw148
+walk <- tw$phase1$walk
+
+goals <- get_all_goal_positions(walk, F)
+start_location <- get_trial_start.twunity(tw$phase1, 1)
+positions <- melt(goals)
+positions <- dcast(positions, L1 ~ variable, value.var = "value")
+positions <- rbind(positions, c("Start", start_location))
+colnames(positions) <- c('name', 'x', 'y')
+positions$x <- as.numeric(positions$x)
+positions$y <- as.numeric(positions$y)
+
+plot_path <- function(points){
+  plt <- navr::create_plot()
+  plt <- brainvr.R::add_limits(plt, BUILDING_LIMITS)
+  plt <- navr::plot_add_path(plt, points$x, points$y)
+  start_end <- list(start = as.numeric(points[1,]), goal = as.numeric(points[nrow(points),]))
+  plt <- navr::plot_add_points(plt, start_end, color = "green")
+  plt <- add_building(plt)
+  return(plt)
+}
+
+ls_positions <-  split(positions[,c(2:3)], seq(nrow(positions)))
+
+plt <- navr::create_plot()
+plt <- brainvr.R::add_limits(plt, BUILDING_LIMITS)
+plt <- navr::plot_add_points(plt, ls_positions, color = "green")
+plt <- add_building(plt)
+plt

### 2018-06-14 Report/progress-report.Rmd
Updates scripts to the almost final poreprocessing stage

@@
 
-sop_all <- read.table("../sop.csv", sep=";", header = T, stringsAsFactors = F)
-walk_all <- read.table("../walk.csv", sep=";", header = T, stringsAsFactors = F)
+sop_all <- read.table("../sop_smoothed.csv", sep=";", header = T, stringsAsFactors = F)
+walk_all <- read.table("../walk_smoothed.csv", sep=";", header = T, stringsAsFactors = F)
 #load("../multi.data")

### 2018-06-14 Report/smoothness-report.Rmd
Updates scripts to the almost final poreprocessing stage

@@
 
-sop_all <- read.table("../sop.csv", sep=";", header = T, stringsAsFactors = F)
+sop_all <- read.table("../sop_smoothed.csv", sep=";", header = T, stringsAsFactors = F)
 walk_all <- read.table("../walk.csv", sep=";", header = T, stringsAsFactors = F)

### 2018-06-14 Report/transfer-report.Rmd
Updates scripts to the almost final poreprocessing stage

--- 2018-06-10 Report/arne-conference.Rmd
@@
 library(ggplot2)
+library(ez)
 library(dplyr)
@@
 # Block measures
-
 ```{r, echo = F}
@@
 ```
-Block 43 diff is now flipped to (block4 - block3)/(block4 + block3) and block41 is the forth and first.
 
-There were several ways how I could calculate it. Eventually, the block 4 data are sums of errors per person and block and average path is the average normalised path for a block per person.
+## Sum of errors
 
@@
 ggplot(melt(df[,1:4]), aes(x = condition, y = value, fill = variable)) + 
-  geom_bar(stat = 'identity', position="dodge") + ylab("Mean normalised distance")
+  geom_bar(stat = 'identity', position="dodge") + ylab("Mean sum of errors")
 
@@
 ```
+
+## Normalised distance
 
@@
 
-## Anova
-This is simple mixed anova (for those three conditions) I know how to do in R.
-Not sure how to do contrasts and within add subject error variability, nor how to run post hocs.
-I expect Mike will be able to do that in his models for the paper.
-
-```{r}
-aov_block_distance <- aov(min_norm_distance~exp_block_id*condition, conference_walk_all)
-summary(aov_block_distance)
-```
-
 ## Independent comparisons
@@
 ```
+
+## Anova
+Simple between subjects anova
+```{r}
+aov_block_distance <- aov(min_norm_distance~exp_block_id*condition, conference_walk_all)
+summary(aov_block_distance)
+```
+
+```{r}
+ez_prepare_block <- function(df, blocks){
+  filtered_data <- df[df$exp_block_id %in% blocks, ]
+  filtered_data <- filtered_data[complete.cases(filtered_data),]
+  has_all <- filtered_data %>% group_by(id) %>% select(id, "exp_block_id") %>% unique() %>% count
+  has_all <- has_all %>% filter(n == 2)
+  filtered_data <- filtered_data %>% filter(id %in% has_all$id)
+  filtered_data$exp_block_id <- factor(filtered_data$exp_block_id)
+  return(filtered_data)
+}
+
+clean_walk_14 <- ez_prepare_block(conference_walk_all, c(1,4))
+clean_walk_34 <- ez_prepare_block(conference_walk_all, c(3,4))
+```
+
+### Normalised distance
+
+Within subjects anova for block 3-4 difference.
+
+```{r, warning = F, echo=F, message=F}
+aov_block_dist_34 <- ezANOVA(
+  data = clean_walk_34,
+  dv = min_norm_distance, 
+  wid = id,
+  within = exp_block_id,
+  between = condition, 
+  type = 3)
+aov_block_dist_34
+
+ggplot(clean_walk_34, aes(x = as.numeric(exp_block_id), y = min_norm_distance, color = factor(condition))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
+  ylab("Average normalised distance") + xlab("Experimental block") + scale_fill_discrete("Condition")
+```
+
+Within subjects anova for block 1-4 difference in minimal path.
+
+```{r, warning = F, echo=F, message=F}
+
+aov_block_dist_14 <- ezANOVA(
+  data = clean_walk_14,
+  dv = min_norm_distance, 
+  wid = id,
+  within = exp_block_id,
+  between = condition,
+  type = 3)
+aov_block_dist_14
+
+ggplot(clean_walk_14, aes(x = as.numeric(exp_block_id), y = min_norm_distance, color = factor(condition))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
+  ylab("Average normalised distance") + xlab("Experimental block") + scale_fill_discrete("Condition")
+```
+
+### NUmber of errors
+Within subjects anova for block 1-4 number of errors.
+
+```{r, warning = F, echo=F, message=F}
+aov_block_err_14 <- ezANOVA(
+  data = clean_walk_14,
+  dv = errors, 
+  wid = id,
+  within = exp_block_id,
+  between = condition,
+  type = 3)
+aov_block_err_14
+
+ggplot(clean_walk_14, aes(x = as.numeric(exp_block_id), y = errors, color = factor(condition))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
+  ylab("Average number of errors") + xlab("Experimental block") + scale_fill_discrete("Condition")
+```
+
+Within subjects anova for block 3-4 number of errors.
+```{r, warning = F, echo=F, message=F}
+aov_block_err_34 <- ezANOVA(
+  data = clean_walk_34,
+  dv = errors, 
+  wid = id,
+  within = exp_block_id,
+  between = condition,
+  type = 3)
+aov_block_err_34
+
+ggplot(clean_walk_34, aes(x = as.numeric(exp_block_id), y = errors, color = factor(condition))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
+  ylab("Average number of errors") + xlab("Experimental block") + scale_fill_discrete("Condition")
+```

### 2018-06-14 Scripts/convert-for-mike.R

@@ -1,0 +1,30 @@
+#[sub id 1]	dv (block 1)	dv (block 2)	dv (block 3)	dv (block 4)	dv (block 5)	dv (block 6)	condition dummy code (0/1/2/3/etc)
+#[sub id 2]	dv (block 1)	dv (block 2)	dv (block 3)	dv (block 4)	dv (block 5)	dv (block 6)	condition dummy code (0/1/2/3/etc)
+#[sub id 3]	dv (block 1)	dv (block 2)	dv (block 3)	dv (block 4)	dv (block 5)	dv (block 6)	condition dummy code (0/1/2/3/etc)
+library(dplyr)
+library(reshape2)
+library(openxlsx)
+
+LEVELS <- c("real-real", "real-vr", "ve-real", "vr-real", "vr-vr")
+
+walk_all <- read.table("walk_smoothed.csv", sep=";", header = T, stringsAsFactors = F)
+df_dist <- walk_all %>% group_by(id, exp_block_id, condition) %>% summarise(mean = mean(min_norm_distance, na.rm = T))
+df_dist <- dcast(df_dist, id + condition ~ exp_block_id, value.var = "mean")
+df_dist$id <- as.numeric(gsub("tw", "", df_dist$id))
+df_dist$condition <- factor(df_dist$condition, levels = LEVELS)
+df_dist$condition <- as.numeric(df_dist$condition)
+wb <- createWorkbook()
+addWorksheet(wb, "Sheet 1")
+writeData(wb, 1, df_dist)
+saveWorkbook(wb, "min_norm_dist.xlsx", overwrite = T)
+
+
+df_error <- walk_all %>% group_by(id, exp_block_id, condition) %>% summarise(mean = mean(min_norm_distance, na.rm = T))
+df_error <- dcast(df_error, id + condition ~ exp_block_id, value.var = "mean")
+df_error$id <- as.numeric(gsub("tw", "", df_error$id))
+df_error$condition <- factor(df_error$condition, levels = LEVELS)
+df_error$condition <- as.numeric(df_error$condition)
+wb <- createWorkbook()
+addWorksheet(wb, "Sheet 1")
+writeData(wb, 1, df_error)
+saveWorkbook(wb, "n_errors.xlsx", overwrite = T)

### 2018-06-14 Scripts/main-preprocess.R

@@ -19,6 +19,7 @@ save(ls, file = "multi_smoothed_2.data")
 sop_all <- multi_sop_results(ls)
 walk_all <- multi_walk_results(ls)
 
+
 ##renames all jesiccas
 walk_all <- walk_all %>% mutate(goal = replace(goal, goal=="Jesicca's office", "Jessica's office"),
                                 start = replace(start, start=="Jesicca's office", "Jessica's office"))
@@ -77,4 +78,5 @@ walk_all_new <- walk_all %>% select(-one_of(drop.cols))
 sop_all$abs_error <- abs(sop_all$error)
 
 write.table(sop_all, file = "sop_smoothed.csv", sep=";")
-write.table(walk_all_new, file = "walk_smoothed.csv", sep=";")
+write.table(walk_all_new, file = "walk_smoothed_errors_06.csv", sep=";")
+

### 2018-06-14 Scripts/smoothin-test.R

@@ -12,7 +12,9 @@ tw <- ls$tw42
 real <- tw$phase1
 vr <- tw$phase2
 
+vr_orig <- ls$tw42$phase2
+
 get_trial_distance(real, 1)
 plot_trial_path(real, 1)
 get_trial_distance(vr$walk, 1)
-get_trial_log(vr$walk, 1)
+plot_trial_path(vr_orig$walk, 1)

### 2018-06-14 TwoWorlds/helpers-results.R

@@ -1,0 +1,9 @@
+ez_prepare <- function(filtered_data, by){
+
+  filtered_data <- filtered_data[complete.cases(filtered_data),]
+  has_all <- filtered_data %>% group_by(id) %>% select(by) %>% count
+  has_all <- has_all %>% filter(n > 4)
+  filtered_data <- filtered_data %>% filter(id %in% has_all$id)
+  filtered_data$id <- factor(filtered_data$id)
+  return(filtered_data)
+}

### 2018-06-14 TwoWorlds/twoworld-optimal-path.R

@@ -1,18 +1,28 @@
 calculate_optimal_path <- function(office1, office2){
   
 }
+prepare_thing <- function(){
+  tw <- ls$tw148
+  walk <- tw$phase1$walk
+  
+  goals <- get_all_goal_positions(walk, F)
+  start_location <- get_trial_start.twunity(tw$phase1, 1)
+  positions <- melt(goals)
+  positions <- dcast(positions, L1 ~ variable, value.var = "value")
+  positions <- rbind(positions, c("Start", start_location))
+  colnames(positions) <- c('name', 'x', 'y')
+  positions$x <- as.numeric(positions$x)
+  positions$y <- as.numeric(positions$y)
+  ls_positions <-  split(positions[,c(2:3)], seq(nrow(positions)))
+  
+  plt <- navr::create_plot()
+  plt <- brainvr.R::add_limits(plt, BUILDING_LIMITS)
+  plt <- navr::plot_add_points(plt, ls_positions, color = "green")
+  plt <- add_building(plt)
+  plt
+  
+}
 
-tw <- ls$tw148
-walk <- tw$phase1$walk
-
-goals <- get_all_goal_positions(walk, F)
-start_location <- get_trial_start.twunity(tw$phase1, 1)
-positions <- melt(goals)
-positions <- dcast(positions, L1 ~ variable, value.var = "value")
-positions <- rbind(positions, c("Start", start_location))
-colnames(positions) <- c('name', 'x', 'y')
-positions$x <- as.numeric(positions$x)
-positions$y <- as.numeric(positions$y)
 
 plot_path <- function(points){
   plt <- navr::create_plot()
@@ -24,10 +34,3 @@ plot_path <- function(points){
   return(plt)
 }
 
-ls_positions <-  split(positions[,c(2:3)], seq(nrow(positions)))
-
-plt <- navr::create_plot()
-plt <- brainvr.R::add_limits(plt, BUILDING_LIMITS)
-plt <- navr::plot_add_points(plt, ls_positions, color = "green")
-plt <- add_building(plt)
-plt

### 2018-06-14 TwoWorlds/twoworld-results-twunity.R

@@ -33,7 +33,6 @@ get_pointed_angle.vr <- function(point_line){
   UNITY_SHIFT <- -90 # WEIRD thing that the controller actually was transformed by this amout, so we need to input it back
   pointed_angle <- navr::angle_to_360(point_line$Rotation.Controller.x) + UNITY_SHIFT
   return(pointed_angle)
-  
 }
 get_pointed_angle.ve <- function(point_line){
   navr::angle_to_360(point_line$Rotation.X)

### 2018-06-18 Report/iNav-data.Rmd
Adds iNav report

---
title: "Transfer report"
author: "Lukáš 'hejtmy' Hejtmánek"
date: "4 June 2018"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(ez)
library(dplyr)
library(reshape2)
library(broom)

source('../Scripts/analysis-helpers.R')
sop_all <- read.table("../sop.csv", sep=";", header = T, stringsAsFactors = F)
walk_all <- read.table("../walk.csv", sep=";", header = T, stringsAsFactors = F)
walk_all$learning.condition <- walk_all$condition
walk_all$learning.condition <- recode(walk_all$learning.condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR")
sop_all$learning.condition <- sop_all$condition
sop_all$learning.condition <- recode(sop_all$learning.condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR")

conference_walk_all <- walk_all %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
conference_sop_all <- sop_all %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
```

```{r setup, include=F}
#green, violet, blue, yellow, pink, dark blue
estimote_pal <- c('#b6d6c1', '#54003d','#93d5f6','#d5d215','#f3aacb','#2d2556')
theme_set(theme_minimal())
```

#Number of participants
```{r}
walk_all %>% group_by(condition) %>% summarise(N = n_distinct(id))
```

## Walking ----

## Per phase separation
```{r warning=F, echo=F}
make_graph(conference_walk_all, "learning.condition", "distance", "phase") + 
  ylab("Average absolute trial distance") + xlab("condition") + 
  scale_fill_manual(values=estimote_pal,"Phase")

make_graph(conference_walk_all, "learning.condition", "min_norm_distance", "phase") + 
  ylab("Average normalised trial distance") + xlab("condition") + 
  scale_fill_manual(values=estimote_pal,"Phase")

make_graph(conference_walk_all, "learning.condition", "time", "phase") + 
  ylab("Average absolute trial duration") + xlab("condition") + 
  scale_fill_manual(values=estimote_pal,"Phase")

make_graph(conference_walk_all, "learning.condition", "min_norm_time", "phase")+ 
  ylab("Average normalised trial duration") + xlab("condition") + 
  scale_fill_manual(values=estimote_pal, "Phase")

make_graph(conference_walk_all, "learning.condition", "errors", "phase")+ 
  ylab("Average number of errors") + xlab("condition") + 
  scale_fill_manual(values=estimote_pal,"Phase")
```

## PER BLOCK
```{r warning=F, echo=F}
make_graph(conference_walk_all, "learning.condition", "distance", "exp_block_id") + 
  ylab("Average absolute trial distance") + xlab("learning.condition") + 
  scale_fill_manual(values=estimote_pal,"Block")

make_graph(conference_walk_all, "learning.condition", "min_norm_distance", "exp_block_id") + 
  ylab("Average normalised trial distance") + xlab("learning.condition") + 
  scale_fill_manual(values=estimote_pal,"Block")

make_graph(conference_walk_all, "learning.condition", "time", "exp_block_id") + 
  ylab("Average absolute trial duration") + xlab("learning.condition") + 
  scale_fill_manual(values=estimote_pal,"Block")

make_graph(conference_walk_all, "learning.condition", "min_norm_time", "exp_block_id") + 
  ylab("Average normalised trial duration") + xlab("learning.condition") + 
  scale_fill_manual(values=estimote_pal,"Block")

make_graph(conference_walk_all, "learning.condition", "errors", "exp_block_id")+ 
  ylab("Average number of errors") + xlab("learning.condition") + 
  scale_fill_manual(values=estimote_pal, "Block")

ggplot(conference_walk_all, aes(x = exp_block_id, y = min_norm_distance, color = factor(learning.condition))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
  ylab("Average normalised trial distance") + xlab("Block") + 
  scale_color_manual(values=estimote_pal,"Condition")

ggplot(conference_walk_all, aes(x = exp_trial_id, y = min_norm_distance, color = factor(learning.condition))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
  ylab("Average normalised trial distance") + xlab("Block") + 
  scale_color_manual(values=estimote_pal,"Condition")

ggplot(conference_walk_all, aes(x = exp_block_id, y = errors, color = factor(learning.condition))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
  ylab("Number of mistakes in office decision") + xlab("Block") + 
  scale_color_manual(values=estimote_pal,"Condition")

```

## Pointing
```{r warning=F, echo=F, message=F}
ggplot(sop_all, aes(x = condition, y = abs(error), fill=factor(phase))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
  ylab("Average pointing error in degrees") + xlab("Condition") + 
  scale_fill_manual(values = estimote_pal, "Phase")

ggplot(sop_all, aes(x = phase, y = abs(error), color = factor(condition))) +
  stat_summary(fun.data=mean_cl_normal, size=1, width = 0.1, geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_identity(), size = 2, geom="line") + 
  ylab("Average pointing error in degrees") + xlab("Phase") + 
  scale_color_manual(values = estimote_pal, "Learning condition")

```

# Block measures
```{r, echo = F}
condition_all_sums <- conference_walk_all %>% 
  group_by(exp_block_id, id, learning.condition) %>% 
  summarise(sum.distance = sum(min_norm_distance, na.rm = T), 
            sum.errors = sum(errors, na.rm = T),
            mean.distance = mean(min_norm_distance, na.rm = T),
            mean.error = mean(errors, na.rm = T))

block_measure <- function(df, value){
  df_wide <- dcast(df, learning.condition + id ~ exp_block_id, value.var = value)
  colnames(df_wide) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")
  df_wide %>% 
    group_by(learning.condition) %>%
    summarise(block1.mean  = mean(block1, na.rm = T),
              block3.mean = mean(block3, na.rm = T),
              block4.mean = mean(block4, na.rm = T),
              block34.diff = mean((block3-block4)/(block4+block3), na.rm = T),
              block34.se  = sqrt(var((block3-block4)/(block4+block3), na.rm = T)/sum(!is.na(block3-block4))),
              block14.diff = mean((block1-block4)/(block1+block4), na.rm = T),
              block14.se = sqrt(var((block1-block4)/(block1+block4), na.rm = T)/sum(!is.na(block1-block4)))
  )
}
```

## Sum of errors

Measure of overall sum of errors. Mean columns define mean sum of errors for all participants in that block
```{r, echo=F, warnings = F}
df <- block_measure(condition_all_sums, "sum.errors")
df
ggplot(melt(df[,1:4]), aes(x = learning.condition, y = value, fill = variable)) + 
  geom_bar(stat = 'identity', position="dodge") + ylab("Mean sum of errors")

ggplot(df, aes(x = learning.condition, y = block34.diff, fill = learning.condition)) + 
  geom_bar(stat = 'identity', position="dodge") + 
  geom_errorbar(aes(ymin=block34.diff-block34.se, ymax=block34.diff+block34.se), 
                width=.2, position=position_dodge(.9)) +
  ylab("Mean sum of errors improvement from block 3 to 4")

ggplot(df, aes(x = learning.condition, y = block14.diff, fill = learning.condition)) + 
  geom_bar(stat = 'identity', position="dodge") + 
  geom_errorbar(aes(ymin=block14.diff-block14.se, ymax=block14.diff+block14.se), 
                width=.2, position=position_dodge(.9)) +
  ylab("Mean sum of errors improvement from block 1 to 4")
```

## Normalised distance

Measure of mean distance. First three columns define mean normalised distance in each block.
```{r, echo=F, warnings = F}
df <- block_measure(condition_all_sums, "mean.distance")
df
ggplot(melt(df[,1:4]), aes(x = learning.condition, y = value, fill = variable)) + geom_bar(stat = 'identity', position="dodge") + ylab("Mean normalised distance")

ggplot(df, aes(x = learning.condition, y = block34.diff, fill = learning.condition)) + 
  geom_bar(stat = 'identity', position="dodge") + 
  geom_errorbar(aes(ymin=block34.diff-block34.se, ymax=block34.diff+block34.se), 
                width=.2, position=position_dodge(.9)) +
  ylab("Mean normalised distance improvement from block 3 to 4")

ggplot(df, aes(x = learning.condition, y = block14.diff, fill = learning.condition)) + 
  geom_bar(stat = 'identity', position="dodge") + 
  geom_errorbar(aes(ymin=block14.diff-block14.se, ymax=block14.diff+block14.se), 
                width=.2, position=position_dodge(.9)) +
  ylab("Mean normalised distance improvement from block 1 to 4")
```

# Stats

## Block T tests

```{r, echo=F}
block_t_test <- function(df, block1, block2, value){
  df <- df %>% filter(exp_block_id %in% c(block1,block2))
  mean.na <- function(data){return(mean(data, na.rm = T))}
  df_wide <- dcast(df, condition + id ~ exp_block_id, value.var = value, fun.aggregate = mean.na)
  colnames(df_wide) <- c("condition","id", "block1","block2")
  df_wide %>% group_by(condition) %>% do(tidy(t.test(.$block1, .$block2, paired = T)))
}
```

### Block 3-4
Paired t tests between min normalised distance difference in blocks 3 and 4.
```{r, echo=F}
block_t_test(conference_walk_all, 4, 3,"min_norm_distance")
```

And for number of errors
```{r, echo=F}
block_t_test(conference_walk_all, 4, 3,"errors")
```

### Block 1-4
Paired t tests between min normalised distance difference in blocks 1 and 4.
```{r, echo=F}
block_t_test(conference_walk_all, 4, 1,"min_norm_distance")
```
And for number of errors
```{r, echo=F}
block_t_test(conference_walk_all, 4, 1,"errors")
```

### Block 1-6
Paired t tests between min normalised distance difference in blocks 1 and 6.
```{r, echo=F}
block_t_test(conference_walk_all, 6, 1,"min_norm_distance")
```
And for number of errors
```{r, echo=F}
block_t_test(conference_walk_all, 6, 1,"errors")
```

## Independent block improvement comparisons between conditions

```{r, echo=F}
df_wide <- dcast(condition_all_sums, learning.condition + id ~ exp_block_id, value.var = "mean.distance")
colnames(df_wide) <- c("learning.condition","id", "block1","block2","block3","block4","block5","block6")
df_wide$diff41 <- (df_wide$block4-df_wide$block1)/(df_wide$block1+df_wide$block3)
df_wide$diff43<- (df_wide$block4-df_wide$block3)/(df_wide$block4+df_wide$block3)
```

### Block 1-4 improvement
VE-real and real-real
```{r, echo=F}
t.test(df_wide$diff41[df_wide$learning.condition=="Desktop"],df_wide$diff41[df_wide$learning.condition=="Real"])
```

VR-real and real-real
```{r, echo=F}
t.test(df_wide$diff41[df_wide$learning.condition=="Treadmill VR"],df_wide$diff41[df_wide$learning.condition=="Real"])  
```

VR-real and VE-real
```{r, echo=F}
t.test(df_wide$diff41[df_wide$learning.condition=="Treadmill VR"],df_wide$diff41[df_wide$learning.condition=="Desktop"])
```

### Block 3-4 improvement
VE-real and real-real
```{r, echo=F}
t.test(df_wide$diff43[df_wide$learning.condition=="Desktop"], df_wide$diff43[df_wide$learning.condition=="Real"])
```

VR-real and real-real
```{r, echo=F}
t.test(df_wide$diff43[df_wide$learning.condition=="Treadmill VR"], df_wide$diff43[df_wide$learning.condition=="Real"])  
```

VR-real and VE-real
```{r, echo=F}
t.test(df_wide$diff43[df_wide$learning.condition=="Treadmill VR"], df_wide$diff43[df_wide$learning.condition=="Desktop"])
```

## Anova
Simple between subjects anova
```{r, echo=F}
aov_block_distance <- aov(min_norm_distance~exp_block_id*learning.condition, conference_walk_all)
summary(aov_block_distance)
```

```{r, echo=F}
ez_prepare_block <- function(df, blocks){
  filtered_data <- df[df$exp_block_id %in% blocks, ]
  filtered_data <- filtered_data[complete.cases(filtered_data),]
  has_all <- filtered_data %>% group_by(id) %>% select(id, "exp_block_id") %>% unique() %>% count
  has_all <- has_all %>% filter(n == 2)
  filtered_data <- filtered_data %>% filter(id %in% has_all$id)
  filtered_data$exp_block_id <- factor(filtered_data$exp_block_id)
  return(filtered_data)
}

clean_walk_14 <- ez_prepare_block(conference_walk_all, c(1,4))
clean_walk_34 <- ez_prepare_block(conference_walk_all, c(3,4))
```

### Normalised distance

Within subjects anova for block 3-4 difference.

```{r, warning = F, echo=F, message=F}
aov_block_dist_34 <- ezANOVA(
  data = clean_walk_34,
  dv = min_norm_distance, 
  wid = id,
  within = exp_block_id,
  between = learning.condition, 
  type = 3)
aov_block_dist_34

ggplot(clean_walk_34, aes(x = as.numeric(exp_block_id), y = min_norm_distance, color = factor(learning.condition))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
  ylab("Average normalised distance") + xlab("Experimental block") + scale_fill_discrete("Condition")
```

Within subjects anova for block 1-4 difference in minimal path.

```{r, warning = F, echo=F, message=F}
aov_block_dist_14 <- ezANOVA(
  data = clean_walk_14,
  dv = min_norm_distance, 
  wid = id,
  within = exp_block_id,
  between = learning.condition,
  type = 3)
aov_block_dist_14

ggplot(clean_walk_14, aes(x = as.numeric(exp_block_id), y = min_norm_distance, color = factor(learning.condition))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
  ylab("Average normalised distance") + xlab("Experimental block") + scale_fill_discrete("Condition")
```

### NUmber of errors
Within subjects anova for block 1-4 number of errors.

```{r, warning = F, echo=F, message=F}
aov_block_err_14 <- ezANOVA(
  data = clean_walk_14,
  dv = errors, 
  wid = id,
  within = exp_block_id,
  between = learning.condition,
  type = 3)
aov_block_err_14

ggplot(clean_walk_14, aes(x = as.numeric(exp_block_id), y = errors, color = factor(learning.condition))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
  ylab("Average number of errors") + xlab("Experimental block") + scale_fill_discrete("Condition")
```

Within subjects anova for block 3-4 number of errors.
```{r, warning = F, echo=F, message=F}
aov_block_err_34 <- ezANOVA(
  data = clean_walk_34,
  dv = errors, 
  wid = id,
  within = exp_block_id,
  between = condition,
  type = 3)
aov_block_err_34

ggplot(clean_walk_34, aes(x = as.numeric(exp_block_id), y = errors, color = factor(condition))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
  ylab("Average number of errors") + xlab("Experimental block") + scale_fill_discrete("Condition")
```

### 2018-06-18 Report/iNav-data.Rmd
Adds new inav things

@@
 library(broom)
-
+library(brainvr.R)
+library(restimoter)
+source("../Scripts/loading.R")
+source_folder("../TwoWorlds")
 source('../Scripts/analysis-helpers.R')
@@
 walk_all$learning.condition <- walk_all$condition
-walk_all$learning.condition <- recode(walk_all$learning.condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR")
+walk_all$learning.condition <- recode(walk_all$learning.condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
 sop_all$learning.condition <- sop_all$condition
-sop_all$learning.condition <- recode(sop_all$learning.condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR")
+sop_all$testing.condition <- sop_all$condition
+sop_all$learning.condition <- recode(sop_all$learning.condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
+sop_all$testing.condition <- recode(sop_all$testing.condition, "real-real" = "Real", "ve-real" = "Real", "vr-real" = "Real", "real-vr" = "Treadmill VR", "vr-vr" = "Treadmill VR")
 
@@
 conference_sop_all <- sop_all %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
+load("../multi_smoothed.data")
 ```
@@
 
-## Walking ----
+# Visualisations
+
+# Walking visualisations
+VR
+
+```{r, echo=F}
+plot_walk_trial(ls$tw67$phase1, 1)
+```
+
+REal world
+```{r, echo=F}
+plot_walk_trial(ls$tw64$phase2, 17)
+```
+
+# Pointing visualisations
+VR
+```{r, echo=F}
+plot_all(ls$tw64$phase1, 1:6, plot_sop_point)
+plot_sop_point(ls$tw64$phase1, 3)
+```
+
+Real wolrd
+```{r, echo=F}
+plot_all(ls$tw64$phase2, 1:6, plot_sop_point)
+plot_sop_point(ls$tw64$phase2, 3)
+```
 
@@
 make_graph(conference_walk_all, "learning.condition", "distance", "exp_block_id") + 
-  ylab("Average absolute trial distance") + xlab("learning.condition") + 
-  scale_fill_manual(values=estimote_pal,"Block")
+  ylab("Average absolute trial distance") + xlab("Learning condition") + 
+  scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F)
 
 make_graph(conference_walk_all, "learning.condition", "min_norm_distance", "exp_block_id") + 
-  ylab("Average normalised trial distance") + xlab("learning.condition") + 
-  scale_fill_manual(values=estimote_pal,"Block")
+  ylab("Average normalised trial distance") +  xlab("Learning condition") + 
+  scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F)
 
 make_graph(conference_walk_all, "learning.condition", "time", "exp_block_id") + 
-  ylab("Average absolute trial duration") + xlab("learning.condition") + 
-  scale_fill_manual(values=estimote_pal,"Block")
+  ylab("Average absolute trial duration") + xlab("Learning condition") + 
+  scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F)
 
 make_graph(conference_walk_all, "learning.condition", "min_norm_time", "exp_block_id") + 
-  ylab("Average normalised trial duration") + xlab("learning.condition") + 
-  scale_fill_manual(values=estimote_pal,"Block")
+  ylab("Average normalised trial duration") + xlab("Learning condition") + 
+  scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F)
 
 make_graph(conference_walk_all, "learning.condition", "errors", "exp_block_id")+ 
-  ylab("Average number of errors") + xlab("learning.condition") + 
-  scale_fill_manual(values=estimote_pal, "Block")
+  ylab("Average number of errors") + xlab("Learning condition") + 
+  scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F)
 
@@
   stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
+  geom_vline(xintercept = 3, linetype=4) +
   ylab("Average normalised trial distance") + xlab("Block") + 
-  scale_color_manual(values=estimote_pal,"Condition")
+  scale_color_manual(values=estimote_pal,"Condition") + guides(fill=F)
 
@@
   stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
-  ylab("Average normalised trial distance") + xlab("Block") + 
-  scale_color_manual(values=estimote_pal,"Condition")
+  geom_vline(xintercept = 18, linetype=4) +
+  ylab("Average normalised trial distance") + xlab("Block number") + 
+  scale_color_manual(values=estimote_pal, "Condition")
 
@@
   stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
-  ylab("Number of mistakes in office decision") + xlab("Block") + 
-  scale_color_manual(values=estimote_pal,"Condition")
+  geom_vline(xintercept = 3, linetype=4) +
+  ylab("Number of mistakes in office decision") + xlab("Block number") + 
+  scale_color_manual(values=estimote_pal, "Condition")
 
@@
 
+```
+
+## Average pointing in all conditions
+```{r, echo=F, warnings = F}
+ggplot(sop_all[sop_all$phase == 1, ], aes(x = condition, y = abs(error), fill=factor(condition))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
+  ylab("Average pointing error in degrees in the first phase") + xlab("Condition") + 
+  scale_fill_manual(values = estimote_pal, "Phase")
+
+ggplot(sop_all[sop_all$phase == 2, ], aes(x = condition, y = abs(error), fill=factor(condition))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
+  ylab("Average pointing error in degrees in second phase") + xlab("Condition") + 
+  scale_fill_manual(values = estimote_pal) + guides(fill=F)
+
+ggplot(sop_all[sop_all$phase == 1, ], aes(x = learning.condition, y = abs(error), fill=factor(learning.condition))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
+  ylab("Average pointing error in degrees in the first phase") + xlab("Condition") + 
+  scale_fill_manual(values = estimote_pal) + guides(fill=F)
+
+ggplot(sop_all[sop_all$phase == 2, ], aes(x = testing.condition, y = abs(error), fill=factor(testing.condition))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
+  ylab("Average pointing error in degrees in second phase") + xlab("Condition") + 
+  scale_fill_manual(values = estimote_pal) + guides(fill=F)
+
+#t.test(abs(error) ~ learning.condition, sop_all[sop_all$phase == 1,])
+t.test(abs(error) ~ testing.condition, sop_all[sop_all$phase == 2,])
 ```
@@
 ## Sum of errors
-
 Measure of overall sum of errors. Mean columns define mean sum of errors for all participants in that block
@@
 ggplot(melt(df[,1:4]), aes(x = learning.condition, y = value, fill = variable)) + 
-  geom_bar(stat = 'identity', position="dodge") + ylab("Mean sum of errors")
+  geom_bar(stat = 'identity', position="dodge") + 
+  scale_fill_manual(values = estimote_pal) +
+  ylab("Mean sum of errors") + xlab("Learning condition")  + guides(fill=F)
 
@@
   geom_bar(stat = 'identity', position="dodge") + 
-  geom_errorbar(aes(ymin=block34.diff-block34.se, ymax=block34.diff+block34.se), 
-                width=.2, position=position_dodge(.9)) +
-  ylab("Mean sum of errors improvement from block 3 to 4")
+  geom_errorbar(aes(ymin=block34.diff-block34.se, ymax=block34.diff+block34.se), width=.2, position=position_dodge(.9)) +
+  scale_fill_manual(values = estimote_pal) +
+  ylab("Mean sum of errors improvement after switch")+ xlab("Learning condition") + guides(fill=F)
 
@@
   geom_bar(stat = 'identity', position="dodge") + 
-  geom_errorbar(aes(ymin=block14.diff-block14.se, ymax=block14.diff+block14.se), 
-                width=.2, position=position_dodge(.9)) +
-  ylab("Mean sum of errors improvement from block 1 to 4")
+  geom_errorbar(aes(ymin=block14.diff-block14.se, ymax=block14.diff+block14.se), width=.2, position=position_dodge(.9)) +
+  scale_fill_manual(values = estimote_pal) +
+  ylab("Mean sum of errors improvement from block 1 to 4") + xlab("Learning condition")  + guides(fill=F)
 ```
@@
 df
-ggplot(melt(df[,1:4]), aes(x = learning.condition, y = value, fill = variable)) + geom_bar(stat = 'identity', position="dodge") + ylab("Mean normalised distance")
+ggplot(melt(df[,1:4]), aes(x = learning.condition, y = value, fill = variable)) + 
+  scale_fill_manual(values = estimote_pal) +
+  geom_bar(stat = 'identity', position="dodge") + ylab("Mean normalised distance")
 
@@
   geom_bar(stat = 'identity', position="dodge") + 
-  geom_errorbar(aes(ymin=block34.diff-block34.se, ymax=block34.diff+block34.se), 
-                width=.2, position=position_dodge(.9)) +
-  ylab("Mean normalised distance improvement from block 3 to 4")
+  geom_errorbar(aes(ymin=block34.diff-block34.se, ymax=block34.diff+block34.se), width=.2, position=position_dodge(.9)) +
+  scale_fill_manual(values = estimote_pal) +
+  ylab("Mean normalised distance improvement from block 3 to 4") + xlab("Learning condition")  + guides(fill=F)
 
@@
   geom_bar(stat = 'identity', position="dodge") + 
-  geom_errorbar(aes(ymin=block14.diff-block14.se, ymax=block14.diff+block14.se), 
-                width=.2, position=position_dodge(.9)) +
-  ylab("Mean normalised distance improvement from block 1 to 4")
+  geom_errorbar(aes(ymin=block14.diff-block14.se, ymax=block14.diff+block14.se), width=.2, position=position_dodge(.9)) +
+  scale_fill_manual(values = estimote_pal) +
+  ylab("Mean normalised distance improvement from block 1 to 4") + xlab("Learning condition")  + guides(fill=F)
 ```
@@
 ## Block T tests
-
+Paired block comparisons for improvement in different condiutions
 ```{r, echo=F}

### 2018-06-19 Report/iNav-data.Rmd
Adds 3-4 bloÃcks

@@
 load("../multi_smoothed.data")
-```
 
-```{r setup, include=F}
 #green, violet, blue, yellow, pink, dark blue
@@
 
-## Independent block improvement comparisons between conditions
+## Independent block improvement comparisons between conditions for distance
 
@@
 ### Block 3-4 improvement
+Desktop to real
+```{r, echo=F}
+t.test(df_wide$diff43[df_wide$learning.condition=="Desktop"], df_wide$diff43[df_wide$learning.condition=="Real"])
+```
+
+VR-real and real-real
+```{r, echo=F}
+t.test(df_wide$diff43[df_wide$learning.condition=="Treadmill VR"], df_wide$diff43[df_wide$learning.condition=="Real"])  
+```
+
+VR-real and VE-real
+```{r, echo=F}
+t.test(df_wide$diff43[df_wide$learning.condition=="Treadmill VR"], df_wide$diff43[df_wide$learning.condition=="Desktop"])
+```
+
+## Independent block improvement comparisons between conditions for number of errors
+
+```{r, echo=F}
+df_wide <- dcast(condition_all_sums, learning.condition + id ~ exp_block_id, value.var = "sum.errors")
+colnames(df_wide) <- c("learning.condition","id", "block1","block2","block3","block4","block5","block6")
+df_wide$diff41 <- (df_wide$block4-df_wide$block1)/(df_wide$block1+df_wide$block3)
+df_wide$diff43<- (df_wide$block4-df_wide$block3)/(df_wide$block4+df_wide$block3)
+```
+
+### Block 1-4 improvement
 VE-real and real-real
+```{r, echo=F}
+t.test(df_wide$diff41[df_wide$learning.condition=="Desktop"],df_wide$diff41[df_wide$learning.condition=="Real"])
+```
+
+VR-real and real-real
+```{r, echo=F}
+t.test(df_wide$diff41[df_wide$learning.condition=="Treadmill VR"],df_wide$diff41[df_wide$learning.condition=="Real"])  
+```
+
+VR-real and VE-real
+```{r, echo=F}
+t.test(df_wide$diff41[df_wide$learning.condition=="Treadmill VR"],df_wide$diff41[df_wide$learning.condition=="Desktop"])
+```
+
+### Block 3-4 improvement
+Desktop to real
 ```{r, echo=F}

### 2018-08-08 Report/paper-data.Rmd
Adds paper data rmd

--- 2018-06-19 Report/iNav-data.Rmd
@@
 library(restimoter)
+
 source("../Scripts/loading.R")

### 2018-08-08 Scripts/analysis.R

@@ -2,8 +2,8 @@ library(ggplot2)
 library(reshape2)
 library(dplyr)
 source('Scripts/analysis-helpers.R')
-sop_all <- read.table("sop_new.csv", sep=";", header = T, stringsAsFactors = F)
-walk_all <- read.table("walk_new.csv", sep=";", header = T, stringsAsFactors = F)
+sop_all <- read.table("sop.csv", sep=";", header = T, stringsAsFactors = F)
+walk_all <- read.table("walk.csv", sep=";", header = T, stringsAsFactors = F)
 
 # Descriptives ----
 df_missing <- walk_all %>% 

### 2018-08-09 Scripts/mike-analyses.R
Adds mike's scrips

@@ -1,0 +1,57 @@
+#################################################
+#################################################
+##### VR-RW transfer (Lukas) ####################
+##### Mixed Effects Modeling ####################
+##### Author: Mike Starrett  ####################
+#################################################
+#################################################
+
+#################################################
+# Set path and load data
+#################################################
+
+# load ; delimited csv files
+walk <- read.csv('walk.csv', header = TRUE, sep = ";")
+sop <- read.csv('sop.csv', header = TRUE, sep = ";")
+
+#################################################
+# SOP Analyses
+#################################################
+
+sop$phase <- as.factor(sop$phase)
+sop <- sop[!is.na(sop$error),]
+
+#---------------------------------
+# ANOVAs
+#---------------------------------
+# aov()
+m1 <- aov(abs_error ~ phase*condition + Error(id/phase), data = sop)
+summary(m1)
+  # phase:  p < .001; condition: p = .166; Interaction: p = .022
+
+# ezANOVA
+library(ez)
+m2 <- ezANOVA(data = sop, dv = abs_error, wid = id, within = phase, between = condition, type = 3)
+m2
+# phase:  p < .001; condition: p = .160; Interaction: p = .021
+
+# lme4
+# Mixed effects  models
+library(lme4)
+m3 <- lmer(abs_error ~ phase*condition + (1 | id/phase), data = sop)
+  # returns error warning about maodel being nearly unidentifiable
+summary(m3)
+
+library(car)
+Anova(m3, type = "III")
+  # phase:  p < .567; condition: p = .005; Interaction: p = .017
+    #NOTE: Using type II vs III flips the Main Effect significance, but doesn't change Interaction p-value
+
+# how about nlm 
+library(nlme)
+m4 <- lme(abs_error ~ phase*condition, random = ~1 | id/phase, data = sop)
+summary(m4)
+Anova(m4, type = "III")
+  # phase:  p < .567; condition: p = .005; Interaction: p = .017
+    #NOTE: Same as lme4 wrt significance of effects and Type II vs III sum of square
+

### 2018-08-09 Report/iNav-data.Rmd
Small changes

@@
 walk_all <- read.table("../walk.csv", sep=";", header = T, stringsAsFactors = F)
+
 walk_all$learning.condition <- walk_all$condition

### 2018-08-09 Report/paper-data.Rmd
Small changes

@@
 conference_sop_all <- sop_all %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
-load("../multi_smoothed.data")
+load("../multi.data")
 

### 2018-08-10 Report/arne-report.Rmd
Adds arne's report

---
title: "Arne report"
author: "Lukáš Hejtmánek"
date: "9 August 2018"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(ez)
library(Hmisc)
library(dplyr)
library(reshape2)
library(broom)
library(knitr)
library(brainvr.R)
library(restimoter)
library(papaja)

source("../Scripts/loading.R")
source_folder("../TwoWorlds")
source('../Scripts/analysis-helpers.R')
sop_all <- read.table("../sop.csv", sep=";", header = T, stringsAsFactors = F)
walk_all <- read.table("../walk.csv", sep=";", header = T, stringsAsFactors = F)
walk_all$learning.condition <- walk_all$condition
walk_all$learning.condition <- recode(walk_all$learning.condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
sop_all$learning.condition <- sop_all$condition
sop_all$testing.condition <- sop_all$condition
sop_all$learning.condition <- recode(sop_all$learning.condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
sop_all$testing.condition <- recode(sop_all$testing.condition, "real-real" = "Real", "ve-real" = "Real", "vr-real" = "Real", "real-vr" = "Treadmill VR", "vr-vr" = "Treadmill VR")

conference_walk_all <- walk_all %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
conference_sop_all <- sop_all %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))

#green, violet, blue, yellow, pink, dark blue
estimote_pal <- c('#b6d6c1', '#54003d','#93d5f6','#d5d215','#f3aacb','#2d2556')
theme_set(theme_minimal())

condition_all_sums <- conference_walk_all %>% 
  group_by(exp_block_id, id, learning.condition) %>% 
  summarise(sum.distance = sum(min_norm_distance, na.rm = T), 
            sum.errors = sum(errors, na.rm = T),
            mean.distance = mean(min_norm_distance, na.rm = T),
            mean.error = mean(errors, na.rm = T))

block_measure <- function(df, value){
  df_wide <- dcast(df, learning.condition + id ~ exp_block_id, value.var = value)
  colnames(df_wide) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")
  df_wide %>% 
    group_by(learning.condition) %>%
    summarise(block1.mean  = mean(block1, na.rm = T),
              block3.mean = mean(block3, na.rm = T),
              block4.mean = mean(block4, na.rm = T),
              block34.diff = mean((block3-block4)/(block4+block3), na.rm = T),
              block34.se  = sqrt(var((block3-block4)/(block4+block3), na.rm = T)/sum(!is.na(block3-block4))),
              block14.diff = mean((block1-block4)/(block1+block4), na.rm = T),
              block14.se = sqrt(var((block1-block4)/(block1+block4), na.rm = T)/sum(!is.na(block1-block4)))
  )
}

```

Unless stated otherwise, I ran all tests on the normalised distance and time statistics. Errors and pointing information is not normalised, as they are the same in all conditions.

I only selected those "important" comparisons - therefore Desktop to Real, Treadmill VR to real and Real to real. Anovas and other comparisons are run only on these.

## Descriptives
In normalised distance we can see similar starting point (not significant starting difference `r apa_print(aov(min_norm_distance ~ learning.condition, conference_walk_all[conference_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`)

```{r, echo=F, warning=F}
make_graph(conference_walk_all, "learning.condition", "min_norm_distance", "exp_block_id") + 
  ylab("Average normalised trial distance") + xlab("condition") + 
  scale_fill_manual(values=estimote_pal,"Block ID")

ggplot(conference_walk_all, aes(x = exp_block_id, y = min_norm_distance, color = factor(learning.condition))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
  geom_vline(xintercept = 3.5, linetype=4) +
  ylab("Average normalised trial distance") + xlab("Block") + 
  scale_color_manual(values=estimote_pal,"Condition") + guides(fill=F)
```

There is difference in starting times in all conditions in the first block (`r apa_print(aov(min_norm_time ~ learning.condition, conference_walk_all[conference_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`)). Probably due to learning issues with the treadmill. Threre is no significant difference in desktop and real (`r apa_print(t.test(min_norm_time ~ learning.condition, conference_walk_all[conference_walk_all$exp_block_id == 1 & conference_walk_all$learning.condition %in% c("Desktop", "Real"), ]))$statistic`))

```{r, echo=F}
make_graph(conference_walk_all, "learning.condition", "min_norm_time", "exp_block_id")+ 
  ylab("Average normalised trial duration") + xlab("condition") + 
  scale_fill_manual(values=estimote_pal, "Block ID")
```

And we don't have any significant difference in number of errors in the first block (`r apa_print(aov(errors ~ learning.condition, conference_walk_all[conference_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`)).

```{r, echo=F}
make_graph(conference_walk_all, "learning.condition", "errors", "exp_block_id")+ 
  ylab("Average number of errors") + xlab("condition") + 
  scale_fill_manual(values=estimote_pal," Block ID")

ggplot(conference_walk_all, aes(x = exp_block_id, y = errors, color = factor(learning.condition))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
  geom_vline(xintercept = 3.5, linetype=4) +
  ylab("Number of mistakes in office decision") + xlab("Block number") + 
  scale_color_manual(values=estimote_pal, "Condition")
```

We can clearly see differences in the progression of the learnign, but as you pointed out many times, as they are not linear, we could benefit from non linear modeling. I hope Mike and professor Ferer will help us with that.

But we can look at either pure changes between block 3-4 or 1-4 or those calculations that you proposed as a (block4-block3)/(block3+block4) measurements.

```{r, echo=F}
df_wide_errors <- dcast(condition_all_sums, learning.condition + id ~ exp_block_id, value.var = "sum.errors")
df_wide_distance <- dcast(condition_all_sums, learning.condition + id ~ exp_block_id, value.var = "mean.distance")
colnames(df_wide_errors) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")
colnames(df_wide_distance) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")

df_wide_errors$block14 <- df_wide_errors$block1-df_wide_errors$block4
df_wide_errors$block34 <- df_wide_errors$block3-df_wide_errors$block4
df_wide_distance$block14 <- df_wide_distance$block1-df_wide_distance$block4
df_wide_distance$block34 <- df_wide_distance$block3-df_wide_distance$block4

df_errors <- block_measure(condition_all_sums, "sum.errors")
df_distance <- block_measure(condition_all_sums, "mean.distance")

block_t_test <- function(df, block1, block2, value){
  df <- df %>% filter(exp_block_id %in% c(block1,block2))
  mean.na <- function(data){return(mean(data, na.rm = T))}
  df_wide <- dcast(df, condition + id ~ exp_block_id, value.var = value, fun.aggregate = mean.na)
  colnames(df_wide) <- c("condition","id", "block1","block2")
  df_wide %>% group_by(condition) %>% do(tidy(t.test(.$block1, .$block2, paired = T)))
}
```

## Block improvement from block 1 to block 4

### Distance
We can see that all conditions improved from the first to forth block in both distance walked and errors in decisions. Paired t-tests show that both Treadmill and Real learning conditions led to improved performance from 1-4, but not Desktop. All conditions show improvemt in number of errors.

```{r, echo=F}
df_dist <- block_t_test(conference_walk_all, 1, 4,"min_norm_distance")[,c(1,4)]
df_error <- block_t_test(conference_walk_all, 1, 4,"errors")[,c(1,4)]
df14 <- merge(df_dist, df_error, by = "condition")
colnames(df14) <- c("condition", "p-value distance improvement", "p-value errors improvement")
kable(df14)
```

But anova shows no significant difference in the improvement between groups for distance (`r apa_print(aov(block14 ~ learning.condition, df_wide_distance))$statistic$learning_condition`)  nor errors (`r apa_print(aov(block14 ~ learning.condition, df_wide_errors))$statistic$learning_condition`). This might be due to quite large variability in the sample.

```{r, echo=F}
ggplot(df_distance, aes(x = learning.condition, y = block14.diff, fill = learning.condition)) + 
  geom_bar(stat = 'identity', position="dodge") + 
  geom_errorbar(aes(ymin=block14.diff-block14.se, ymax=block14.diff+block14.se), width=.2, position=position_dodge(.9)) +
  scale_fill_manual(values = estimote_pal) +
  ylab("Mean normalised distance improvement from block 1 to 4") + xlab("Learning condition")  + guides(fill=F)
```

### Errors
```{r, echo=F}
ggplot(df_errors, aes(x = learning.condition, y = block14.diff, fill = learning.condition)) + 
  geom_bar(stat = 'identity', position="dodge") + 
  geom_errorbar(aes(ymin=block14.diff-block14.se, ymax=block14.diff+block14.se), width=.2, position=position_dodge(.9)) +
  scale_fill_manual(values = estimote_pal) +
  ylab("Mean sum of errors improvement from block 1 to 4") + xlab("Learning condition")  + guides(fill=F)
```

## Block improvement from block 3 to block 4

We can see that everybody (including Real to Real condition) gets a little bit worse from 3 to 4. But this "decrease of performance" is not significant when computed using paired t-tests except marginally worse in the desktop version. All conditions show imprvement in the number of errors in paired t-tests.

```{r, echo=F}
df_dist <- block_t_test(conference_walk_all, 3, 4,"min_norm_distance")[,c(1,4)]
df_error <- block_t_test(conference_walk_all, 3, 4,"errors")[,c(1,4)]
df34 <- merge(df_dist, df_error, by = "condition")
colnames(df34) <- c("condition", "p-value distance improvement", "p-value errors improvement")
kable(df34)
```

Comparing the block improvement in different conditions using anovas, we see significant difference in the improvement between groups for distance (`r apa_print(aov(block34 ~ learning.condition, df_wide_distance))$statistic$learning_condition`)  and for errors (`r apa_print(aov(block34 ~ learning.condition, df_wide_errors))$statistic$learning_condition`).

Runnign Tukey's post hoc on the distance shows marginally significant difference between Desktop from the other two conditions for distance.

```{r}
kable(TukeyHSD(x=aov(block34 ~ learning.condition, df_wide_distance), conf.level=0.95)$learning.condition)
```

And si
```{r}
kable(TukeyHSD(x=aov(block34 ~ learning.condition, df_wide_errors), conf.level=0.95)$learning.condition)
```

### Distance
```{r, echo=F}
ggplot(df_distance, aes(x = learning.condition, y = block34.diff, fill = learning.condition)) + 
  geom_bar(stat = 'identity', position="dodge") + 
  geom_errorbar(aes(ymin=block34.diff-block34.se, ymax=block34.diff+block34.se), width=.2, position=position_dodge(.9)) +
  scale_fill_manual(values = estimote_pal) +
  ylab("Mean normalised distance improvement from block 3 to 4") + xlab("Learning condition")  + guides(fill=F)
```

### Errors
```{r, echo=F}
ggplot(df_errors, aes(x = learning.condition, y = block34.diff, fill = learning.condition)) + 
  geom_bar(stat = 'identity', position="dodge") + 
  geom_errorbar(aes(ymin=block34.diff-block34.se, ymax=block34.diff+block34.se), width=.2, position=position_dodge(.9)) +
  scale_fill_manual(values = estimote_pal) +
  ylab("Mean sum of errors improvement after switch")+ xlab("Learning condition") + guides(fill=F)
```

### 2018-08-10 Report/paper-data.Rmd
Adds arne's report

@@
 library(ez)
+library(Hmisc)
+library(knitr)
 library(dplyr)
@@
 library(brainvr.R)
+library(xtable)
 library(restimoter)
@@
 conference_sop_all <- sop_all %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
-load("../multi.data")
 
@@
 theme_set(theme_minimal())
-```
-
-#Number of participants
-```{r}
-walk_all %>% group_by(condition) %>% summarise(N = n_distinct(id))
-```
-
-# Visualisations
-
-# Walking visualisations
-VR
-
-```{r, echo=F}
-plot_walk_trial(ls$tw67$phase1, 1)
-```
-
-REal world
-```{r, echo=F}
-plot_walk_trial(ls$tw64$phase2, 17)
-```
-
-# Pointing visualisations
-VR
-```{r, echo=F}
-plot_all(ls$tw64$phase1, 1:6, plot_sop_point)
-plot_sop_point(ls$tw64$phase1, 3)
-```
-
-Real wolrd
-```{r, echo=F}
-plot_all(ls$tw64$phase2, 1:6, plot_sop_point)
-plot_sop_point(ls$tw64$phase2, 3)
 ```
@@
   stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
-  geom_vline(xintercept = 3, linetype=4) +
+  geom_vline(xintercept = 3.5, linetype=4) +
   ylab("Average normalised trial distance") + xlab("Block") + 
@@
   stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
-  geom_vline(xintercept = 18, linetype=4) +
+  geom_vline(xintercept = 18.5, linetype=4) +
   ylab("Average normalised trial distance") + xlab("Block number") + 
@@
   stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
-  geom_vline(xintercept = 3, linetype=4) +
+  geom_vline(xintercept = 3.5, linetype=4) +
   ylab("Number of mistakes in office decision") + xlab("Block number") + 
@@
 
-```
-
-## Average pointing in all conditions
-```{r, echo=F, warnings = F}
-ggplot(sop_all[sop_all$phase == 1, ], aes(x = condition, y = abs(error), fill=factor(condition))) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
-  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
-  ylab("Average pointing error in degrees in the first phase") + xlab("Condition") + 
-  scale_fill_manual(values = estimote_pal, "Phase")
-
-ggplot(sop_all[sop_all$phase == 2, ], aes(x = condition, y = abs(error), fill=factor(condition))) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
-  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
-  ylab("Average pointing error in degrees in second phase") + xlab("Condition") + 
-  scale_fill_manual(values = estimote_pal) + guides(fill=F)
-
-ggplot(sop_all[sop_all$phase == 1, ], aes(x = learning.condition, y = abs(error), fill=factor(learning.condition))) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
-  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
-  ylab("Average pointing error in degrees in the first phase") + xlab("Condition") + 
-  scale_fill_manual(values = estimote_pal) + guides(fill=F)
-
-ggplot(sop_all[sop_all$phase == 2, ], aes(x = testing.condition, y = abs(error), fill=factor(testing.condition))) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
-  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
-  ylab("Average pointing error in degrees in second phase") + xlab("Condition") + 
-  scale_fill_manual(values = estimote_pal) + guides(fill=F)
-
-#t.test(abs(error) ~ learning.condition, sop_all[sop_all$phase == 1,])
-t.test(abs(error) ~ testing.condition, sop_all[sop_all$phase == 2,])
 ```
@@
 df <- block_measure(condition_all_sums, "sum.errors")
-df
+kable(df)
 ggplot(melt(df[,1:4]), aes(x = learning.condition, y = value, fill = variable)) + 
@@
 df <- block_measure(condition_all_sums, "mean.distance")
-df
+kable(df)
 ggplot(melt(df[,1:4]), aes(x = learning.condition, y = value, fill = variable)) + 
@@
 ```{r, echo=F}
-block_t_test(conference_walk_all, 4, 3,"min_norm_distance")
+kable(block_t_test(conference_walk_all, 4, 3,"min_norm_distance"))
 ```
@@
 ```{r, echo=F}
-block_t_test(conference_walk_all, 4, 3,"errors")
+kable(block_t_test(conference_walk_all, 4, 3,"errors"))
 ```
@@
 ```{r, echo=F}
-block_t_test(conference_walk_all, 4, 1,"min_norm_distance")
+kable(block_t_test(conference_walk_all, 4, 1,"min_norm_distance"))
 ```
@@
 ```{r, echo=F}
-block_t_test(conference_walk_all, 4, 1,"errors")
+kable(block_t_test(conference_walk_all, 4, 1,"errors"))
 ```
@@
 Paired t tests between min normalised distance difference in blocks 1 and 6.
+
 ```{r, echo=F}
-block_t_test(conference_walk_all, 6, 1,"min_norm_distance")
+kable(block_t_test(conference_walk_all, 6, 1,"min_norm_distance"))
 ```
+
 And for number of errors
+
 ```{r, echo=F}
-block_t_test(conference_walk_all, 6, 1,"errors")
+kable(block_t_test(conference_walk_all, 6, 1,"errors"))
 ```
@@
 ## Independent block improvement comparisons between conditions for distance
+```{r, echo=F}
+report_independent_comparisons <- function(df_wide, column){
+  desktop_real <- t.test(df_wide[df_wide$learning.condition=="Desktop", column],df_wide[df_wide$learning.condition=="Real", column])
+  treadmill_real <- t.test(df_wide[df_wide$learning.condition=="Treadmill VR", column],df_wide[df_wide$learning.condition=="Real", column])  
+  treadmill_desktop <- t.test(df_wide[df_wide$learning.condition=="Treadmill VR", column],df_wide[df_wide$learning.condition=="Desktop", column])
+  results <- list(desktop_real, treadmill_real, treadmill_desktop)
+  df <- data.frame(`Comparison`=c("Desktop - real", "Treadmil - real", "Treadmil- Desktop"), mean.first=sapply(results, function(x){x$estimate[1]}), mean.second=sapply(results, function(x){x$estimate[2]}), `p-value`=sapply(results, function(x){x$p.value}))
+  return(df)
+}
+```
 
@@
 ### Block 1-4 improvement
-VE-real and real-real
 ```{r, echo=F}
-t.test(df_wide$diff41[df_wide$learning.condition=="Desktop"],df_wide$diff41[df_wide$learning.condition=="Real"])
-```
-
-VR-real and real-real
-```{r, echo=F}
-t.test(df_wide$diff41[df_wide$learning.condition=="Treadmill VR"],df_wide$diff41[df_wide$learning.condition=="Real"])  
-```
-
-VR-real and VE-real
-```{r, echo=F}
-t.test(df_wide$diff41[df_wide$learning.condition=="Treadmill VR"],df_wide$diff41[df_wide$learning.condition=="Desktop"])
+kable(report_independent_comparisons(df_wide, "diff41"))
 ```
@@
 ### Block 3-4 improvement
-Desktop to real
 ```{r, echo=F}
-t.test(df_wide$diff43[df_wide$learning.condition=="Desktop"], df_wide$diff43[df_wide$learning.condition=="Real"])
-```
-
-VR-real and real-real
-```{r, echo=F}
-t.test(df_wide$diff43[df_wide$learning.condition=="Treadmill VR"], df_wide$diff43[df_wide$learning.condition=="Real"])  
-```
-
-VR-real and VE-real
-```{r, echo=F}
-t.test(df_wide$diff43[df_wide$learning.condition=="Treadmill VR"], df_wide$diff43[df_wide$learning.condition=="Desktop"])
+kable(report_independent_comparisons(df_wide, "diff43"))
 ```
@@
 aov_block_distance <- aov(min_norm_distance~exp_block_id*learning.condition, conference_walk_all)
-summary(aov_block_distance)
+kable(anova_table(aov_block_distance))
 ```
@@
   type = 3)
-aov_block_dist_34
+kable(aov_block_dist_34)
 
@@
   type = 3)
-aov_block_dist_14
+kable(aov_block_dist_14)
 
@@
 
-### NUmber of errors
+### Number of errors
 Within subjects anova for block 1-4 number of errors.
@@
   type = 3)
-aov_block_err_14
+kable(aov_block_err_14)
 
@@
   type = 3)
-aov_block_err_34
+kable(aov_block_err_34)
 

### 2018-08-10 Scripts/analysis-helpers.R

@@ -1,6 +1,6 @@
 make_graph <- function(data, x_name, y_name, fill_name){
   ggplot(data, aes(x = data[, x_name], y = data[, y_name], fill=factor(data[, fill_name]))) +
-    stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
+    stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
     stat_summary(fun.y=mean,position=position_dodge(width=0.95), geom="bar") 
 }
 
@@ -8,4 +8,8 @@ make_line_graph <- function(data, x_name, y_name, fill_name){
   ggplot(data, aes(x = data[, x_name], y = data[, y_name], color=factor(data[, fill_name]))) +
     stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
     stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line")
+}
+
+anova_table <- function(aov){
+  return(summary(aov)[[1]])
 }

### 2018-08-10 Report/arne-report.Rmd
Modifies arnes report

@@
 ## Descriptives
-In normalised distance we can see similar starting point (not significant starting difference `r apa_print(aov(min_norm_distance ~ learning.condition, conference_walk_all[conference_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`)
+In normalised distance we can see that people in different conditions start fairly at the same level of performance (difference between learnign conditions is not significant `r apa_print(aov(min_norm_distance ~ learning.condition, conference_walk_all[conference_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`)
 
@@
 
-There is difference in starting times in all conditions in the first block (`r apa_print(aov(min_norm_time ~ learning.condition, conference_walk_all[conference_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`)).
+There is difference in normalised times in all conditions in the first block (`r apa_print(aov(min_norm_time ~ learning.condition, conference_walk_all[conference_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`)).
 Probably due to learning issues with the treadmill.
-Threre is no significant difference in desktop and real (`r apa_print(t.test(min_norm_time ~ learning.condition, conference_walk_all[conference_walk_all$exp_block_id == 1 & conference_walk_all$learning.condition %in% c("Desktop", "Real"), ]))$statistic`))
+Threre is no significant difference between desktop and real world learning in normalised times (`r apa_print(t.test(min_norm_time ~ learning.condition, conference_walk_all[conference_walk_all$exp_block_id == 1 & conference_walk_all$learning.condition %in% c("Desktop", "Real"), ]))$statistic`))
 
@@
 
-We can clearly see differences in the progression of the learnign, but as you pointed out many times, as they are not linear, we could benefit from non linear modeling.
+We can clearly see differences in the progression of the learning slopes, but as you pointed out many times, as they are not linear, we could benefit from non linear modeling.
 I hope Mike and professor Ferer will help us with that.
 
-But we can look at either pure changes between block 3-4 or 1-4 or those calculations that you proposed as a (block4-block3)/(block3+block4) measurements.
+But I can look at either pure changes between block 3-4 or 1-4 or those calculations that you proposed as a (block4-block3)/(block3+block4) measurements.
 
@@
 
-df_wide_errors$block14 <- df_wide_errors$block1-df_wide_errors$block4
-df_wide_errors$block34 <- df_wide_errors$block3-df_wide_errors$block4
-df_wide_distance$block14 <- df_wide_distance$block1-df_wide_distance$block4
-df_wide_distance$block34 <- df_wide_distance$block3-df_wide_distance$block4
+df_wide_errors$block14 <- (df_wide_errors$block1-df_wide_errors$block4)/(df_wide_errors$block1+df_wide_errors$block4)
+df_wide_errors$block34 <- (df_wide_errors$block3-df_wide_errors$block4)/(df_wide_errors$block3+df_wide_errors$block4)
+df_wide_distance$block14 <- (df_wide_distance$block1-df_wide_distance$block4)/(df_wide_distance$block1+df_wide_distance$block4)
+df_wide_distance$block34 <- (df_wide_distance$block3-df_wide_distance$block4)/(df_wide_distance$block3+df_wide_distance$block4)
 
@@
   mean.na <- function(data){return(mean(data, na.rm = T))}
-  df_wide <- dcast(df, condition + id ~ exp_block_id, value.var = value, fun.aggregate = mean.na)
-  colnames(df_wide) <- c("condition","id", "block1","block2")
-  df_wide %>% group_by(condition) %>% do(tidy(t.test(.$block1, .$block2, paired = T)))
+  df_wide <- dcast(df, learning.condition + id ~ exp_block_id, value.var = value, fun.aggregate = mean.na)
+  colnames(df_wide) <- c("learning.condition","id", "block1","block2")
+  df_wide %>% group_by(learning.condition) %>% do(tidy(t.test(.$block1, .$block2, paired = T)))
 }
+
+ez_prepare_block <- function(df, blocks){
+  filtered_data <- df[df$exp_block_id %in% blocks, ]
+  filtered_data <- filtered_data[complete.cases(filtered_data),]
+  has_all <- filtered_data %>% group_by(id) %>% select(id, "exp_block_id") %>% unique() %>% count
+  has_all <- has_all %>% filter(n == 2)
+  filtered_data <- filtered_data %>% filter(id %in% has_all$id)
+  filtered_data$exp_block_id <- factor(filtered_data$exp_block_id)
+  return(filtered_data)
+}
+
+clean_walk_14 <- ez_prepare_block(conference_walk_all, c(1,4))
+clean_walk_34 <- ez_prepare_block(conference_walk_all, c(3,4))
 ```
@@
 ### Distance
-We can see that all conditions improved from the first to forth block in both distance walked and errors in decisions.
-Paired t-tests show that both Treadmill and Real learning conditions led to improved performance from 1-4, but not Desktop.
-All conditions show improvemt in number of errors.
+All conditions improved from the first to fourth block in both distance walked and errors in decisions.
+Paired t-tests show that both Treadmill and Real world learning conditions led to improved performance from 1-4, but Desktop did not.
+All conditions show significant improvement in number of errors from block 1 to 4.
 
@@
 df_error <- block_t_test(conference_walk_all, 1, 4,"errors")[,c(1,4)]
-df14 <- merge(df_dist, df_error, by = "condition")
-colnames(df14) <- c("condition", "p-value distance improvement", "p-value errors improvement")
+df14 <- merge(df_dist, df_error, by = "learning.condition")
+colnames(df14) <- c("learning condition", "p-value distance improvement", "p-value errors improvement")
 kable(df14)
@@
 
-But anova shows no significant difference in the improvement between groups for distance (`r apa_print(aov(block14 ~ learning.condition, df_wide_distance))$statistic$learning_condition`)  nor errors (`r apa_print(aov(block14 ~ learning.condition, df_wide_errors))$statistic$learning_condition`).
+But anova shows no significant difference in the improvement between groups for distance (`r apa_print(aov(block14 ~ learning.condition, df_wide_distance))$statistic$learning_condition`).
 This might be due to quite large variability in the sample.
@@
 ### Errors
+Unlike for distance, anova reveals significant difference in number of errors improvement from block 1 to block 4 between different learning conditions (`r apa_print(aov(block14 ~ learning.condition, df_wide_errors))$statistic$learning_condition`).
+
+Runnign Tukey's post hoc on the distance shows that this is originating from the difference between Desktop and the Real world, with Treadmill not being significantly different from either of them.
+
+```{r, echo=F}
+kable(TukeyHSD(x=aov(block14 ~ learning.condition, df_wide_errors), conf.level=0.95)$learning.condition[, c(1,4)])
+```
+
 ```{r, echo=F}
@@
 
-## Block improvement from block 3 to block 4
+## Block performance improvement from block 3 to block 4
 
-We can see that everybody (including Real to Real condition) gets a little bit worse from 3 to 4.
-But this "decrease of performance" is not significant when computed using paired t-tests except marginally worse in the desktop version.
-All conditions show imprvement in the number of errors in paired t-tests.
+Everybody's distance performance (including Real to Real condition) gets a little bit worse from 3 to 4.
+But this is not significant when computed using paired t-tests, although it is marginally worse in the desktop version.
+But all conditions show decrease in the number of errors from block 3 to 4.
 
@@
 df_error <- block_t_test(conference_walk_all, 3, 4,"errors")[,c(1,4)]
-df34 <- merge(df_dist, df_error, by = "condition")
-colnames(df34) <- c("condition", "p-value distance improvement", "p-value errors improvement")
+df34 <- merge(df_dist, df_error, by = "learning.condition")
+colnames(df34) <- c("learning.condition", "p-value distance improvement", "p-value errors improvement")
 kable(df34)
@@
 
-Comparing the block improvement in different conditions using anovas, we see significant difference in the improvement between groups for distance (`r apa_print(aov(block34 ~ learning.condition, df_wide_distance))$statistic$learning_condition`)  and for errors (`r apa_print(aov(block34 ~ learning.condition, df_wide_errors))$statistic$learning_condition`).
+### Distance
 
-Runnign Tukey's post hoc on the distance shows marginally significant difference between Desktop from the other two conditions for distance.
+Comparing the block improvement in different conditions using anovas, we see marginally significant difference in the improvement between groups for distance (`r apa_print(aov(block34 ~ learning.condition, df_wide_distance))$statistic$learning_condition`).
 
-```{r}
-kable(TukeyHSD(x=aov(block34 ~ learning.condition, df_wide_distance), conf.level=0.95)$learning.condition)
-```
-
-And si
-```{r}
-kable(TukeyHSD(x=aov(block34 ~ learning.condition, df_wide_errors), conf.level=0.95)$learning.condition)
-```
-
-### Distance
 ```{r, echo=F}
@@
 
+Within subject anova for distance improvement between 3 and 4 shows significant interaction effect and marginally significant effect of learning condition and block on distance.
+```{r, warning = F, echo=F, message=F}
+aov_block_dist_34 <- ezANOVA(
+  data = clean_walk_34,
+  dv = min_norm_distance, 
+  wid = id,
+  within = exp_block_id,
+  between = learning.condition, 
+  type = 3)
+kable(aov_block_dist_34)
+
+ggplot(clean_walk_34, aes(x = as.numeric(exp_block_id), y = min_norm_distance, color = factor(learning.condition))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
+  ylab("Average normalised distance") + xlab("Experimental block") + 
+  scale_color_manual(values = estimote_pal, "Learning condition")
+```
+
 ### Errors
+
+We can see significant differences between different conditions in error rate improvement from block 3 to 4 when running independent anova (`r apa_print(aov(block34 ~ learning.condition, df_wide_errors))$statistic$learning_condition`).
+
+Post hoc tests show significant difference between Real and Desktop and marginally for Treadmill and Desktop.
+```{r, echo=F}
+kable(TukeyHSD(x=aov(block34 ~ learning.condition, df_wide_errors), conf.level=0.95)$learning.condition[, c(1,4)])
+```
+
 ```{r, echo=F}
@@
 ```
+
+Running within subjects anova for block 3-4 number of errors shows significant effect of both condition and block, but unlike in distance improvement, their interaction doesn't seem to have an effect on error rate improvement.
+```{r, warning = F, echo=F, message=F}
+aov_block_err_34 <- ezANOVA(
+  data = clean_walk_34,
+  dv = errors, 
+  wid = id,
+  within = exp_block_id,
+  between = condition,
+  type = 3)
+kable(aov_block_err_34)
+
+ggplot(clean_walk_34, aes(x = as.numeric(exp_block_id), y = errors, color = factor(condition))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
+  ylab("Average number of errors") + xlab("Experimental block") + 
+  scale_color_manual(values = estimote_pal, "Learning condition")
+```
+
+## Summary
+We see effect of learning condition (blocks 1-3) on performance in block 4. We also see
+
+Participant's performance in distance from block 3-4 gets slightly worse after the switch, but their number of errors decrease.
+It can mean that they know which door are correct, but they are unsure in how to get to them, showing some level of memory transfer, but not necessarily spatial knowledge transfer.

### 2018-08-10 Report/arne-report.Rmd
Updates report

@@
 
-Unless stated otherwise, I ran all tests on the normalised distance and time statistics. Errors and pointing information is not normalised, as they are the same in all conditions.
+I ran all tests on the normalised distance, errors are not normalised, as they are the same in all conditions.
 
-I only selected those "important" comparisons - therefore Desktop to Real, Treadmill VR to real and Real to real. Anovas and other comparisons are run only on these.
+I only selected those "important" comparisons - therefore Desktop to Real world, Treadmill VR to real world and Real world to real world.
+Anovas and other comparisons are run only on these.
+Learning conditions (block 1-3) are the different modalities, and blocks 3-4 are all real world testing.
 
@@
 ### Errors
-Unlike for distance, anova reveals significant difference in number of errors improvement from block 1 to block 4 between different learning conditions (`r apa_print(aov(block14 ~ learning.condition, df_wide_errors))$statistic$learning_condition`).
+Unlike for distance, anova reveals significant difference between learning conditions in number of errors improvement from block 1 to block 4 (`r apa_print(aov(block14 ~ learning.condition, df_wide_errors))$statistic$learning_condition`).
 
@@
 ## Summary
-We see effect of learning condition (blocks 1-3) on performance in block 4. We also see
+We see effect of learning condition (blocks 1-3) on performance in block 4.
 
-Participant's performance in distance from block 3-4 gets slightly worse after the switch, but their number of errors decrease.
+We see positive improvement in distance from block 1 to 4 in all conditions and this improvement is not significantly different across learning groups.
+We see similar improvement in number of errors, but Desktop improves significantly worse than real world training.
+
+Participant's performance in distance from block 3-4 gets slightly worse after the switch distance-wise (but not significanly), but their number of errors decrease across all conditions.
 It can mean that they know which door are correct, but they are unsure in how to get to them, showing some level of memory transfer, but not necessarily spatial knowledge transfer.
+
+We can see marginally significant difference in block 3-4 distance improvement acrosss learning conditions, with desktop doing the worst but not significantly different from the rest.
+Within subject anova for distance improvement in blocks 3-4 shows interaction between learning condition and experimental block - only desktop gets "worse" after the switch of environment.
+
+But we can see clear error rate improvement differences across conditions - all participants improved, but desktop learning shows least amount of improvement from block 3-4, participants doing significantly worse than those with real world training and marginally worse than those with Treadmill training.

### 2018-08-11 Report/arne-report.Rmd
Update arne-report.Rmd
Removes unused packages

@@
 library(knitr)
-library(brainvr.R)
-library(restimoter)
 library(papaja)

### 2018-09-17 Scripts/main-preprocess.R
Changes preprocessing for non R loading

@@ -28,8 +28,8 @@ walk_all <- walk_all %>% mutate(goal = replace(goal, goal=="Jesicca's office", "
 walk_all$exp_trial_id <- walk_all$trial_id + (walk_all$phase-1)*18
 walk_all$exp_block_id <- walk_all$block_id + (walk_all$phase-1)*3
 
-sop_all$exp_trial_id <- sop_all$trial_id + (sop_all$phase-1)*18
-sop_all$exp_block_id <- sop_all$block_id + (sop_all$phase-1)*3
+sop_all$exp_trial_id <- sop_all$trial_id + (sop_all$phase-1)*12
+sop_all$exp_block_id <- sop_all$block_id + (sop_all$phase-1)*2
 
 ### Conditions ----
 conditions <- data.frame(id = settings$participants$Code,
@@ -77,6 +77,7 @@ walk_all_new <- walk_all %>% select(-one_of(drop.cols))
 
 sop_all$abs_error <- abs(sop_all$error)
 
-write.table(sop_all, file = "sop_smoothed.csv", sep=";")
-write.table(walk_all_new, file = "walk_smoothed_errors_06.csv", sep=";")
+write.table(sop_all, file = "sop_smoothed.csv", sep=";", row.names = F)
+write.table(walk_all_new, file = "walk_smoothed_errors_06.csv", sep=";", row.names = F)
 
+write.table(walk_all, file = "walk.csv", sep=";", row.names = F)

### 2018-09-21 Report/paper-all-data.Rmd
Updates the report

--- 2018-08-10 Report/paper-data.Rmd
@@
 
-source("../Scripts/loading.R")
-source_folder("../TwoWorlds")
-source('../Scripts/analysis-helpers.R')
-sop_all <- read.table("../sop.csv", sep=";", header = T, stringsAsFactors = F)
-walk_all <- read.table("../walk.csv", sep=";", header = T, stringsAsFactors = F)
-walk_all$learning.condition <- walk_all$condition
-walk_all$learning.condition <- recode(walk_all$learning.condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
-sop_all$learning.condition <- sop_all$condition
-sop_all$testing.condition <- sop_all$condition
-sop_all$learning.condition <- recode(sop_all$learning.condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
-sop_all$testing.condition <- recode(sop_all$testing.condition, "real-real" = "Real", "ve-real" = "Real", "vr-real" = "Real", "real-vr" = "Treadmill VR", "vr-vr" = "Treadmill VR")
-
-conference_walk_all <- walk_all %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
-conference_sop_all <- sop_all %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
+source("../Scripts/paper-prepare.R")
 
@@
 estimote_pal <- c('#b6d6c1', '#54003d','#93d5f6','#d5d215','#f3aacb','#2d2556')
+estimote_pal_mono <- function(n){return(rep(estimote_pal[n],6))}
 theme_set(theme_minimal())
@@
 ```{r warning=F, echo=F}
-make_graph(conference_walk_all, "learning.condition", "distance", "phase") + 
+make_graph(sub_walk_all, "learning.condition", "distance", "phase") + 
   ylab("Average absolute trial distance") + xlab("condition") + 
@@
 
-make_graph(conference_walk_all, "learning.condition", "min_norm_distance", "phase") + 
+make_graph(sub_walk_all, "learning.condition", "min_norm_distance", "phase") + 
   ylab("Average normalised trial distance") + xlab("condition") + 
@@
 
-make_graph(conference_walk_all, "learning.condition", "time", "phase") + 
+make_graph(sub_walk_all, "learning.condition", "time", "phase") + 
   ylab("Average absolute trial duration") + xlab("condition") + 
@@
 
-make_graph(conference_walk_all, "learning.condition", "min_norm_time", "phase")+ 
+make_graph(sub_walk_all, "learning.condition", "min_norm_time", "phase")+ 
   ylab("Average normalised trial duration") + xlab("condition") + 
@@
 
-make_graph(conference_walk_all, "learning.condition", "errors", "phase")+ 
+make_graph(sub_walk_all, "learning.condition", "errors", "phase")+ 
   ylab("Average number of errors") + xlab("condition") + 
@@
 ```{r warning=F, echo=F}
-make_graph(conference_walk_all, "learning.condition", "distance", "exp_block_id") + 
+make_graph(walk_all, "learning.condition", "distance", "exp_block_id") + 
   ylab("Average absolute trial distance") + xlab("Learning condition") + 
@@
 
-make_graph(conference_walk_all, "learning.condition", "min_norm_distance", "exp_block_id") + 
+make_graph(sub_walk_all, "learning.condition", "min_norm_distance", "exp_block_id") + 
   ylab("Average normalised trial distance") +  xlab("Learning condition") + 
@@
 
-make_graph(conference_walk_all, "learning.condition", "time", "exp_block_id") + 
+make_graph(sub_walk_all, "learning.condition", "time", "exp_block_id") + 
   ylab("Average absolute trial duration") + xlab("Learning condition") + 
@@
 
-make_graph(conference_walk_all, "learning.condition", "min_norm_time", "exp_block_id") + 
+make_graph(sub_walk_all, "learning.condition", "min_norm_time", "exp_block_id") + 
   ylab("Average normalised trial duration") + xlab("Learning condition") + 
@@
 
-make_graph(conference_walk_all, "learning.condition", "errors", "exp_block_id")+ 
+make_graph(sub_walk_all, "learning.condition", "errors", "exp_block_id")+ 
   ylab("Average number of errors") + xlab("Learning condition") + 
@@
 
-ggplot(conference_walk_all, aes(x = exp_block_id, y = min_norm_distance, color = factor(learning.condition))) +
+ggplot(sub_walk_all, aes(x = exp_block_id, y = min_norm_distance, color = factor(learning.condition))) +
   stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
@@
 
-ggplot(conference_walk_all, aes(x = exp_trial_id, y = min_norm_distance, color = factor(learning.condition))) +
+ggplot(sub_walk_all, aes(x = exp_trial_id, y = min_norm_distance, color = factor(learning.condition))) +
   stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
@@
 
-ggplot(conference_walk_all, aes(x = exp_block_id, y = errors, color = factor(learning.condition))) +
+ggplot(sub_walk_all, aes(x = exp_block_id, y = errors, color = factor(learning.condition))) +
   stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
@@
 ```{r, echo = F}
-condition_all_sums <- conference_walk_all %>% 
+condition_all_sums <- sub_walk_all %>% 
   group_by(exp_block_id, id, learning.condition) %>% 
@@
 ## Sum of errors
+
 Measure of overall sum of errors. Mean columns define mean sum of errors for all participants in that block
@@
 ```{r, echo=F}
-kable(block_t_test(conference_walk_all, 4, 3,"min_norm_distance"))
+kable(block_t_test(sub_walk_all, 4, 3,"min_norm_distance"))
 ```
@@
 ```{r, echo=F}
-kable(block_t_test(conference_walk_all, 4, 3,"errors"))
+kable(block_t_test(sub_walk_all, 4, 3,"errors"))
 ```
@@
 ```{r, echo=F}
-kable(block_t_test(conference_walk_all, 4, 1,"min_norm_distance"))
+kable(block_t_test(sub_walk_all, 4, 1,"min_norm_distance"))
 ```
@@
 ```{r, echo=F}
-kable(block_t_test(conference_walk_all, 4, 1,"errors"))
+kable(block_t_test(sub_walk_all, 4, 1,"errors"))
 ```
@@
 ```{r, echo=F}
-kable(block_t_test(conference_walk_all, 6, 1,"min_norm_distance"))
+kable(block_t_test(sub_walk_all, 6, 1,"min_norm_distance"))
 ```
@@
 ```{r, echo=F}
-kable(block_t_test(conference_walk_all, 6, 1,"errors"))
+kable(block_t_test(sub_walk_all, 6, 1,"errors"))
 ```
@@
 ```{r, echo=F}
-aov_block_distance <- aov(min_norm_distance~exp_block_id*learning.condition, conference_walk_all)
+aov_block_distance <- aov(min_norm_distance~exp_block_id*learning.condition, sub_walk_all)
 kable(anova_table(aov_block_distance))
@@
 
-clean_walk_14 <- ez_prepare_block(conference_walk_all, c(1,4))
-clean_walk_34 <- ez_prepare_block(conference_walk_all, c(3,4))
+clean_walk_14 <- ez_prepare_block(sub_walk_all, c(1,4))
+clean_walk_34 <- ez_prepare_block(sub_walk_all, c(3,4))
 ```

### 2018-09-21 Report/paper-story.Rmd
Updates the report

--- 2018-08-11 Report/arne-report.Rmd
@@
 ---
-title: "Arne report"
+title: "Transfer report"
 author: "Lukáš Hejtmánek"
@@
 
-source("../Scripts/loading.R")
-source_folder("../TwoWorlds")
-source('../Scripts/analysis-helpers.R')
-sop_all <- read.table("../sop.csv", sep=";", header = T, stringsAsFactors = F)
-walk_all <- read.table("../walk.csv", sep=";", header = T, stringsAsFactors = F)
-walk_all$learning.condition <- walk_all$condition
-walk_all$learning.condition <- recode(walk_all$learning.condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
-sop_all$learning.condition <- sop_all$condition
-sop_all$testing.condition <- sop_all$condition
-sop_all$learning.condition <- recode(sop_all$learning.condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
-sop_all$testing.condition <- recode(sop_all$testing.condition, "real-real" = "Real", "ve-real" = "Real", "vr-real" = "Real", "real-vr" = "Treadmill VR", "vr-vr" = "Treadmill VR")
-
-conference_walk_all <- walk_all %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
-conference_sop_all <- sop_all %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
-
+source("../Scripts/paper-prepare.R")
 #green, violet, blue, yellow, pink, dark blue
@@
 
-condition_all_sums <- conference_walk_all %>% 
+condition_all_sums <- sub_walk_all %>% 
   group_by(exp_block_id, id, learning.condition) %>% 
@@
 }
+```
+ There will be smaller spatial knowledge transfer in those conditions which use less
 
-```
+## Hypotheses
+1. There will be some transfer in all learning conditions.
+2. Transfer will be higher in the real world learning condition than in desktop treadmill condition.
+3. Transfer will be higher in the real world learning condition than in VR treadmill condition.
+4. Transfer will be higher in the VR treadmill condition that in the desktop condition.
 
+- transfer is operationalised as a change in performance in navigational tasks following the environment/modality switch
+- navigational performance change can be measured in different ways
+
+Specific tests (hypothesis in parenthesis)
+- Overrall average walking performance in phase 1 vs phase 2 
+- Performance in block 4 vs performance in block 1 and 2
+
+- Overrall average pointing performance in phase 1 vs phase 2 
+- Average walking perfomrance in block 3 vs 4 (block directly before vs directly after the switch)
+- Personal walking performance change in block 3 vs 4 (calculated as an average of individuals (block3-block4)/(block3+block4))
+- Personal walking performance change in block 1 vs 4 (calculated as an average of individuals (block1-block4)/(block3+block4))
+- level of correlation of personal performance in block 1,2,3 and 4 - higher transfer should show higher correlation in these comparisons
+- Difference in performance between groups in blocks 5 and 6 (e.g. not immediately after the switch, but after all groups have been subjected the real world environment for some time)
+- Difference in performance between groups in block 4-6 with inclusion of VR-VR group nad real-VR group (to see if the final performance is subjected to the necessity of a real world experience, or not and to see if the swith has detrimental effect on real-VR as well).
+
+## Controlling analyses
+1. Every condition starts on a same level of performance (there isn't a bias in the first block)
+
+## Procedure description
 I ran all tests on the normalised distance, errors are not normalised, as they are the same in all conditions.
@@
 Anovas and other comparisons are run only on these.
-Learning conditions (block 1-3) are the different modalities, and blocks 3-4 are all real world testing.
+Learning conditions (block 1-3) are the different modalities, and blocks 4-6 are all real world testing.
 
 ## Descriptives
-In normalised distance we can see that people in different conditions start fairly at the same level of performance (difference between learnign conditions is not significant `r apa_print(aov(min_norm_distance ~ learning.condition, conference_walk_all[conference_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`)
+In normalised distance we can see that people in different conditions start fairly at the same level of performance (difference between learnign conditions is not significant `r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`).
+Nor there is significant difference between all 5 starting groups `r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$condition`.
 
 ```{r, echo=F, warning=F}
-make_graph(conference_walk_all, "learning.condition", "min_norm_distance", "exp_block_id") + 
+make_graph(sub_walk_all, "learning.condition", "min_norm_distance", "exp_block_id") + 
   ylab("Average normalised trial distance") + xlab("condition") + 
@@
 
-There is difference in normalised times in all conditions in the first block (`r apa_print(aov(min_norm_time ~ learning.condition, conference_walk_all[conference_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`)).
+There is difference in normalised times in all conditions in the first block (`r apa_print(aov(min_norm_time ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`)).
 Probably due to learning issues with the treadmill.
-Threre is no significant difference between desktop and real world learning in normalised times (`r apa_print(t.test(min_norm_time ~ learning.condition, conference_walk_all[conference_walk_all$exp_block_id == 1 & conference_walk_all$learning.condition %in% c("Desktop", "Real"), ]))$statistic`))
+Threre is no significant difference between desktop and real world learning in normalised times (`r apa_print(t.test(min_norm_time ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1 & sub_walk_all$learning.condition %in% c("Desktop", "Real"), ]))$statistic`))
 
 ```{r, echo=F}
-make_graph(conference_walk_all, "learning.condition", "min_norm_time", "exp_block_id")+ 
+make_graph(walk_all, "condition", "min_norm_time", "exp_block_id")+ 
   ylab("Average normalised trial duration") + xlab("condition") + 
@@
 
-And we don't have any significant difference in number of errors in the first block (`r apa_print(aov(errors ~ learning.condition, conference_walk_all[conference_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`)).
+And we don't have any significant difference in number of errors in the first block (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`)) nor for the all 5 condition (`r apa_print(aov(errors ~ condition, walk_all[walk_all$exp_block_id == 1, ]))$statistic$condition`).
 
 ```{r, echo=F}
-make_graph(conference_walk_all, "learning.condition", "errors", "exp_block_id")+ 
-  ylab("Average number of errors") + xlab("condition") + 
+make_graph(walk_all, "condition", "errors", "exp_block_id")+ 
+  ylab("Average number of errors") + xlab("Condition") + 
   scale_fill_manual(values=estimote_pal," Block ID")
 
-ggplot(conference_walk_all, aes(x = exp_block_id, y = errors, color = factor(learning.condition))) +
+ggplot(sub_walk_all, aes(x = exp_block_id, y = errors, color = factor(learning.condition))) +
   stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
@@
 
-We can clearly see differences in the progression of the learning slopes, but as you pointed out many times, as they are not linear, we could benefit from non linear modeling.
-I hope Mike and professor Ferer will help us with that.
+## Specific predictions testing
+
+### Performance in block 4 is better than in block 1 and 2
+If we assume some level of transfer, than the average or individual performance in block 4 should be better in all conditions than in block 1 or potentially in block 2.
+I don't test for block 3 because that immediately compares the switch performance.
+```{r, echo=F}
+df_dist <- block_t_test(sub_walk_all, 1, 4,"min_norm_distance")[,c(1,4)]
+df_error <- block_t_test(sub_walk_all, 1, 4,"errors")[,c(1,4)]
+df14 <- merge(df_dist, df_error, by = "learning.condition")
+colnames(df14) <- c("learning condition", "p-value distance improvement", "p-value errors improvement")
+kable(df14, caption="Paired T test comparing infividual performance improvement from block 1-4 in different conditions")
+```
+We see taht both real world and treadmill VR learning leads to improvement from block 1 to block 4 in both errors and path travelled.
+But desktop only in number of errors and marginally in distance improvement.
+
+### Overrall average walking performance in phase 1 vs phase 2
+Given that all participants had 3 blocks of trials in the second phase done in the real world, We can assume, if one learning method leads to better transfer than another
+```{r, echo=F}
+
+```
+
+### Overrall average pointing performance in phase 1 vs phase 2
+```{r, echo=F}
+
+```
+### Average walking perfomrance in block 3 vs 4 
+(block directly before vs directly after the switch)
+```{r, echo=F}
+
+```
+
+### Average walkkig perfomance in block 1 vs 4 
+(starting baseline vs switch)
+```{r, echo=F}
+
+```
+
+### Personal walking performance change in block 3 vs 4 
+(calculated as an average of individuals (block3-block4)/(block3+block4))
+```{r, echo=F}
+
+```
+
+### Personal walking performance change in block 1 vs 4 
+(calculated as an average of individuals (block1-block4)/(block3+block4))
+```{r, echo=F}
+
+```
+
+### level of correlation of personal performance in block 1,2,3 and 4
+
+- higher transfer should show higher correlation in these comparisons
+```{r, echo=F}
+
+```
+
+POtential followups
+### difference in performance between groups in blocks 5 and 6
+
+(e.g. not immediately after the switch, but after all groups have been subjected the real world environment for some time)
+```{r, echo=F}
+
+```
+### difference in performance between groups in block 4-6 with inclusion of VR-VR group nad real-VR group
+
+(to see if the final performance is subjected to the necessity of a real world experience, or not and to see if the swith has detrimental effect on real-VR as well).
+```{r, echo=F}
+
+```
 
@@
 
-clean_walk_14 <- ez_prepare_block(conference_walk_all, c(1,4))
-clean_walk_34 <- ez_prepare_block(conference_walk_all, c(3,4))
+clean_walk_14 <- ez_prepare_block(sub_walk_all, c(1,4))
+clean_walk_34 <- ez_prepare_block(sub_walk_all, c(3,4))
 ```
-
 ## Block improvement from block 1 to block 4
-
-### Distance
-All conditions improved from the first to fourth block in both distance walked and errors in decisions.
-Paired t-tests show that both Treadmill and Real world learning conditions led to improved performance from 1-4, but Desktop did not.
-All conditions show significant improvement in number of errors from block 1 to 4.
-
-```{r, echo=F}
-df_dist <- block_t_test(conference_walk_all, 1, 4,"min_norm_distance")[,c(1,4)]
-df_error <- block_t_test(conference_walk_all, 1, 4,"errors")[,c(1,4)]
-df14 <- merge(df_dist, df_error, by = "learning.condition")
-colnames(df14) <- c("learning condition", "p-value distance improvement", "p-value errors improvement")
-kable(df14)
-```
 
@@
 ```{r, echo=F}
-df_dist <- block_t_test(conference_walk_all, 3, 4,"min_norm_distance")[,c(1,4)]
-df_error <- block_t_test(conference_walk_all, 3, 4,"errors")[,c(1,4)]
+df_dist <- block_t_test(sub_walk_all, 3, 4,"min_norm_distance")[,c(1,4)]
+df_error <- block_t_test(sub_walk_all, 3, 4,"errors")[,c(1,4)]
 df34 <- merge(df_dist, df_error, by = "learning.condition")

### 2018-09-21 Scripts/paper-prepare.R

@@ -1,0 +1,14 @@
+source("../Scripts/loading.R")
+source_folder("../TwoWorlds")
+source('../Scripts/analysis-helpers.R')
+sop_all <- read.table("../sop.csv", sep=";", header = T, stringsAsFactors = F)
+walk_all <- read.table("../walk.csv", sep=";", header = T, stringsAsFactors = F)
+walk_all$learning.condition <- walk_all$condition
+walk_all$learning.condition <- recode(walk_all$learning.condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
+sop_all$learning.condition <- sop_all$condition
+sop_all$testing.condition <- sop_all$condition
+sop_all$learning.condition <- recode(sop_all$learning.condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
+sop_all$testing.condition <- recode(sop_all$testing.condition, "real-real" = "Real", "ve-real" = "Real", "vr-real" = "Real", "real-vr" = "Treadmill VR", "vr-vr" = "Treadmill VR")
+
+sub_walk_all <- walk_all %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
+sub_sop_all <- sop_all %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))

### 2018-09-21 TwoWorlds/helpers-report.R
Adds new helper methods

@@ -1,0 +1,18 @@
+apa_p <- function(values){
+  output <- c()
+  for (value in values){
+    if(value < 0.01){ output <- c(output, "p < .001"); next}
+    if(value < 0.05) { output <- c(output, "p < .0.05"); next}
+    output <- c(output, paste0("p = ", round(value, 2)))
+  }
+  return(output)
+}
+block_t_test <- function(df, block1, block2, value, paired = T){
+  df <- df %>% filter(exp_block_id %in% c(block1,block2))
+  mean.na <- function(data){return(mean(data, na.rm = T))}
+  df_wide <- dcast(df, learning.condition + id ~ exp_block_id, value.var = value, fun.aggregate = mean.na)
+  colnames(df_wide) <- c("learning.condition","id", "block1","block2")
+  df_out <- df_wide %>% group_by(learning.condition) %>% do(tidy(t.test(.$block1, .$block2, paired = paired)))
+  df_out$p.value <- apa_p(df_out$p.value)
+  return(df_out)
+}

### 2018-09-21 TwoWorlds/helpers-results.R

@@ -1,9 +1,19 @@
 ez_prepare <- function(filtered_data, by){
-
   filtered_data <- filtered_data[complete.cases(filtered_data),]
   has_all <- filtered_data %>% group_by(id) %>% select(by) %>% count
   has_all <- has_all %>% filter(n > 4)
   filtered_data <- filtered_data %>% filter(id %in% has_all$id)
   filtered_data$id <- factor(filtered_data$id)
+  return(filtered_data)
+}
+
+
+ez_prepare_block <- function(df, blocks){
+  filtered_data <- df[df$exp_block_id %in% blocks, ]
+  filtered_data <- filtered_data[complete.cases(filtered_data),]
+  has_all <- filtered_data %>% group_by(id) %>% select(id, "exp_block_id") %>% unique() %>% count
+  has_all <- has_all %>% filter(n == 2)
+  filtered_data <- filtered_data %>% filter(id %in% has_all$id)
+  filtered_data$exp_block_id <- factor(filtered_data$exp_block_id)
   return(filtered_data)
 }

### 2018-09-21 Report/paper-story.Rmd
Updates the report

@@
 1. There will be some transfer in all learning conditions.
-2. Transfer will be higher in the real world learning condition than in desktop treadmill condition.
+2. Transfer will be higher in the real world learning condition than in desktop condition.
 3. Transfer will be higher in the real world learning condition than in VR treadmill condition.
@@
 
-- transfer is operationalised as a change in performance in navigational tasks following the environment/modality switch
-- navigational performance change can be measured in different ways
+- transfer is operationalised as a change in performance in navigational tasks following the environment/modality switch.
 
@@
 
-### Performance in block 4 is better than in block 1 and 2
+### Performance in block 4 is better than in block 1
 If we assume some level of transfer, than the average or individual performance in block 4 should be better in all conditions than in block 1 or potentially in block 2.
 I don't test for block 3 because that immediately compares the switch performance.
+```{r, echo=F,message=F}
+make_graph(sub_walk_all[sub_walk_all$exp_block_id %in% c(1,4),], "learning.condition", "min_norm_distance", "exp_block_id") + 
+  ylab("Average normalised trial distance") + xlab("condition") + 
+  scale_fill_manual(values=estimote_pal,"Block ID")
+make_graph(sub_walk_all[sub_walk_all$exp_block_id %in% c(1,4),], "learning.condition", "errors", "exp_block_id") + 
+  ylab("Average number of errors") + xlab("condition") + 
+  scale_fill_manual(values=estimote_pal,"Block ID")
+```
+
 ```{r, echo=F}
@@
 ```
-We see taht both real world and treadmill VR learning leads to improvement from block 1 to block 4 in both errors and path travelled.
+
+We see that both real world and treadmill VR learning leads to improvement from block 1 to block 4 in both errors and path travelled.
 But desktop only in number of errors and marginally in distance improvement.
 
-### Overrall average walking performance in phase 1 vs phase 2
-Given that all participants had 3 blocks of trials in the second phase done in the real world, We can assume, if one learning method leads to better transfer than another
+### Overrall average walking and error performance in phase 1 vs phase 2
+Given that all participants had 3 blocks of trials in the second phase done in the real world, We can assume, if one learning method leads to better transfer than another, than the performance improvement between phases could be different.
+```{r, echo=F, warning=F}
+make_graph(sub_walk_all, "learning.condition", "min_norm_distance", "phase") + 
+  ylab("Average normalised trial distance") + xlab("condition") + 
+  scale_fill_manual(values=estimote_pal,"Phase")
+make_graph(sub_walk_all, "learning.condition", "errors", "phase")+ 
+  ylab("Average number of errors") + xlab("condition") + 
+  scale_fill_manual(values=estimote_pal,"Phase")
+```
+
+Anova shows a significant effect of phase and condition on overall performance, but not of their interaction.
 ```{r, echo=F}
+apa_print(aov(min_norm_distance ~ phase*learning.condition, sub_walk_all))
+pair_t_norm_distance_phase_2 <- pairwise.t.test(sub_walk_all$min_norm_distance, sub_walk_all$learning.condition, p.adj = "holm")
+```
+Looking solely at the performance differences in the second phase differs significantly between groups (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$phase == 2,]))$statistic$learning_condition`).
+Tukey test shows significant difference between real adn a desktop versions
 
+```{r, echo=F}
+kable(TukeyHSD(x=aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$phase == 2,]), conf.level=0.95)$learning.condition[, c(1,4)])
 ```
@@
 ### Overrall average pointing performance in phase 1 vs phase 2
-```{r, echo=F}
-
-```
-### Average walking perfomrance in block 3 vs 4 
-(block directly before vs directly after the switch)
-```{r, echo=F}
-
+```{r, echo=F, warning=F}
+make_graph(sub_sop_all, "learning.condition", "abs_error", "phase") + 
+  ylab("Average normalised trial distance") + xlab("condition") + 
+  scale_fill_manual(values=estimote_pal,"Phase")
 ```
 
-### Average walkkig perfomance in block 1 vs 4 
-(starting baseline vs switch)
+Anova shows significant effect of both phase, learning condition and their interaction on pointing performance.
 ```{r, echo=F}
+kable(apa_print(aov(abs_error ~ phase*learning.condition, sub_sop_all))$table)
+```
 
+### Average walking perfomrance in block 3 vs 4
+
+Looking directly at the performance after the switch, we can see if there is a sudden drop in performance.
+We can study this both suing individual paired t.tests as well as non paired t-tests.
+Looking at the graphs we see differences in the level to which people have learned the environemtn in the 3rd block.
+```{r, echo=F, message=F}
+make_graph(sub_walk_all[sub_walk_all$exp_block_id %in% c(3,4),], "learning.condition", "min_norm_distance", "exp_block_id") + 
+  ylab("Average normalised trial distance") + xlab("condition") + 
+  scale_fill_manual(values=estimote_pal,"Block ID")
+make_graph(sub_walk_all[sub_walk_all$exp_block_id %in% c(3,4),], "learning.condition", "errors", "exp_block_id") + 
+  ylab("Average number of errors") + xlab("condition") + 
+  scale_fill_manual(values=estimote_pal,"Block ID")
 ```
+
+Anova doesn¨t show shows a significant difference in the block 3 (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 3,]))$statistic$learning_condition`), but there is a significant difference between path performance in the block 4 (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`).
+
+There is significant difference in the number of errors in both block 3 (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 3,]))$statistic$learning_condition`), and the block 4 (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`).
+
+Running pairwise t-tests to see performance change from block 3 to 4 gives the following results.
+```{r, echo=F}
+df_dist <- block_t_test(sub_walk_all, 3, 4,"min_norm_distance")[,c(1,4)]
+df_error <- block_t_test(sub_walk_all, 3, 4,"errors")[,c(1,4)]
+df34 <- merge(df_dist, df_error, by = "learning.condition")
+colnames(df34) <- c("learning condition", "p-value distance improvement", "p-value errors improvement")
+kable(df34, caption="Paired T test comparing infividual performance improvement from block 3-4 in different conditions")
+```
+WE can see that all participants improved in the errors made, but neigher group is significantly worse between block 3 and 4 in the path travelled (although participants learning on desktop perform slightly worse in 4th block, but the significance is only marginal)
 
@@
 ```
-
 But I can look at either pure changes between block 3-4 or 1-4 or those calculations that you proposed as a (block4-block3)/(block3+block4) measurements.
@@
 df_distance <- block_measure(condition_all_sums, "mean.distance")
-
-block_t_test <- function(df, block1, block2, value){
-  df <- df %>% filter(exp_block_id %in% c(block1,block2))
-  mean.na <- function(data){return(mean(data, na.rm = T))}
-  df_wide <- dcast(df, learning.condition + id ~ exp_block_id, value.var = value, fun.aggregate = mean.na)
-  colnames(df_wide) <- c("learning.condition","id", "block1","block2")
-  df_wide %>% group_by(learning.condition) %>% do(tidy(t.test(.$block1, .$block2, paired = T)))
-}
-
-ez_prepare_block <- function(df, blocks){
-  filtered_data <- df[df$exp_block_id %in% blocks, ]
-  filtered_data <- filtered_data[complete.cases(filtered_data),]
-  has_all <- filtered_data %>% group_by(id) %>% select(id, "exp_block_id") %>% unique() %>% count
-  has_all <- has_all %>% filter(n == 2)
-  filtered_data <- filtered_data %>% filter(id %in% has_all$id)
-  filtered_data$exp_block_id <- factor(filtered_data$exp_block_id)
-  return(filtered_data)
-}
 

### 2018-09-21 Report/paper-story.Rmd
Updates the paper report

@@
 }
+df_errors <- block_measure(condition_all_sums, "sum.errors")
+df_distance <- block_measure(condition_all_sums, "mean.distance")
+
+df_wide_errors <- dcast(condition_all_sums, learning.condition + id ~ exp_block_id, value.var = "sum.errors")
+df_wide_distance <- dcast(condition_all_sums, learning.condition + id ~ exp_block_id, value.var = "mean.distance")
+colnames(df_wide_errors) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")
+colnames(df_wide_distance) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")
 ```
- There will be smaller spatial knowledge transfer in those conditions which use less
 
@@
 ## Descriptives
-In normalised distance we can see that people in different conditions start fairly at the same level of performance (difference between learnign conditions is not significant `r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`).
-Nor there is significant difference between all 5 starting groups `r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$condition`.
+In normalised distance we can see that people in different conditions start fairly at the same level of performance (difference between learning conditions is not significant `r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`).
+Nor there is significant difference between all 5 starting groups `r apa_print(aov(min_norm_distance ~ condition, walk_all[walk_all$exp_block_id == 1, ]))$statistic$condition`.
 
-```{r, echo=F, warning=F}
+```{r, echo=F, warning=F, message=F}
 make_graph(sub_walk_all, "learning.condition", "min_norm_distance", "exp_block_id") + 
@@
 But desktop only in number of errors and marginally in distance improvement.
+All this suggests transfer is happening across all learning conditions.
 
 ### Overrall average walking and error performance in phase 1 vs phase 2
-Given that all participants had 3 blocks of trials in the second phase done in the real world, We can assume, if one learning method leads to better transfer than another, than the performance improvement between phases could be different.
-```{r, echo=F, warning=F}
+Given that all participants had 3 blocks of trials in the second phase done in the real world, We can assume, if one learning method leads to better transfer than another, than the performance improvement between phases could be different as well.
+```{r, echo=F, warning=F,message=F}
 make_graph(sub_walk_all, "learning.condition", "min_norm_distance", "phase") + 
@@
 ```{r, echo=F}
-apa_print(aov(min_norm_distance ~ phase*learning.condition, sub_walk_all))
+kable(apa_print(aov(min_norm_distance ~ phase*learning.condition, sub_walk_all))$table)
 pair_t_norm_distance_phase_2 <- pairwise.t.test(sub_walk_all$min_norm_distance, sub_walk_all$learning.condition, p.adj = "holm")
 ```
-Looking solely at the performance differences in the second phase differs significantly between groups (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$phase == 2,]))$statistic$learning_condition`).
-Tukey test shows significant difference between real adn a desktop versions
+Performance in the second phase differs significantly between groups (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$phase == 2,]))$statistic$learning_condition`), suggesting that the learnign modality has an effect on the overall performance.
+Tukey test shows significant difference between real adn a desktop versions.
+although this can be only because of bad performance in block 4 after switch, not due to blocks 5 and 6 (will test later.)
 
 ```{r, echo=F}
-kable(TukeyHSD(x=aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$phase == 2,]), conf.level=0.95)$learning.condition[, c(1,4)])
+df_temp <- TukeyHSD(x=aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$phase == 2,]), conf.level=0.95)$learning.condition[, c(1,4)]
+kable(tukey_report(df_temp))
 ```
@@
 ### Overrall average pointing performance in phase 1 vs phase 2
-```{r, echo=F, warning=F}
+The issue with comparing pointing is that people point after learning condition and then only after being subjected to the new condition for 18 trials.
+Which means that each participant has experienced the real world for half an hour before pointing.
+As we see laterr, the biggest difference in transfer seems to occur between blocks 3-4, with most participants achieving same performance by the block 6.
+That means that the comparison of pointing is not very informative.
+
+```{r, echo=F, warning=F, message=F}
 make_graph(sub_sop_all, "learning.condition", "abs_error", "phase") + 
@@
 Anova shows significant effect of both phase, learning condition and their interaction on pointing performance.
+Probably due to the fact that participants pointed less acurately on the treadmill and evn less on the desktop, but all then achieved same performance by the end of the 6th block.
 ```{r, echo=F}
@@
 
-Looking directly at the performance after the switch, we can see if there is a sudden drop in performance.
-We can study this both suing individual paired t.tests as well as non paired t-tests.
+Looking directly at the performance after the switch, we can see if there is a sudden drop in path length desktop condition.
+We can study this using paired t.tests as well as non paired t-tests.
 Looking at the graphs we see differences in the level to which people have learned the environemtn in the 3rd block.
-```{r, echo=F, message=F}
+
+```{r, echo=F, warning=F, message=F}
 make_graph(sub_walk_all[sub_walk_all$exp_block_id %in% c(3,4),], "learning.condition", "min_norm_distance", "exp_block_id") + 
@@
 
-Anova doesn¨t show shows a significant difference in the block 3 (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 3,]))$statistic$learning_condition`), but there is a significant difference between path performance in the block 4 (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`).
+Anova doesn't show shows a significant difference between groups in the block 3 (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 3,]))$statistic$learning_condition`), but there is a significant difference between distance performance in the block 4 (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`).
 
-There is significant difference in the number of errors in both block 3 (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 3,]))$statistic$learning_condition`), and the block 4 (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`).
+There is significant difference in the number of errors in both block 3 (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 3,]))$statistic$learning_condition`), and the block 4 (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`)., with treadmill and desktop performing worse than real world.
 
@@
 ```
+
 WE can see that all participants improved in the errors made, but neigher group is significantly worse between block 3 and 4 in the path travelled (although participants learning on desktop perform slightly worse in 4th block, but the significance is only marginal)
@@
 ### Personal walking performance change in block 3 vs 4
-(calculated as an average of individuals (block3-block4)/(block3+block4))
-```{r, echo=F}
-
-```
-
-### Personal walking performance change in block 1 vs 4 
-(calculated as an average of individuals (block1-block4)/(block3+block4))
-```{r, echo=F}
-
-```
-
-### level of correlation of personal performance in block 1,2,3 and 4
-
-- higher transfer should show higher correlation in these comparisons
-```{r, echo=F}
-
-```
-
-POtential followups
-### difference in performance between groups in blocks 5 and 6
-
-(e.g. not immediately after the switch, but after all groups have been subjected the real world environment for some time)
-```{r, echo=F}
-
-```
-### difference in performance between groups in block 4-6 with inclusion of VR-VR group nad real-VR group
-
-(to see if the final performance is subjected to the necessity of a real world experience, or not and to see if the swith has detrimental effect on real-VR as well).
-```{r, echo=F}
-
-```
-But I can look at either pure changes between block 3-4 or 1-4 or those calculations that you proposed as a (block4-block3)/(block3+block4) measurements.
 
 ```{r, echo=F}
-df_wide_errors <- dcast(condition_all_sums, learning.condition + id ~ exp_block_id, value.var = "sum.errors")
-df_wide_distance <- dcast(condition_all_sums, learning.condition + id ~ exp_block_id, value.var = "mean.distance")
-colnames(df_wide_errors) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")
-colnames(df_wide_distance) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")
-
-df_wide_errors$block14 <- (df_wide_errors$block1-df_wide_errors$block4)/(df_wide_errors$block1+df_wide_errors$block4)
 df_wide_errors$block34 <- (df_wide_errors$block3-df_wide_errors$block4)/(df_wide_errors$block3+df_wide_errors$block4)
-df_wide_distance$block14 <- (df_wide_distance$block1-df_wide_distance$block4)/(df_wide_distance$block1+df_wide_distance$block4)
 df_wide_distance$block34 <- (df_wide_distance$block3-df_wide_distance$block4)/(df_wide_distance$block3+df_wide_distance$block4)
-
-df_errors <- block_measure(condition_all_sums, "sum.errors")
-df_distance <- block_measure(condition_all_sums, "mean.distance")
-
-clean_walk_14 <- ez_prepare_block(sub_walk_all, c(1,4))
 clean_walk_34 <- ez_prepare_block(sub_walk_all, c(3,4))
 ```
-## Block improvement from block 1 to block 4
+The other approach is using the "performance change" coefficient as you proposed calculated as an average of individual's (block3-block4)/(block3+block4))
 
-But anova shows no significant difference in the improvement between groups for distance (`r apa_print(aov(block14 ~ learning.condition, df_wide_distance))$statistic$learning_condition`).
-This might be due to quite large variability in the sample.
-
-```{r, echo=F}
-ggplot(df_distance, aes(x = learning.condition, y = block14.diff, fill = learning.condition)) + 
-  geom_bar(stat = 'identity', position="dodge") + 
-  geom_errorbar(aes(ymin=block14.diff-block14.se, ymax=block14.diff+block14.se), width=.2, position=position_dodge(.9)) +
-  scale_fill_manual(values = estimote_pal) +
-  ylab("Mean normalised distance improvement from block 1 to 4") + xlab("Learning condition")  + guides(fill=F)
-```
-
-### Errors
-Unlike for distance, anova reveals significant difference between learning conditions in number of errors improvement from block 1 to block 4 (`r apa_print(aov(block14 ~ learning.condition, df_wide_errors))$statistic$learning_condition`).
-
-Runnign Tukey's post hoc on the distance shows that this is originating from the difference between Desktop and the Real world, with Treadmill not being significantly different from either of them.
-
-```{r, echo=F}
-kable(TukeyHSD(x=aov(block14 ~ learning.condition, df_wide_errors), conf.level=0.95)$learning.condition[, c(1,4)])
-```
-
-```{r, echo=F}
-ggplot(df_errors, aes(x = learning.condition, y = block14.diff, fill = learning.condition)) + 
-  geom_bar(stat = 'identity', position="dodge") + 
-  geom_errorbar(aes(ymin=block14.diff-block14.se, ymax=block14.diff+block14.se), width=.2, position=position_dodge(.9)) +
-  scale_fill_manual(values = estimote_pal) +
-  ylab("Mean sum of errors improvement from block 1 to 4") + xlab("Learning condition")  + guides(fill=F)
-```
-
-## Block performance improvement from block 3 to block 4
-
-Everybody's distance performance (including Real to Real condition) gets a little bit worse from 3 to 4.
-But this is not significant when computed using paired t-tests, although it is marginally worse in the desktop version.
-But all conditions show decrease in the number of errors from block 3 to 4.
-
-```{r, echo=F}
-df_dist <- block_t_test(sub_walk_all, 3, 4,"min_norm_distance")[,c(1,4)]
-df_error <- block_t_test(sub_walk_all, 3, 4,"errors")[,c(1,4)]
-df34 <- merge(df_dist, df_error, by = "learning.condition")
-colnames(df34) <- c("learning.condition", "p-value distance improvement", "p-value errors improvement")
-kable(df34)
-```
-
-### Distance
-
-Comparing the block improvement in different conditions using anovas, we see marginally significant difference in the improvement between groups for distance (`r apa_print(aov(block34 ~ learning.condition, df_wide_distance))$statistic$learning_condition`).
+Comparing the block perfomance change in different conditions using anovas, we see marginally significant difference in the improvement between groups for distance (`r apa_print(aov(block34 ~ learning.condition, df_wide_distance))$statistic$learning_condition`).
 
@@
 Within subject anova for distance improvement between 3 and 4 shows significant interaction effect and marginally significant effect of learning condition and block on distance.
+
 ```{r, warning = F, echo=F, message=F}
@@
 ```
-
-### Errors
 
@@
 ```
+### Personal walking performance change in block 1 vs 4 
+(calculated as an average of individuals (block1-block4)/(block3+block4))
+```{r, echo=F}
+df_wide_errors$block14 <- (df_wide_errors$block1-df_wide_errors$block4)/(df_wide_errors$block1+df_wide_errors$block4)
+df_wide_distance$block14 <- (df_wide_distance$block1-df_wide_distance$block4)/(df_wide_distance$block1+df_wide_distance$block4)
+clean_walk_14 <- ez_prepare_block(sub_walk_all, c(1,4))
+```
+
+Anova shows no significant difference in the improvement between groups for distance (`r apa_print(aov(block14 ~ learning.condition, df_wide_distance))$statistic$learning_condition`).
+This might be due to quite large variability in the sample.
+
+```{r, echo=F}
+ggplot(df_distance, aes(x = learning.condition, y = block14.diff, fill = learning.condition)) + 
+  geom_bar(stat = 'identity', position="dodge") + 
+  geom_errorbar(aes(ymin=block14.diff-block14.se, ymax=block14.diff+block14.se), width=.2, position=position_dodge(.9)) +
+  scale_fill_manual(values = estimote_pal) +
+  ylab("Mean normalised distance improvement from block 1 to 4") + xlab("Learning condition")  + guides(fill=F)
+```
+
+But it reveals significant difference between learning conditions in number of errors improvement from block 1 to block 4 (`r apa_print(aov(block14 ~ learning.condition, df_wide_errors))$statistic$learning_condition`).
+
+Runnign Tukey's post hoc on the distance shows that this is originating from the difference between Desktop and the Real world, with Treadmill not being significantly different from either of them.
+
+```{r, echo=F}
+kable(TukeyHSD(x=aov(block14 ~ learning.condition, df_wide_errors), conf.level=0.95)$learning.condition[, c(1,4)])
+```
+```{r, echo=F}
+ggplot(df_errors, aes(x = learning.condition, y = block14.diff, fill = learning.condition)) + 
+  geom_bar(stat = 'identity', position="dodge") + 
+  geom_errorbar(aes(ymin=block14.diff-block14.se, ymax=block14.diff+block14.se), width=.2, position=position_dodge(.9)) +
+  scale_fill_manual(values = estimote_pal) +
+  ylab("Mean sum of errors improvement from block 1 to 4") + xlab("Learning condition")  + guides(fill=F)
+```
+
+### level of correlation of personal performance in block 1,2,3 and 4
+
+- higher transfer should show higher correlation in these comparisons
+```{r, echo=F}
+
+```
+
+POtential followups
+### difference in performance between groups in blocks 5 and 6
+
+(e.g. not immediately after the switch, but after all groups have been subjected the real world environment for some time)
+```{r, echo=F}
+
+```
+### difference in performance between groups in block 4-6 with inclusion of VR-VR group nad real-VR group
+
+(to see if the final performance is subjected to the necessity of a real world experience, or not and to see if the swith has detrimental effect on real-VR as well).
+```{r, echo=F}
+
+```
 

### 2018-09-21 TwoWorlds/helpers-report.R

@@ -16,3 +16,10 @@ block_t_test <- function(df, block1, block2, value, paired = T){
   df_out$p.value <- apa_p(df_out$p.value)
   return(df_out)
 }
+
+tukey_report <- function(tukey){
+  df <- as.data.frame(tukey)
+  colnames(df) <- c("diff", "p-value")
+  df$`p-value` <- apa_p(df$`p-value`)
+  return(df)
+}

### 2018-09-24 Report/paper-story.Rmd
Updates the paper with mixed models instead of between subject anovas

@@
 library(knitr)
+library(ez)
+library(lme4)
+library(car)
 library(papaja)
@@
 
-Specific tests (hypothesis in parenthesis)
+Specific tests
 - Overrall average walking performance in phase 1 vs phase 2 
@@
 - Personal walking performance change in block 1 vs 4 (calculated as an average of individuals (block1-block4)/(block3+block4))
-- level of correlation of personal performance in block 1,2,3 and 4 - higher transfer should show higher correlation in these comparisons
+- Learning modality has an effect on the learning curve (slope)
+- Level of correlation of personal performance in block 1,2,3 and 4 - higher transfer should show higher correlation in these comparisons
 - Difference in performance between groups in blocks 5 and 6 (e.g. not immediately after the switch, but after all groups have been subjected the real world environment for some time)
@@
 ### Performance in block 4 is better than in block 1
+
 If we assume some level of transfer, than the average or individual performance in block 4 should be better in all conditions than in block 1 or potentially in block 2.
@@
 
-Anova shows a significant effect of phase and condition on overall performance, but not of their interaction.
+Within subject anova shows a significant effect of phase and condition on overall performance, but not of their interaction.
 ```{r, echo=F}
-kable(apa_print(aov(min_norm_distance ~ phase*learning.condition, sub_walk_all))$table)
+fit_dist_phase <- lmer(min_norm_distance ~ phase*learning.condition + (1 | id/phase), data = sub_walk_all)
+kable(mixed_table_report(fit_dist_phase))
 pair_t_norm_distance_phase_2 <- pairwise.t.test(sub_walk_all$min_norm_distance, sub_walk_all$learning.condition, p.adj = "holm")
 ```
+
 Performance in the second phase differs significantly between groups (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$phase == 2,]))$statistic$learning_condition`), suggesting that the learnign modality has an effect on the overall performance.
@@
 Probably due to the fact that participants pointed less acurately on the treadmill and evn less on the desktop, but all then achieved same performance by the end of the 6th block.
+
 ```{r, echo=F}
-kable(apa_print(aov(abs_error ~ phase*learning.condition, sub_sop_all))$table)
+fit_angle_phase <- lmer(abs_error ~ phase*learning.condition + (1 | id/phase), data = sub_sop_all)
+kable(mixed_table_report(fit_dist_phase))
 ```
 
-### Average walking perfomrance in block 3 vs 4
+### Average walking performance in block 3 vs 4
 
@@
 
-Running pairwise t-tests to see performance change from block 3 to 4 gives the following results.
+Within subject anova shows significant effect of both block and its interaction on distance walked.
+```{r, echo=F}
+fit_dist_34 <- lmer(min_norm_distance ~ exp_block_id*learning.condition + (1 | id/phase), data = sub_walk_all[sub_walk_all$exp_block_id %in% c(3,4),])
+kable(mixed_table_report(fit_dist_34))
+```
+
+And significant effect all both conditoin, block and their interaction on the number of errors.
+```{r, echo=F}
+fit_dist_34 <- lmer(errors ~ exp_block_id*learning.condition + (1 | id/phase), data = sub_walk_all[sub_walk_all$exp_block_id %in% c(3,4),])
+kable(mixed_table_report(fit_dist_34))
+```
+
+Running separate pairwise t-tests to see individual performance change from block 3 to 4 gives the following results.
 ```{r, echo=F}
@@
 ```
+
 ### Personal walking performance change in block 1 vs 4 
@@
 
-### level of correlation of personal performance in block 1,2,3 and 4
+### Learning modality has an effect on the learning curve (slope)
+I hope mike or professor Ferer will help with this
+
+### Level of correlation of personal performance in block 1,2,3 and 4
 
@@
 POtential followups
-### difference in performance between groups in blocks 5 and 6
+
+### Difference in performance between groups in blocks 5 and 6
 
@@
 ```
-### difference in performance between groups in block 4-6 with inclusion of VR-VR group nad real-VR group
+
+### Difference in performance between groups in block 4-6 with inclusion of VR-VR group and real-VR group
 

### 2018-09-24 TwoWorlds/helpers-report.R

@@ -23,3 +23,10 @@ tukey_report <- function(tukey){
   df$`p-value` <- apa_p(df$`p-value`)
   return(df)
 }
+
+# NEeds fitted lme model
+mixed_table_report <- function(fit_lme){
+  aov_result <- as.data.frame(car::Anova(fit_lme, type = "III"))
+  aov_result[,3] <- apa_p(aov_result[,3])
+  return(aov_result)
+} 

### 2018-09-24 Report/paper-story.Rmd
Updates the report

@@
 ```{r setup, include=FALSE}
+#needs lme and car in the top because of the recode
+library(car)
+library(lme4)
 library(ggplot2)
@@
 library(knitr)
-library(ez)
-library(lme4)
-library(car)
 library(papaja)
@@
 Specific tests
+
 - Overrall average walking performance in phase 1 vs phase 2 
 - Performance in block 4 vs performance in block 1 and 2 
-
 - Overrall average pointing performance in phase 1 vs phase 2 
-- Average walking perfomrance in block 3 vs 4 (block directly before vs directly after the switch)
+- Average walking performance in block 3 vs 4 (block directly before vs directly after the switch)
 - Personal walking performance change in block 3 vs 4 (calculated as an average of individuals (block3-block4)/(block3+block4))
@@
 I only selected those "important" comparisons - therefore Desktop to Real world, Treadmill VR to real world and Real world to real world.
-Anovas and other comparisons are run only on these.
+Anovas, mixed models and other comparisons are run only on these.
 Learning conditions (block 1-3) are the different modalities, and blocks 4-6 are all real world testing.
@@
 I don't test for block 3 because that immediately compares the switch performance.
-```{r, echo=F,message=F}
+```{r, echo=F, message=F, warning=F}
 make_graph(sub_walk_all[sub_walk_all$exp_block_id %in% c(1,4),], "learning.condition", "min_norm_distance", "exp_block_id") + 
@@
 Given that all participants had 3 blocks of trials in the second phase done in the real world, We can assume, if one learning method leads to better transfer than another, than the performance improvement between phases could be different as well.
-```{r, echo=F, warning=F,message=F}
+```{r, echo=F, warning=F, message=F}
 make_graph(sub_walk_all, "learning.condition", "min_norm_distance", "phase") + 
@@
 
-Within subject anova shows a significant effect of phase and condition on overall performance, but not of their interaction.
+Mixed models show a significant effect of phase on overall performance, but not of their interaction.
 ```{r, echo=F}
@@
 
-Anova shows significant effect of both phase, learning condition and their interaction on pointing performance.
+Mixed models show effect of both phase, learning condition and their interaction on pointing performance.
 Probably due to the fact that participants pointed less acurately on the treadmill and evn less on the desktop, but all then achieved same performance by the end of the 6th block.
@@
 fit_angle_phase <- lmer(abs_error ~ phase*learning.condition + (1 | id/phase), data = sub_sop_all)
-kable(mixed_table_report(fit_dist_phase))
+kable(mixed_table_report(fit_angle_phase))
 ```
 
-### Average walking performance in block 3 vs 4
+### Comparing walking performance in block 3 vs 4
 
@@
 
-Within subject anova shows significant effect of both block and its interaction on distance walked.
+Mixed model shows significant effect of both block and its interaction on distance walked.
 ```{r, echo=F}
-fit_dist_34 <- lmer(min_norm_distance ~ exp_block_id*learning.condition + (1 | id/phase), data = sub_walk_all[sub_walk_all$exp_block_id %in% c(3,4),])
+fit_dist_34 <- lmer(min_norm_distance ~ exp_block_id*learning.condition + (1 | id/exp_block_id), data = sub_walk_all[sub_walk_all$exp_block_id %in% c(3,4),])
 kable(mixed_table_report(fit_dist_34))
@@
 ```{r, echo=F}
-fit_dist_34 <- lmer(errors ~ exp_block_id*learning.condition + (1 | id/phase), data = sub_walk_all[sub_walk_all$exp_block_id %in% c(3,4),])
+fit_dist_34 <- lmer(errors ~ exp_block_id*learning.condition + (1 | id/exp_block_id), data = sub_walk_all[sub_walk_all$exp_block_id %in% c(3,4),])
 kable(mixed_table_report(fit_dist_34))
@@
 
-Within subject anova for distance improvement between 3 and 4 shows significant interaction effect and marginally significant effect of learning condition and block on distance.
-
 ```{r, warning = F, echo=F, message=F}
-aov_block_dist_34 <- ezANOVA(
-  data = clean_walk_34,
-  dv = min_norm_distance, 
-  wid = id,
-  within = exp_block_id,
-  between = learning.condition, 
-  type = 3)
-kable(aov_block_dist_34)
-
-ggplot(clean_walk_34, aes(x = as.numeric(exp_block_id), y = min_norm_distance, color = factor(learning.condition))) +
+ggplot(sub_walk_all[sub_walk_all$exp_block_id %in% c(3,4),], aes(x = as.numeric(exp_block_id), y = min_norm_distance, color = factor(learning.condition))) +
   stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
@@
 ```{r, echo=F}
-kable(TukeyHSD(x=aov(block34 ~ learning.condition, df_wide_errors), conf.level=0.95)$learning.condition[, c(1,4)])
+kable(tukey_report(TukeyHSD(x=aov(block34 ~ learning.condition, df_wide_errors), conf.level=0.95)$learning.condition[, c(1,4)]))
 ```
@@
 
-Running within subjects anova for block 3-4 number of errors shows significant effect of both condition and block, but unlike in distance improvement, their interaction doesn't seem to have an effect on error rate improvement.
 ```{r, warning = F, echo=F, message=F}
-aov_block_err_34 <- ezANOVA(
-  data = clean_walk_34,
-  dv = errors, 
-  wid = id,
-  within = exp_block_id,
-  between = condition,
-  type = 3)
-kable(aov_block_err_34)
-
-ggplot(clean_walk_34, aes(x = as.numeric(exp_block_id), y = errors, color = factor(condition))) +
+ggplot(sub_walk_all[sub_walk_all$exp_block_id %in% c(3,4),], aes(x = as.numeric(exp_block_id), y = errors, color = factor(condition))) +
   stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
@@
 (calculated as an average of individuals (block1-block4)/(block3+block4))
+
 ```{r, echo=F}
@@
 ```{r, echo=F}
-kable(TukeyHSD(x=aov(block14 ~ learning.condition, df_wide_errors), conf.level=0.95)$learning.condition[, c(1,4)])
+kable(tukey_report(TukeyHSD(x=aov(block14 ~ learning.condition, df_wide_errors), conf.level=0.95)$learning.condition[, c(1,4)]))
 ```
+
 ```{r, echo=F}
@@
 
-- higher transfer should show higher correlation in these comparisons
-```{r, echo=F}
+This is a bit weird way of comparing things and it should be somewhat duplicit to the mixed models, but we can measure how much performance from 3-3 correlates between different conditions.
+I don't know if it isn't actually measured in the mixed models, Mike should know this better.
 
+Table for 
+```{r, echo=F, warning=F}
+df_dist<-block_correlation(sub_walk_all, 3,4, "min_norm_distance")
+df_err<-block_correlation(sub_walk_all, 3,4, "errors")
+df_out <- merge(df_dist, df_err, by = "learning.condition")
+colnames(df_out) <- c("Learning condition", "Correlation of normalised distance", "Correlation of number of errors")
+kable(df_out)
 ```
 
-POtential followups
+I don't know how to interpret this properly, but it looks as the data before suggested - that the least correlation is between desktop in 3rd block and real in 4th.
+whereas in real to real, the performance correlates very well.
+The number of errors seems to improve in all cases, so that is why we see medium correlation is across the board.
 
 ### Difference in performance between groups in blocks 5 and 6 
-
-(e.g. not immediately after the switch, but after all groups have been subjected the real world environment for some time)
-```{r, echo=F}
-
-```
+Not sure if this is interesting or we should focus mainly on the slope analyses.
 
@@
 
-(to see if the final performance is subjected to the necessity of a real world experience, or not and to see if the swith has detrimental effect on real-VR as well).
-```{r, echo=F}
-
-```
+And here I am not sure if the inclusion of the other groups affects the results in some way - especially in mixed models, Mike said we should pick a model and stick with it.
+I have an intuition in how to use the other groups to "confirm" or elaborate on our results, but I am unsure how to formulate it or how specifically to run the analyses.
 
 ## Summary
-We see effect of learning condition (blocks 1-3) on performance in block 4.
+We see effect of learning condition (blocks 1-3) on performance in block 4 by t-tests in all conditions separately.
 
@@
 
-Participant's performance in distance from block 3-4 gets slightly worse after the switch distance-wise (but not significanly), but their number of errors decrease across all conditions.
-It can mean that they know which door are correct, but they are unsure in how to get to them, showing some level of memory transfer, but not necessarily spatial knowledge transfer.
+We see improvement from phase 1-2 in both walking, errors and pointing precision.
 
-We can see marginally significant difference in block 3-4 distance improvement acrosss learning conditions, with desktop doing the worst but not significantly different from the rest.
-Within subject anova for distance improvement in blocks 3-4 shows interaction between learning condition and experimental block - only desktop gets "worse" after the switch of environment.
+Participant's performance in distance from block 3-4 gets slightly worse after the switch distance-wise (but not significanly).
 
-But we can see clear error rate improvement differences across conditions - all participants improved, but desktop learning shows least amount of improvement from block 3-4, participants doing significantly worse than those with real world training and marginally worse than those with Treadmill training.
+number of errors decreases (improves) across all  conditions., but desktop learning shows least amount of improvement from block 3-4 and 1-4, participants doing significantly worse than those with real world training and marginally worse than those with Treadmill training.
+It may mean that they know which door are correct, but they are unsure in how to get to them, showing some level of memory transfer, but not necessarily spatial knowledge transfer.
+
+Only desktop shows worse distance performance in block 4 (marginal significance),. treadmil-real and real-real stays relatively the same (good transfer - not improvement).
+
+There is a significant effect in mixed models of learning condition and interaction on distance (marginal effect of learning condition) and errors from block 3-4.
+
+Personal distance performance in blocks 3 and 4 in real world and treadmil doesn't correlate much, but real to real does. Correlation of error performance is similar across all conditions.
+
+We see somewhat clearly the trend Real >= Treadmill VR > Desktop, but often only marginally, due to large variability of the result.
+Treadmil and real are often different in an expected way but not significantly different.
+We usually get significant differences only in real to desktop comparisons.

### 2018-09-24 TwoWorlds/helpers-report.R

@@ -17,6 +17,16 @@ block_t_test <- function(df, block1, block2, value, paired = T){
   return(df_out)
 }
 
+block_correlation <- function(df, block1, block2, value){
+  df <- df %>% filter(exp_block_id %in% c(block1,block2))
+  mean.na <- function(data){return(mean(data, na.rm = T))}
+  df_wide <- dcast(df, learning.condition + id ~ exp_block_id, value.var = value, fun.aggregate = mean.na)
+  colnames(df_wide) <- c("learning.condition","id", "block1","block2")
+  df_out <- df_wide %>% group_by(learning.condition) %>% do(tidy(cor(.$block1, .$block2)))
+  df_out$x <- round(df_out$x, 2)
+  return(df_out)
+}
+
 tukey_report <- function(tukey){
   df <- as.data.frame(tukey)
   colnames(df) <- c("diff", "p-value")

### 2018-09-24 TwoWorlds/helpers-results.R

@@ -17,3 +17,5 @@ ez_prepare_block <- function(df, blocks){
   filtered_data$exp_block_id <- factor(filtered_data$exp_block_id)
   return(filtered_data)
 }
+
+mean.na <- function(data){return(mean(data, na.rm = T))}

### 2018-10-15 Report/descriptives.Rmd
removes old reports, adds descriptives

---
title: "R Notebook"
output: github_document
---

```{r setup}
#load the participant data
library(ggplot2)
library(dplyr)
library(knitr)
library(reshape2)
library(googlesheets)
source('../Scripts/paper-prepare.R')
settings <- load_google_sheets()
df_participants <- settings$participants
df_participants <- df_participants %>% mutate(condition=paste0(First.phase,"-",Second.phase))
overview <- settings$versions
overview <- overview %>% mutate(condition=paste0(First.phase,"-",Second.phase))
```

## Dropout ratings
```{r}
finished <- overview %>% group_by(condition) %>% count(finished)
finished <- dcast(finished, condition~finished)
finished$percent <- round(finished$yes/(finished$yes+finished$no), 2)
kable(finished)

is_ok <- overview %>% filter(finished=="yes") %>% group_by(condition) %>% count(is_ok)
is_ok <- dcast(is_ok, condition~is_ok)
is_ok$percent <- round(is_ok$yes/(is_ok$yes+is_ok$no), 2)
kable(is_ok)
```
A total of `r sum(finished$no)` participants didn't finish due to motion sickness, with about 50% not finishing the VR to VR condition.

## Female/male

## Age

## Descriptives for the VR/real/desktop - time and path traveled

### 2018-10-15 TwoWorlds/twoworld-loading.R
CHanges renaming of the participant data

@@ -139,9 +139,16 @@ load_google_sheets <- function(){
   tw_participants <- fetch_sheet('TW-Participants', 'Overview')
   tw_participants <- tw_participants$Overview
   ls$participants <- merge(tw_participants, tw_questionnaire, by = 'Code')
-  
-  colnames(ls$participants)[18:20] <- c("age", "weight", "hours_sleep")
-  colnames(ls$participants)[44:45] <- c("sex", "ethnicity")
+  colnames(ls$participants)[18:45] <- c("age", "weight", "hours_sleep", 
+                                        "last_eaten", "video_game_experience", 
+                                        "vr_experience", "motion_sickness", 
+                                        "dizzy_history", "stress_level", "stressors", 
+                                        "general_discomfort", "fatigue", "headache", 
+                                        "eye_strain", "focus_difficulty", "salivation", 
+                                        "sweating", "nausea", "concentrate_difficulty", 
+                                        "full_head", "blurred_vision", "dizzy_eyes_open", 
+                                        "dizzy_eyes_closed", "vertigo", 
+                                        "stomach_awareness", "burping", "sex", "ethnicity")
   
   return(ls)
 }

### 2018-10-15 Scripts/analysis.R
Some changes to the participant loading

@@ -27,7 +27,6 @@ summary(aov_distance_condition)
 aov_time_condition <- aov(time~condition*phase, data = walk_all)
 summary(aov_time_condition)
 
-
 ## Arne ----
 #' for each dependent measure, (block 4 - block 3) / (block 4+block3) 
 #' so we can compare transfer for ve vs vr vs reality (ie all 5 conditions)

### 2018-10-15 Scripts/main-preprocess.R

@@ -7,8 +7,7 @@ library(dplyr)
 library(reshape2)
 library(googlesheets)
 source("Scripts/loading.R")
-dir <- "D:/OneDrive/Vyzkum/Davis/Transfer/Data/"
-dir <- "C:/Users/hejtm/OneDrive/Vyzkum/Davis/Transfer/Data/"
+dir <- "M:/OneDrive/Vyzkum/Davis/Transfer/Data/"
 
 settings <- load_google_sheets()
 save(settings, file = "settings.data")
@@ -18,7 +17,6 @@ save(ls, file = "multi_smoothed_2.data")
 
 sop_all <- multi_sop_results(ls)
 walk_all <- multi_walk_results(ls)
-
 
 ##renames all jesiccas
 walk_all <- walk_all %>% mutate(goal = replace(goal, goal=="Jesicca's office", "Jessica's office"),
@@ -78,6 +76,4 @@ walk_all_new <- walk_all %>% select(-one_of(drop.cols))
 sop_all$abs_error <- abs(sop_all$error)
 
 write.table(sop_all, file = "sop_smoothed.csv", sep=";", row.names = F)
-write.table(walk_all_new, file = "walk_smoothed_errors_06.csv", sep=";", row.names = F)
-
-write.table(walk_all, file = "walk.csv", sep=";", row.names = F)
+write.table(walk_all_new, file = "walk_smoothed.csv", sep=";", row.names = F)

### 2018-10-15 Scripts/missing-data-control.R

@@ -1,18 +1,0 @@
-### CHECKINIG MISSING DATA
-walk_all <- walk_all %>% group_by(id, phase) %>% mutate(na_values_phase = sum(is.na(distance)))
-walk_all <- walk_all %>% group_by(id) %>% mutate(na_values = sum(is.na(distance)))
-
-qplot(walk_all$na_values_phase[seq(1, nrow(walk_all), 18)] , bins=18)
-plot(walk_all %>% group_by(na_values) %>% summarise(mean = mean(time)))
-summary(lm(walk_all$time~walk_all$na_values))
-summary(lm(walk_all$time~walk_all$na_values_phase))
-
-summary(lm(walk_all$errors~walk_all$na_values))
-summary(lm(walk_all$errors~walk_all$na_values_phase))
-
-lm_walk_distance <- walk_all %>% filter(!is.na(distance))
-summary(lm(lm_walk_distance$time~lm_walk_distance$na_values))
-summary(lm(lm_walk_distance$time~lm_walk_distance$na_values_phase))
-
-
-        

### 2018-10-15 Scripts/paper-prepare.R

@@ -1,3 +1,4 @@
+library(dplyr)
 source("../Scripts/loading.R")
 source_folder("../TwoWorlds")
 source('../Scripts/analysis-helpers.R')

### 2018-10-15 Scripts/participant-check.R

@@ -5,22 +5,22 @@ library(dplyr)
 library(reshape2)
 library(googlesheets)
 source("Scripts/loading.R")
-dir <- "D:/OneDrive/Vyzkum/Davis/Transfer/Data/"
+dir <- "M:/OneDrive/Vyzkum/Davis/Transfer/Data/"
 #dir <- "C:/Users/hejtm/OneDrive/Vyzkum/Davis/Transfer/Data/"
 
 settings <- load_google_sheets()
-
-code <- "tw42"
+#101, 104
+code <- "tw101"
 ls <- load_participant(code, settings, dir)
 sop_results(ls$phase1)
 sop_results(ls$phase2)
 walk_results(ls$phase1) #takes long time and is correct if sop is correct
 walk_results(ls$phase2)
 
-plot_walk_trial(ls$phase1, 5)
+plot_walk_trial(ls$phase2, )
 
 plot_all(ls$phase1, 1:12, plot_sop_point)
-plot_all(ls$phase1, 1:18, plot_walk_trial)
+plot_all(ls$phase2, 1:18, plot_walk_trial)
 
-diff(get_action_times.restimote(ls$phase1, "New trial"))
+diff(get_action_times.restimote(ls$phase2, "New trial"))
 ls$phase2$log[ls$phase2$log$Action == "pointed","Time"]

### 2018-10-15 Report/descriptives.Rmd
new reports

@@
 df_participants <- df_participants %>% mutate(condition=paste0(First.phase,"-",Second.phase))
+df_participants <- df_participants %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
 overview <- settings$versions
 overview <- overview %>% mutate(condition=paste0(First.phase,"-",Second.phase))
+overview <- overview %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
 ```
@@
 
-## Age
+# Report
+A total of `r nrow(df_participants)` (M = `r round(mean(df_participants$age, na.rm=T),1)`, SD = `r round(sd(df_participants$age, na.rm=T), 1)`) undergraduate students at UC Davis particiapted in the study in exchange for a study credit.
+Each students was randomly assigned a condition and a randomised set of goals.
+`r sum(finished$no)` participants didn't finish due to motion sickness and `r sum(is_ok$no)` were removed due to a technical failure of the real world tracking systen.
+Only `r finished[finished$condition == "vr-real","percent"] * 100` percent the students being able to finish VR to Real world condition.
 
 ## Descriptives for the VR/real/desktop - time and path traveled
+
+```{r}
+kable(walk_all %>% group_by(type, exp_block_id) %>% summarize(mean_dist = mean(distance, na.rm=T), mean_time = mean(time, na.rm=T)))
+```

### 2018-10-15 Scripts/main-preprocess.R

@@ -12,7 +12,7 @@ dir <- "M:/OneDrive/Vyzkum/Davis/Transfer/Data/"
 settings <- load_google_sheets()
 save(settings, file = "settings.data")
 ls <- load_all(settings, dir, only_ok = T)
-save(ls, file = "multi_smoothed_2.data")
+save(ls, file = "multi_smoothed.data")
 #load(file = "multi_smoothed.data")
 
 sop_all <- multi_sop_results(ls)

### 2018-10-20 Report/descriptives.Rmd
Moves functions to helpers, creates new report part

@@
 ## Female/male
+```{r, echo=F}
+kable(df_participants %>% filter(finished=="yes") %>% group_by(condition, sex) %>% count)
+```
 
 # Report
-A total of `r nrow(df_participants)` (M = `r round(mean(df_participants$age, na.rm=T),1)`, SD = `r round(sd(df_participants$age, na.rm=T), 1)`) undergraduate students at UC Davis particiapted in the study in exchange for a study credit.
+A total of `r sum(overview$is_ok == "yes", na.rm=T)` (M = `r round(mean(df_participants[df_participants$finished == "yes",]$age, na.rm=T),1)`, SD = `r round(sd(df_participants[df_participants$finished == "yes",]$age, na.rm=T), 1)`) undergraduate students at UC Davis particiapted in the study in exchange for a study credit.
 Each students was randomly assigned a condition and a randomised set of goals.
@@
 ```{r}
-kable(walk_all %>% group_by(type, exp_block_id) %>% summarize(mean_dist = mean(distance, na.rm=T), mean_time = mean(time, na.rm=T)))
+df_desc <- walk_all %>% group_by(type, exp_block_id) %>% summarize(mean_dist = mean(distance, na.rm=T), mean_time = mean(time, na.rm=T))
+ggplot(df_desc, aes(exp_block_id, mean_dist, fill=type)) + geom_bar(stat="identity", position = "dodge")
+ggplot(df_desc, aes(exp_block_id, mean_time, fill=type)) + geom_bar(stat="identity", position = "dodge")
 ```

### 2018-10-20 Report/paper-results.Rmd
Moves functions to helpers, creates new report part

---
title: "R Notebook"
output: github_document
---

```{r setup, include=FALSE}
#needs lme and car in the top because of the recode
library(car)
library(lme4)
library(ggplot2)
library(ez)
library(Hmisc)
library(dplyr)
library(reshape2)
library(broom)
library(knitr)
library(papaja)
library(googlesheets)

source("../Scripts/paper-prepare.R")
source("helpers/block-t-test-improvement.R")
source("helpers/block-t-test.R")
source("helpers/block-measure.R")

settings <- load_google_sheets()
df_participants <- settings$participants
df_participants <- df_participants %>% mutate(condition=paste0(First.phase,"-",Second.phase))
df_participants <- df_participants %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
overview <- settings$versions
overview <- overview %>% mutate(condition=paste0(First.phase,"-",Second.phase))
overview <- overview %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))

#green, violet, blue, yellow, pink, dark blue
estimote_pal <- c('#b6d6c1', '#54003d','#93d5f6','#d5d215','#f3aacb','#2d2556')
theme_set(theme_minimal())

condition_all_sums <- sub_walk_all %>% 
  group_by(exp_block_id, id, learning.condition) %>% 
  summarise(sum.distance = sum(min_norm_distance, na.rm = T), 
            sum.errors = sum(errors, na.rm = T),
            mean.distance = mean(min_norm_distance, na.rm = T),
            mean.error = mean(errors, na.rm = T))

df_errors <- block_measure(condition_all_sums, "sum.errors")
df_distance <- block_measure(condition_all_sums, "mean.distance")

df_wide_errors <- dcast(condition_all_sums, learning.condition + id ~ exp_block_id, value.var = "sum.errors")
df_wide_distance <- dcast(condition_all_sums, learning.condition + id ~ exp_block_id, value.var = "mean.distance")
colnames(df_wide_errors) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")
colnames(df_wide_distance) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")

knitr::opts_chunk$set(echo=F, warning = F, message = F)

```

# Methods

## Measures
We calculated participants’ walked distance in the building, time spent in each task and number of incorrectly visited doors. The conditions were not perfectly comparable in their measured distance and time, with the VR trials taking a little bit longer in both time and distance than the real world condition. We therefore min-normalised the distance traveled and the time for each task and environment  by dividing the measure by any participant's performance for that trial. We calculated all statistics on these normalised measures.

There was no difference between conditions at the start of the experiment in normalised distance (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`), nor in number errors (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`), but there is a significant difference in normalised time (`r apa_print(aov(min_norm_time ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`)), due to VR taking a longer times to get used to (and therefore having longer normalised trial times at the start).

## Participants
```{r}
finished <- overview %>% group_by(condition) %>% count(finished)
finished <- dcast(finished, condition~finished)
finished$percent <- round(finished$yes/(finished$yes+finished$no), 2)
is_ok <- overview %>% filter(finished=="yes") %>% group_by(condition) %>% count(is_ok)
is_ok <- dcast(is_ok, condition~is_ok)
is_ok$percent <- round(is_ok$yes/(is_ok$yes+is_ok$no), 2)
```
A total of `r sum(overview$is_ok == "yes", na.rm=T)` (M = `r round(mean(df_participants[df_participants$finished == "yes",]$age, na.rm=T),1)`, SD = `r round(sd(df_participants[df_participants$finished == "yes",]$age, na.rm=T), 1)`) undergraduate students at UC Davis particiapted in the study in exchange for a study credit. Each students was randomly assigned a condition and a randomised set of goals. `r sum(finished$no)` participants didn't finish due to motion sickness and `r sum(is_ok$no)` were removed due to a technical failure of the real world tracking systen. Only `r finished[finished$condition == "vr-real","percent"] * 100` percent the students being able to finish VR to Real world condition.

# Results

First we wanted to assess that there is going to be some level of transfer in all conditions. All conditions (except desktop in distance) show better performance after the switch (see table XXX), suggesting some level of transfer from either learnign condition.

```{r,echo=F}
df <- data.frame(condition=rep("",3),dist=rep("",3), err=rep("",3), stringsAsFactors = F)
df_dist <- block_t_test_improvement(sub_walk_all, 1, 4,"min_norm_distance")
df_error <- block_t_test_improvement(sub_walk_all, 1, 4,"errors")
for(i in 1:length(names(df_dist))){
  name <- names(df_dist)[i]
  df[i,"condition"] <- name
  df[i, "dist"] <- apa_print(df_dist[[name]])$full_result
  df[i, "err"] <- apa_print(df_error[[name]])$full_result
}

colnames(df) <- c("Learning condition", "Distance improvement", "Errors improvement")
kable(df, caption="Paired T test comparing individual performance improvement from block 1-4 in different conditions")
```

### 2018-10-20 Report/paper-story.Rmd
new reports
Moves functions to helpers, creates new report part

@@
 date: "9 August 2018"
-output: html_document
+output: github_document
 ---
@@
 colnames(df14) <- c("learning condition", "p-value distance improvement", "p-value errors improvement")
-kable(df14, caption="Paired T test comparing infividual performance improvement from block 1-4 in different conditions")
+kable(df14, caption="Paired T test comparing individual performance improvement from block 1-4 in different conditions")
 ```

### 2018-10-20 Report/helpers/block-measure.R

@@ -1,0 +1,15 @@
+block_measure <- function(df, value){
+  df_wide <- dcast(df, learning.condition + id ~ exp_block_id, value.var = value)
+  colnames(df_wide) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")
+  
+  df_wide %>% 
+    group_by(learning.condition) %>%
+    summarise(block1.mean  = mean(block1, na.rm = T),
+              block3.mean = mean(block3, na.rm = T),
+              block4.mean = mean(block4, na.rm = T),
+              block34.diff = mean((block3-block4)/(block4+block3), na.rm = T),
+              block34.se  = sqrt(var((block3-block4)/(block4+block3), na.rm = T)/sum(!is.na(block3-block4))),
+              block14.diff = mean((block1-block4)/(block1+block4), na.rm = T),
+              block14.se = sqrt(var((block1-block4)/(block1+block4), na.rm = T)/sum(!is.na(block1-block4)))
+    )
+}

### 2018-10-20 Report/helpers/block-t-test-improvement.R

@@ -1,0 +1,12 @@
+block_t_test_improvement <- function(df, block1, block2, value){
+  df <- df %>% filter(exp_block_id %in% c(block1,block2))
+  mean.na <- function(data){return(mean(data, na.rm = T))}
+  df_wide <- dcast(df, condition + id ~ exp_block_id, value.var = value, fun.aggregate = mean.na)
+  colnames(df_wide) <- c("condition","id", "block1","block2")
+  conditions <- unique(df_wide$condition)
+  ls <- list()
+  for(cond in conditions){
+    ls[[cond]] <- t.test(df_wide[df_wide$condition == cond,"block1"], df_wide[df_wide$condition == cond,"block2"], paired=T)
+  }
+  return(ls)
+}

### 2018-10-20 Report/helpers/block-t-test.R

@@ -1,0 +1,7 @@
+block_t_test <- function(df, block1, block2, value){
+  df <- df %>% filter(exp_block_id %in% c(block1,block2))
+  mean.na <- function(data){return(mean(data, na.rm = T))}
+  df_wide <- dcast(df, condition + id ~ exp_block_id, value.var = value, fun.aggregate = mean.na)
+  colnames(df_wide) <- c("condition","id", "block1","block2")
+  df_wide %>% group_by(condition) %>% do(tidy(t.test(.$block1, .$block2, paired = T)))
+}

### 2018-10-20 Scripts/paper-prepare.R

@@ -5,11 +5,11 @@ source('../Scripts/analysis-helpers.R')
 sop_all <- read.table("../sop.csv", sep=";", header = T, stringsAsFactors = F)
 walk_all <- read.table("../walk.csv", sep=";", header = T, stringsAsFactors = F)
 walk_all$learning.condition <- walk_all$condition
-walk_all$learning.condition <- recode(walk_all$learning.condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
+walk_all$learning.condition <- dplyr::recode(walk_all$learning.condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
 sop_all$learning.condition <- sop_all$condition
 sop_all$testing.condition <- sop_all$condition
-sop_all$learning.condition <- recode(sop_all$learning.condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
-sop_all$testing.condition <- recode(sop_all$testing.condition, "real-real" = "Real", "ve-real" = "Real", "vr-real" = "Real", "real-vr" = "Treadmill VR", "vr-vr" = "Treadmill VR")
+sop_all$learning.condition <- dplyr::recode(sop_all$learning.condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
+sop_all$testing.condition <- dplyr::recode(sop_all$testing.condition, "real-real" = "Real", "ve-real" = "Real", "vr-real" = "Real", "real-vr" = "Treadmill VR", "vr-vr" = "Treadmill VR")
 
 sub_walk_all <- walk_all %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
 sub_sop_all <- sop_all %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))

### 2018-10-21 Report/paper-all-data.Rmd
new reports
Updates reports, moves functions

@@
 date: "4 June 2018"
-output: html_document
+output: github_document
 ---
@@
 ```
-
-# Block measures
-```{r, echo = F}
-condition_all_sums <- sub_walk_all %>% 
-  group_by(exp_block_id, id, learning.condition) %>% 
-  summarise(sum.distance = sum(min_norm_distance, na.rm = T), 
-            sum.errors = sum(errors, na.rm = T),
-            mean.distance = mean(min_norm_distance, na.rm = T),
-            mean.error = mean(errors, na.rm = T))
-
-block_measure <- function(df, value){
-  df_wide <- dcast(df, learning.condition + id ~ exp_block_id, value.var = value)
-  colnames(df_wide) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")
-  df_wide %>% 
-    group_by(learning.condition) %>%
-    summarise(block1.mean  = mean(block1, na.rm = T),
-              block3.mean = mean(block3, na.rm = T),
-              block4.mean = mean(block4, na.rm = T),
-              block34.diff = mean((block3-block4)/(block4+block3), na.rm = T),
-              block34.se  = sqrt(var((block3-block4)/(block4+block3), na.rm = T)/sum(!is.na(block3-block4))),
-              block14.diff = mean((block1-block4)/(block1+block4), na.rm = T),
-              block14.se = sqrt(var((block1-block4)/(block1+block4), na.rm = T)/sum(!is.na(block1-block4)))
-  )
-}
-```
-
 ## Sum of errors
@@
 df_wide <- dcast(condition_all_sums, learning.condition + id ~ exp_block_id, value.var = "sum.errors")
-colnames(df_wide) <- c("learning.condition","id", "block1","block2","block3","block4","block5","block6")
+colnames(df_wide) <- c("learning.condition", "id", "block1", "block2", "block3", "block4", "block5", "block6")
 df_wide$diff41 <- (df_wide$block4-df_wide$block1)/(df_wide$block1+df_wide$block3)
@@
   type = 3)
+
 kable(aov_block_dist_34)
@@
   type = 3)
+
 kable(aov_block_dist_14)
@@
   type = 3)
+
 kable(aov_block_err_14)
@@
   type = 3)
+
 kable(aov_block_err_34)
@@
 ```
+
+## Office specific learning
+```{r, warning=F}
+df_corr_err <- data.frame(learning.condition=c("Desktop", "Real", "Treadmill VR"), stringsAsFactors = F)
+df_corr_dist <- data.frame(learning.condition=c("Desktop", "Real", "Treadmill VR"), stringsAsFactors = F)
+for(i in 1:5){
+  df_err <- block_correlation(sub_walk_all, i,i+1, "errors", learning.condition + id + goal ~ exp_block_id, columns=c("learning.condition","id", "goal"))
+  colnames(df_err) <- c("learning.condition", paste0("block", i+1))
+  df_corr_err <- merge(df_corr_err, data.frame(df_err), by="learning.condition")
+
+  df_dist <- block_correlation(sub_walk_all, i,i+1, "min_norm_distance", learning.condition + id + goal ~ exp_block_id, columns=c("learning.condition","id", "goal"))
+  colnames(df_dist) <- c("learning.condition", paste0("block", i+1))
+  df_corr_dist <- merge(df_corr_dist, data.frame(df_dist), by="learning.condition")
+}
+df_corr_err
+df_corr_dist
+```

### 2018-10-21 Report/paper-results.Rmd
Updates reports, moves functions

@@
 library(lme4)
+library(lmerTest)
 library(ggplot2)
@@
 colnames(df_wide_distance) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")
+df_wide_errors$block34 <- (df_wide_errors$block3-df_wide_errors$block4)/(df_wide_errors$block3+df_wide_errors$block4)
+df_wide_distance$block34 <- (df_wide_distance$block3-df_wide_distance$block4)/(df_wide_distance$block3+df_wide_distance$block4)
 
@@
 
-There was no difference between conditions at the start of the experiment in normalised distance (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`), nor in number errors (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`), but there is a significant difference in normalised time (`r apa_print(aov(min_norm_time ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`)), due to VR taking a longer times to get used to (and therefore having longer normalised trial times at the start).
+There was no difference between conditions at the start of the experiment in normalised distance (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`), nor in number errors (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`), but there is a significant difference in normalised time (`r apa_print(aov(min_norm_time ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`)), due to VR taking longer times to get used to (and therefore having longer normalised trial times at the start).
 
@@
 First we wanted to assess that there is going to be some level of transfer in all conditions.
-All conditions (except desktop in distance) show better performance after the switch (see table XXX), suggesting some level of transfer from either learnign condition.
+If we would assume no transfer to happen, we should observe same performance in the real world testing in 1st, 2nd and 3rd block after the switch to be similar in performance as the 1st, 2nd and 3rd block in the real world learning group.
+But all conditions (except desktop in distance) show better performance after the switch (see table XXX) than in the first trial, suggesting some level of transfer from all learning conditions.
+This is not the case with distance improvement in the desktop learning group.
 
-```{r,echo=F}
+```{r}
 df <- data.frame(condition=rep("",3),dist=rep("",3), err=rep("",3), stringsAsFactors = F)
@@
 }
-
 colnames(df) <- c("Learning condition", "Distance improvement", "Errors improvement")
@@
 ```
+
+Looking at the progression in distance improvements and error improvements, we can see that all conditions show improvement over the experiment progression which is only hindered by the switch in the desktop learning condition (see fig.
+XXX and fig.
+XXX), but otherwise we see constant improvement unobstructed by the environemnt switch.
+```{r}
+ggplot(sub_walk_all, aes(learning.condition, min_norm_distance, fill=factor(exp_block_id))) +
+  stat_summary(fun.data=mean_cl_normal, position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean, position=position_dodge(width=0.95), geom="bar") +
+  ylab("Average normalised trial distance") + xlab("Learning condition") + 
+  scale_fill_manual(values=estimote_pal,"Block ID") + guides(fill=F)
+
+ggplot(sub_walk_all, aes(learning.condition, errors, fill=factor(exp_block_id))) +
+  stat_summary(fun.data=mean_cl_normal, position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean, position=position_dodge(width=0.95), geom="bar") +
+  ylab("Average number of errors") + xlab("Learning condition") + 
+  scale_fill_manual(values=estimote_pal,"Block ID") + guides(fill=F)
+```
+
+Participant's performance in the first pre-switch phase shows different rates of learning rate for errors for our learning conditions (see table XX).
+There is no significant difference between different modalities in distance performance before the switch `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 3)))$full_result`, but the performance across groups differs in errors `r  apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 3)))$full_result`.
+
+```{r}
+learning_slopes_distance <- lmer(min_norm_distance ~ learning.condition*exp_block_id + (1 | id/exp_block_id), data = sub_walk_all[sub_walk_all$exp_block_id %in% 1:3,])
+
+tab_distance <- mixed_table_report(learning_slopes_distance)
+
+learning_slopes_errors <- lmer(errors ~ learning.condition*exp_block_id + (1 | id/exp_block_id), data = sub_walk_all[sub_walk_all$exp_block_id %in% 1:3,])
+
+tab_errors <- mixed_table_report(learning_slopes_distance)
+tab <- merge(tab_distance[,c(3,4)], tab_errors[,c(3,4)], by = "names")
+colnames(tab) <- c("Predictor", "Significance for distance improvement", "Significance for error improvement")
+tab$names <- c("Intercept", "Block number", "Learning condition", "Block - Learning condition interaction")
+kable(tab)
+#ggplot(sub_walk_all[sub_walk_all$exp_block_id %in% 1:3,], aes(exp_block_id, min_norm_distance, color=learning.condition)) + geom_point() + geom_smooth(method="lm", se=F)
+
+#ggplot(sub_walk_all[sub_walk_all$exp_block_id %in% 1:3,], aes(exp_block_id, errors, color=learning.condition)) + geom_point() + geom_smooth(method="lm", se=F)
+
+```
+
+### After modality switch performance change
+Anova doesn't show shows a significant difference between groups in the block 3 (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 3,]))$statistic$learning_condition`), but there is a significant difference between distance performance in the block 4 (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`).
+
+After the switch, we can see a significant difference between distance performance (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`) and errors as well (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`), with groups learning on the treadmill or desktop performing worse than those learning in the real world.
+
+Mixed effect models with individual random effect show significant effect of the of the learning condition on rate of improvement from phase 1 to pahase 2 (See table XXX )
+
+```{r}
+learning_slopes_distance <- lmer(min_norm_distance ~ learning.condition*exp_block_id + (1 | id/exp_block_id), data = only_blocks(sub_walk_all, 3:4))
+
+tab_distance <- mixed_table_report(learning_slopes_distance)
+
+learning_slopes_errors <- lmer(errors ~ learning.condition*exp_block_id + (1 | id/exp_block_id), data = only_blocks(sub_walk_all, 3:4))
+
+tab_errors <- mixed_table_report(learning_slopes_distance)
+tab <- merge(tab_distance[,c(3,4)], tab_errors[,c(3,4)], by = "names")
+colnames(tab) <- c("Predictor", "Significance for distance improvement", "Significance for error improvement")
+tab$names <- c("Intercept", "Block number", "Learning condition", "Block - Learning condition interaction")
+kable(tab)
+```
+
+Running separate pairwise t-tests to see individual performance change from block 3 to 4 gives the following results.
+```{r, echo=F}
+df_dist <- block_t_test(sub_walk_all, 3, 4,"min_norm_distance")[,c(1,4)]
+df_error <- block_t_test(sub_walk_all, 3, 4,"errors")[,c(1,4)]
+df34 <- merge(df_dist, df_error, by = "condition")
+colnames(df34) <- c("learning condition", "p-value distance improvement", "p-value errors improvement")
+kable(df34, caption="Paired T test comparing infividual performance improvement from block 3-4 in different conditions")
+```
+
+We can see that all participants improved in the errors made, but neigher group is significantly worse between block 3 and 4 in the path travelled (although participants learning on desktop perform slightly worse in 4th block, but the significance is only marginal)
+
+### Personal 
+To assess the level of performance change form directly before and after the switch, we calculated personal improvement as (block3-block4)/(block3+block4)´.
+
+```{r, echo=F}
+ggplot(df_distance, aes(x = learning.condition, y = block34.diff, fill = learning.condition)) + 
+  geom_bar(stat = 'identity', position="dodge") + 
+  geom_errorbar(aes(ymin=block34.diff-block34.se, ymax=block34.diff+block34.se), width=.2, position=position_dodge(.9)) +
+  scale_fill_manual(values = estimote_pal) +
+  ylab("Mean normalised distance improvement from block 3 to 4") + xlab("Learning condition")  + guides(fill=F)
+
+ggplot(df_errors, aes(x = learning.condition, y = block34.diff, fill = learning.condition)) + 
+  geom_bar(stat = 'identity', position="dodge") + 
+  geom_errorbar(aes(ymin=block34.diff-block34.se, ymax=block34.diff+block34.se), width=.2, position=position_dodge(.9)) +
+  scale_fill_manual(values = estimote_pal) +
+  ylab("Mean sum of errors improvement after switch")+ xlab("Learning condition") + guides(fill=F)
+```
+
+Comparing the block perfomance change in different conditions using anovas, we see marginally significant difference in the improvement between groups for distance (`r apa_print(aov(block34 ~ learning.condition, df_wide_distance))$statistic$learning_condition`), but we see significant differences in error rate improvement (`r apa_print(aov(block34 ~ learning.condition, df_wide_errors))$statistic$learning_condition`).
+This is consistent with the mixed models which poited at interaction between block and learning condition.
+
+Tukey post-hoc tests show significant difference between error rate improvement between the group that leardned in the real world and that which learned on the desktop.
+```{r, echo=F}
+kable(tukey_report(TukeyHSD(x=aov(block34 ~ learning.condition, df_wide_errors), conf.level=0.95)$learning.condition[, c(1,4)]))
+```

### 2018-10-21 Report/paper-story.Rmd
Updates reports, moves functions

@@
 source("../Scripts/paper-prepare.R")
+source("helpers/block-measure.R")
+source("helpers/block-t-test.R")
 #green, violet, blue, yellow, pink, dark blue
@@
 
-block_measure <- function(df, value){
-  df_wide <- dcast(df, learning.condition + id ~ exp_block_id, value.var = value)
-  colnames(df_wide) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")
-  df_wide %>% 
-    group_by(learning.condition) %>%
-    summarise(block1.mean  = mean(block1, na.rm = T),
-              block3.mean = mean(block3, na.rm = T),
-              block4.mean = mean(block4, na.rm = T),
-              block34.diff = mean((block3-block4)/(block4+block3), na.rm = T),
-              block34.se  = sqrt(var((block3-block4)/(block4+block3), na.rm = T)/sum(!is.na(block3-block4))),
-              block14.diff = mean((block1-block4)/(block1+block4), na.rm = T),
-              block14.se = sqrt(var((block1-block4)/(block1+block4), na.rm = T)/sum(!is.na(block1-block4)))
-  )
-}
 df_errors <- block_measure(condition_all_sums, "sum.errors")
@@
 df_error <- block_t_test(sub_walk_all, 1, 4,"errors")[,c(1,4)]
-df14 <- merge(df_dist, df_error, by = "learning.condition")
-colnames(df14) <- c("learning condition", "p-value distance improvement", "p-value errors improvement")
+df14 <- merge(df_dist, df_error, by = "condition")
+colnames(df14) <- c("condition", "p-value distance improvement", "p-value errors improvement")
 kable(df14, caption="Paired T test comparing individual performance improvement from block 1-4 in different conditions")
@@
 
+### Different rate of learning in blocks 1-3
+Running the mixed models to see if thre is a different linear learning rate in learning blocks (1-3), we can see that there is a significant interaction in learning condition and block for errors, but not fordistance.
+```{r}
+learning_slopes_distance <- lmer(min_norm_distance ~ learning.condition*exp_block_id + (1 | id/exp_block_id), data = sub_walk_all[sub_walk_all$exp_block_id %in% 1:3,])
+kable(mixed_table_report(learning_slopes_distance))
+
+learning_slopes_errors <- lmer(errors ~ learning.condition*exp_block_id + (1 | id/exp_block_id), data = sub_walk_all[sub_walk_all$exp_block_id %in% 1:3,])
+
+kable(mixed_table_report(learning_slopes_errors))
+
+ggplot(sub_walk_all[sub_walk_all$exp_block_id %in% 1:3,], aes(exp_block_id, min_norm_distance, color=learning.condition)) + 
+  geom_point() + geom_smooth(method="lm", se=F)
+
+ggplot(sub_walk_all[sub_walk_all$exp_block_id %in% 1:3,], aes(exp_block_id, errors, color=learning.condition)) + 
+  geom_point() + geom_smooth(method="lm", se=F)
+```
+
 ### Comparing walking performance in block 3 vs 4
@@
 
-And significant effect all both conditoin, block and their interaction on the number of errors.
+And significant effect all both condition, block and their interaction on the number of errors.
 ```{r, echo=F}
@@
 df_error <- block_t_test(sub_walk_all, 3, 4,"errors")[,c(1,4)]
-df34 <- merge(df_dist, df_error, by = "learning.condition")
-colnames(df34) <- c("learning condition", "p-value distance improvement", "p-value errors improvement")
+df34 <- merge(df_dist, df_error, by = "condition")
+colnames(df34) <- c("condition", "p-value distance improvement", "p-value errors improvement")
 kable(df34, caption="Paired T test comparing infividual performance improvement from block 3-4 in different conditions")

### 2018-10-21 Report/helpers/block-correlation.R

@@ -1,0 +1,9 @@
+block_correlation <- function(df, block1, block2, value, formula=learning.condition + id ~ exp_block_id, columns=c("learning.condition","id")){
+  df <- df %>% filter(exp_block_id %in% c(block1,block2))
+  mean.na <- function(data){return(mean(data, na.rm = T))}
+  df_wide <- dcast(df, formula, value.var = value, fun.aggregate = mean.na)
+  colnames(df_wide) <- c(columns, "block1", "block2")
+  df_out <- df_wide %>% group_by(learning.condition) %>% do(tidy(cor(.$block1, .$block2, use = "complete.obs")))
+  df_out$x <- round(df_out$x, 2)
+  return(df_out)
+}

### 2018-10-21 Report/helpers/block-measure.R

@@ -2,7 +2,7 @@ block_measure <- function(df, value){
   df_wide <- dcast(df, learning.condition + id ~ exp_block_id, value.var = value)
   colnames(df_wide) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")
   
-  df_wide %>% 
+  df_wide <- df_wide %>% 
     group_by(learning.condition) %>%
     summarise(block1.mean  = mean(block1, na.rm = T),
               block3.mean = mean(block3, na.rm = T),
@@ -12,4 +12,5 @@ block_measure <- function(df, value){
               block14.diff = mean((block1-block4)/(block1+block4), na.rm = T),
               block14.se = sqrt(var((block1-block4)/(block1+block4), na.rm = T)/sum(!is.na(block1-block4)))
     )
+  return(df_wide)
 }

### 2018-10-21 TwoWorlds/helpers-report.R

@@ -7,25 +7,6 @@ apa_p <- function(values){
   }
   return(output)
 }
-block_t_test <- function(df, block1, block2, value, paired = T){
-  df <- df %>% filter(exp_block_id %in% c(block1,block2))
-  mean.na <- function(data){return(mean(data, na.rm = T))}
-  df_wide <- dcast(df, learning.condition + id ~ exp_block_id, value.var = value, fun.aggregate = mean.na)
-  colnames(df_wide) <- c("learning.condition","id", "block1","block2")
-  df_out <- df_wide %>% group_by(learning.condition) %>% do(tidy(t.test(.$block1, .$block2, paired = paired)))
-  df_out$p.value <- apa_p(df_out$p.value)
-  return(df_out)
-}
-
-block_correlation <- function(df, block1, block2, value){
-  df <- df %>% filter(exp_block_id %in% c(block1,block2))
-  mean.na <- function(data){return(mean(data, na.rm = T))}
-  df_wide <- dcast(df, learning.condition + id ~ exp_block_id, value.var = value, fun.aggregate = mean.na)
-  colnames(df_wide) <- c("learning.condition","id", "block1","block2")
-  df_out <- df_wide %>% group_by(learning.condition) %>% do(tidy(cor(.$block1, .$block2)))
-  df_out$x <- round(df_out$x, 2)
-  return(df_out)
-}
 
 tukey_report <- function(tukey){
   df <- as.data.frame(tukey)
@@ -38,5 +19,11 @@ tukey_report <- function(tukey){
 mixed_table_report <- function(fit_lme){
   aov_result <- as.data.frame(car::Anova(fit_lme, type = "III"))
   aov_result[,3] <- apa_p(aov_result[,3])
+  aov_result$names <- rownames(aov_result)
+  rownames(aov_result) <- NULL
   return(aov_result)
 } 
+
+only_blocks <- function(df, ...){
+  return(df[df$exp_block_id %in% c(...),])
+}

### 2018-10-21 Report/paper-results.Rmd
New version of the results

@@
 If we would assume no transfer to happen, we should observe same performance in the real world testing in 1st, 2nd and 3rd block after the switch to be similar in performance as the 1st, 2nd and 3rd block in the real world learning group.
-But all conditions (except desktop in distance) show better performance after the switch (see table XXX) than in the first trial, suggesting some level of transfer from all learning conditions.
-This is not the case with distance improvement in the desktop learning group.
-
+But all conditions (except desktop in distance) show better performance after the switch (see table XXX) than in the first trial, suggesting some level of transfer happening in all learning conditions.
 ```{r}
@@
 colnames(df) <- c("Learning condition", "Distance improvement", "Errors improvement")
+df$`Learning condition` <- dplyr::recode(df$`Learning condition`, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
 kable(df, caption="Paired T test comparing individual performance improvement from block 1-4 in different conditions")
@@
 
-Looking at the progression in distance improvements and error improvements, we can see that all conditions show improvement over the experiment progression which is only hindered by the switch in the desktop learning condition (see fig.
+Looking at the progression in distance improvements and error improvements over the course of the experiment, we can see that all conditions show progressive improvement which is only hindered by the switch in the desktop learning condition (see fig.
 XXX and fig.
-XXX), but otherwise we see constant improvement unobstructed by the environemnt switch.
+XXX), but in the Treadmill condition we see constant improvement unobstructed by the environmrnt switch.
 ```{r}
@@
 
-Participant's performance in the first pre-switch phase shows different rates of learning rate for errors for our learning conditions (see table XX).
+Participant's performance in the first pre-switch phase shows different rates of learning rate (see table XX).
 There is no significant difference between different modalities in distance performance before the switch `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 3)))$full_result`, but the performance across groups differs in errors `r  apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 3)))$full_result`.
@@
 learning_slopes_distance <- lmer(min_norm_distance ~ learning.condition*exp_block_id + (1 | id/exp_block_id), data = sub_walk_all[sub_walk_all$exp_block_id %in% 1:3,])
-
 tab_distance <- mixed_table_report(learning_slopes_distance)
@@
 learning_slopes_errors <- lmer(errors ~ learning.condition*exp_block_id + (1 | id/exp_block_id), data = sub_walk_all[sub_walk_all$exp_block_id %in% 1:3,])
+tab_errors <- mixed_table_report(learning_slopes_errors)
 
-tab_errors <- mixed_table_report(learning_slopes_distance)
 tab <- merge(tab_distance[,c(3,4)], tab_errors[,c(3,4)], by = "names")
+tab$names <- c("Intercept", "Block number", "Learning condition", "Block - Learning condition interaction")
 colnames(tab) <- c("Predictor", "Significance for distance improvement", "Significance for error improvement")
-tab$names <- c("Intercept", "Block number", "Learning condition", "Block - Learning condition interaction")
 kable(tab)
 #ggplot(sub_walk_all[sub_walk_all$exp_block_id %in% 1:3,], aes(exp_block_id, min_norm_distance, color=learning.condition)) + geom_point() + geom_smooth(method="lm", se=F)
-
 #ggplot(sub_walk_all[sub_walk_all$exp_block_id %in% 1:3,], aes(exp_block_id, errors, color=learning.condition)) + geom_point() + geom_smooth(method="lm", se=F)
@@
 ### After modality switch performance change
-Anova doesn't show shows a significant difference between groups in the block 3 (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 3,]))$statistic$learning_condition`), but there is a significant difference between distance performance in the block 4 (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`).
+Anova doesn't show a significant difference between groups in the pre-switch block (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 3,]))$statistic$learning_condition`), but there is a significant difference between distance performance in the post-switch block (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`).
 
-After the switch, we can see a significant difference between distance performance (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`) and errors as well (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`), with groups learning on the treadmill or desktop performing worse than those learning in the real world.
+After the modality switch, we can see a significant difference between distance performance (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`) and errors as well (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`), with groups learning on the treadmill or desktop performing worse than those learning in the real world.
+This difference among conditions post-switch is still apparent in the 2nd block after the switch in number of errors `r  apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 5)))$full_result` and marginally in distance `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 5)))$full_result`, but disappears in the last testing block for both errors `r apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 6)))$full_result` and distance `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 6)))$full_result`.
 
-Mixed effect models with individual random effect show significant effect of the of the learning condition on rate of improvement from phase 1 to pahase 2 (See table XXX )
+Mixed effect models with individual random effect also show significant effect of the of the learning condition on rate of improvement from pre-switch to post-switch block (See table XXX ).
 
@@
 learning_slopes_distance <- lmer(min_norm_distance ~ learning.condition*exp_block_id + (1 | id/exp_block_id), data = only_blocks(sub_walk_all, 3:4))
-
 tab_distance <- mixed_table_report(learning_slopes_distance)
@@
 learning_slopes_errors <- lmer(errors ~ learning.condition*exp_block_id + (1 | id/exp_block_id), data = only_blocks(sub_walk_all, 3:4))
+tab_errors <- mixed_table_report(learning_slopes_errors)
 
-tab_errors <- mixed_table_report(learning_slopes_distance)
 tab <- merge(tab_distance[,c(3,4)], tab_errors[,c(3,4)], by = "names")
+tab$names <- c("Intercept", "Block number", "Learning condition", "Block - Learning condition interaction")
 colnames(tab) <- c("Predictor", "Significance for distance improvement", "Significance for error improvement")
-tab$names <- c("Intercept", "Block number", "Learning condition", "Block - Learning condition interaction")
 kable(tab)
@@
 
-Running separate pairwise t-tests to see individual performance change from block 3 to 4 gives the following results.
+Running separate pairwise t-tests to see individual performance change from pre-switch to post-switch block, we see significant improvement in errors made across all conditions, and marginally significant worsening in distance performance in the desktop condition.
 ```{r, echo=F}
-df_dist <- block_t_test(sub_walk_all, 3, 4,"min_norm_distance")[,c(1,4)]
-df_error <- block_t_test(sub_walk_all, 3, 4,"errors")[,c(1,4)]
-df34 <- merge(df_dist, df_error, by = "condition")
-colnames(df34) <- c("learning condition", "p-value distance improvement", "p-value errors improvement")
-kable(df34, caption="Paired T test comparing infividual performance improvement from block 3-4 in different conditions")
+df <- data.frame(condition=rep("",3),dist=rep("",3), err=rep("",3), stringsAsFactors = F)
+df_dist <- block_t_test_improvement(sub_walk_all, 3, 4,"min_norm_distance")
+df_error <- block_t_test_improvement(sub_walk_all, 3, 4,"errors")
+for(i in 1:length(names(df_dist))){
+  name <- names(df_dist)[i]
+  df[i,"condition"] <- name
+  df[i, "dist"] <- apa_print(df_dist[[name]])$full_result
+  df[i, "err"] <- apa_print(df_error[[name]])$full_result
+}
+colnames(df) <- c("Learning condition", "Distance improvement", "Errors improvement")
+df$`Learning condition` <- dplyr::recode(df$`Learning condition`, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
+kable(df, caption="Paired T test comparing individual performance improvement from pre switch block to post-switch block in different conditions")
 ```
 
-We can see that all participants improved in the errors made, but neigher group is significantly worse between block 3 and 4 in the path travelled (although participants learning on desktop perform slightly worse in 4th block, but the significance is only marginal)
+We can see that all participants improved in the errors made, but neigher group is significantly worse in the post-switch block in the path travelled (although participants who learned on the desktop performed slightly worse, but the significance is only marginal).
 
 ### Personal 
-To assess the level of performance change form directly before and after the switch, we calculated personal improvement as (block3-block4)/(block3+block4)´.
+To assess the level of performance change form directly before and after the switch, we calculated a score personal improvement as (block3-block4)/(block3+block4)´.
+This allowed us to directly comapre the percentual improvement or deterioration from pre-switch block to the post-switch one.
 
@@
 
-Comparing the block perfomance change in different conditions using anovas, we see marginally significant difference in the improvement between groups for distance (`r apa_print(aov(block34 ~ learning.condition, df_wide_distance))$statistic$learning_condition`), but we see significant differences in error rate improvement (`r apa_print(aov(block34 ~ learning.condition, df_wide_errors))$statistic$learning_condition`).
-This is consistent with the mixed models which poited at interaction between block and learning condition.
+Comparing the perfomance change in different conditions using anovas, we see marginally significant difference between groups in the distance improvement (`r apa_print(aov(block34 ~ learning.condition, df_wide_distance))$statistic$learning_condition`), but significant differences in error rate improvement (`r apa_print(aov(block34 ~ learning.condition, df_wide_errors))$statistic$learning_condition`).
+This is consistent with the mixed model result which poited at interaction between block and learning condition.
 
@@
 ```
+
+### Something about pointing?
+
+## Summary
+We can see different rates of learning speed across different conditions, with significan differences between groups in their performance after the first phase witt the real world learning group performing the best, but only significanly different from group learning on the desktop.
+All conditions show some level of transfer, demonstrated both by general improvement from the first block to the fourth, as well as decreased errors in desktop and treadmill VR group after the switch to the real environment.

### 2018-10-21 Report/paper-results.Rmd
Changed the graphs to line

@@
 ```{r}
-ggplot(sub_walk_all, aes(learning.condition, min_norm_distance, fill=factor(exp_block_id))) +
-  stat_summary(fun.data=mean_cl_normal, position=position_dodge(0.95), geom="errorbar") + 
-  stat_summary(fun.y=mean, position=position_dodge(width=0.95), geom="bar") +
-  ylab("Average normalised trial distance") + xlab("Learning condition") + 
-  scale_fill_manual(values=estimote_pal,"Block ID") + guides(fill=F)
+ggplot(sub_walk_all, aes(x = exp_block_id, y = min_norm_distance, color = factor(learning.condition))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.5, geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_identity(), size = 1.25, geom="line") + 
+  geom_vline(xintercept = 3.5, linetype=4) +
+  ylab("Average normalised trial distance") + xlab("Block") + 
+  scale_color_manual(values=estimote_pal,"Learning condition") + guides(fill=F)
 
-ggplot(sub_walk_all, aes(learning.condition, errors, fill=factor(exp_block_id))) +
-  stat_summary(fun.data=mean_cl_normal, position=position_dodge(0.95), geom="errorbar") + 
-  stat_summary(fun.y=mean, position=position_dodge(width=0.95), geom="bar") +
-  ylab("Average number of errors") + xlab("Learning condition") + 
-  scale_fill_manual(values=estimote_pal,"Block ID") + guides(fill=F)
+ggplot(sub_walk_all, aes(x = exp_block_id, y = errors, color = factor(learning.condition))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.5, geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_identity(), size = 1.25, geom="line") + 
+  geom_vline(xintercept = 3.5, linetype=4) +
+  ylab("Number of mistakes in office decision") + xlab("Block number") + 
+  scale_color_manual(values=estimote_pal, "Learning condition")
 ```

### 2018-10-22 Report/paper-all-data.Rmd
Adds new charts, new build and longer results

@@
   scale_color_manual(values=estimote_pal, "Condition")
+
+ggplot(walk_all, aes(x = exp_block_id, y = min_norm_distance, color = factor(condition))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
+  geom_vline(xintercept = 3.5, linetype=4) +
+  ylab("Average normalised trial distance") + xlab("Block") + 
+  scale_color_manual(values=estimote_pal,"Condition") + guides(fill=F)
+
+ggplot(walk_all, aes(x = exp_block_id, y = errors, color = factor(condition))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
+  geom_vline(xintercept = 3.5, linetype=4) +
+  ylab("Average normalised trial distance") + xlab("Block") + 
+  scale_color_manual(values=estimote_pal,"Condition") + guides(fill=F)
 

### 2018-10-22 Report/paper-results.Rmd
Adds new charts, new build and longer results

@@
 
-Participant's performance in the first pre-switch phase shows different rates of learning rate (see table XX).
-There is no significant difference between different modalities in distance performance before the switch `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 3)))$full_result`, but the performance across groups differs in errors `r  apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 3)))$full_result`.
+Participant's performance in the first pre-switch phase shows different rates of learning rate, demonstrated by an interaction between block bumber and learning condition for error rate performance (see table XX).
 
@@
 ### After modality switch performance change
-Anova doesn't show a significant difference between groups in the pre-switch block (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 3,]))$statistic$learning_condition`), but there is a significant difference between distance performance in the post-switch block (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`).
+We found no significant difference between different modalities in distance performance before the switch (although are marginally significnat suggetsting there might be some difference) `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 3)))$statistic`, but the error rate in the pre-switch block differs across groups `r  apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 3)))$statistic`.
 
 After the modality switch, we can see a significant difference between distance performance (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`) and errors as well (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`), with groups learning on the treadmill or desktop performing worse than those learning in the real world.
-This difference among conditions post-switch is still apparent in the 2nd block after the switch in number of errors `r  apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 5)))$full_result` and marginally in distance `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 5)))$full_result`, but disappears in the last testing block for both errors `r apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 6)))$full_result` and distance `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 6)))$full_result`.
+This difference post-switch is still apparent in the 2nd block after the switch (block 5) in error rate `r  apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 5)))$statistic` and marginally in distance `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 5)))$statistic`.
+In the last testing block (block 6), we found no differences among the groups in either error rate `r apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 6)))$statistic` nor distance `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 6)))$statistic`.
 
-Mixed effect models with individual random effect also show significant effect of the of the learning condition on rate of improvement from pre-switch to post-switch block (See table XXX ).
+Mixed effect models with individual random effect show significant effect of the of the learning condition on rate of improvement from pre-switch to post-switch block (See table XXX ).
 
@@
 
-We can see that all participants improved in the errors made, but neigher group is significantly worse in the post-switch block in the path travelled (although participants who learned on the desktop performed slightly worse, but the significance is only marginal).
+We can see that all participants improved in the errors made, but neigher group is significantly worse in the post-switch block in the path travelled (although participants who learned on the desktop performed slightly worse, the significance is only marginal).
 
-### Personal 
-To assess the level of performance change form directly before and after the switch, we calculated a score personal improvement as (block3-block4)/(block3+block4)´.
+### Performance change 
+To assess the level of performance change form directly before and after the switch, we calculated a personal improvement score as (block3-block4)/(block3+block4)´.
 This allowed us to directly comapre the percentual improvement or deterioration from pre-switch block to the post-switch one.
@@
 
-Comparing the perfomance change in different conditions using anovas, we see marginally significant difference between groups in the distance improvement (`r apa_print(aov(block34 ~ learning.condition, df_wide_distance))$statistic$learning_condition`), but significant differences in error rate improvement (`r apa_print(aov(block34 ~ learning.condition, df_wide_errors))$statistic$learning_condition`).
+Comparing the perfomance change in different conditions using ANOVA, we see marginally significant difference between groups in the distance improvement (`r apa_print(aov(block34 ~ learning.condition, df_wide_distance))$statistic$learning_condition`), but significant differences in error rate improvement (`r apa_print(aov(block34 ~ learning.condition, df_wide_errors))$statistic$learning_condition`).
 This is consistent with the mixed model result which poited at interaction between block and learning condition.
 
-Tukey post-hoc tests show significant difference between error rate improvement between the group that leardned in the real world and that which learned on the desktop.
+Tukey post-hoc tests show significant difference between error rate improvement between the group that learned in the real world and that which learned on the desktop.
 ```{r, echo=F}
@@
 
+### Improvement from the pre-switch block to the final block
+Assuming the performance platoes around the 6the block, looking at participants performance in the pre-switch block and the last block we can deduce what margin of improvement is still possible.
+We see that participants learning in the real world no longer improve in error rate `r apa_print(block_t_test_improvement(sub_walk_all, 4, 6, "errors")[["real-real"]])$statistic`, but they are continuously imrpoving in distance performance `r apa_print(block_t_test_improvement(sub_walk_all, 4, 6, "min_norm_distance")[["real-real"]])$statistic`.
+
 ### Something about pointing?
+Overall, we found
 
 ## Summary
-We can see different rates of learning speed across different conditions, with significan differences between groups in their performance after the first phase witt the real world learning group performing the best, but only significanly different from group learning on the desktop.
+We can see different rates of learning speed across different conditions, with significan differences between groups in their performance after the first phase with the real world learning group performing the best, but only significanly different from group learning on the desktop.
 All conditions show some level of transfer, demonstrated both by general improvement from the first block to the fourth, as well as decreased errors in desktop and treadmill VR group after the switch to the real environment.
+All groups get to the same level of performance after the three blocks post switch in the real world, not differing in errors made nor distance.
+But we also observed that whilst real world training leads to almost perfect performance in the 3rd block, not differing from the last block in error rate and only slightly in distance, it takes another three blocks of learning in the real world for the groups which learned on the treadmill or the desktop to achieve same performance.
+
+*I am unsure if it is worth adding the rest of the dataset in (the VR - VR and real - VR groups), but the performance of Treadmill VR groups in the distance traveled gets exactly to the same level as the real world performance in the 6th block (so the VR - VR group performs the same after the 6 blocks as VR - real and real - VR).
+Which is nice comparison to the results we get here about the distance performance in the real world being almost perfect in the 3rd block and the rest of the groups slowly "getting there".*
+
+*The VR groups still demonstrate larger error difference in the 4-6th block, making more errors overall than the real world groups (about 2 errors per trial more).
+I believe it can be explained by either social difference - participants not wanting to "ask as much" in the real world, or accidental walking into a door due to treadmill control issues.
+Also, particiapnts told me that the VR is more "inviting" to test all the doors that are being passed (just in case), because it is more tiring to get back to one in case of a mistake.
+Anyway, not sure it it warranties adding more groups to the study, but at least the data are consistent and distance performance gets to the same level even in the VR*
+
+We found difference

### 2018-10-22 Report/paper-results.Rmd
Pre bed commit
updates results

@@
 
+### Try running it without outliers
+sub_walk_all[!is.na(sub_walk_all$norm_distance) & sub_walk_all$norm_distance > 3, c("distance", "min_norm_distance")] <- NA
+
 condition_all_sums <- sub_walk_all %>% 
@@
 There was no difference between conditions at the start of the experiment in normalised distance (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`), nor in number errors (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`), but there is a significant difference in normalised time (`r apa_print(aov(min_norm_time ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`)), due to VR taking longer times to get used to (and therefore having longer normalised trial times at the start).
+We removed distance measurements larger than 3 standard deviations from the mean.
 
@@
 If we would assume no transfer to happen, we should observe same performance in the real world testing in 1st, 2nd and 3rd block after the switch to be similar in performance as the 1st, 2nd and 3rd block in the real world learning group.
-But all conditions (except desktop in distance) show better performance after the switch (see table XXX) than in the first trial, suggesting some level of transfer happening in all learning conditions.
+But all conditions show better performance after the switch (see table XXX) than in the first trial, suggesting some level of transfer happening in all learning conditions.
 ```{r}
@@
 XXX and fig.
-XXX), but in the Treadmill condition we see constant improvement unobstructed by the environmrnt switch.
+XXX), but in the Treadmill condition we see constant improvement unobstructed by the environment switch.
 ```{r}
@@
 
+**this is a mixed model with id/block id random term.
+Significance was measuyred with Anova from car package, but I read it is not a good way to measure significance and we shouldn't even use significance in mixed models.
+I'll check on past cognition papers to see how and if they do it.**
+
 ```{r}
@@
 ### After modality switch performance change
-We found no significant difference between different modalities in distance performance before the switch (although are marginally significnat suggetsting there might be some difference) `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 3)))$statistic`, but the error rate in the pre-switch block differs across groups `r  apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 3)))$statistic`.
+We found significant difference between different modalities  in the pre-switch block in distance performance `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 3)))$statistic`, and the error rate as well `r  apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 3)))$statistic`.
 
-After the modality switch, we can see a significant difference between distance performance (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`) and errors as well (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`), with groups learning on the treadmill or desktop performing worse than those learning in the real world.
-This difference post-switch is still apparent in the 2nd block after the switch (block 5) in error rate `r  apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 5)))$statistic` and marginally in distance `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 5)))$statistic`.
+After the modality switch, we can see a significant difference between groups in distance performance (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`) and errors as well (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`), with groups learning on the treadmill or desktop performing worse than those learning in the real world.
+This between group difference is still present in the 2nd block of the second phase (block 5) in error rate `r  apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 5)))$statistic` and marginally in distance `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 5)))$statistic`.
 In the last testing block (block 6), we found no differences among the groups in either error rate `r apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 6)))$statistic` nor distance `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 6)))$statistic`.
@@
 Assuming the performance platoes around the 6the block, looking at participants performance in the pre-switch block and the last block we can deduce what margin of improvement is still possible.
-We see that participants learning in the real world no longer improve in error rate `r apa_print(block_t_test_improvement(sub_walk_all, 4, 6, "errors")[["real-real"]])$statistic`, but they are continuously imrpoving in distance performance `r apa_print(block_t_test_improvement(sub_walk_all, 4, 6, "min_norm_distance")[["real-real"]])$statistic`.
+We see that participants learning in the real world no longer improve in error rate after the 3rd block `r apa_print(block_t_test_improvement(sub_walk_all, 4, 6, "errors")[["real-real"]])$statistic`, but they are continuously improving in distance performance `r apa_print(block_t_test_improvement(sub_walk_all, 4, 6, "min_norm_distance")[["real-real"]])$statistic`.
+
+*Not sure about the mixed models here. I think they are OK, but I haven't published anything with them before*
+
+Looking at the linear mixed model prediction for the distnace and error improvement in the second phase (post-switch), we can see that both learning condition and
+
+```{r}
+learning_slopes_distance <- lmer(min_norm_distance ~ learning.condition*exp_block_id + (1 | id/exp_block_id), data = only_blocks(sub_walk_all[sub_walk_all$min_norm_distance < 10, ], 4:6))
+tab_distance <- mixed_table_report(learning_slopes_distance)
+
+learning_slopes_errors <- lmer(errors ~ learning.condition*exp_block_id + (1 | id/exp_block_id), data = only_blocks(sub_walk_all[sub_walk_all$min_norm_distance < 10, ], 4:6))
+tab_errors <- mixed_table_report(learning_slopes_errors)
+
+tab <- merge(tab_distance[,c(3,4)], tab_errors[,c(3,4)], by = "names")
+tab$names <- c("Intercept", "Block number", "Learning condition", "Block - Learning condition interaction")
+colnames(tab) <- c("Predictor", "Significance for distance improvement", "Significance for error improvement")
+kable(tab)
+```
 
 ### Something about pointing?
-Overall, we found
 
 ## Summary
-We can see different rates of learning speed across different conditions, with significan differences between groups in their performance after the first phase with the real world learning group performing the best, but only significanly different from group learning on the desktop.
+We can see different rates of learning speed across different conditions, with significant differences between groups in their performance in the post-switch block.
+Participants learning in the real world are performing the best after the switch, but only significanly different from the group learning on the desktop.
 All conditions show some level of transfer, demonstrated both by general improvement from the first block to the fourth, as well as decreased errors in desktop and treadmill VR group after the switch to the real environment.
-All groups get to the same level of performance after the three blocks post switch in the real world, not differing in errors made nor distance.
-But we also observed that whilst real world training leads to almost perfect performance in the 3rd block, not differing from the last block in error rate and only slightly in distance, it takes another three blocks of learning in the real world for the groups which learned on the treadmill or the desktop to achieve same performance.
+All groups get to the same level of performance after the three blocks post-switch moving in the real world, not differing in errors made nor in distance in the last block.
+But we also observed that whilst real world training leads to almost perfect performance in the pre-switch block, not differing from the last block in error rate and only slightly in distance.
+It takes another 3 blocks in the real world for the groups which learned on the treadmill or the desktop to achieve same performance, again pointing to qualitative difference between different modalities.
 
@@
 
-We found difference
+We found difference in relative improvement rate

### 2018-10-28 Report/paper-results.Rmd
changes how tukey reporting works

@@
 
+**Running the tukey on these???**
+There was also a difference between conditions in their pointing performance at the end of the first phase (`r apa_print(aov(abs_error ~ learning.condition, data = sub_sop_all))$statistic$learning_condition`), with both Treadmill VR group and the Desktop group performend significantly worse than the group in the real world.
+```{r}
+TukeyHSD(aov(abs_error ~ learning.condition, data = sub_sop_all))
+```
+
 After the modality switch, we can see a significant difference between groups in distance performance (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`) and errors as well (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`), with groups learning on the treadmill or desktop performing worse than those learning in the real world.
@@
 
-Mixed effect models with individual random effect show significant effect of the of the learning condition on rate of improvement from pre-switch to post-switch block (See table XXX ).
+Mixed effect models with individual random effect show significant effect of the of the learning condition on rate of improvement from pre-switch to post-switch block (See table XXX).
 
@@
 ```{r, echo=F}
-kable(tukey_report(TukeyHSD(x=aov(block34 ~ learning.condition, df_wide_errors), conf.level=0.95)$learning.condition[, c(1,4)]))
+kable(tukey_report_table(TukeyHSD(x=aov(block34 ~ learning.condition, df_wide_errors), conf.level=0.95), "learning.condition"))
 ```
@@
 ### Something about pointing?
+```{r}
+err_phase <- lmer(abs_error ~ phase*learning.condition + (1 | id/phase), data = sub_sop_all)
+car::Anova(err_phase, type = "III")
+
+ggplot(sub_sop_all, aes(phase, abs_error, color = learning.condition)) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.5, geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_identity(), size = 1.25, geom="line")
+
+aov(abs_error ~ learning.condition, data = sub_sop_all)
+```
 

### 2018-10-28 Report/paper-story.Rmd
changes how tukey reporting works

@@
 ```{r, echo=F}
-df_temp <- TukeyHSD(x=aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$phase == 2,]), conf.level=0.95)$learning.condition[, c(1,4)]
-kable(tukey_report(df_temp))
+df_temp <- TukeyHSD(x=aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$phase == 2,]), conf.level=0.95)
+kable(tukey_report_table(df_temp,"learning.condition"))
 ```
@@
 ```{r, echo=F}
-kable(tukey_report(TukeyHSD(x=aov(block34 ~ learning.condition, df_wide_errors), conf.level=0.95)$learning.condition[, c(1,4)]))
+kable(tukey_report_table(TukeyHSD(x=aov(block34 ~ learning.condition, df_wide_errors), conf.level=0.95),"learning.condition"))
 ```
@@
 ```{r, echo=F}
-kable(tukey_report(TukeyHSD(x=aov(block14 ~ learning.condition, df_wide_errors), conf.level=0.95)$learning.condition[, c(1,4)]))
+kable(tukey_report_table(TukeyHSD(x=aov(block14 ~ learning.condition, df_wide_errors), conf.level=0.95),"learning.condition"))
 ```

### 2018-10-28 TwoWorlds/helpers-report.R

@@ -8,11 +8,23 @@ apa_p <- function(values){
   return(output)
 }
 
-tukey_report <- function(tukey){
-  df <- as.data.frame(tukey)
-  colnames(df) <- c("diff", "p-value")
+tukey_report_table <- function(tukey, predictor=NULL){
+  if(is.null(predictor)){
+    df <-  tukey[[1]]
+  } else {
+    df <- tukey[[predictor]]
+  }
+  df <- as.data.frame(df)
+  colnames(df) <- c("diff", "lwr", "upr", "p-value")
   df$`p-value` <- apa_p(df$`p-value`)
-  return(df)
+  return(df[, c(1,4)])
+}
+
+tukey_report <- function(tukey, condition,delta_name = "difference", predictor = NULL){
+  df <- tukey_report_table(tukey, predictor)
+  line <- df[condition,]
+  report <- paste0('$\\Delta$', delta_name,' = ', round(line$diff, 2), ', ', line$`p-value`)
+  return(report)
 }
 
 # NEeds fitted lme model

### 2018-10-28 Report/paper-results.Rmd
updated results

@@
 **Running the tukey on these???**
-There was also a difference between conditions in their pointing performance at the end of the first phase (`r apa_print(aov(abs_error ~ learning.condition, data = sub_sop_all))$statistic$learning_condition`), with both Treadmill VR group and the Desktop group performend significantly worse than the group in the real world.
 ```{r}
-TukeyHSD(aov(abs_error ~ learning.condition, data = sub_sop_all))
+aov_point_phase1 <- aov(abs_error ~ learning.condition, data = sub_sop_all[sub_sop_all$phase == 1,])
 ```
+There was also a difference between conditions in their pointing performance at the end of the first phase (`r apa_print(aov_point_phase1)$statistic$learning_condition`), with real group performing significantly better than the Treadmill VR group `r tukey_report(TukeyHSD(aov_point_phase1), condition = "Treadmill VR-Real", delta_name = "error")` and the Desktop group `r tukey_report(TukeyHSD(aov_point_phase1), condition = "Real-Desktop", delta_name = "error")`.
 
@@
 This between group difference is still present in the 2nd block of the second phase (block 5) in error rate `r  apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 5)))$statistic` and marginally in distance `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 5)))$statistic`.
-In the last testing block (block 6), we found no differences among the groups in either error rate `r apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 6)))$statistic` nor distance `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 6)))$statistic`.
+In the last testing block (block 6), we found no differences among the groups in either error rate `r apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 6)))$statistic`, nor pointing performance `r apa_print(aov(abs_error ~ learning.condition, data = sub_sop_all[sub_sop_all$phase == 2,]))$statistic$learning_condition`) but we still see small difference in distance performance `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 6)))$statistic`.
 
@@
 ### Something about pointing?
-```{r}
+```{r, echo=F, results='hide'}
 err_phase <- lmer(abs_error ~ phase*learning.condition + (1 | id/phase), data = sub_sop_all)

### 2018-10-29 Report/paper-results.Rmd
Another update

--- 2018-10-28 Report/paper-results.Rmd
@@
 ## Measures
-We calculated participants’ walked distance in the building, time spent in each task and number of incorrectly visited doors.
-The conditions were not perfectly comparable in their measured distance and time, with the VR trials taking a little bit longer in both time and distance than the real world condition.
-We therefore min-normalised the distance traveled and the time for each task and environment  by dividing the measure by any participant's performance for that trial.
+We calculated participants’ walked distance in the building, time spent in each task, number of incorrectly visited doors during walking trials and pointing precision during pointing trials.
+The conditions were not perfectly comparable in measured distance and time, with the VR trials taking a little bit longer in both time and distance than the real world condition.
+We therefore min-normalised the distance traveled and the time for each task and environment by dividing the measure by any participant's best performance for that trial (therefore the result of 2 in min-normalised distance would mean the participant traveled double the distance than was necessary).
 We calculated all statistics on these normalised measures.
 
-There was no difference between conditions at the start of the experiment in normalised distance (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`), nor in number errors (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`), but there is a significant difference in normalised time (`r apa_print(aov(min_norm_time ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`)), due to VR taking longer times to get used to (and therefore having longer normalised trial times at the start).
-We removed distance measurements larger than 3 standard deviations from the mean.
+There was no difference between conditions at the start of the experiment in normalised distance (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`), nor in number errors (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`), but we observed a significant difference in normalised time (`r apa_print(aov(min_norm_time ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`), due to the treadmill VR participants taking longer times to get used to the device and therefore having longer trial times at the start.
+We removed distance or time measurements larger than 3 standard deviations from the mean (a total of `r round((nrow(sub_walk_all[!is.na(sub_walk_all$norm_distance) & sub_walk_all$norm_distance > 3, ])/nrow(sub_walk_all))* 100, 1)` percent of trials).
 
@@
 ```
-A total of `r sum(overview$is_ok == "yes", na.rm=T)` (M = `r round(mean(df_participants[df_participants$finished == "yes",]$age, na.rm=T),1)`, SD = `r round(sd(df_participants[df_participants$finished == "yes",]$age, na.rm=T), 1)`) undergraduate students at UC Davis particiapted in the study in exchange for a study credit.
-Each students was randomly assigned a condition and a randomised set of goals.
+A total of `r sum(overview$is_ok == "yes", na.rm=T) + sum(finished$no)` undergraduate students at UC Davis (M = `r round(mean(df_participants[df_participants$finished == "yes",]$age, na.rm=T),1)`, SD = `r round(sd(df_participants[df_participants$finished == "yes",]$age, na.rm=T), 1)`) particiapted in the study in exchange for a study credit.
 `r sum(finished$no)` participants didn't finish due to motion sickness and `r sum(is_ok$no)` were removed due to a technical failure of the real world tracking systen.
-Only `r finished[finished$condition == "vr-real","percent"] * 100` percent the students being able to finish VR to Real world condition.
+Only `r finished[finished$condition == "vr-real","percent"] * 100` percent the students being able to finish Treadmill VR learning condition.
+Each students was randomly assigned a condition and a randomised set of goals before arrival.
 
@@
 
-First we wanted to assess that there is going to be some level of transfer in all conditions.
-If we would assume no transfer to happen, we should observe same performance in the real world testing in 1st, 2nd and 3rd block after the switch to be similar in performance as the 1st, 2nd and 3rd block in the real world learning group.
-But all conditions show better performance after the switch (see table XXX) than in the first trial, suggesting some level of transfer happening in all learning conditions.
+First we wanted to assess that there was some level of transfer in all conditions.
+All conditions demonstrate better performance in the post-switch block than in the first block (see table XXX) , suggesting some level of transfer happening in all learning conditions.
 ```{r}
@@
 
-Looking at the progression in distance improvements and error improvements over the course of the experiment, we can see that all conditions show progressive improvement which is only hindered by the switch in the desktop learning condition (see fig.
+Looking at the progression of distance and error rate performance over the course of the experiment, we can see that all conditions show incremental improvement which is only hindered by the switch in the desktop learning condition (see fig.
 XXX and fig.
-XXX), but in the Treadmill condition we see constant improvement unobstructed by the environment switch.
+XXX), but in the Treadmill VR condition we see constant improvement unobstructed by the environment switch.
 ```{r}
@@
 **this is a mixed model with id/block id random term.
-Significance was measuyred with Anova from car package, but I read it is not a good way to measure significance and we shouldn't even use significance in mixed models.
+Significance was measured with Anova from car package, but I read it is not a good way to measure significance and we shouldn't even use significance in mixed models.
 I'll check on past cognition papers to see how and if they do it.**
@@
 
-### After modality switch performance change
-We found significant difference between different modalities  in the pre-switch block in distance performance `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 3)))$statistic`, and the error rate as well `r  apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 3)))$statistic`.
-
-**Running the tukey on these???**
-There was also a difference between conditions in their pointing performance at the end of the first phase (`r apa_print(aov(abs_error ~ learning.condition, data = sub_sop_all))$statistic$learning_condition`), with both Treadmill VR group and the Desktop group performend significantly worse than the group in the real world.
+### Performance change after modality switch
 ```{r}
-TukeyHSD(aov(abs_error ~ learning.condition, data = sub_sop_all))
+aov_point_phase1 <- aov(abs_error ~ learning.condition, data = sub_sop_all[sub_sop_all$phase == 1,])
 ```
-
-After the modality switch, we can see a significant difference between groups in distance performance (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`) and errors as well (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`), with groups learning on the treadmill or desktop performing worse than those learning in the real world.
-This between group difference is still present in the 2nd block of the second phase (block 5) in error rate `r  apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 5)))$statistic` and marginally in distance `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 5)))$statistic`.
-In the last testing block (block 6), we found no differences among the groups in either error rate `r apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 6)))$statistic` nor distance `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 6)))$statistic`.
+We found significant difference between different modalities  in the pre-switch block in distance performance (`r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 3)))$statistic`), and the error rate as well (`r  apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 3)))$statistic`).
+We also obseved a difference between conditions in their pointing performance at the end of the first phase (`r apa_print(aov_point_phase1)$statistic$learning_condition`), with real group performing significantly better than the Treadmill VR group (`r tukey_report(TukeyHSD(aov_point_phase1), condition = "Treadmill VR-Real", delta_name = "angle error")`) and the Desktop group (`r tukey_report(TukeyHSD(aov_point_phase1), condition = "Real-Desktop", delta_name = "angle error")`).
 
@@
 
-We can see that all participants improved in the errors made, but neigher group is significantly worse in the post-switch block in the path travelled (although participants who learned on the desktop performed slightly worse, the significance is only marginal).
+We can see that all participants improved in the number of errors made, but neigher group is significantly worse or better in the post-switch block in the path travelled (although participants who learned on the desktop performed slightly worse, the significance is only marginal).
 
-### Performance change 
-To assess the level of performance change form directly before and after the switch, we calculated a personal improvement score as (block3-block4)/(block3+block4)´.
-This allowed us to directly comapre the percentual improvement or deterioration from pre-switch block to the post-switch one.
+### Relative performance change after switch
+To assess the level of performance change form directly before and after the switch, we calculated personal improvement score as (block3-block4)/(block3+block4)´.
+This allowed us to directly compare the relative improvement or deterioration from pre-switch block to the post-switch one.
 
@@
 
-Comparing the perfomance change in different conditions using ANOVA, we see marginally significant difference between groups in the distance improvement (`r apa_print(aov(block34 ~ learning.condition, df_wide_distance))$statistic$learning_condition`), but significant differences in error rate improvement (`r apa_print(aov(block34 ~ learning.condition, df_wide_errors))$statistic$learning_condition`).
-This is consistent with the mixed model result which poited at interaction between block and learning condition.
+Comparing the performance change in different conditions using ANOVA, we see marginally significant difference between groups in the distance improvement (`r apa_print(aov(block34 ~ learning.condition, df_wide_distance))$statistic$learning_condition`), but  there are significant differences in error rate improvement (`r apa_print(aov(block34 ~ learning.condition, df_wide_errors))$statistic$learning_condition`).
+This is consistent with the mixed model result which pointed at an interaction between block number and learning condition.
 
@@
 ```
+
+### Post Switch performance
+
+In the post-switch block, we can still see a significant difference between groups in their distance performance (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`) and error rate as well (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`), with groups learning on the treadmill or desktop performing worse than those learning in the real world.
+This between group difference is still present in the 2nd block of the second phase (block 5) in error rate `r  apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 5)))$statistic` and marginally in distance `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 5)))$statistic`.
+In the last testing block (block 6), we found no differences among the groups in either error rate (`r apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 6)))$statistic`), nor pointing performance (`r apa_print(aov(abs_error ~ learning.condition, data = sub_sop_all[sub_sop_all$phase == 2,]))$statistic$learning_condition`) but we still see small difference in distance performance (`r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 6)))$statistic`).
 
@@
 
-Looking at the linear mixed model prediction for the distnace and error improvement in the second phase (post-switch), we can see that both learning condition and
+Looking at the linear mixed model prediction for the distance and error improvement in the second phase (post-switch), we can see that both learning condition and interaction with the experiment progression is significant, showing that there are different learning rates for different learning conditions even after the switch to the real world.
 
@@
 
-### Something about pointing?
-```{r}
+```{r, echo=F, results='hide'}
 err_phase <- lmer(abs_error ~ phase*learning.condition + (1 | id/phase), data = sub_sop_all)
@@
 
-We found difference in relative improvement rate
+If we assume no transfer to happen, we should observe the same performance in the real world testing in the second phase to be similar to the real world learning group's performance in the first phase.
+In other terms, the switch groups should demonstrate same performance in the post-switch block as the real world group in the first exposition.
+But we clearly observe that the post-switch performance is better than the first block performance across all conditions.

### 2018-10-29 Report/paper-results.Rmd
adds plot of walking

@@
 
+```{r}
+#plot for walking demonstration - need to load 
+#load(file = "multi_smoothed.data")
+#tw102 <- ls$tw102
+#save(tw102, file = "tw102.data"
+#load(file='tw102.data')
+#plot_walk_trial(tw102$phase2,12)
+#plot_walk_trial(tw102$phase1,12)
+``
 ## Measures

### 2018-12-09 TwoWorlds/helpers-results.R
Adds option to prepare block based on variable presence

@@ -7,10 +7,15 @@ ez_prepare <- function(filtered_data, by){
   return(filtered_data)
 }
 
-
-ez_prepare_block <- function(df, blocks){
+### Removes all cases where there is missing data in blocks
+ez_prepare_block <- function(df, blocks, variable=NULL){
   filtered_data <- df[df$exp_block_id %in% blocks, ]
-  filtered_data <- filtered_data[complete.cases(filtered_data),]
+  if(is.null(variable)){
+    filtered_data <- filtered_data[complete.cases(filtered_data),]
+  } else {
+    filtered_data <- filtered_data[!is.na(filtered_data[,variable]),]
+  }
+  
   has_all <- filtered_data %>% group_by(id) %>% select(id, "exp_block_id") %>% unique() %>% count
   has_all <- has_all %>% filter(n == 2)
   filtered_data <- filtered_data %>% filter(id %in% has_all$id)

### 2018-12-09 Report/final-results.Rmd
Rewriting the results into ezANOVA

--- 2018-10-28 Report/paper-results.Rmd
@@
 ---
-title: "R Notebook"
+title: "Transfer report"
+author: "Lukáš Hejtmánek"
+date: "9 August 2018"
 output: github_document
@@
 
+sub_walk_all_factor <- sub_walk_all
+sub_walk_all_factor$exp_block_id <- factor(sub_walk_all_factor$exp_block_id)
+sub_walk_all_factor$id <- factor(sub_walk_all_factor$id)
+sub_walk_all_factor$learning.condition <- factor(sub_walk_all_factor$learning.condition)
 ```
 
-# Methods
+```{r}
+fit_mm_dist_block <- lmer(min_norm_distance ~ learning.condition*exp_block_id + (1 | id), data = ez_data)
+summary(fit_mm_dist_block)
+anova(fit_mm_dist_block, type=3)
 
-## Measures
-We calculated participants’ walked distance in the building, time spent in each task and number of incorrectly visited doors.
-The conditions were not perfectly comparable in their measured distance and time, with the VR trials taking a little bit longer in both time and distance than the real world condition.
-We therefore min-normalised the distance traveled and the time for each task and environment  by dividing the measure by any participant's performance for that trial.
-We calculated all statistics on these normalised measures.
-
-There was no difference between conditions at the start of the experiment in normalised distance (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`), nor in number errors (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`), but there is a significant difference in normalised time (`r apa_print(aov(min_norm_time ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`)), due to VR taking longer times to get used to (and therefore having longer normalised trial times at the start).
-We removed distance measurements larger than 3 standard deviations from the mean.
-
-## Participants
-```{r}
-finished <- overview %>% group_by(condition) %>% count(finished)
-finished <- dcast(finished, condition~finished)
-finished$percent <- round(finished$yes/(finished$yes+finished$no), 2)
-is_ok <- overview %>% filter(finished=="yes") %>% group_by(condition) %>% count(is_ok)
-is_ok <- dcast(is_ok, condition~is_ok)
-is_ok$percent <- round(is_ok$yes/(is_ok$yes+is_ok$no), 2)
-```
-A total of `r sum(overview$is_ok == "yes", na.rm=T)` (M = `r round(mean(df_participants[df_participants$finished == "yes",]$age, na.rm=T),1)`, SD = `r round(sd(df_participants[df_participants$finished == "yes",]$age, na.rm=T), 1)`) undergraduate students at UC Davis particiapted in the study in exchange for a study credit.
-Each students was randomly assigned a condition and a randomised set of goals.
-`r sum(finished$no)` participants didn't finish due to motion sickness and `r sum(is_ok$no)` were removed due to a technical failure of the real world tracking systen.
-Only `r finished[finished$condition == "vr-real","percent"] * 100` percent the students being able to finish VR to Real world condition.
-
-# Results
-
-First we wanted to assess that there is going to be some level of transfer in all conditions.
-If we would assume no transfer to happen, we should observe same performance in the real world testing in 1st, 2nd and 3rd block after the switch to be similar in performance as the 1st, 2nd and 3rd block in the real world learning group.
-But all conditions show better performance after the switch (see table XXX) than in the first trial, suggesting some level of transfer happening in all learning conditions.
-```{r}
-df <- data.frame(condition=rep("",3),dist=rep("",3), err=rep("",3), stringsAsFactors = F)
-df_dist <- block_t_test_improvement(sub_walk_all, 1, 4,"min_norm_distance")
-df_error <- block_t_test_improvement(sub_walk_all, 1, 4,"errors")
-for(i in 1:length(names(df_dist))){
-  name <- names(df_dist)[i]
-  df[i,"condition"] <- name
-  df[i, "dist"] <- apa_print(df_dist[[name]])$full_result
-  df[i, "err"] <- apa_print(df_error[[name]])$full_result
-}
-colnames(df) <- c("Learning condition", "Distance improvement", "Errors improvement")
-df$`Learning condition` <- dplyr::recode(df$`Learning condition`, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
-kable(df, caption="Paired T test comparing individual performance improvement from block 1-4 in different conditions")
+ez_data <- ez_prepare_block(sub_walk_all_factor, c(1,4))
+ez_fit <- ezANOVA(ez_data, 
+        dv = min_norm_distance,
+        wid = id,
+        within = exp_block_id,
+        between = learning.condition,
+        return_aov = T,
+        type = 3)
 ```
 
-Looking at the progression in distance improvements and error improvements over the course of the experiment, we can see that all conditions show progressive improvement which is only hindered by the switch in the desktop learning condition (see fig.
-XXX and fig.
-XXX), but in the Treadmill condition we see constant improvement unobstructed by the environment switch.
-```{r}
-ggplot(sub_walk_all, aes(x = exp_block_id, y = min_norm_distance, color = factor(learning.condition))) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.5, geom="errorbar") + 
-  stat_summary(fun.y=mean,position=position_identity(), size = 1.25, geom="line") + 
-  geom_vline(xintercept = 3.5, linetype=4) +
-  ylab("Average normalised trial distance") + xlab("Block") + 
-  scale_color_manual(values=estimote_pal,"Learning condition") + guides(fill=F)
+We hypothesized that the degree of transfer to the real-world building would vary as a function of immersion.
+As a first test of this issue, we wished to determine whether all modalities resulted in some degree of transfer.
+Data for distance walked and number of door visitation errors were each entered into a 3 (Learning Condition: real, desktop, treadmill VR) × 2 (Block: block 1, block 4) mixed-model analysis of variance (ANOVA).
+For distance walked, there was a main effect for Learning Condition, `r apa_print(ez_fit$aov)$statistic$learning_condition`.
+There was a also a main effect for Block, `r apa_print(ez_fit$aov)$statistic$learning_condition`.
+The Learning Condition by Block interaction effect did not reach significance `r apa_print(ez_fit$aov)$statistic$learning_condition_exp_block_id`.
+Follow-up comparisons showed that for walked distance all three modalities showed an improvement from block 1 (first exposure to the modality) to block 4 (first exposure to the real building for immersed/impoverished conditions, fourth exposure for real-world).
+Descriptive and inferential statistics for these comparisons are shown in Table X and Table X+1 for distance and errors, respectively.
 
-ggplot(sub_walk_all, aes(x = exp_block_id, y = errors, color = factor(learning.condition))) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.5, geom="errorbar") + 
-  stat_summary(fun.y=mean,position=position_identity(), size = 1.25, geom="line") + 
-  geom_vline(xintercept = 3.5, linetype=4) +
-  ylab("Number of mistakes in office decision") + xlab("Block number") + 
-  scale_color_manual(values=estimote_pal, "Learning condition")
-```
-
-Participant's performance in the first pre-switch phase shows different rates of learning rate, demonstrated by an interaction between block bumber and learning condition for error rate performance (see table XX).
-
-**this is a mixed model with id/block id random term.
-Significance was measuyred with Anova from car package, but I read it is not a good way to measure significance and we shouldn't even use significance in mixed models.
-I'll check on past cognition papers to see how and if they do it.**
-
-```{r}
-learning_slopes_distance <- lmer(min_norm_distance ~ learning.condition*exp_block_id + (1 | id/exp_block_id), data = sub_walk_all[sub_walk_all$exp_block_id %in% 1:3,])
-tab_distance <- mixed_table_report(learning_slopes_distance)
-
-learning_slopes_errors <- lmer(errors ~ learning.condition*exp_block_id + (1 | id/exp_block_id), data = sub_walk_all[sub_walk_all$exp_block_id %in% 1:3,])
-tab_errors <- mixed_table_report(learning_slopes_errors)
-
-tab <- merge(tab_distance[,c(3,4)], tab_errors[,c(3,4)], by = "names")
-tab$names <- c("Intercept", "Block number", "Learning condition", "Block - Learning condition interaction")
-colnames(tab) <- c("Predictor", "Significance for distance improvement", "Significance for error improvement")
-kable(tab)
-#ggplot(sub_walk_all[sub_walk_all$exp_block_id %in% 1:3,], aes(exp_block_id, min_norm_distance, color=learning.condition)) + geom_point() + geom_smooth(method="lm", se=F)
-#ggplot(sub_walk_all[sub_walk_all$exp_block_id %in% 1:3,], aes(exp_block_id, errors, color=learning.condition)) + geom_point() + geom_smooth(method="lm", se=F)
-
-```
-
-### After modality switch performance change
-We found significant difference between different modalities  in the pre-switch block in distance performance `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 3)))$statistic`, and the error rate as well `r  apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 3)))$statistic`.
-
-**Running the tukey on these???**
-```{r}
-aov_point_phase1 <- aov(abs_error ~ learning.condition, data = sub_sop_all[sub_sop_all$phase == 1,])
-```
-There was also a difference between conditions in their pointing performance at the end of the first phase (`r apa_print(aov_point_phase1)$statistic$learning_condition`), with real group performing significantly better than the Treadmill VR group `r tukey_report(TukeyHSD(aov_point_phase1), condition = "Treadmill VR-Real", delta_name = "error")` and the Desktop group `r tukey_report(TukeyHSD(aov_point_phase1), condition = "Real-Desktop", delta_name = "error")`.
-
-After the modality switch, we can see a significant difference between groups in distance performance (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`) and errors as well (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`), with groups learning on the treadmill or desktop performing worse than those learning in the real world.
-This between group difference is still present in the 2nd block of the second phase (block 5) in error rate `r  apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 5)))$statistic` and marginally in distance `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 5)))$statistic`.
-In the last testing block (block 6), we found no differences among the groups in either error rate `r apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 6)))$statistic`, nor pointing performance `r apa_print(aov(abs_error ~ learning.condition, data = sub_sop_all[sub_sop_all$phase == 2,]))$statistic$learning_condition`) but we still see small difference in distance performance `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 6)))$statistic`.
-
-Mixed effect models with individual random effect show significant effect of the of the learning condition on rate of improvement from pre-switch to post-switch block (See table XXX).
-
-```{r}
-learning_slopes_distance <- lmer(min_norm_distance ~ learning.condition*exp_block_id + (1 | id/exp_block_id), data = only_blocks(sub_walk_all, 3:4))
-tab_distance <- mixed_table_report(learning_slopes_distance)
-
-learning_slopes_errors <- lmer(errors ~ learning.condition*exp_block_id + (1 | id/exp_block_id), data = only_blocks(sub_walk_all, 3:4))
-tab_errors <- mixed_table_report(learning_slopes_errors)
-
-tab <- merge(tab_distance[,c(3,4)], tab_errors[,c(3,4)], by = "names")
-tab$names <- c("Intercept", "Block number", "Learning condition", "Block - Learning condition interaction")
-colnames(tab) <- c("Predictor", "Significance for distance improvement", "Significance for error improvement")
-kable(tab)
-```
-
-Running separate pairwise t-tests to see individual performance change from pre-switch to post-switch block, we see significant improvement in errors made across all conditions, and marginally significant worsening in distance performance in the desktop condition.
-```{r, echo=F}
-df <- data.frame(condition=rep("",3),dist=rep("",3), err=rep("",3), stringsAsFactors = F)
-df_dist <- block_t_test_improvement(sub_walk_all, 3, 4,"min_norm_distance")
-df_error <- block_t_test_improvement(sub_walk_all, 3, 4,"errors")
-for(i in 1:length(names(df_dist))){
-  name <- names(df_dist)[i]
-  df[i,"condition"] <- name
-  df[i, "dist"] <- apa_print(df_dist[[name]])$full_result
-  df[i, "err"] <- apa_print(df_error[[name]])$full_result
-}
-colnames(df) <- c("Learning condition", "Distance improvement", "Errors improvement")
-df$`Learning condition` <- dplyr::recode(df$`Learning condition`, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
-kable(df, caption="Paired T test comparing individual performance improvement from pre switch block to post-switch block in different conditions")
-```
-
-We can see that all participants improved in the errors made, but neigher group is significantly worse in the post-switch block in the path travelled (although participants who learned on the desktop performed slightly worse, the significance is only marginal).
-
-### Performance change 
-To assess the level of performance change form directly before and after the switch, we calculated a personal improvement score as (block3-block4)/(block3+block4)´.
-This allowed us to directly comapre the percentual improvement or deterioration from pre-switch block to the post-switch one.
-
-```{r, echo=F}
-ggplot(df_distance, aes(x = learning.condition, y = block34.diff, fill = learning.condition)) + 
-  geom_bar(stat = 'identity', position="dodge") + 
-  geom_errorbar(aes(ymin=block34.diff-block34.se, ymax=block34.diff+block34.se), width=.2, position=position_dodge(.9)) +
-  scale_fill_manual(values = estimote_pal) +
-  ylab("Mean normalised distance improvement from block 3 to 4") + xlab("Learning condition")  + guides(fill=F)
-
-ggplot(df_errors, aes(x = learning.condition, y = block34.diff, fill = learning.condition)) + 
-  geom_bar(stat = 'identity', position="dodge") + 
-  geom_errorbar(aes(ymin=block34.diff-block34.se, ymax=block34.diff+block34.se), width=.2, position=position_dodge(.9)) +
-  scale_fill_manual(values = estimote_pal) +
-  ylab("Mean sum of errors improvement after switch")+ xlab("Learning condition") + guides(fill=F)
-```
-
-Comparing the perfomance change in different conditions using ANOVA, we see marginally significant difference between groups in the distance improvement (`r apa_print(aov(block34 ~ learning.condition, df_wide_distance))$statistic$learning_condition`), but significant differences in error rate improvement (`r apa_print(aov(block34 ~ learning.condition, df_wide_errors))$statistic$learning_condition`).
-This is consistent with the mixed model result which poited at interaction between block and learning condition.
-
-Tukey post-hoc tests show significant difference between error rate improvement between the group that learned in the real world and that which learned on the desktop.
-```{r, echo=F}
-kable(tukey_report_table(TukeyHSD(x=aov(block34 ~ learning.condition, df_wide_errors), conf.level=0.95), "learning.condition"))
-```
-
-### Improvement from the pre-switch block to the final block
-Assuming the performance platoes around the 6the block, looking at participants performance in the pre-switch block and the last block we can deduce what margin of improvement is still possible.
-We see that participants learning in the real world no longer improve in error rate after the 3rd block `r apa_print(block_t_test_improvement(sub_walk_all, 4, 6, "errors")[["real-real"]])$statistic`, but they are continuously improving in distance performance `r apa_print(block_t_test_improvement(sub_walk_all, 4, 6, "min_norm_distance")[["real-real"]])$statistic`.
-
-*Not sure about the mixed models here. I think they are OK, but I haven't published anything with them before*
-
-Looking at the linear mixed model prediction for the distnace and error improvement in the second phase (post-switch), we can see that both learning condition and
-
-```{r}
-learning_slopes_distance <- lmer(min_norm_distance ~ learning.condition*exp_block_id + (1 | id/exp_block_id), data = only_blocks(sub_walk_all[sub_walk_all$min_norm_distance < 10, ], 4:6))
-tab_distance <- mixed_table_report(learning_slopes_distance)
-
-learning_slopes_errors <- lmer(errors ~ learning.condition*exp_block_id + (1 | id/exp_block_id), data = only_blocks(sub_walk_all[sub_walk_all$min_norm_distance < 10, ], 4:6))
-tab_errors <- mixed_table_report(learning_slopes_errors)
-
-tab <- merge(tab_distance[,c(3,4)], tab_errors[,c(3,4)], by = "names")
-tab$names <- c("Intercept", "Block number", "Learning condition", "Block - Learning condition interaction")
-colnames(tab) <- c("Predictor", "Significance for distance improvement", "Significance for error improvement")
-kable(tab)
-```
-
-### Something about pointing?
-```{r, echo=F, results='hide'}
-err_phase <- lmer(abs_error ~ phase*learning.condition + (1 | id/phase), data = sub_sop_all)
-car::Anova(err_phase, type = "III")
-
-ggplot(sub_sop_all, aes(phase, abs_error, color = learning.condition)) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.5, geom="errorbar") + 
-  stat_summary(fun.y=mean,position=position_identity(), size = 1.25, geom="line")
-
-aov(abs_error ~ learning.condition, data = sub_sop_all)
-```
-
-## Summary
-We can see different rates of learning speed across different conditions, with significant differences between groups in their performance in the post-switch block.
-Participants learning in the real world are performing the best after the switch, but only significanly different from the group learning on the desktop.
-All conditions show some level of transfer, demonstrated both by general improvement from the first block to the fourth, as well as decreased errors in desktop and treadmill VR group after the switch to the real environment.
-All groups get to the same level of performance after the three blocks post-switch moving in the real world, not differing in errors made nor in distance in the last block.
-But we also observed that whilst real world training leads to almost perfect performance in the pre-switch block, not differing from the last block in error rate and only slightly in distance.
-It takes another 3 blocks in the real world for the groups which learned on the treadmill or the desktop to achieve same performance, again pointing to qualitative difference between different modalities.
-
-*I am unsure if it is worth adding the rest of the dataset in (the VR - VR and real - VR groups), but the performance of Treadmill VR groups in the distance traveled gets exactly to the same level as the real world performance in the 6th block (so the VR - VR group performs the same after the 6 blocks as VR - real and real - VR).
-Which is nice comparison to the results we get here about the distance performance in the real world being almost perfect in the 3rd block and the rest of the groups slowly "getting there".*
-
-*The VR groups still demonstrate larger error difference in the 4-6th block, making more errors overall than the real world groups (about 2 errors per trial more).
-I believe it can be explained by either social difference - participants not wanting to "ask as much" in the real world, or accidental walking into a door due to treadmill control issues.
-Also, particiapnts told me that the VR is more "inviting" to test all the doors that are being passed (just in case), because it is more tiring to get back to one in case of a mistake.
-Anyway, not sure it it warranties adding more groups to the study, but at least the data are consistent and distance performance gets to the same level even in the VR*
-
-We found difference in relative improvement rate
+For visitation errors, there was a (was no) main effect for Learning Condition, F(DFn, DFd) = ##.##, p = .###.
+There was a (was no) main effect for Block, F(DFn, DFd) = ##.##, p = .###.
+The Learning Condition by Block interaction effect was significant (did not reach significance), F(DFn, DFd) = ##.##, p = .###.
+Follow-up comparisons showed that all three modalities showed an improvement in visitation errors from block 1 (first exposure to the modality) to block 4 (first exposure to the real building for immersed/impoverished conditions, fourth exposure for real-world).
+Descriptive and inferential statistics for these comparisons are shown in Table X+1.

### 2018-12-09 Report/paper-results.Rmd
Rewriting the results into ezANOVA

--- 2018-10-29 Report/paper-results.Rmd
@@
 
+```{r}
+#plot for walking demonstration - need to load 
+#load(file = "multi_smoothed.data")
+#tw102 <- ls$tw102
+#save(tw102, file = "tw102.data"
+#load(file='tw102.data')
+#plot_walk_trial(tw102$phase2,12)
+#plot_walk_trial(tw102$phase1,12)
+```
+
 ## Measures
@@
   geom_vline(xintercept = 3.5, linetype=4) +
-  ylab("Number of mistakes in office decision") + xlab("Block number") + 
+  ylab("Visitation errors") + xlab("Block number") + 
   scale_color_manual(values=estimote_pal, "Learning condition")

### 2018-12-09 TwoWorlds/helpers-report.R
adds new helpers to calculate results

@@ -37,5 +37,17 @@ mixed_table_report <- function(fit_lme){
 } 
 
 only_blocks <- function(df, ...){
-  return(df[df$exp_block_id %in% c(...),])
+  df <- df[df$exp_block_id %in% c(...),]
+  return(df)
+}
+
+only_condition <- function(df, condition){
+  df <- df[df$learning.condition == condition, ]
+  return(df)
+}
+
+only_condition_blocks <- function(df, condition, ...){
+  df <- only_condition(df, condition)
+  df <- only_blocks(df, ...)
+  return(df)
 }

### 2018-12-09 TwoWorlds/helpers-results.R

@@ -24,3 +24,9 @@ ez_prepare_block <- function(df, blocks, variable=NULL){
 }
 
 mean.na <- function(data){return(mean(data, na.rm = T))}
+
+m_sd_report <- function(vec, digits=2){
+  m <- round(mean(vec, na.rm=T), digits)
+  s <- round(sd(vec, na.rm=T), digits)
+  return(paste("M = ", m, ", SD = ", s, collapse=""))
+}

### 2018-12-09 Report/helpers/block-t-test-improvement.R
Adds t-test report table helper

@@ -6,7 +6,32 @@ block_t_test_improvement <- function(df, block1, block2, value){
   conditions <- unique(df_wide$condition)
   ls <- list()
   for(cond in conditions){
-    ls[[cond]] <- t.test(df_wide[df_wide$condition == cond,"block1"], df_wide[df_wide$condition == cond,"block2"], paired=T)
+    ls[[cond]] <- t.test(df_wide[df_wide$condition == cond, "block1"], df_wide[df_wide$condition == cond,"block2"], paired=T)
+    ls[[cond]]$cohen.d <- effsize::cohen.d(df_wide[df_wide$condition == cond, "block1"], df_wide[df_wide$condition == cond,"block2"], paired=T, na.rm=T)
   }
   return(ls)
+}
+
+block_t_test_improvement_report <- function(ls){
+  n_conditions <- length(names(ls))
+  df <- data.frame(condition=rep(NA_character_,n_conditions), 
+                   Mdiff=rep(NA_real_,n_conditions), 
+                   df=rep(NA_real_,n_conditions), 
+                   t=rep(NA_real_,n_conditions), 
+                   p=rep(NA_character_,n_conditions), 
+                   d=rep(NA_real_,n_conditions), 
+                   CI=rep(NA_character_,n_conditions), 
+                   stringsAsFactors = F)
+  for(i in 1:n_conditions){
+    name <- names(ls)[i]
+    cond <- ls[[i]]
+    df[i,"condition"] <- name
+    df[i, "Mdiff"] <- cond$estimate
+    df[i, "df"] <- cond$parameter
+    df[i, "t"] <- cond$statistic
+    df[i, "p"] <- apa_p(cond$p.value)
+    df[i, "d"] <- cond$cohen.d$estimate
+    df[i, "CI"] <- paste(round(cond$conf.int[1:2], 2), collapse = ", ")
+  }
+  return(df)
 }

### 2018-12-09 Report/final-results.Rmd
adds new reports

@@
 
-settings <- load_google_sheets()
-df_participants <- settings$participants
-df_participants <- df_participants %>% mutate(condition=paste0(First.phase,"-",Second.phase))
-df_participants <- df_participants %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
-overview <- settings$versions
-overview <- overview %>% mutate(condition=paste0(First.phase,"-",Second.phase))
-overview <- overview %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
+#settings <- load_google_sheets()
+#df_participants <- settings$participants
+#df_participants <- df_participants %>% mutate(condition=paste0(First.phase,"-",Second.phase))
+#df_participants <- df_participants %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
+#overview <- settings$versions
+#overview <- overview %>% mutate(condition=paste0(First.phase,"-",Second.phase))
+#overview <- overview %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
 
@@
 ```{r}
+ez_data<- ez_prepare_block(sub_walk_all_factor, c(1,4), 'min_norm_distance')
 fit_mm_dist_block <- lmer(min_norm_distance ~ learning.condition*exp_block_id + (1 | id), data = ez_data)
-summary(fit_mm_dist_block)
-anova(fit_mm_dist_block, type=3)
-
-ez_data <- ez_prepare_block(sub_walk_all_factor, c(1,4))
-ez_fit <- ezANOVA(ez_data, 
+ez_fit_distance <- ezANOVA(ez_data, 
         dv = min_norm_distance,
@@
         type = 3)
+
+ez_data <- ez_prepare_block(sub_walk_all_factor, c(1,4), 'errors')
+ez_fit_errors <- ezANOVA(ez_data, 
+        dv = errors,
+        wid = id,
+        within = exp_block_id,
+        between = learning.condition,
+        return_aov = T,
+        type = 3)
+
 ```
@@
 Data for distance walked and number of door visitation errors were each entered into a 3 (Learning Condition: real, desktop, treadmill VR) × 2 (Block: block 1, block 4) mixed-model analysis of variance (ANOVA).
-For distance walked, there was a main effect for Learning Condition, `r apa_print(ez_fit$aov)$statistic$learning_condition`.
-There was a also a main effect for Block, `r apa_print(ez_fit$aov)$statistic$learning_condition`.
-The Learning Condition by Block interaction effect did not reach significance `r apa_print(ez_fit$aov)$statistic$learning_condition_exp_block_id`.
+For distance walked, there was a main effect for Learning Condition, `r apa_print(ez_fit_distance$aov)$statistic$learning_condition`.
+There was a also a main effect for Block, `r apa_print(ez_fit_distance$aov)$statistic$exp_block_id`.
+The Learning Condition by Block interaction effect did not reach significance `r apa_print(ez_fit_distance$aov)$statistic$learning_condition_exp_block_id`.
 Follow-up comparisons showed that for walked distance all three modalities showed an improvement from block 1 (first exposure to the modality) to block 4 (first exposure to the real building for immersed/impoverished conditions, fourth exposure for real-world).
@@
 
+For visitation errors, there was no main effect for Learning Condition,`r apa_print(ez_fit_errors$aov)$statistic$learning_condition`.
+There was a main effect for Block, `r apa_print(ez_fit_errors$aov)$statistic$exp_block_id`.
+The Learning Condition by Block interaction effect did not reach significance, `r apa_print(ez_fit_errors$aov)$statistic$learning_condition_exp_block_id`.
+Follow-up comparisons showed that all three modalities showed an improvement in visitation errors from block 1 (first exposure to the modality) to block 4 (first exposure to the real building for immersed/impoverished conditions, fourth exposure for real-world).
+Descriptive and inferential statistics for these comparisons are shown in Table X+1.
+
+#### TABLES
+```{r}
+#descriptive table
+kable(sub_walk_all_factor %>% filter(exp_block_id %in% c(1,4)) %>% group_by(learning.condition, exp_block_id) %>% summarise(M = mean(min_norm_distance, na.rm=T),
+                                                                                                                     Sd = sd(min_norm_distance, na.rm=T)), digits = 2)
+df_dist <- block_t_test_improvement(sub_walk_all, 1, 4,"min_norm_distance")
+df <- block_t_test_improvement_report(df_dist)
+df$condition <- dplyr::recode(df$condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
+kable(df, caption="Paired T test comparing individual distance performance improvement from pre switch block to post-switch block", digits = 2)
+
+kable(sub_walk_all_factor %>% filter(exp_block_id %in% c(1,4)) %>% group_by(learning.condition, exp_block_id) %>% summarise(M = mean(errors, na.rm=T),
+                                                                                                                     Sd = sd(errors, na.rm=T)), digits = 2)
+df_error <- block_t_test_improvement(sub_walk_all, 1, 4,"errors")
+df <- block_t_test_improvement_report(df_error)
+df$condition <- dplyr::recode(df$condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
+kable(df, caption="Paired T test comparing individual error rate improvement from pre switch block to post-switch block", digits = 2)
+```
+
+Figure X shows the time course of each learning condition over the 6 blocks of exposure to the environments, in terms of walked distance and visitation errors over the course of the experiment.
+As can be seen in Figure XX, all modalities showed an incremental improvement in terms of both distance and visitation errors (see fig.
+XXX and fig.
+XXX).
+At block 4, however, an uptick can be observed for the desktop modality switch group and slight for no modality switch (i.e.
+real to real).
+To better quantify these effects, we took two different approaches. First, we analyzed the slope of the learning curves from phase 1 (blocks 1-3) and compared between different Learning Conditions.
+
+```{r}
+### graphs
+ggplot(sub_walk_all, aes(x = exp_block_id, y = min_norm_distance, color = factor(learning.condition))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.5, geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_identity(), size = 1.25, geom="line") + 
+  geom_vline(xintercept = 3.5, linetype=4) +
+  ylab("Average normalised trial distance") + xlab("Block") + 
+  scale_color_manual(values=estimote_pal,"Learning condition") + guides(fill=F)
+
+ggplot(sub_walk_all, aes(x = exp_block_id, y = errors, color = factor(learning.condition))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.5, geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_identity(), size = 1.25, geom="line") + 
+  geom_vline(xintercept = 3.5, linetype=4) +
+  ylab("Visitation errors") + xlab("Block number") + 
+  scale_color_manual(values=estimote_pal, "Learning condition")
+```
+Figure X.
+Performance of each learning condition across time is shown for average normalized distance (a) and visitation errors (b).
+Vertical, dashed line indicates when the second phase of testing (always in the real world) began.
+Error bars represent [95% CI’s].
+
+## Quantifying what happens before transfer
+To better understand the effect of modality (i.e., real world vs. treadmill VR vs. desktop), we assessed the change in walked distance and visitation errors prior to transfer. 
+```{r}
+point_data_1 <- sub_sop_all[sub_sop_all$phase == 1,]
+aov_point_phase1 <- aov(abs_error ~ learning.condition, data = point_data_1)
+```
+
+Mixed-effect models revealed a main effect of block on distance performance, suggesting that participants improved their paths in all three conditions, F(DFn, DFd) = ##.##, p = .###.
+The interaction effect with modality was not significant , F(DFn, DFd) = ##.##, p = .###, suggesting that learning rate did not differ as a function of modality.
+In contrast, for visitation errors, we again found a main effect of block but, importantly, a block X modality interaction effect.
+This finding suggested that visitation errors improved differently as a function of modality.
+
+### Tables?
+
+For the last block of learning (block 3), we conducted separate one-way ANOVAs to compare learning conditions for walked distance and visitation errors.
+There were a main effects of for both walked distance, `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 3)))$statistic`, and visitation errors, `r  apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 3)))$statistic`).
+
+We also observed significant differences between conditions in pointing performance at the end of the first phase, `r apa_print(aov_point_phase1)$statistic$learning_condition`.
+Specifically, participants in the real-world learning group  performed significantly better compared to both the immersive VR group `r m_sd_report(only_condition(point_data_1, "Real")$abs_error)`, t(df) = ##.##, p < .001) and the desktop group (`r m_sd_report(only_condition(point_data_1, "Desktop")$abs_error)`), t(df) = ##.##, p < .001).
+Overall, these findings suggested that participants learned the building the best as a result of real-world navigation, followed by the immersive condition, with the desktop condition showing the slowest acquisition.
+
+## Quantifying what happens after/during?  transfer
+We next considered the effects of transferring from either the treadmill or desktop to the real-world building compared with continuing to navigate the real-world building.
+We did this by directly comparing walked distance and visitation errors on block 3 (the last block before the transfer phase) with the first block of the transfer phase (block 4).
+Data for distance walked and number of door visitation errors were each entered into a 3 (Learning Condition: real, desktop, treadmill VR) × 2 (Block: block 1, block 4) mixed-model ANOVA.
 For visitation errors, there was a (was no) main effect for Learning Condition, F(DFn, DFd) = ##.##, p = .###.
@@
 The Learning Condition by Block interaction effect was significant (did not reach significance), F(DFn, DFd) = ##.##, p = .###.
-Follow-up comparisons showed that all three modalities showed an improvement in visitation errors from block 1 (first exposure to the modality) to block 4 (first exposure to the real building for immersed/impoverished conditions, fourth exposure for real-world).
-Descriptive and inferential statistics for these comparisons are shown in Table X+1.
+For walked distance, there was no main effect of Learning Condition, F(DFn, DFd) = ##.##, p = .###, or Block, F(DFn, DFd) = ##.##, p = .###, and the interaction also failed to reach significance, , F(DFn, DFd) = ##.##, p = .### (See table XXX).

### 2018-12-09 TwoWorlds/helpers-report.R
Changes the selection helper

@@ -41,8 +41,8 @@ only_blocks <- function(df, ...){
   return(df)
 }
 
-only_condition <- function(df, condition){
-  df <- df[df$learning.condition == condition, ]
+only_conditions <- function(df, conditions){
+  df <- df[df$learning.condition %in% conditions, ]
   return(df)
 }
 

### 2018-12-09 Report/final-results.Rmd
Adds to the report

@@
 To better understand the effect of modality (i.e., real world vs. treadmill VR vs. desktop), we assessed the change in walked distance and visitation errors prior to transfer.
-```{r}
-point_data_1 <- sub_sop_all[sub_sop_all$phase == 1,]
-aov_point_phase1 <- aov(abs_error ~ learning.condition, data = point_data_1)
-```
 
@@
 
+```{r}
+point_data_1 <- sub_sop_all[sub_sop_all$phase == 1,]
+aov_point_phase1 <- aov(abs_error ~ learning.condition, data = point_data_1)
+```
+
 We also observed significant differences between conditions in pointing performance at the end of the first phase, `r apa_print(aov_point_phase1)$statistic$learning_condition`.
-Specifically, participants in the real-world learning group  performed significantly better compared to both the immersive VR group `r m_sd_report(only_condition(point_data_1, "Real")$abs_error)`, t(df) = ##.##, p < .001) and the desktop group (`r m_sd_report(only_condition(point_data_1, "Desktop")$abs_error)`), t(df) = ##.##, p < .001).
+Specifically, participants in the real-world learning group  (`r m_sd_report(only_conditions(point_data_1, "Real")$abs_error)`) performed significantly better compared to both the immersive VR group (`r m_sd_report(only_conditions(point_data_1, "Treadmill VR")$abs_error)`), `r apa_print(t.test(abs_error ~ learning.condition, only_conditions(point_data_1, c('Real', 'Treadmill VR'))))$statistic`, and the desktop group (`r m_sd_report(only_conditions(point_data_1, "Desktop")$abs_error)`), `r apa_print(t.test(abs_error ~ learning.condition, only_conditions(point_data_1, c('Real', 'Desktop'))))$statistic`.
 Overall, these findings suggested that participants learned the building the best as a result of real-world navigation, followed by the immersive condition, with the desktop condition showing the slowest acquisition.

### 2018-12-10 Report/final-results.Rmd
Updates the reposrt

@@
 
-#### TABLES
 ```{r}
@@
 df$condition <- dplyr::recode(df$condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
-kable(df, caption="Paired T test comparing individual distance performance improvement from pre switch block to post-switch block", digits = 2)
+kable(df, caption="Paired T test comparing individual distance performance improvement from the first block to post-switch block", digits = 2)
 
@@
 df$condition <- dplyr::recode(df$condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
-kable(df, caption="Paired T test comparing individual error rate improvement from pre switch block to post-switch block", digits = 2)
+kable(df, caption="Paired T test comparing individual error rate improvement from the first block to post-switch block", digits = 2)
 ```
@@
 ```
+
 Figure X.
@@
 ## Quantifying what happens before transfer
-To better understand the effect of modality (i.e., real world vs. treadmill VR vs. desktop), we assessed the change in walked distance and visitation errors prior to transfer.
+To better understand the effect of modality (i.e., real world vs. treadmill VR vs. desktop), we assessed the change in walked distance and visitation errors prior to transfer from block 1 to block 3.
 
-Mixed-effect models revealed a main effect of block on distance performance, suggesting that participants improved their paths in all three conditions, F(DFn, DFd) = ##.##, p = .###.
-The interaction effect with modality was not significant , F(DFn, DFd) = ##.##, p = .###, suggesting that learning rate did not differ as a function of modality.
-In contrast, for visitation errors, we again found a main effect of block but, importantly, a block X modality interaction effect.
+```{r}
+ez_data <- ez_prepare_block(sub_walk_all_factor, c(1,3), 'min_norm_distance')
+fit_mm_dist_block <- lmer(min_norm_distance ~ learning.condition*exp_block_id + (1 | id/exp_block_id), data = only_blocks(sub_walk_all, c(1,3)))
+ez_fit_distance <- ezANOVA(
+        ez_data, 
+        dv = min_norm_distance,
+        wid = id,
+        within = exp_block_id,
+        between = learning.condition,
+        return_aov = T,
+        type = 3)
+
+ez_data <- ez_prepare_block(sub_walk_all_factor, c(1,3), 'errors')
+ez_fit_errors <- ezANOVA(
+        ez_data, 
+        dv = errors,
+        wid = id,
+        within = exp_block_id,
+        between = learning.condition,
+        return_aov = T,
+        type = 3)
+```
+
+3x2 repeated measures ANOVA revealed a main effect of block on distance performance, suggesting that participants improved their paths in all three conditions, `r apa_print(ez_fit_distance$aov)$statistic$exp_block_id`.
+The interaction effect with modality was not significant , `r apa_print(ez_fit_distance$aov)$statistic$learning_condition_exp_block_id`, suggesting that learning rate did not differ as a function of modality.
+In contrast, for visitation errors, we again found a main effect of block (`r apa_print(ez_fit_errors$aov)$statistic$exp_block_id`) but, importantly, a block X modality interaction effect, `r apa_print(ez_fit_errors$aov)$statistic$learning_condition_exp_block_id`.
 This finding suggested that visitation errors improved differently as a function of modality.
 
-### Tables?
-
-For the last block of learning (block 3), we conducted separate one-way ANOVAs to compare learning conditions for walked distance and visitation errors.
-There were a main effects of for both walked distance, `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 3)))$statistic`, and visitation errors, `r  apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 3)))$statistic`).
+To investigate further, we conducted separate one-way ANOVAs to compare learning conditions for walked distance and visitation errors for the last block of learning (block 3).
+There were main effects of learning condition for both walked distance, `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 3)))$statistic`, and visitation errors, `r  apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 3)))$statistic`).
+Comparing the conditions closer, we find that there is a significant difference in visitations error performance between real world and desktop learning (`r apa_print(t.test(errors ~ learning.condition, only_conditions(sub_walk_all, c('Real', 'Desktop'))))$statistic`) and real world nad treadmill VR learning (`r apa_print(t.test(errors ~ learning.condition, only_conditions(sub_walk_all, c('Real', 'Treadmill VR'))))$statistic`), but no difference between Desktop and Treadmill learning (`r apa_print(t.test(errors ~ learning.condition, only_conditions(sub_walk_all, c('Desktop', 'Treadmill VR'))))$statistic`).
+The situation is the same for distance comparisons.
 
@@
 ## Quantifying what happens after/during? transfer
+```{r}
+ez_data <- ez_prepare_block(sub_walk_all_factor, c(3,4), 'min_norm_distance')
+fit_mm_dist_block <- lmer(min_norm_distance ~ learning.condition*exp_block_id + (1 | id/exp_block_id), data = only_blocks(sub_walk_all, 3:4))
+ez_fit_distance <- ezANOVA(
+        ez_data, 
+        dv = min_norm_distance,
+        wid = id,
+        within = exp_block_id,
+        between = learning.condition,
+        return_aov = T,
+        type = 3)
+
+ez_data <- ez_prepare_block(sub_walk_all_factor, c(3,4), 'errors')
+ez_fit_errors <- ezANOVA(
+        ez_data, 
+        dv = errors,
+        wid = id,
+        within = exp_block_id,
+        between = learning.condition,
+        return_aov = T,
+        type = 3)
+```
+
 We next considered the effects of transferring from either the treadmill or desktop to the real-world building compared with continuing to navigate the real-world building.
@@
 Data for distance walked and number of door visitation errors were each entered into a 3 (Learning Condition: real, desktop, treadmill VR) × 2 (Block: block 1, block 4) mixed-model ANOVA.
-For visitation errors, there was a (was no) main effect for Learning Condition, F(DFn, DFd) = ##.##, p = .###.
-There was a (was no) main effect for Block, F(DFn, DFd) = ##.##, p = .###.
-The Learning Condition by Block interaction effect was significant (did not reach significance), F(DFn, DFd) = ##.##, p = .###.
-For walked distance, there was no main effect of Learning Condition, F(DFn, DFd) = ##.##, p = .###, or Block, F(DFn, DFd) = ##.##, p = .###, and the interaction also failed to reach significance, , F(DFn, DFd) = ##.##, p = .### (See table XXX).
+For visitation errors, there was a main effect for Learning Condition,`r apa_print(ez_fit_errors$aov)$statistic$learning_condition`.
+There was a main effect for Block, `r apa_print(ez_fit_errors$aov)$statistic$exp_block_id`.
+The Learning Condition by Block interaction effect was significant, `r apa_print(ez_fit_errors$aov)$statistic$learning_condition_exp_block_id`.
+For walked distance, there was a main effect of Learning Condition,`r apa_print(ez_fit_distance$aov)$statistic$learning_condition`, but no effect of Block, `r apa_print(ez_fit_distance$aov)$statistic$exp_block_id`, and the interaction also failed to reach significance, `r apa_print(ez_fit_distance$aov)$statistic$learning_condition_exp_block_id`.
+
+To better understand these differences, we ran separate pairwise t-tests to assess the change from block 3 (last block of learning phase) to block 4 (first block of transfer phase).
+For distance, we found no significant differences from block 3 (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Real", 3)$min_norm_distance)`) to block 4 (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Real", 4)$min_norm_distance)`) for the real-world learning condition, `r apa_print(t.test(min_norm_distance ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Real'), 3:4)))$statistic`.
+Similarly, there was no difference between block 3 (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Treadmill VR", 3)$min_norm_distance)`) and block 4 (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Treadmill VR", 4)$min_norm_distance)`) for the treadmill VR condition, `r apa_print(t.test(min_norm_distance ~ exp_block_id, only_conditions_blocks(sub_walk_all, 'Treadmill VR', 3:4)))$statistic`.
+For visitation errors, we found an improvement in all conditions, although the effects sizes were greater for the treadmill compared to the desktop conditions.
+Together, these findings suggested that participants maintained or improved for both distance and visitation errors, although these effects were generally smaller, and even tending toward being worse, for the desktop condition.
+For summary please see table XXX.
+
+```{r}
+#descriptive table
+kable(sub_walk_all_factor %>% filter(exp_block_id %in% c(3,4)) %>% group_by(learning.condition, exp_block_id) %>% summarise(M = mean(min_norm_distance, na.rm=T),
+                                                                                                                     Sd = sd(min_norm_distance, na.rm=T)), digits = 2)
+df_dist <- block_t_test_improvement(sub_walk_all_factor, 3, 4,"min_norm_distance")
+df <- block_t_test_improvement_report(df_dist)
+df$condition <- dplyr::recode(df$condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
+kable(df, caption="Paired T test comparing individual distance performance improvement from pre switch block to post-switch block", digits = 2)
+
+kable(sub_walk_all_factor %>% filter(exp_block_id %in% c(3,4)) %>% group_by(learning.condition, exp_block_id) %>% summarise(M = mean(errors, na.rm=T),
+                                                                                                                     Sd = sd(errors, na.rm=T)), digits = 2)
+df_error <- block_t_test_improvement(sub_walk_all_factor, 3, 4, "errors")
+df <- block_t_test_improvement_report(df_error)
+df$condition <- dplyr::recode(df$condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
+kable(df, caption="Paired T test comparing individual error rate improvement from pre switch block to post-switch block", digits = 2)
+```
+
+## Quantifying transfer 
+To assess the level of performance change form directly before and after the switch, we calculated percent improvement score as
+(pre-switch performance - post-switch performance)/(pre-switch performance + post-switch performance)
+This allowed us to directly compare the relative improvement or deterioration from pre-switch block to the post-switch one.
+These findings are shown in Figure XX for normalized distance (Figure XXa) and visitation errors (Figure XXb).
+
+```{r, echo=F}
+ggplot(df_distance, aes(x = learning.condition, y = block34.diff, fill = learning.condition)) + 
+  geom_bar(stat = 'identity', position="dodge") + 
+  geom_errorbar(aes(ymin=block34.diff-block34.se, ymax=block34.diff+block34.se), width=.2, position=position_dodge(.9)) +
+  scale_fill_manual(values = estimote_pal) +
+  ylab("Mean normalised distance improvement from block 3 to 4") + xlab("Learning condition")  + guides(fill=F)
+
+ggplot(df_errors, aes(x = learning.condition, y = block34.diff, fill = learning.condition)) + 
+  geom_bar(stat = 'identity', position="dodge") + 
+  geom_errorbar(aes(ymin=block34.diff-block34.se, ymax=block34.diff+block34.se), width=.2, position=position_dodge(.9)) +
+  scale_fill_manual(values = estimote_pal) +
+  ylab("Mean sum of errors improvement after switch")+ xlab("Learning condition") + guides(fill=F)
+```
+
+Comparing the improvement in walked distance using one-way ANOVA, we see did not find any  significant difference between groups `r apa_print(aov(block34 ~ learning.condition, df_wide_distance))$statistic$learning_condition `.
+Visual inspection of the data, however, suggested that the desktop condition again showed the least transfer from block 3 to block 4.
+For visitation errors, there were significant group differences in error rate improvement, `r apa_print(aov(block34 ~ learning.condition, df_wide_errors))$statistic$learning_condition`.
+Comparing the conditions pairwise, we can asses that there is a significant difference in improvement between real world learnign and desktop (`r apa_print(t.test(block34 ~ learning.condition, only_conditions(df_wide_errors, c('Real', 'Desktop'))))$statistic`), as well as Treadmill VR, `r apa_print(t.test(block34 ~ learning.condition, only_conditions(df_wide_errors, c('Real', 'Treadmill VR'))))$statistic`, but we find no difference in improvement betweem desktop learning and Treadmill VR learning, `r apa_print(t.test(block34 ~ learning.condition, only_conditions(df_wide_errors, c('Treadmill VR', 'Desktop'))))$statistic`.
+
+Together, these findings suggest that transfer from desktop to real-world environments resulted in an increase in visitation errors, while no change was present for immersive and real-world conditions.
+
+## Quantifying what happens after transfer
+
+```{r}
+ez_data <- ez_prepare_block(sub_walk_all_factor, c(4,6), 'min_norm_distance')
+fit_mm_dist_block <- lmer(min_norm_distance ~ learning.condition*exp_block_id + (1 | id/exp_block_id), data = only_blocks(sub_walk_all, c(4,6)))
+ez_fit_distance <- ezANOVA(
+        ez_data, 
+        dv = min_norm_distance,
+        wid = id,
+        within = exp_block_id,
+        between = learning.condition,
+        return_aov = T,
+        type = 3)
+
+ez_data <- ez_prepare_block(sub_walk_all_factor, c(4,6), 'errors')
+ez_fit_errors <- ezANOVA(
+        ez_data, 
+        dv = errors,
+        wid = id,
+        within = exp_block_id,
+        between = learning.condition,
+        return_aov = T,
+        type = 3)
+```
+
+As an additional assay of transfer, we wished to determine whether learning post modality switch was faster in the group that had been exposed to the real-world in the learning blocks compared to the VR groups.
+Comparing walked distance and visitation errors across between blocks 4 and 6 by running repeated measures ANOVA, we found a significant effect of learning condition on distance,`r apa_print(ez_fit_distance$aov)$statistic$learning_condition` and visitation errors `r apa_print(ez_fit_errors$aov)$statistic$learning_condition`.
+We also wound significant interaction effect between learning condition and block for both distance `r apa_print(ez_fit_distance$aov)$statistic$learning_condition` and visitation errors `r apa_print(ez_fit_errors$aov)$statistic$learning_condition`.
+
+Looking at the improvement in visitations errors from block 4 to block 6 in separate learning conditions using t-tests, we can assess that only Desktop group (`r apa_print(t.test(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Desktop'), c(4,6))))$statistic`) and Treadmill VR group (`r apa_print(t.test(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Treadmill VR'), c(4,6))))$statistic`) have improved, whereas group learned in the real world did not improve any more in the transfer phase (`r apa_print(t.test(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Real'), c(4,6))))$statistic`).
+In fact, the real world group retains its visitation errors performance even from the last learning phase block does not significantly improve in visitation errors until the end (`r apa_print(t.test(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Real'), c(3,6))))$statistic`).
+
+The story is similar with distance performance, with both Desktop (`r apa_print(t.test(min_norm_distance ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Desktop'), c(4,6))))$statistic`) and Treadmill VR (`r apa_print(t.test(min_norm_distance ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Treadmill VR'), c(4,6))))$statistic`) improve from the 4th to the 6th block, but so does the real world learning group (`r apa_print(t.test(min_norm_distance ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Real'), c(4,6))))$statistic`).
+
+In the last testing block (block 6), we found no differences among the groups in either error rate (`r apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 6)))$statistic`), nor pointing performance (`r apa_print(aov(abs_error ~ learning.condition, data = sub_sop_all[sub_sop_all$phase == 2,]))$statistic$learning_condition`).
+However we still see small difference in distance performance (`r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 6)))$statistic`), with Treadmill VR group (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Real", 6)$min_norm_distance)`) performing slightly worse than the real world learning group (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Real", 6)$min_norm_distance)`), `r apa_print(t.test(min_norm_distance ~ learning.condition, only_conditions(sub_walk_all, c('Treadmill VR', 'Real'))))$statistic`.
+
+## Could effects be driven by difficulty with the immersive interface?
+
+To attempt to address this issue, we compared the three different modalities, navigating in the real-world (real-world), navigating on the treadmill with the head-mounted display (immersed), and navigating with a joystick in desktop VR (impoverished) on the first block of the experiment.
+Comparing normalized distances, there were no difference between (`r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 1)))$statistic$learning_condition`), nor in visitation errors (`r apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 1)))$statistic$learning_condition`).
+This suggests that participants were readily able to navigate in the different modalities regardless of the interface.
+We did, however, find a significant difference in normalized time (`r apa_print(aov(min_norm_time ~ learning.condition, only_blocks(sub_walk_all, 1)))$statistic$learning_condition`).
+This effect appeared to be driven largely by the fact that treadmill VR participants took to get used to the interface and therefore having longer trial times at the start.
+
+## Experiment 1b: Comparing transfer from desktop to real and treadmill to desktop
+
+Based on our findings above, we might predict that transferring knowledge from desktop to the real world was most difficult for subjects, with the immersive group somewhere in between.
+We also found, that despite the groups differ at the end of the learning phase, their performance is the same after three blocks in the transfer phase.
+It is unclear, whether this equalisation of performance is solely due to exposure to the real world, and wihout it participants could never perform perfectly, or if the tradmill VR modality could achieve the same performance as the real world training, but it only needs more time.
+
+To test this idea, we added two more conditions.
+One group learned the UC Center for Neuroscience building on the treadmill VR while the second group learned in the real world, but both were tested in the tradmill VR in the transfer phase.
+This allowed us to assess if the treadmill VR can achieve the same performance given additional time and also investigate the process of opposite transfer.
+
+### participants

### 2018-12-10 Report/paper-results.Rmd
adds new reports
Updates the reposrt

@@
 ### Improvement from the pre-switch block to the final block
+
 Assuming the performance platoes around the 6the block, looking at participants performance in the pre-switch block and the last block we can deduce what margin of improvement is still possible.
@@
 
-Looking at the linear mixed model prediction for the distance and error improvement in the second phase (post-switch), we can see that both learning condition and interaction with the experiment progression is significant, showing that there are different learning rates for different learning conditions even after the switch to the real world.
+Looking at the linear mixed model prediction for the distance and error improvement in the second phase, we can see that both learning condition and interaction with the experiment progression is significant, showing that there are different learning rates for different learning conditions even after the switch to the real world.
 

### 2018-12-10 TwoWorlds/helpers-report.R

@@ -46,8 +46,8 @@ only_conditions <- function(df, conditions){
   return(df)
 }
 
-only_condition_blocks <- function(df, condition, ...){
-  df <- only_condition(df, condition)
+only_conditions_blocks <- function(df, condition, ...){
+  df <- only_conditions(df, condition)
   df <- only_blocks(df, ...)
   return(df)
 }

### 2018-12-10 TwoWorlds/helpers-results.R

@@ -17,7 +17,7 @@ ez_prepare_block <- function(df, blocks, variable=NULL){
   }
   
   has_all <- filtered_data %>% group_by(id) %>% select(id, "exp_block_id") %>% unique() %>% count
-  has_all <- has_all %>% filter(n == 2)
+  has_all <- has_all %>% filter(n >= 2)
   filtered_data <- filtered_data %>% filter(id %in% has_all$id)
   filtered_data$exp_block_id <- factor(filtered_data$exp_block_id)
   return(filtered_data)

### 2018-12-10 Report/descriptives.Rmd
Updates the reports

--- 2018-10-15 Report/descriptives.Rmd
@@
 df_participants <- df_participants %>% mutate(condition=paste0(First.phase,"-",Second.phase))
+df_participants <- df_participants %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
 overview <- settings$versions
 overview <- overview %>% mutate(condition=paste0(First.phase,"-",Second.phase))
+overview <- overview %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
+
+all <- df_participants %>% left_join(overview, by="Code")
 ```
@@
 ## Female/male
+```{r, echo=F}
+kable(df_participants %>% right_join(overview, by='Code') %>% filter(is_ok=="yes") %>% group_by(condition, sex) %>% count)
+```
 
-## Age
+# Report
+A total of `r sum(overview$finished == "yes", na.rm=T)` undergraduate students at UC Davis particiapted in the study in exchange for a course credit.
+Each participant was randomly assigned a condition and a randomized set of goals before arrival.
+`r sum(finished$no)` participants didn't finish due to motion sickness and `r sum(is_ok$no)` were removed due to a technical failure of the real world tracking systen.
+
+Eighteen participants didn't finish due to motion sickness and three were removed due to a technical problems with the real world tracking system.
+This resulted in a final sample size of `r sum(overview$is_ok == "yes", na.rm=T)` participants (M = `r round(mean(df_participants[df_participants$is_ok == "yes",]$age, na.rm=T),1)`, SD = `r round(sd(df_participants[df_participants$finished == "yes",]$age, na.rm=T), 1)`) () with a mean age of ## used for all reported analyses.All procedures were approved by the UC Davis Institutional Review Board (IRB    ).
 
 ## Descriptives for the VR/real/desktop - time and path traveled
+
+```{r}
+df_desc <- walk_all %>% group_by(type, exp_block_id) %>% summarize(mean_dist = mean(distance, na.rm=T), mean_time = mean(time, na.rm=T))
+ggplot(df_desc, aes(exp_block_id, mean_dist, fill=type)) + geom_bar(stat="identity", position = "dodge")
+ggplot(df_desc, aes(exp_block_id, mean_time, fill=type)) + geom_bar(stat="identity", position = "dodge")
+```

### 2018-12-10 Report/final-results.Rmd
Updates the reports

@@
 
-#settings <- load_google_sheets()
-#df_participants <- settings$participants
-#df_participants <- df_participants %>% mutate(condition=paste0(First.phase,"-",Second.phase))
-#df_participants <- df_participants %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
-#overview <- settings$versions
-#overview <- overview %>% mutate(condition=paste0(First.phase,"-",Second.phase))
-#overview <- overview %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
+settings <- load_google_sheets()
+df_participants <- settings$participants
+df_participants <- df_participants %>% mutate(condition=paste0(First.phase,"-",Second.phase))
+df_participants <- df_participants %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
+overview <- settings$versions
+overview <- overview %>% mutate(condition=paste0(First.phase,"-",Second.phase))
+overview <- overview %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
 
@@
 
-knitr::opts_chunk$set(echo=F, warning = F, message = F)
-
 sub_walk_all_factor <- sub_walk_all
@@
 sub_walk_all_factor$learning.condition <- factor(sub_walk_all_factor$learning.condition)
+
+knitr::opts_chunk$set(echo=F, warning = F, message = F)
 ```
+
+## Participants
+```{r}
+# need revising 
+# is_ok == data well measured
+# finished == did finsih, regardless of potentially missing real world data
+finished <- overview %>% group_by(condition) %>% count(finished)
+finished <- dcast(finished, condition~finished)
+finished$percent <- round(finished$yes/(finished$yes+finished$no), 2)
+is_ok <- overview %>% filter(finished=="yes") %>% group_by(condition) %>% count(is_ok)
+is_ok <- dcast(is_ok, condition~is_ok)
+is_ok$percent <- round(is_ok$yes/(is_ok$yes+is_ok$no), 2)
+```
+A total of `r sum(overview$is_ok == "yes", na.rm=T) + sum(finished$no)` undergraduate students at UC Davis (M = `r round(mean(df_participants[df_participants$finished == "yes",]$age, na.rm=T),1)`, SD = `r round(sd(df_participants[df_participants$finished == "yes",]$age, na.rm=T), 1)`) particiapted in the study in exchange for a study credit.
+`r sum(finished$no)` participants didn't finish due to motion sickness and `r sum(is_ok$no)` were removed due to a technical failure of the real world tracking systen.
+Only `r finished[finished$condition == "vr-real","percent"] * 100` percent the students being able to finish Treadmill VR learning condition.
 
@@
 Looking at the improvement in visitations errors from block 4 to block 6 in separate learning conditions using t-tests, we can assess that only Desktop group (`r apa_print(t.test(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Desktop'), c(4,6))))$statistic`) and Treadmill VR group (`r apa_print(t.test(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Treadmill VR'), c(4,6))))$statistic`) have improved, whereas group learned in the real world did not improve any more in the transfer phase (`r apa_print(t.test(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Real'), c(4,6))))$statistic`).
-In fact, the real world group retains its visitation errors performance even from the last learning phase block does not significantly improve in visitation errors until the end (`r apa_print(t.test(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Real'), c(3,6))))$statistic`).
 
@@
 ## Experiment 1b: Comparing transfer from desktop to real and treadmill to desktop
+```{r}
+### Try running it without outliers
+sub_walk_all_2[!is.na(sub_walk_all_2$norm_distance) & sub_walk_all_2$norm_distance > 3, c("distance", "min_norm_distance")] <- NA
+
+sub_walk_all_2_factor <- sub_walk_all_2
+sub_walk_all_2_factor$exp_block_id <- factor(sub_walk_all_2_factor$exp_block_id)
+sub_walk_all_2_factor$id <- factor(sub_walk_all_2_factor$id)
+sub_walk_all_2_factor$learning.condition <- factor(sub_walk_all_2_factor$learning.condition)
+```
 
@@
 
-### participants
+### Participants
+```{r}
+df_participants <- settings$participants
+df_participants <- df_participants %>% mutate(condition=paste0(First.phase,"-",Second.phase))
+df_participants <- df_participants %>% filter(condition %in% c("real-vr", "vr-vr"))
+overview <- settings$versions
+overview <- overview %>% mutate(condition=paste0(First.phase,"-",Second.phase))
+overview <- overview %>% filter(condition %in% c("real-vr", "vr-vr"))
+all_participant <- df_participants %>% left_join(overview, by="Code")
+
+finished <- overview %>% group_by(condition) %>% count(finished)
+finished <- dcast(finished, condition~finished)
+finished$percent <- round(finished$yes/(finished$yes+finished$no), 2)
+is_ok <- overview %>% filter(finished=="yes") %>% group_by(condition) %>% count(is_ok)
+is_ok <- dcast(is_ok, condition~is_ok)
+is_ok$percent <- round(is_ok$yes/(is_ok$yes+is_ok$no), 2)
+```
+A total of `r sum(overview$is_ok == "yes", na.rm=T) + sum(finished$no)` undergraduate students at UC Davis particiapted in the second experiment.
+`r sum(finished$no)` participants didn't finish due to motion sickness and `r sum(is_ok$no, na.rm=T)` were removed due to a technical failure of the real world tracking system.
+Analyses were then conducted on a final set of `r sum(overview$is_ok == "yes", na.rm=T) ` (14 female) with mean age of `r round(mean(all_participant[all_participant$is_ok == "yes",]$age, na.rm=T),1)` (SD = `r round(sd(all_participant[all_participant$is_ok == "yes",]$age, na.rm=T), 1)`).
+
+```{r}
+kable(finished)
+all_participant %>% filter(is_ok=="yes") %>% group_by(sex) %>% count(sex)
+```
+
+## Quantifying performance in the last block
+First we assessed difference between the two learning conditions in the last block performance usint t-tests.
+The conditions did not differ in the walking performance (`r apa_print(t.test(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all_2, c(6))))$statistic`) nor visiting errors (`r apa_print(t.test(errors ~ learning.condition, only_blocks(sub_walk_all_2, c(6))))$statistic`), therefore showing that both treadmill VR and real world learning leads to the same performance in the VR.
+
+## Comparing to previous conditions
+Running a one-way ANOVA togeher together with other conditions from the first experiment (real to real, Vr - real and desktop to real), shows a significant effect of condition on final walking performance (`r apa_print(aov(min_norm_distance ~ condition, only_blocks(walk_all, c(6))))$statistic`), but Tukey post hoc tests failed to find any significant differences due to correction for multiple comparisons.
+Looking at the final walking performance in the Table XXX, we can see that the differences between conditions are truly negligible.
+Interestingly, we found no significant difference between the conditions when we analysed absolute distance, rather than normalised (`r apa_print(aov(distance ~ condition, only_blocks(walk_all, c(6))))$statistic`).
+This means that all aprticipants walked the same distance in the end on average, regardless of them being tested on a treadmill or in the real building.
+We also found no difference between any of the conditions in their final pointing performance, `r apa_print(aov(abs_error ~ condition, sop_all[sop_all$phase==2,]))$statistic`.
+
+```{r}
+df <- only_blocks(walk_all, c(6)) %>% group_by(condition) %>% summarize(dist = mean(min_norm_distance, na.rm=T), errors= mean(errors, na.rm=T))
+kable(df, caption='Average performance in all conditions in the last block in visitation errors and walking disntace', digits=2)
+```
+
+A different story is in comparing the visitation errors across conditions in the last block, where one-way ANOVA shows significant difference (`r apa_print(aov(min_norm_distance ~ condition, only_blocks(walk_all, c(6))))$statistic`) and Tukey's post hoc tests demonstrates significant differences between all conditions that differs in transfer phase modality, e.g.
+participants in the real world always perform better than participants on the treadmill VR.
+Looking at the table XXX, we can clearly see that conditions with transfer phase happening in VR, despite having comparable walked distance, have higher number of visitation errors.
+We think that this may have emerged from two different sources: firstly, it can point to higher difficulty in recognizing proper doors because of lesser visual accuracy in the VR, where all doors look similar unlike real doors which may demonstrate unique signs of wear or use.
+The second reason for the increased number of visitation errors in VR conditions can be explained by the lesser "social impact" of wrong choice.
+Whereas in the real conditions, participants had to actively inquire the experimenter, there was technically no penalty for "trying out" wrong door in the VR.
+
+```{r}
+ggplot(walk_all, aes(x = exp_block_id, y = min_norm_distance, color = factor(condition))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
+  geom_vline(xintercept = 3.5, linetype=4) +
+  ylab("Average normalised trial distance") + xlab("Block") + 
+  scale_color_manual(values=estimote_pal,"Condition") + guides(fill=F)
+
+ggplot(walk_all, aes(x = exp_block_id, y = errors, color = factor(condition))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
+  geom_vline(xintercept = 3.5, linetype=4) +
+  ylab("Average visitation errors") + xlab("Block") + 
+  scale_color_manual(values=estimote_pal,"Condition") + guides(fill=F)
+```
+
+### Quantifying transfer
+```{r}
+ez_data <- ez_prepare_block(sub_walk_all_2_factor, c(3,4), 'min_norm_distance')
+fit_mm_dist_block <- lmer(min_norm_distance ~ learning.condition*exp_block_id + (1 | id/exp_block_id), data = only_blocks(sub_walk_all, 3:4))
+ez_fit_distance <- ezANOVA(
+        ez_data, 
+        dv = min_norm_distance,
+        wid = id,
+        within = exp_block_id,
+        between = learning.condition,
+        return_aov = T,
+        type = 3)
+
+ez_data <- ez_prepare_block(sub_walk_all_factor, c(3,4), 'errors')
+ez_fit_errors <- ezANOVA(
+        ez_data, 
+        dv = errors,
+        wid = id,
+        within = exp_block_id,
+        between = learning.condition,
+        return_aov = T,
+        type = 3)
+```
+
+Assessing the effects of trasfering from the real world to the Treadmill VR, we ran 2 (Learning condition: real vs treadmill VR) x 2 (block 3 vs blcok 4) ANOVA on data for distance walked and visitation errors.
+There was no effect of block `r apa_print(ez_fit_distance$aov)$statistic$learning_condition` nor learning condition (`r apa_print(ez_fit_distance$aov)$statistic$learning_condition`) on distance walked and we observed no interaction effects (`r apa_print(ez_fit_distance$aov)$statistic$learning_condition_exp_block_id`).
+But we found significant interaction effect between block and learning condition in visitation errors, `r apa_print(ez_fit_errors$aov)$statistic$learning_condition_exp_block_id`.
+
+Pairwise t-tests comparing change in error rate from 3rd to 4th block show, that the group learning on the treadmill VR does show significant difference (`r apa_print(t.test(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all_2, 'Treadmill VR', 3:4)))$statistic`) as it effectively didn't change the environment, but the group switching to the VR from the real world demonstrates increase in visitations errors from block 3 (`r m_sd_report(only_conditions_blocks(sub_walk_all_2, "Real", 3)$errors)`) to block 4 (`r m_sd_report(only_conditions_blocks(sub_walk_all_2, "Real", 4)$errors)`), `r apa_print(t.test(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all_2, 'Real', 3:4)))$statistic`.
+
+### Summary
+In the last block, all conditios perform similarly as far as the walking distance goes, regardless of them being tested on the VR or in the real building.
+VR leads to overall higher error rates, but same walking performance, even when considering absolute distances.
+There are no significant differences in pointing performance between conditions, regardless of pointing being tested in the real world or in the VR.

### 2018-12-10 Report/paper-all-data.Rmd
Updates the reports

@@
   geom_vline(xintercept = 3.5, linetype=4) +
-  ylab("Average normalised trial distance") + xlab("Block") + 
+  ylab("Average visitation errors") + xlab("Block") + 
   scale_color_manual(values=estimote_pal,"Condition") + guides(fill=F)

### 2018-12-10 Scripts/paper-prepare.R

@@ -13,3 +13,6 @@ sop_all$testing.condition <- dplyr::recode(sop_all$testing.condition, "real-real
 
 sub_walk_all <- walk_all %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
 sub_sop_all <- sop_all %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
+
+sub_walk_all_2 <- walk_all %>% filter(condition %in% c("vr-vr", "real-vr"))
+sub_sop_all_2 <- sop_all %>% filter(condition %in% c("vr-vr", "real-vr"))

### 2018-12-11 Report/final-results.Rmd
Small change to the talbe output

@@
 ```{r}
-df <- only_blocks(walk_all, c(6)) %>% group_by(condition) %>% summarize(dist = mean(min_norm_distance, na.rm=T), errors= mean(errors, na.rm=T))
+df <- only_blocks(walk_all, c(6)) %>% group_by(condition) %>% summarize(dist =m_sd_report(min_norm_distance), errors= m_sd_report(errors))
 kable(df, caption='Average performance in all conditions in the last block in visitation errors and walking disntace', digits=2)

### 2019-01-12 Report/paper-all-data.Rmd
updates paper all data

@@
 library(restimoter)
+library(papaja)
+library(googlesheets)
 
@@
 
+source("helpers/block-t-test-improvement.R")
+source("helpers/block-t-test.R")
+source("helpers/block-measure.R")
+source("helpers/block-correlation.R")
+
+settings <- load_google_sheets()
+df_participants <- settings$participants
+df_participants <- df_participants %>% mutate(condition=paste0(First.phase,"-",Second.phase))
+df_participants <- df_participants %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
+overview <- settings$versions
+overview <- overview %>% mutate(condition=paste0(First.phase,"-",Second.phase))
+overview <- overview %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
+
 #green, violet, blue, yellow, pink, dark blue
 estimote_pal <- c('#b6d6c1', '#54003d','#93d5f6','#d5d215','#f3aacb','#2d2556')
-estimote_pal_mono <- function(n){return(rep(estimote_pal[n],6))}
 theme_set(theme_minimal())
+
+### Try running it without outliers
+sub_walk_all[!is.na(sub_walk_all$norm_distance) & sub_walk_all$norm_distance > 3, c("distance", "min_norm_distance")] <- NA
+
+condition_all_sums <- sub_walk_all %>% 
+  group_by(exp_block_id, id, learning.condition) %>% 
+  summarise(sum.distance = sum(min_norm_distance, na.rm = T), 
+            sum.errors = sum(errors, na.rm = T),
+            mean.distance = mean(min_norm_distance, na.rm = T),
+            mean.error = mean(errors, na.rm = T))
+
+df_errors <- block_measure(condition_all_sums, "sum.errors")
+df_distance <- block_measure(condition_all_sums, "mean.distance")
+
+df_wide_errors <- dcast(condition_all_sums, learning.condition + id ~ exp_block_id, value.var = "sum.errors")
+df_wide_distance <- dcast(condition_all_sums, learning.condition + id ~ exp_block_id, value.var = "mean.distance")
+colnames(df_wide_errors) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")
+colnames(df_wide_distance) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")
+df_wide_errors$block34 <- (df_wide_errors$block3-df_wide_errors$block4)/(df_wide_errors$block3+df_wide_errors$block4)
+df_wide_distance$block34 <- (df_wide_distance$block3-df_wide_distance$block4)/(df_wide_distance$block3+df_wide_distance$block4)
+
+sub_walk_all_factor <- sub_walk_all
+sub_walk_all_factor$exp_block_id <- factor(sub_walk_all_factor$exp_block_id)
+sub_walk_all_factor$id <- factor(sub_walk_all_factor$id)
+sub_walk_all_factor$learning.condition <- factor(sub_walk_all_factor$learning.condition)
+
+knitr::opts_chunk$set(echo=F, warning = F, message = F)
 ```
@@
 }
+```
+
+### Block 1-4
+Paired t tests between min normalised distance difference in blocks 1 and 4.
+```{r, echo=F}
+kable(block_t_test(sub_walk_all, 3, 1,"min_norm_distance"))
+```
+And for number of errors
+```{r, echo=F}
+kable(block_t_test(sub_walk_all, 3, 1,"errors"))
 ```

### 2019-01-12 Report/final-results.Rmd
Updates results

@@
 source("helpers/block-measure.R")
+source("../TwoWorlds/helpers-report.R")
 
@@
 There were main effects of learning condition for both walked distance, `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 3)))$statistic`, and visitation errors, `r  apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 3)))$statistic`).
-Comparing the conditions closer, we find that there is a significant difference in visitations error performance between real world and desktop learning (`r apa_print(t.test(errors ~ learning.condition, only_conditions(sub_walk_all, c('Real', 'Desktop'))))$statistic`) and real world nad treadmill VR learning (`r apa_print(t.test(errors ~ learning.condition, only_conditions(sub_walk_all, c('Real', 'Treadmill VR'))))$statistic`), but no difference between Desktop and Treadmill learning (`r apa_print(t.test(errors ~ learning.condition, only_conditions(sub_walk_all, c('Desktop', 'Treadmill VR'))))$statistic`).
+Comparing the conditions closer, we find that there is a significant difference in visitations error performance between real world and desktop learning (`r apa_print(t.test(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Desktop'), c(1,3))))$statistic`) and real world nad treadmill VR learning (`r apa_print(t.test(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Treadmill VR'), c(1,3))))$statistic`), but no difference between Desktop and Treadmill learning (`r apa_print(t.test(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Desktop', 'Treadmill VR'), c(1,3))))$statistic`).
 The situation is the same for distance comparisons.

### 2019-01-12 Report/paper-all-data.Rmd
Updates results

@@
 
-### Block 1-4
+### Block 1-3
 Paired t tests between min normalised distance difference in blocks 1 and 4.
@@
 ```
+
 And for number of errors

### 2019-01-13 Report/final-results.Rmd
Updates final report

@@
 There were main effects of learning condition for both walked distance, `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 3)))$statistic`, and visitation errors, `r  apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 3)))$statistic`).
-Comparing the conditions closer, we find that there is a significant difference in visitations error performance between real world and desktop learning (`r apa_print(t.test(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Desktop'), c(1,3))))$statistic`) and real world nad treadmill VR learning (`r apa_print(t.test(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Treadmill VR'), c(1,3))))$statistic`), but no difference between Desktop and Treadmill learning (`r apa_print(t.test(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Desktop', 'Treadmill VR'), c(1,3))))$statistic`).
+Comparing the conditions closer, we find that there is a significant difference in visitations error performance between real world and desktop learning (`r apa_print(t.test(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Desktop'), 3)))$statistic`) and real world nad treadmill VR learning (`r apa_print(t.test(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Treadmill VR'), 3)))$statistic`), but no difference between Desktop and Treadmill learning (`r apa_print(t.test(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Desktop', 'Treadmill VR'), 3)))$statistic`).
 The situation is the same for distance comparisons.
+real world and desktop learning (`r apa_print(t.test(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Desktop'), 3)))$statistic`) and real world nad treadmill VR learning (`r apa_print(t.test(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Treadmill VR'), 3)))$statistic`), but no difference between Desktop and Treadmill learning (`r apa_print(t.test(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Desktop', 'Treadmill VR'), 3)))$statistic`).
 
@@
 Overall, these findings suggested that participants learned the building the best as a result of real-world navigation, followed by the immersive condition, with the desktop condition showing the slowest acquisition.
+
+```{r}
+# Pre transfer plot of normalised distances
+make_graph(only_blocks(sub_walk_all, 1:3), "learning.condition", "min_norm_distance", "exp_block_id") + 
+  ylab("Average normalised trial distance") + xlab("Learning condition") + 
+  scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F)
+
+make_graph(only_blocks(sub_walk_all, 1:3), "learning.condition", "errors", "exp_block_id") + 
+  ylab("Average normalised trial distance") + xlab("Learning condition") + 
+  scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F)
+
+ggplot(point_data_1, aes(x = learning.condition, y = abs_error, fill = learning.condition)) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
+  ylab("Average pointing error in degrees after learning phase") + xlab("Learning condition") + 
+  scale_fill_manual(values = estimote_pal, "Learning condition")  + guides(fill=F)
+
+```
 
@@
 In the last testing block (block 6), we found no differences among the groups in either error rate (`r apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 6)))$statistic`), nor pointing performance (`r apa_print(aov(abs_error ~ learning.condition, data = sub_sop_all[sub_sop_all$phase == 2,]))$statistic$learning_condition`).
-However we still see small difference in distance performance (`r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 6)))$statistic`), with Treadmill VR group (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Real", 6)$min_norm_distance)`) performing slightly worse than the real world learning group (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Real", 6)$min_norm_distance)`), `r apa_print(t.test(min_norm_distance ~ learning.condition, only_conditions(sub_walk_all, c('Treadmill VR', 'Real'))))$statistic`.
+However we still see small difference in distance performance (`r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 6)))$statistic`), with real group (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Real", 6)$min_norm_distance)`) performing better than Treadmill VR group (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Treadmill VR", 6)$min_norm_distance)`)(`r apa_print(t.test(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Treadmill VR', 'Real'), 6)))$statistic`) and desktop group (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Desktop", 6)$min_norm_distance)`)(`r apa_print(t.test(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Desktop'), 6)))$statistic`).
 
@@
 
-A different story is in comparing the visitation errors across conditions in the last block, where one-way ANOVA shows significant difference (`r apa_print(aov(min_norm_distance ~ condition, only_blocks(walk_all, c(6))))$statistic`) and Tukey's post hoc tests demonstrates significant differences between all conditions that differs in transfer phase modality, e.g.
+A different story is in comparing the visitation errors across conditions in the last block, where one-way ANOVA shows significant difference (`r apa_print(aov(errors ~ condition, only_blocks(walk_all, c(6))))$statistic`) and Tukey's post hoc tests demonstrates significant differences between all conditions that differs in transfer phase modality, e.g.
 participants in the real world always perform better than participants on the treadmill VR.
@@
 Assessing the effects of trasfering from the real world to the Treadmill VR, we ran 2 (Learning condition: real vs treadmill VR) x 2 (block 3 vs blcok 4) ANOVA on data for distance walked and visitation errors.
-There was no effect of block `r apa_print(ez_fit_distance$aov)$statistic$learning_condition` nor learning condition (`r apa_print(ez_fit_distance$aov)$statistic$learning_condition`) on distance walked and we observed no interaction effects (`r apa_print(ez_fit_distance$aov)$statistic$learning_condition_exp_block_id`).
+There was no effect of block `r apa_print(ez_fit_distance$aov)$statistic$exp_block_id` nor learning condition (`r apa_print(ez_fit_distance$aov)$statistic$learning_condition`) on distance walked and we observed no interaction effects (`r apa_print(ez_fit_distance$aov)$statistic$learning_condition_exp_block_id`).
 But we found significant interaction effect between block and learning condition in visitation errors, `r apa_print(ez_fit_errors$aov)$statistic$learning_condition_exp_block_id`.
@@
 Pairwise t-tests comparing change in error rate from 3rd to 4th block show, that the group learning on the treadmill VR does show significant difference (`r apa_print(t.test(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all_2, 'Treadmill VR', 3:4)))$statistic`) as it effectively didn't change the environment, but the group switching to the VR from the real world demonstrates increase in visitations errors from block 3 (`r m_sd_report(only_conditions_blocks(sub_walk_all_2, "Real", 3)$errors)`) to block 4 (`r m_sd_report(only_conditions_blocks(sub_walk_all_2, "Real", 4)$errors)`), `r apa_print(t.test(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all_2, 'Real', 3:4)))$statistic`.
+
+```{r}
+
+walk_all_filtered <- rbind(sub_walk_all, sub_walk_all_2)
+ggplot(only_blocks(walk_all_filtered, 6), aes(x = condition, y = min_norm_distance, fill = learning.condition)) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
+  ylab("Average path distance in the last block of the transfer phase") + xlab("Condition") + 
+  scale_fill_manual(values = estimote_pal, "Condition")  + guides(fill=F)
+
+ggplot(only_blocks(walk_all_filtered, 6), aes(x = condition, y = errors, fill = learning.condition)) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
+  ylab("Number of visiting errors in the last block of the transfer phase") + xlab("Condition") + 
+  scale_fill_manual(values = estimote_pal, "Condition")  + guides(fill=F)
+```
 

### 2019-02-05 Report/final-results.Rmd
updated results

@@
 sub_walk_all[!is.na(sub_walk_all$norm_distance) & sub_walk_all$norm_distance > 3, c("distance", "min_norm_distance")] <- NA
+walk_all[!is.na(walk_all$norm_distance) & walk_all$norm_distance > 3, c("distance", "min_norm_distance")] <- NA
 
@@
 ## Comparing to previous conditions
-Running a one-way ANOVA togeher together with other conditions from the first experiment (real to real, Vr - real and desktop to real), shows a significant effect of condition on final walking performance (`r apa_print(aov(min_norm_distance ~ condition, only_blocks(walk_all, c(6))))$statistic`), but Tukey post hoc tests failed to find any significant differences due to correction for multiple comparisons.
+Running a one-way ANOVA togeher together with other conditions from the first experiment (real to real, Vr - real and desktop to real), shows a significant effect of condition on final walking performance (`r apa_print(aov(min_norm_distance ~ condition, only_blocks(walk_all, c(6))))$statistic`).
+
+```{r}
+kable(tukey_report_table(TukeyHSD(aov(min_norm_distance ~ condition, only_blocks(walk_all, c(6))))))
+```
+
 Looking at the final walking performance in the Table XXX, we can see that the differences between conditions are truly negligible.
@@
 ```{r}
-df <- only_blocks(walk_all, c(6)) %>% group_by(condition) %>% summarize(dist =m_sd_report(min_norm_distance), errors= m_sd_report(errors))
+df <- only_blocks(walk_all, c(6)) %>% group_by(condition) %>% summarize(dist = m_sd_report(min_norm_distance), errors= m_sd_report(errors))
 kable(df, caption='Average performance in all conditions in the last block in visitation errors and walking disntace', digits=2)

### 2019-02-06 Report/final-results.Rmd
Updated results

@@
 ```{r}
-
-walk_all_filtered <- rbind(sub_walk_all, sub_walk_all_2)
-ggplot(only_blocks(walk_all_filtered, 6), aes(x = condition, y = min_norm_distance, fill = learning.condition)) +
+ggplot(only_blocks(walk_all, 6), aes(x = condition, y = min_norm_distance, fill = learning.condition)) +
   stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
@@
 
-ggplot(only_blocks(walk_all_filtered, 6), aes(x = condition, y = errors, fill = learning.condition)) +
+ggplot(only_blocks(walk_all, 6), aes(x = condition, y = errors, fill = learning.condition)) +
   stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 

### 2019-02-08 Report/final-results.Rmd
Updates the graphs

@@
 library(ggplot2)
+library(gridExtra)
 library(ez)
@@
 estimote_pal <- c('#b6d6c1', '#54003d','#93d5f6','#d5d215','#f3aacb','#2d2556')
-theme_set(theme_minimal())
+theme_set(theme_minimal() + theme(axis.text=element_text(size=14),
+        axis.title=element_text(size=16)))
 
@@
 df_distance <- block_measure(condition_all_sums, "mean.distance")
+df_distance$graph.learning.condition <- dplyr::recode(df_distance$learning.condition, "Real" = "Real-world", "Treadmill VR" = "Immersive VR")
+df_errors$graph.learning.condition <- dplyr::recode(df_errors$learning.condition, "Real" = "Real-world", "Treadmill VR" = "Immersive VR")
 
@@
 colnames(df_wide_distance) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")
+df_wide_distance$graph.learning.condition <- dplyr::recode(df_wide_distance$learning.condition, "Real" = "Real-world", "Treadmill VR" = "Immersive VR")
+df_wide_errors$graph.learning.condition <- dplyr::recode(df_wide_errors$learning.condition, "Real" = "Real-world", "Treadmill VR" = "Immersive VR")
+
 df_wide_errors$block34 <- (df_wide_errors$block3-df_wide_errors$block4)/(df_wide_errors$block3+df_wide_errors$block4)
@@
 # Pre transfer plot of normalised distances
-make_graph(only_blocks(sub_walk_all, 1:3), "learning.condition", "min_norm_distance", "exp_block_id") + 
+ggplot(only_blocks(sub_walk_all_factor, 1:3), aes(x = graph.exp_block_id, y = min_norm_distance, fill = graph.exp_block_id)) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
   ylab("Average normalised trial distance") + xlab("Learning condition") + 
-  scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F)
+  scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F) +
+  facet_wrap(~graph.learning.condition, strip.position = "bottom", scales = "free_x") + 
+  theme(panel.spacing = unit(0, "lines"), 
+         strip.background = element_blank(),
+         strip.placement = "outside")
 
-make_graph(only_blocks(sub_walk_all, 1:3), "learning.condition", "errors", "exp_block_id") + 
-  ylab("Average normalised trial distance") + xlab("Learning condition") + 
-  scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F)
+ggplot(only_blocks(sub_walk_all_factor, 1:3), aes(x = graph.exp_block_id, y = errors, fill = graph.exp_block_id)) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
+  ylab("Average visitation errors") + xlab("Learning condition") + 
+  scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F) +
+  facet_wrap(~graph.learning.condition, strip.position = "bottom", scales = "free_x") + 
+  theme(panel.spacing = unit(0, "lines"), 
+         strip.background = element_blank(),
+         strip.placement = "outside")
 
-ggplot(point_data_1, aes(x = learning.condition, y = abs_error, fill = learning.condition)) +
+ggplot(point_data_1, aes(x = graph.learning.condition, y = abs_error, fill = graph.learning.condition)) +
   stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
@@
 ```{r, echo=F}
-ggplot(df_distance, aes(x = learning.condition, y = block34.diff, fill = learning.condition)) + 
+ggplot(df_distance, aes(x = graph.learning.condition, y = block34.diff, fill = graph.learning.condition)) + 
   geom_bar(stat = 'identity', position="dodge") + 
@@
   scale_fill_manual(values = estimote_pal) +
-  ylab("Mean normalised distance improvement from block 3 to 4") + xlab("Learning condition")  + guides(fill=F)
+  ylab("Level of transfer for normalised distance") + 
+  xlab("Learning condition") + guides(fill=F) + 
+  theme(axis.title.y = element_text(size = rel(0.9)))
 
-ggplot(df_errors, aes(x = learning.condition, y = block34.diff, fill = learning.condition)) + 
+ggplot(df_errors, aes(x = graph.learning.condition, y = block34.diff, fill = graph.learning.condition)) + 
   geom_bar(stat = 'identity', position="dodge") + 
@@
   scale_fill_manual(values = estimote_pal) +
-  ylab("Mean sum of errors improvement after switch")+ xlab("Learning condition") + guides(fill=F)
+  ylab("Level of transfer for visitation errors")+ xlab("Learning condition") + 
+  guides(fill=F) + theme(axis.title.y = element_text(size = rel(0.9)))
 ```
@@
 
-```{r}
-ggplot(walk_all, aes(x = exp_block_id, y = min_norm_distance, color = factor(condition))) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
-  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
+```{r, fig.width=10, fig.height=8}
+ggplot(walk_all, aes(x = exp_block_id, y = min_norm_distance, group = factor(condition))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar", aes(color=graph.learning.condition, linetype=testing.condition)) + 
+  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line", aes(color=graph.learning.condition, linetype=testing.condition)) + 
   geom_vline(xintercept = 3.5, linetype=4) + 
   ylab("Average normalised trial distance") + xlab("Block") + 
-  scale_color_manual(values=estimote_pal,"Condition") + guides(fill=F)
+  scale_color_manual(values=estimote_pal, "Learning condition") + 
+  scale_linetype_manual(values=c("dashed", "solid"), "Transfer condition") + 
+  guides(fill=F)
 
 ggplot(walk_all, aes(x = exp_block_id, y = errors, color = factor(condition))) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
-  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
+  stat_summary(fun.data=mean_cl_normal, position=position_dodge(width=0.1), width = 0.2, geom="errorbar", aes(color=graph.learning.condition, linetype=testing.condition)) + 
+  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line", aes(color=graph.learning.condition, linetype=testing.condition)) + 
   geom_vline(xintercept = 3.5, linetype=4) +
-  ylab("Average visitation errors") + xlab("Block") + 
-  scale_color_manual(values=estimote_pal,"Condition") + guides(fill=F)
+  ylab("Average number of visitation errors") + xlab("Block") + 
+  scale_color_manual(values=estimote_pal, "Learning condition") + 
+  scale_linetype_manual(values=c("dashed", "solid"), "Transfer condition") + 
+  guides(fill=F)
 ```
@@
 
-```{r}
-ggplot(only_blocks(walk_all, 6), aes(x = condition, y = min_norm_distance, fill = learning.condition)) +
+```{r, fig.width=10, fig.height=8}
+ggplot(only_blocks(walk_all, 6), aes(x = testing.condition, y = min_norm_distance, fill = graph.learning.condition)) +
   stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
   stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
-  ylab("Average path distance in the last block of the transfer phase") + xlab("Condition") + 
-  scale_fill_manual(values = estimote_pal, "Condition")  + guides(fill=F)
+  ylab("Average path distance in the last block of the transfer phase") + xlab("Transfer condition") + 
+  scale_fill_manual(values = estimote_pal, "Condition")  + guides(fill=F) + 
+  facet_wrap(~graph.learning.condition, strip.position = "top", scales = "free_x") + 
+  ggtitle("Learning condition")+
+  theme(panel.spacing = unit(0, "lines"), 
+         strip.background = element_blank(),
+         strip.placement = "outside",
+         strip.text = element_text(size = 16),
+          plot.title = element_text(hjust = 0.5))
 
-ggplot(only_blocks(walk_all, 6), aes(x = condition, y = errors, fill = learning.condition)) +
+ggplot(only_blocks(walk_all, 6), aes(x = testing.condition, y = errors, fill = graph.learning.condition)) +
   stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
   stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
-  ylab("Number of visiting errors in the last block of the transfer phase") + xlab("Condition") + 
-  scale_fill_manual(values = estimote_pal, "Condition")  + guides(fill=F)
+  ylab("Average number of visitation errors in the last block of the transfer phase") + xlab("Transfer condition") + 
+  scale_fill_manual(values = estimote_pal, "Condition")  + guides(fill=F) + 
+  facet_wrap(~graph.learning.condition, strip.position = "top", scales = "free_x") + 
+  ggtitle("Learning condition")+
+  theme(panel.spacing = unit(0, "lines"), 
+         strip.background = element_blank(),
+         strip.placement = "outside",
+         strip.text = element_text(size = 16),
+          plot.title = element_text(hjust = 0.5))
 ```

### 2019-02-08 Scripts/paper-prepare.R

@@ -6,10 +6,17 @@ sop_all <- read.table("../sop.csv", sep=";", header = T, stringsAsFactors = F)
 walk_all <- read.table("../walk.csv", sep=";", header = T, stringsAsFactors = F)
 walk_all$learning.condition <- walk_all$condition
 walk_all$learning.condition <- dplyr::recode(walk_all$learning.condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
+walk_all$graph.condition <- dplyr::recode(walk_all$condition, "real-real" = "Real-world to Real-world", "ve-real" = "Desktop to Real-world", "vr-real" = "Immersive VR to Real-world", "real-vr" = "Real-world to Immersive VR", "vr-vr" = "Immersive VR to Immersive VR")
+walk_all$testing.condition <- dplyr::recode(walk_all$condition, "real-real" = "Real-world", "ve-real" = "Real-world", "vr-real" = "Real-world", "real-vr" = "Immersive VR", "vr-vr" = "Immersive VR")
+
+walk_all$graph.learning.condition <- dplyr::recode(walk_all$learning.condition, "Real" = "Real-world", "Treadmill VR" = "Immersive VR")
+walk_all$graph.exp_block_id <- paste0("Block ", walk_all$exp_block_id)
+
 sop_all$learning.condition <- sop_all$condition
 sop_all$testing.condition <- sop_all$condition
 sop_all$learning.condition <- dplyr::recode(sop_all$learning.condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
 sop_all$testing.condition <- dplyr::recode(sop_all$testing.condition, "real-real" = "Real", "ve-real" = "Real", "vr-real" = "Real", "real-vr" = "Treadmill VR", "vr-vr" = "Treadmill VR")
+sop_all$graph.learning.condition <- dplyr::recode(sop_all$learning.condition, "Real" = "Real-world", "Treadmill VR" = "Immersive VR")
 
 sub_walk_all <- walk_all %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
 sub_sop_all <- sop_all %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))

### 2019-02-08 Report/final-results.Rmd
Updates pointing chart

@@
   stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
-  ylab("Average pointing error in degrees after learning phase") + xlab("Learning condition") + 
-  scale_fill_manual(values = estimote_pal, "Learning condition")  + guides(fill=F)
+  ylab("Average pointing error after learning phase") + xlab("Learning condition") + 
+  scale_fill_manual(values = estimote_pal, "Learning condition") +
+  guides(fill=F) + theme(axis.title.y = element_text(size = rel(0.9)))
 

### 2019-02-15 Report/final-results.Rmd
Updates final report

@@
 is_ok$percent <- round(is_ok$yes/(is_ok$yes+is_ok$no), 2)
+kable(finished)
 ```
-A total of `r sum(overview$is_ok == "yes", na.rm=T) + sum(finished$no)` undergraduate students at UC Davis (M = `r round(mean(df_participants[df_participants$finished == "yes",]$age, na.rm=T),1)`, SD = `r round(sd(df_participants[df_participants$finished == "yes",]$age, na.rm=T), 1)`) particiapted in the study in exchange for a study credit.
+
+A total of `r nrow(overview)` undergraduate students at UC Davis (`r m_sd_report(df_participants[df_participants$finished == "yes",]$age)`) particiapted in the study in exchange for a study credit.
 `r sum(finished$no)` participants didn't finish due to motion sickness and `r sum(is_ok$no)` were removed due to a technical failure of the real world tracking systen.
@@
 
-3x2 repeated measures ANOVA revealed a main effect of block on distance performance, suggesting that participants improved their paths in all three conditions, `r apa_print(ez_fit_distance$aov)$statistic$exp_block_id`.
-The interaction effect with modality was not significant , `r apa_print(ez_fit_distance$aov)$statistic$learning_condition_exp_block_id`, suggesting that learning rate did not differ as a function of modality.
-In contrast, for visitation errors, we again found a main effect of block (`r apa_print(ez_fit_errors$aov)$statistic$exp_block_id`) but, importantly, a block X modality interaction effect, `r apa_print(ez_fit_errors$aov)$statistic$learning_condition_exp_block_id`.
+3x2 repeated measures ANOVA revealed a main effect of block on distance performance `r apa_print(ez_fit_distance$aov)$statistic$exp_block_id`, suggesting that participants improved their paths in all three conditions.
+Although the effect of the learning modality was significant (`r apa_print(ez_fit_distance$aov)$statistic$learning_condition`), the interaction between block and modality was not significant, `r apa_print(ez_fit_distance$aov)$statistic$learning_condition_exp_block_id`, suggesting that learning rate did not differ as a function of modality.
+In contrast, for visitation errors,  we again found a main effect of block (`r apa_print(ez_fit_errors$aov)$statistic$exp_block_id`) adn modality (`r apa_print(ez_fit_errors$aov)$statistic$learning_condition`), but, importantly, a block X modality interaction effect, `r apa_print(ez_fit_errors$aov)$statistic$learning_condition_exp_block_id`.
 This finding suggested that visitation errors improved differently as a function of modality.
@@
 ```
-A total of `r sum(overview$is_ok == "yes", na.rm=T) + sum(finished$no)` undergraduate students at UC Davis particiapted in the second experiment.
-`r sum(finished$no)` participants didn't finish due to motion sickness and `r sum(is_ok$no, na.rm=T)` were removed due to a technical failure of the real world tracking system.
-Analyses were then conducted on a final set of `r sum(overview$is_ok == "yes", na.rm=T) ` (14 female) with mean age of `r round(mean(all_participant[all_participant$is_ok == "yes",]$age, na.rm=T),1)` (SD = `r round(sd(all_participant[all_participant$is_ok == "yes",]$age, na.rm=T), 1)`).
+A total of `r nrow(overview)` undergraduate students at UC Davis particiapted in the second experiment.
+`r sum(finished$no)` participants didn't finish due to motion sickness.
+Analyses were then conducted on a final set of `r sum(overview$is_ok == "yes", na.rm=T)` (14 female) participants(`r m_sd_report(all_participant[all_participant$is_ok == "yes",]$age)`).
 

### 2019-02-15 TwoWorlds/helpers-report.R

@@ -2,8 +2,7 @@ apa_p <- function(values){
   output <- c()
   for (value in values){
     if(value < 0.01){ output <- c(output, "p < .001"); next}
-    if(value < 0.05) { output <- c(output, "p < .0.05"); next}
-    output <- c(output, paste0("p = ", round(value, 2)))
+    output <- c(output, paste0("p = ", round(value, 3)))
   }
   return(output)
 }

### 2019-02-15 TwoWorlds/helpers-results.R

@@ -25,7 +25,7 @@ ez_prepare_block <- function(df, blocks, variable=NULL){
 
 mean.na <- function(data){return(mean(data, na.rm = T))}
 
-m_sd_report <- function(vec, digits=2){
+m_sd_report <- function(vec, digits=3){
   m <- round(mean(vec, na.rm=T), digits)
   s <- round(sd(vec, na.rm=T), digits)
   return(paste("M = ", m, ", SD = ", s, collapse=""))

### 2019-03-20 Report/final-results.Rmd
Adds first revision markdown

@@
 df_participants <- df_participants %>% mutate(condition=paste0(First.phase,"-",Second.phase))
+all_participant_experiment <- df_participants
 df_participants <- df_participants %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
+
 overview <- settings$versions
 overview <- overview %>% mutate(condition=paste0(First.phase,"-",Second.phase))
+all_participant_experiment <- overview %>% left_join(all_participant_experiment, by="Code")
 overview <- overview %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
+overview <- overview %>% left_join(df_participants[, c('Code', 'sex')], by="Code")
 
@@
 kable(finished)
+
+finished_gender <- overview %>% group_by(condition, sex) %>% count(finished)
+finished_gender <- dcast(finished_gender, condition+sex~finished)
+finished_gender$no[is.na(finished_gender$no)] <- 0
+finished_gender$percent <- round(finished_gender$yes/(finished_gender$yes+finished_gender$no), 2)
 ```
@@
 is_ok$percent <- round(is_ok$yes/(is_ok$yes+is_ok$no), 2)
+
 ```

### 2019-03-20 Report/first-revision.Rmd
Adds first revision markdown

--- 2019-03-20 Report/final-results.Rmd
@@
 ---
-title: "Transfer report"
+title: "1st Revision"
 author: "Lukáš Hejtmánek"
-date: "9 August 2018"
-output: github_document
+date: "20 March 2019"
+output: html_document
 ---
@@
 overview <- overview %>% mutate(condition=paste0(First.phase,"-",Second.phase))
-all_participant_experiment <- overview %>% left_join(all_participant_experiment, by="Code")
+all_participant_experiment <- overview %>% left_join(all_participant_experiment[, c('Code', 'sex')], by="Code")
 overview <- overview %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
@@
 finished$percent <- round(finished$yes/(finished$yes+finished$no), 2)
+
 is_ok <- overview %>% filter(finished=="yes") %>% group_by(condition) %>% count(is_ok)
@@
 finished_gender$percent <- round(finished_gender$yes/(finished_gender$yes+finished_gender$no), 2)
-```
+kable(finished_gender)
 
-A total of `r nrow(overview)` undergraduate students at UC Davis (`r m_sd_report(df_participants[df_participants$finished == "yes",]$age)`) particiapted in the study in exchange for a study credit.
-`r sum(finished$no)` participants didn't finish due to motion sickness and `r sum(is_ok$no)` were removed due to a technical failure of the real world tracking systen.
-Only `r finished[finished$condition == "vr-real","percent"] * 100` percent the students being able to finish Treadmill VR learning condition.
-
-```{r}
-ez_data<- ez_prepare_block(sub_walk_all_factor, c(1,4), 'min_norm_distance')
-fit_mm_dist_block <- lmer(min_norm_distance ~ learning.condition*exp_block_id + (1 | id), data = ez_data)
-ez_fit_distance <- ezANOVA(ez_data, 
-        dv = min_norm_distance,
-        wid = id,
-        within = exp_block_id,
-        between = learning.condition,
-        return_aov = T,
-        type = 3)
-
-ez_data <- ez_prepare_block(sub_walk_all_factor, c(1,4), 'errors')
-ez_fit_errors <- ezANOVA(ez_data, 
-        dv = errors,
-        wid = id,
-        within = exp_block_id,
-        between = learning.condition,
-        return_aov = T,
-        type = 3)
+### All experiment
+finished_gender <- all_participant_experiment %>% group_by(condition, sex) %>% count(finished)
+finished_gender <- dcast(finished_gender, condition+sex~finished)
+finished_gender$no[is.na(finished_gender$no)] <- 0
+finished_gender$percent <- round(finished_gender$yes/(finished_gender$yes+finished_gender$no), 2)
+kable(finished_gender)
 
 ```
-
-We hypothesized that the degree of transfer to the real-world building would vary as a function of immersion.
-As a first test of this issue, we wished to determine whether all modalities resulted in some degree of transfer.
-Data for distance walked and number of door visitation errors were each entered into a 3 (Learning Condition: real, desktop, treadmill VR) × 2 (Block: block 1, block 4) mixed-model analysis of variance (ANOVA).
-For distance walked, there was a main effect for Learning Condition, `r apa_print(ez_fit_distance$aov)$statistic$learning_condition`.
-There was a also a main effect for Block, `r apa_print(ez_fit_distance$aov)$statistic$exp_block_id`.
-The Learning Condition by Block interaction effect did not reach significance `r apa_print(ez_fit_distance$aov)$statistic$learning_condition_exp_block_id`.
-Follow-up comparisons showed that for walked distance all three modalities showed an improvement from block 1 (first exposure to the modality) to block 4 (first exposure to the real building for immersed/impoverished conditions, fourth exposure for real-world).
-Descriptive and inferential statistics for these comparisons are shown in Table X and Table X+1 for distance and errors, respectively.
-
-For visitation errors, there was no main effect for Learning Condition,`r apa_print(ez_fit_errors$aov)$statistic$learning_condition`.
-There was a main effect for Block, `r apa_print(ez_fit_errors$aov)$statistic$exp_block_id`.
-The Learning Condition by Block interaction effect did not reach significance, `r apa_print(ez_fit_errors$aov)$statistic$learning_condition_exp_block_id`.
-Follow-up comparisons showed that all three modalities showed an improvement in visitation errors from block 1 (first exposure to the modality) to block 4 (first exposure to the real building for immersed/impoverished conditions, fourth exposure for real-world).
-Descriptive and inferential statistics for these comparisons are shown in Table X+1.
-
-```{r}
-#descriptive table
-kable(sub_walk_all_factor %>% filter(exp_block_id %in% c(1,4)) %>% group_by(learning.condition, exp_block_id) %>% summarise(M = mean(min_norm_distance, na.rm=T),
-                                                                                                                     Sd = sd(min_norm_distance, na.rm=T)), digits = 2)
-df_dist <- block_t_test_improvement(sub_walk_all, 1, 4,"min_norm_distance")
-df <- block_t_test_improvement_report(df_dist)
-df$condition <- dplyr::recode(df$condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
-kable(df, caption="Paired T test comparing individual distance performance improvement from the first block to post-switch block", digits = 2)
-
-kable(sub_walk_all_factor %>% filter(exp_block_id %in% c(1,4)) %>% group_by(learning.condition, exp_block_id) %>% summarise(M = mean(errors, na.rm=T),
-                                                                                                                     Sd = sd(errors, na.rm=T)), digits = 2)
-df_error <- block_t_test_improvement(sub_walk_all, 1, 4,"errors")
-df <- block_t_test_improvement_report(df_error)
-df$condition <- dplyr::recode(df$condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
-kable(df, caption="Paired T test comparing individual error rate improvement from the first block to post-switch block", digits = 2)
-```
-
-Figure X shows the time course of each learning condition over the 6 blocks of exposure to the environments, in terms of walked distance and visitation errors over the course of the experiment.
-As can be seen in Figure XX, all modalities showed an incremental improvement in terms of both distance and visitation errors (see fig.
-XXX and fig.
-XXX).
-At block 4, however, an uptick can be observed for the desktop modality switch group and slight for no modality switch (i.e.
-real to real).
-To better quantify these effects, we took two different approaches. First, we analyzed the slope of the learning curves from phase 1 (blocks 1-3) and compared between different Learning Conditions.
-
-```{r}
-### graphs
-ggplot(sub_walk_all, aes(x = exp_block_id, y = min_norm_distance, color = factor(learning.condition))) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.5, geom="errorbar") + 
-  stat_summary(fun.y=mean,position=position_identity(), size = 1.25, geom="line") + 
-  geom_vline(xintercept = 3.5, linetype=4) +
-  ylab("Average normalised trial distance") + xlab("Block") + 
-  scale_color_manual(values=estimote_pal,"Learning condition") + guides(fill=F)
-
-ggplot(sub_walk_all, aes(x = exp_block_id, y = errors, color = factor(learning.condition))) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.5, geom="errorbar") + 
-  stat_summary(fun.y=mean,position=position_identity(), size = 1.25, geom="line") + 
-  geom_vline(xintercept = 3.5, linetype=4) +
-  ylab("Visitation errors") + xlab("Block number") + 
-  scale_color_manual(values=estimote_pal, "Learning condition")
-```
-
-Figure X.
-Performance of each learning condition across time is shown for average normalized distance (a) and visitation errors (b).
-Vertical, dashed line indicates when the second phase of testing (always in the real world) began.
-Error bars represent [95% CI’s].
-
-## Quantifying what happens before transfer
-To better understand the effect of modality (i.e., real world vs. treadmill VR vs. desktop), we assessed the change in walked distance and visitation errors prior to transfer from block 1 to block 3.
-
-```{r}
-ez_data <- ez_prepare_block(sub_walk_all_factor, c(1,3), 'min_norm_distance')
-fit_mm_dist_block <- lmer(min_norm_distance ~ learning.condition*exp_block_id + (1 | id/exp_block_id), data = only_blocks(sub_walk_all, c(1,3)))
-ez_fit_distance <- ezANOVA(
-        ez_data, 
-        dv = min_norm_distance,
-        wid = id,
-        within = exp_block_id,
-        between = learning.condition,
-        return_aov = T,
-        type = 3)
-
-ez_data <- ez_prepare_block(sub_walk_all_factor, c(1,3), 'errors')
-ez_fit_errors <- ezANOVA(
-        ez_data, 
-        dv = errors,
-        wid = id,
-        within = exp_block_id,
-        between = learning.condition,
-        return_aov = T,
-        type = 3)
-```
-
-3x2 repeated measures ANOVA revealed a main effect of block on distance performance `r apa_print(ez_fit_distance$aov)$statistic$exp_block_id`, suggesting that participants improved their paths in all three conditions.
-Although the effect of the learning modality was significant (`r apa_print(ez_fit_distance$aov)$statistic$learning_condition`), the interaction between block and modality was not significant, `r apa_print(ez_fit_distance$aov)$statistic$learning_condition_exp_block_id`, suggesting that learning rate did not differ as a function of modality.
-In contrast, for visitation errors,  we again found a main effect of block (`r apa_print(ez_fit_errors$aov)$statistic$exp_block_id`) adn modality (`r apa_print(ez_fit_errors$aov)$statistic$learning_condition`), but, importantly, a block X modality interaction effect, `r apa_print(ez_fit_errors$aov)$statistic$learning_condition_exp_block_id`.
-This finding suggested that visitation errors improved differently as a function of modality.
-
-To investigate further, we conducted separate one-way ANOVAs to compare learning conditions for walked distance and visitation errors for the last block of learning (block 3).
-There were main effects of learning condition for both walked distance, `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 3)))$statistic`, and visitation errors, `r apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 3)))$statistic`).
-Comparing the conditions closer, we find that there is a significant difference in visitations error performance between real world and desktop learning (`r apa_print(t.test(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Desktop'), 3)))$statistic`) and real world nad treadmill VR learning (`r apa_print(t.test(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Treadmill VR'), 3)))$statistic`), but no difference between Desktop and Treadmill learning (`r apa_print(t.test(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Desktop', 'Treadmill VR'), 3)))$statistic`).
-The situation is the same for distance comparisons.
-real world and desktop learning (`r apa_print(t.test(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Desktop'), 3)))$statistic`) and real world nad treadmill VR learning (`r apa_print(t.test(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Treadmill VR'), 3)))$statistic`), but no difference between Desktop and Treadmill learning (`r apa_print(t.test(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Desktop', 'Treadmill VR'), 3)))$statistic`).
-
-```{r}
-point_data_1 <- sub_sop_all[sub_sop_all$phase == 1,]
-aov_point_phase1 <- aov(abs_error ~ learning.condition, data = point_data_1)
-```
-
-We also observed significant differences between conditions in pointing performance at the end of the first phase, `r apa_print(aov_point_phase1)$statistic$learning_condition`.
-Specifically, participants in the real-world learning group  (`r m_sd_report(only_conditions(point_data_1, "Real")$abs_error)`) performed significantly better compared to both the immersive VR group (`r m_sd_report(only_conditions(point_data_1, "Treadmill VR")$abs_error)`), `r apa_print(t.test(abs_error ~ learning.condition, only_conditions(point_data_1, c('Real', 'Treadmill VR'))))$statistic`, and the desktop group (`r m_sd_report(only_conditions(point_data_1, "Desktop")$abs_error)`), `r apa_print(t.test(abs_error ~ learning.condition, only_conditions(point_data_1, c('Real', 'Desktop'))))$statistic`.
-Overall, these findings suggested that participants learned the building the best as a result of real-world navigation, followed by the immersive condition, with the desktop condition showing the slowest acquisition.
-
-```{r}
-# Pre transfer plot of normalised distances
-ggplot(only_blocks(sub_walk_all_factor, 1:3), aes(x = graph.exp_block_id, y = min_norm_distance, fill = graph.exp_block_id)) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
-  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
-  ylab("Average normalised trial distance") + xlab("Learning condition") + 
-  scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F) +
-  facet_wrap(~graph.learning.condition, strip.position = "bottom", scales = "free_x") + 
-  theme(panel.spacing = unit(0, "lines"), 
-         strip.background = element_blank(),
-         strip.placement = "outside")
-
-ggplot(only_blocks(sub_walk_all_factor, 1:3), aes(x = graph.exp_block_id, y = errors, fill = graph.exp_block_id)) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
-  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
-  ylab("Average visitation errors") + xlab("Learning condition") + 
-  scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F) +
-  facet_wrap(~graph.learning.condition, strip.position = "bottom", scales = "free_x") + 
-  theme(panel.spacing = unit(0, "lines"), 
-         strip.background = element_blank(),
-         strip.placement = "outside")
-
-ggplot(point_data_1, aes(x = graph.learning.condition, y = abs_error, fill = graph.learning.condition)) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
-  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
-  ylab("Average pointing error after learning phase") + xlab("Learning condition") + 
-  scale_fill_manual(values = estimote_pal, "Learning condition") +
-  guides(fill=F) + theme(axis.title.y = element_text(size = rel(0.9)))
-
-```
-
-## Quantifying what happens after/during? transfer
-```{r}
-ez_data <- ez_prepare_block(sub_walk_all_factor, c(3,4), 'min_norm_distance')
-fit_mm_dist_block <- lmer(min_norm_distance ~ learning.condition*exp_block_id + (1 | id/exp_block_id), data = only_blocks(sub_walk_all, 3:4))
-ez_fit_distance <- ezANOVA(
-        ez_data, 
-        dv = min_norm_distance,
-        wid = id,
-        within = exp_block_id,
-        between = learning.condition,
-        return_aov = T,
-        type = 3)
-
-ez_data <- ez_prepare_block(sub_walk_all_factor, c(3,4), 'errors')
-ez_fit_errors <- ezANOVA(
-        ez_data, 
-        dv = errors,
-        wid = id,
-        within = exp_block_id,
-        between = learning.condition,
-        return_aov = T,
-        type = 3)
-```
-
-We next considered the effects of transferring from either the treadmill or desktop to the real-world building compared with continuing to navigate the real-world building.
-We did this by directly comparing walked distance and visitation errors on block 3 (the last block before the transfer phase) with the first block of the transfer phase (block 4).
-Data for distance walked and number of door visitation errors were each entered into a 3 (Learning Condition: real, desktop, treadmill VR) × 2 (Block: block 1, block 4) mixed-model ANOVA.
-For visitation errors, there was a main effect for Learning Condition,`r apa_print(ez_fit_errors$aov)$statistic$learning_condition`.
-There was a main effect for Block, `r apa_print(ez_fit_errors$aov)$statistic$exp_block_id`.
-The Learning Condition by Block interaction effect was significant, `r apa_print(ez_fit_errors$aov)$statistic$learning_condition_exp_block_id`.
-For walked distance, there was a main effect of Learning Condition,`r apa_print(ez_fit_distance$aov)$statistic$learning_condition`, but no effect of Block, `r apa_print(ez_fit_distance$aov)$statistic$exp_block_id`, and the interaction also failed to reach significance, `r apa_print(ez_fit_distance$aov)$statistic$learning_condition_exp_block_id`.
-
-To better understand these differences, we ran separate pairwise t-tests to assess the change from block 3 (last block of learning phase) to block 4 (first block of transfer phase).
-For distance, we found no significant differences from block 3 (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Real", 3)$min_norm_distance)`) to block 4 (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Real", 4)$min_norm_distance)`) for the real-world learning condition, `r apa_print(t.test(min_norm_distance ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Real'), 3:4)))$statistic`.
-Similarly, there was no difference between block 3 (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Treadmill VR", 3)$min_norm_distance)`) and block 4 (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Treadmill VR", 4)$min_norm_distance)`) for the treadmill VR condition, `r apa_print(t.test(min_norm_distance ~ exp_block_id, only_conditions_blocks(sub_walk_all, 'Treadmill VR', 3:4)))$statistic`.
-For visitation errors, we found an improvement in all conditions, although the effects sizes were greater for the treadmill compared to the desktop conditions.
-Together, these findings suggested that participants maintained or improved for both distance and visitation errors, although these effects were generally smaller, and even tending toward being worse, for the desktop condition.
-For summary please see table XXX.
-
-```{r}
-#descriptive table
-kable(sub_walk_all_factor %>% filter(exp_block_id %in% c(3,4)) %>% group_by(learning.condition, exp_block_id) %>% summarise(M = mean(min_norm_distance, na.rm=T),
-                                                                                                                     Sd = sd(min_norm_distance, na.rm=T)), digits = 2)
-df_dist <- block_t_test_improvement(sub_walk_all_factor, 3, 4,"min_norm_distance")
-df <- block_t_test_improvement_report(df_dist)
-df$condition <- dplyr::recode(df$condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
-kable(df, caption="Paired T test comparing individual distance performance improvement from pre switch block to post-switch block", digits = 2)
-
-kable(sub_walk_all_factor %>% filter(exp_block_id %in% c(3,4)) %>% group_by(learning.condition, exp_block_id) %>% summarise(M = mean(errors, na.rm=T),
-                                                                                                                     Sd = sd(errors, na.rm=T)), digits = 2)
-df_error <- block_t_test_improvement(sub_walk_all_factor, 3, 4, "errors")
-df <- block_t_test_improvement_report(df_error)
-df$condition <- dplyr::recode(df$condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
-kable(df, caption="Paired T test comparing individual error rate improvement from pre switch block to post-switch block", digits = 2)
-```
-
-## Quantifying transfer 
-To assess the level of performance change form directly before and after the switch, we calculated percent improvement score as
-(pre-switch performance - post-switch performance)/(pre-switch performance + post-switch performance)
-This allowed us to directly compare the relative improvement or deterioration from pre-switch block to the post-switch one.
-These findings are shown in Figure XX for normalized distance (Figure XXa) and visitation errors (Figure XXb).
-
-```{r, echo=F}
-ggplot(df_distance, aes(x = graph.learning.condition, y = block34.diff, fill = graph.learning.condition)) + 
-  geom_bar(stat = 'identity', position="dodge") + 
-  geom_errorbar(aes(ymin=block34.diff-block34.se, ymax=block34.diff+block34.se), width=.2, position=position_dodge(.9)) +
-  scale_fill_manual(values = estimote_pal) +
-  ylab("Level of transfer for normalised distance") + 
-  xlab("Learning condition") + guides(fill=F) + 
-  theme(axis.title.y = element_text(size = rel(0.9)))
-
-ggplot(df_errors, aes(x = graph.learning.condition, y = block34.diff, fill = graph.learning.condition)) + 
-  geom_bar(stat = 'identity', position="dodge") + 
-  geom_errorbar(aes(ymin=block34.diff-block34.se, ymax=block34.diff+block34.se), width=.2, position=position_dodge(.9)) +
-  scale_fill_manual(values = estimote_pal) +
-  ylab("Level of transfer for visitation errors")+ xlab("Learning condition") + 
-  guides(fill=F) + theme(axis.title.y = element_text(size = rel(0.9)))
-```
-
-Comparing the improvement in walked distance using one-way ANOVA, we see did not find any  significant difference between groups `r apa_print(aov(block34 ~ learning.condition, df_wide_distance))$statistic$learning_condition `.
-Visual inspection of the data, however, suggested that the desktop condition again showed the least transfer from block 3 to block 4.
-For visitation errors, there were significant group differences in error rate improvement, `r apa_print(aov(block34 ~ learning.condition, df_wide_errors))$statistic$learning_condition`.
-Comparing the conditions pairwise, we can asses that there is a significant difference in improvement between real world learnign and desktop (`r apa_print(t.test(block34 ~ learning.condition, only_conditions(df_wide_errors, c('Real', 'Desktop'))))$statistic`), as well as Treadmill VR, `r apa_print(t.test(block34 ~ learning.condition, only_conditions(df_wide_errors, c('Real', 'Treadmill VR'))))$statistic`, but we find no difference in improvement betweem desktop learning and Treadmill VR learning, `r apa_print(t.test(block34 ~ learning.condition, only_conditions(df_wide_errors, c('Treadmill VR', 'Desktop'))))$statistic`.
-
-Together, these findings suggest that transfer from desktop to real-world environments resulted in an increase in visitation errors, while no change was present for immersive and real-world conditions.
-
-## Quantifying what happens after transfer
-
-```{r}
-ez_data <- ez_prepare_block(sub_walk_all_factor, c(4,6), 'min_norm_distance')
-fit_mm_dist_block <- lmer(min_norm_distance ~ learning.condition*exp_block_id + (1 | id/exp_block_id), data = only_blocks(sub_walk_all, c(4,6)))
-ez_fit_distance <- ezANOVA(
-        ez_data, 
-        dv = min_norm_distance,
-        wid = id,
-        within = exp_block_id,
-        between = learning.condition,
-        return_aov = T,
-        type = 3)
-
-ez_data <- ez_prepare_block(sub_walk_all_factor, c(4,6), 'errors')
-ez_fit_errors <- ezANOVA(
-        ez_data, 
-        dv = errors,
-        wid = id,
-        within = exp_block_id,
-        between = learning.condition,
-        return_aov = T,
-        type = 3)
-```
-
-As an additional assay of transfer, we wished to determine whether learning post modality switch was faster in the group that had been exposed to the real-world in the learning blocks compared to the VR groups.
-Comparing walked distance and visitation errors across between blocks 4 and 6 by running repeated measures ANOVA, we found a significant effect of learning condition on distance,`r apa_print(ez_fit_distance$aov)$statistic$learning_condition` and visitation errors `r apa_print(ez_fit_errors$aov)$statistic$learning_condition`.
-We also wound significant interaction effect between learning condition and block for both distance `r apa_print(ez_fit_distance$aov)$statistic$learning_condition` and visitation errors `r apa_print(ez_fit_errors$aov)$statistic$learning_condition`.
-
-Looking at the improvement in visitations errors from block 4 to block 6 in separate learning conditions using t-tests, we can assess that only Desktop group (`r apa_print(t.test(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Desktop'), c(4,6))))$statistic`) and Treadmill VR group (`r apa_print(t.test(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Treadmill VR'), c(4,6))))$statistic`) have improved, whereas group learned in the real world did not improve any more in the transfer phase (`r apa_print(t.test(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Real'), c(4,6))))$statistic`).
-
-The story is similar with distance performance, with both Desktop (`r apa_print(t.test(min_norm_distance ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Desktop'), c(4,6))))$statistic`) and Treadmill VR (`r apa_print(t.test(min_norm_distance ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Treadmill VR'), c(4,6))))$statistic`) improve from the 4th to the 6th block, but so does the real world learning group (`r apa_print(t.test(min_norm_distance ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Real'), c(4,6))))$statistic`).
-
-In the last testing block (block 6), we found no differences among the groups in either error rate (`r apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 6)))$statistic`), nor pointing performance (`r apa_print(aov(abs_error ~ learning.condition, data = sub_sop_all[sub_sop_all$phase == 2,]))$statistic$learning_condition`).
-However we still see small difference in distance performance (`r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 6)))$statistic`), with real group (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Real", 6)$min_norm_distance)`) performing better than Treadmill VR group (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Treadmill VR", 6)$min_norm_distance)`)(`r apa_print(t.test(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Treadmill VR', 'Real'), 6)))$statistic`) and desktop group (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Desktop", 6)$min_norm_distance)`)(`r apa_print(t.test(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Desktop'), 6)))$statistic`).
-
-## Could effects be driven by difficulty with the immersive interface?
-
-To attempt to address this issue, we compared the three different modalities, navigating in the real-world (real-world), navigating on the treadmill with the head-mounted display (immersed), and navigating with a joystick in desktop VR (impoverished) on the first block of the experiment.
-Comparing normalized distances, there were no difference between (`r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 1)))$statistic$learning_condition`), nor in visitation errors (`r apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 1)))$statistic$learning_condition`).
-This suggests that participants were readily able to navigate in the different modalities regardless of the interface.
-We did, however, find a significant difference in normalized time (`r apa_print(aov(min_norm_time ~ learning.condition, only_blocks(sub_walk_all, 1)))$statistic$learning_condition`).
-This effect appeared to be driven largely by the fact that treadmill VR participants took to get used to the interface and therefore having longer trial times at the start.
-
-## Experiment 1b: Comparing transfer from desktop to real and treadmill to desktop
-```{r}
-### Try running it without outliers
-sub_walk_all_2[!is.na(sub_walk_all_2$norm_distance) & sub_walk_all_2$norm_distance > 3, c("distance", "min_norm_distance")] <- NA
-
-sub_walk_all_2_factor <- sub_walk_all_2
-sub_walk_all_2_factor$exp_block_id <- factor(sub_walk_all_2_factor$exp_block_id)
-sub_walk_all_2_factor$id <- factor(sub_walk_all_2_factor$id)
-sub_walk_all_2_factor$learning.condition <- factor(sub_walk_all_2_factor$learning.condition)
-```
-
-Based on our findings above, we might predict that transferring knowledge from desktop to the real world was most difficult for subjects, with the immersive group somewhere in between.
-We also found, that despite the groups differ at the end of the learning phase, their performance is the same after three blocks in the transfer phase.
-It is unclear, whether this equalisation of performance is solely due to exposure to the real world, and wihout it participants could never perform perfectly, or if the tradmill VR modality could achieve the same performance as the real world training, but it only needs more time.
-
-To test this idea, we added two more conditions.
-One group learned the UC Center for Neuroscience building on the treadmill VR while the second group learned in the real world, but both were tested in the tradmill VR in the transfer phase.
-This allowed us to assess if the treadmill VR can achieve the same performance given additional time and also investigate the process of opposite transfer.
-
-### Participants
-```{r}
-df_participants <- settings$participants
-df_participants <- df_participants %>% mutate(condition=paste0(First.phase,"-",Second.phase))
-df_participants <- df_participants %>% filter(condition %in% c("real-vr", "vr-vr"))
-overview <- settings$versions
-overview <- overview %>% mutate(condition=paste0(First.phase,"-",Second.phase))
-overview <- overview %>% filter(condition %in% c("real-vr", "vr-vr"))
-all_participant <- df_participants %>% left_join(overview, by="Code")
-
-finished <- overview %>% group_by(condition) %>% count(finished)
-finished <- dcast(finished, condition~finished)
-finished$percent <- round(finished$yes/(finished$yes+finished$no), 2)
-is_ok <- overview %>% filter(finished=="yes") %>% group_by(condition) %>% count(is_ok)
-is_ok <- dcast(is_ok, condition~is_ok)
-is_ok$percent <- round(is_ok$yes/(is_ok$yes+is_ok$no), 2)
-
-```
-A total of `r nrow(overview)` undergraduate students at UC Davis particiapted in the second experiment.
-`r sum(finished$no)` participants didn't finish due to motion sickness.
-Analyses were then conducted on a final set of `r sum(overview$is_ok == "yes", na.rm=T)` (14 female) participants(`r m_sd_report(all_participant[all_participant$is_ok == "yes",]$age)`).
-
-```{r}
-kable(finished)
-all_participant %>% filter(is_ok=="yes") %>% group_by(sex) %>% count(sex)
-```
-
-## Quantifying performance in the last block
-First we assessed difference between the two learning conditions in the last block performance usint t-tests.
-The conditions did not differ in the walking performance (`r apa_print(t.test(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all_2, c(6))))$statistic`) nor visiting errors (`r apa_print(t.test(errors ~ learning.condition, only_blocks(sub_walk_all_2, c(6))))$statistic`), therefore showing that both treadmill VR and real world learning leads to the same performance in the VR.
-
-## Comparing to previous conditions
-Running a one-way ANOVA togeher together with other conditions from the first experiment (real to real, Vr - real and desktop to real), shows a significant effect of condition on final walking performance (`r apa_print(aov(min_norm_distance ~ condition, only_blocks(walk_all, c(6))))$statistic`).
-
-```{r}
-kable(tukey_report_table(TukeyHSD(aov(min_norm_distance ~ condition, only_blocks(walk_all, c(6))))))
-```
-
-Looking at the final walking performance in the Table XXX, we can see that the differences between conditions are truly negligible.
-Interestingly, we found no significant difference between the conditions when we analysed absolute distance, rather than normalised (`r apa_print(aov(distance ~ condition, only_blocks(walk_all, c(6))))$statistic`).
-This means that all aprticipants walked the same distance in the end on average, regardless of them being tested on a treadmill or in the real building.
-We also found no difference between any of the conditions in their final pointing performance, `r apa_print(aov(abs_error ~ condition, sop_all[sop_all$phase==2,]))$statistic`.
-
-```{r}
-df <- only_blocks(walk_all, c(6)) %>% group_by(condition) %>% summarize(dist = m_sd_report(min_norm_distance), errors= m_sd_report(errors))
-kable(df, caption='Average performance in all conditions in the last block in visitation errors and walking disntace', digits=2)
-```
-
-A different story is in comparing the visitation errors across conditions in the last block, where one-way ANOVA shows significant difference (`r apa_print(aov(errors ~ condition, only_blocks(walk_all, c(6))))$statistic`) and Tukey's post hoc tests demonstrates significant differences between all conditions that differs in transfer phase modality, e.g.
-participants in the real world always perform better than participants on the treadmill VR.
-Looking at the table XXX, we can clearly see that conditions with transfer phase happening in VR, despite having comparable walked distance, have higher number of visitation errors.
-We think that this may have emerged from two different sources: firstly, it can point to higher difficulty in recognizing proper doors because of lesser visual accuracy in the VR, where all doors look similar unlike real doors which may demonstrate unique signs of wear or use.
-The second reason for the increased number of visitation errors in VR conditions can be explained by the lesser "social impact" of wrong choice.
-Whereas in the real conditions, participants had to actively inquire the experimenter, there was technically no penalty for "trying out" wrong door in the VR.
-
-```{r, fig.width=10, fig.height=8}
-ggplot(walk_all, aes(x = exp_block_id, y = min_norm_distance, group = factor(condition))) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar", aes(color=graph.learning.condition, linetype=testing.condition)) + 
-  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line", aes(color=graph.learning.condition, linetype=testing.condition)) + 
-  geom_vline(xintercept = 3.5, linetype=4) + 
-  ylab("Average normalised trial distance") + xlab("Block") + 
-  scale_color_manual(values=estimote_pal, "Learning condition") + 
-  scale_linetype_manual(values=c("dashed", "solid"), "Transfer condition") + 
-  guides(fill=F)
-
-ggplot(walk_all, aes(x = exp_block_id, y = errors, color = factor(condition))) +
-  stat_summary(fun.data=mean_cl_normal, position=position_dodge(width=0.1), width = 0.2, geom="errorbar", aes(color=graph.learning.condition, linetype=testing.condition)) + 
-  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line", aes(color=graph.learning.condition, linetype=testing.condition)) + 
-  geom_vline(xintercept = 3.5, linetype=4) +
-  ylab("Average number of visitation errors") + xlab("Block") + 
-  scale_color_manual(values=estimote_pal, "Learning condition") + 
-  scale_linetype_manual(values=c("dashed", "solid"), "Transfer condition") + 
-  guides(fill=F)
-```
-
-### Quantifying transfer
-```{r}
-ez_data <- ez_prepare_block(sub_walk_all_2_factor, c(3,4), 'min_norm_distance')
-fit_mm_dist_block <- lmer(min_norm_distance ~ learning.condition*exp_block_id + (1 | id/exp_block_id), data = only_blocks(sub_walk_all, 3:4))
-ez_fit_distance <- ezANOVA(
-        ez_data, 
-        dv = min_norm_distance,
-        wid = id,
-        within = exp_block_id,
-        between = learning.condition,
-        return_aov = T,
-        type = 3)
-
-ez_data <- ez_prepare_block(sub_walk_all_factor, c(3,4), 'errors')
-ez_fit_errors <- ezANOVA(
-        ez_data, 
-        dv = errors,
-        wid = id,
-        within = exp_block_id,
-        between = learning.condition,
-        return_aov = T,
-        type = 3)
-```
-
-Assessing the effects of trasfering from the real world to the Treadmill VR, we ran 2 (Learning condition: real vs treadmill VR) x 2 (block 3 vs blcok 4) ANOVA on data for distance walked and visitation errors.
-There was no effect of block `r apa_print(ez_fit_distance$aov)$statistic$exp_block_id` nor learning condition (`r apa_print(ez_fit_distance$aov)$statistic$learning_condition`) on distance walked and we observed no interaction effects (`r apa_print(ez_fit_distance$aov)$statistic$learning_condition_exp_block_id`).
-But we found significant interaction effect between block and learning condition in visitation errors, `r apa_print(ez_fit_errors$aov)$statistic$learning_condition_exp_block_id`.
-
-Pairwise t-tests comparing change in error rate from 3rd to 4th block show, that the group learning on the treadmill VR does show significant difference (`r apa_print(t.test(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all_2, 'Treadmill VR', 3:4)))$statistic`) as it effectively didn't change the environment, but the group switching to the VR from the real world demonstrates increase in visitations errors from block 3 (`r m_sd_report(only_conditions_blocks(sub_walk_all_2, "Real", 3)$errors)`) to block 4 (`r m_sd_report(only_conditions_blocks(sub_walk_all_2, "Real", 4)$errors)`), `r apa_print(t.test(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all_2, 'Real', 3:4)))$statistic`.
-
-```{r, fig.width=10, fig.height=8}
-ggplot(only_blocks(walk_all, 6), aes(x = testing.condition, y = min_norm_distance, fill = graph.learning.condition)) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
-  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
-  ylab("Average path distance in the last block of the transfer phase") + xlab("Transfer condition") + 
-  scale_fill_manual(values = estimote_pal, "Condition")  + guides(fill=F) + 
-  facet_wrap(~graph.learning.condition, strip.position = "top", scales = "free_x") + 
-  ggtitle("Learning condition")+
-  theme(panel.spacing = unit(0, "lines"), 
-         strip.background = element_blank(),
-         strip.placement = "outside",
-         strip.text = element_text(size = 16),
-          plot.title = element_text(hjust = 0.5))
-
-ggplot(only_blocks(walk_all, 6), aes(x = testing.condition, y = errors, fill = graph.learning.condition)) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
-  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
-  ylab("Average number of visitation errors in the last block of the transfer phase") + xlab("Transfer condition") + 
-  scale_fill_manual(values = estimote_pal, "Condition")  + guides(fill=F) + 
-  facet_wrap(~graph.learning.condition, strip.position = "top", scales = "free_x") + 
-  ggtitle("Learning condition")+
-  theme(panel.spacing = unit(0, "lines"), 
-         strip.background = element_blank(),
-         strip.placement = "outside",
-         strip.text = element_text(size = 16),
-          plot.title = element_text(hjust = 0.5))
-```
-
-### Summary
-In the last block, all conditios perform similarly as far as the walking distance goes, regardless of them being tested on the VR or in the real building.
-VR leads to overall higher error rates, but same walking performance, even when considering absolute distances.
-There are no significant differences in pointing performance between conditions, regardless of pointing being tested in the real world or in the VR.

### 2019-03-31 Report/first-revision.Rmd
Adds simple pointing results

@@
 library(lme4)
-library(lmerTest)
 library(ggplot2)
@@
 source("../Scripts/paper-prepare.R")
-source("helpers/block-t-test-improvement.R")
-source("helpers/block-t-test.R")
-source("helpers/block-measure.R")
 source("../TwoWorlds/helpers-report.R")
@@
 df_participants <- df_participants %>% mutate(condition=paste0(First.phase,"-",Second.phase))
-all_participant_experiment <- df_participants
-df_participants <- df_participants %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
 
@@
 overview <- overview %>% mutate(condition=paste0(First.phase,"-",Second.phase))
-all_participant_experiment <- overview %>% left_join(all_participant_experiment[, c('Code', 'sex')], by="Code")
-overview <- overview %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
 overview <- overview %>% left_join(df_participants[, c('Code', 'sex')], by="Code")
 
-#green, violet, blue, yellow, pink, dark blue
-estimote_pal <- c('#b6d6c1', '#54003d','#93d5f6','#d5d215','#f3aacb','#2d2556')
-theme_set(theme_minimal() + theme(axis.text=element_text(size=14),
-        axis.title=element_text(size=16)))
-
 ### Try running it without outliers
-sub_walk_all[!is.na(sub_walk_all$norm_distance) & sub_walk_all$norm_distance > 3, c("distance", "min_norm_distance")] <- NA
 walk_all[!is.na(walk_all$norm_distance) & walk_all$norm_distance > 3, c("distance", "min_norm_distance")] <- NA
-
-condition_all_sums <- sub_walk_all %>% 
-  group_by(exp_block_id, id, learning.condition) %>% 
-  summarise(sum.distance = sum(min_norm_distance, na.rm = T), 
-            sum.errors = sum(errors, na.rm = T),
-            mean.distance = mean(min_norm_distance, na.rm = T),
-            mean.error = mean(errors, na.rm = T))
-
-df_errors <- block_measure(condition_all_sums, "sum.errors")
-df_distance <- block_measure(condition_all_sums, "mean.distance")
-df_distance$graph.learning.condition <- dplyr::recode(df_distance$learning.condition, "Real" = "Real-world", "Treadmill VR" = "Immersive VR")
-df_errors$graph.learning.condition <- dplyr::recode(df_errors$learning.condition, "Real" = "Real-world", "Treadmill VR" = "Immersive VR")
-
-df_wide_errors <- dcast(condition_all_sums, learning.condition + id ~ exp_block_id, value.var = "sum.errors")
-df_wide_distance <- dcast(condition_all_sums, learning.condition + id ~ exp_block_id, value.var = "mean.distance")
-colnames(df_wide_errors) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")
-colnames(df_wide_distance) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")
-df_wide_distance$graph.learning.condition <- dplyr::recode(df_wide_distance$learning.condition, "Real" = "Real-world", "Treadmill VR" = "Immersive VR")
-df_wide_errors$graph.learning.condition <- dplyr::recode(df_wide_errors$learning.condition, "Real" = "Real-world", "Treadmill VR" = "Immersive VR")
-
-df_wide_errors$block34 <- (df_wide_errors$block3-df_wide_errors$block4)/(df_wide_errors$block3+df_wide_errors$block4)
-df_wide_distance$block34 <- (df_wide_distance$block3-df_wide_distance$block4)/(df_wide_distance$block3+df_wide_distance$block4)
-
-sub_walk_all_factor <- sub_walk_all
-sub_walk_all_factor$exp_block_id <- factor(sub_walk_all_factor$exp_block_id)
-sub_walk_all_factor$id <- factor(sub_walk_all_factor$id)
-sub_walk_all_factor$learning.condition <- factor(sub_walk_all_factor$learning.condition)
 
@@
 
-## Participants
+## Dropout rate per gender and condition
 ```{r}
@@
 is_ok$percent <- round(is_ok$yes/(is_ok$yes+is_ok$no), 2)
+kable(is_ok)
 kable(finished)
@@
 kable(finished_gender)
+```
 
-### All experiment
-finished_gender <- all_participant_experiment %>% group_by(condition, sex) %>% count(finished)
-finished_gender <- dcast(finished_gender, condition+sex~finished)
-finished_gender$no[is.na(finished_gender$no)] <- 0
-finished_gender$percent <- round(finished_gender$yes/(finished_gender$yes+finished_gender$no), 2)
-kable(finished_gender)
+## POinting results
 
+### Experiment 1 phase 1
+```{r}
+summary(aov(abs_error ~ learning.condition, data = sub_sop_all[sub_sop_all$phase == 1, ]))
+ggplot(sub_sop_all[sub_sop_all$phase == 1, ], aes(learning.condition, abs_error)) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
 ```
+
+### Experiment 1 phase 2
+```{r}
+summary(aov(abs_error ~ learning.condition, data = sub_sop_all[sub_sop_all$phase == 2, ]))
+ggplot(sub_sop_all[sub_sop_all$phase == 2, ], aes(learning.condition, abs_error)) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
+```
+
+### Experiment 2 phase 1
+```{r}
+summary(aov(abs_error ~ learning.condition, data = sub_sop_all_2[sub_sop_all_2$phase == 2, ]))
+ggplot(sub_sop_all_2[sub_sop_all_2$phase == 1, ], aes(learning.condition, abs_error)) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
+```
+
+### Experiment 2 phase 2
+```{r}
+summary(aov(abs_error ~ learning.condition, data = sub_sop_all_2[sub_sop_all_2$phase == 2, ]))
+ggplot(sub_sop_all_2[sub_sop_all_2$phase == 2, ], aes(learning.condition, abs_error)) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
+```
+
+### Both experiments split by final condition
+```{r}
+summary(aov(abs_error ~ type, data = sop_all[sop_all$phase == 2, ]))
+ggplot(sop_all[sop_all$phase == 2, ], aes(type, abs_error)) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
+```
+
+### Plot of type to type evolution
+```{r}
+ggplot(sop_all, aes(phase, abs_error, color=condition)) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.25), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.25),geom="line")
+```

### 2019-03-31 Scripts/main-preprocess.R

@@ -19,8 +19,8 @@ sop_all <- multi_sop_results(ls)
 walk_all <- multi_walk_results(ls)
 
 ##renames all jesiccas
-walk_all <- walk_all %>% mutate(goal = replace(goal, goal=="Jesicca's office", "Jessica's office"),
-                                start = replace(start, start=="Jesicca's office", "Jessica's office"))
+walk_all <- walk_all %>% mutate(goal = replace(goal, goal == "Jesicca's office", "Jessica's office"),
+                                start = replace(start, start == "Jesicca's office", "Jessica's office"))
 
 ### Total trial ----
 walk_all$exp_trial_id <- walk_all$trial_id + (walk_all$phase-1)*18

### 2019-03-31 Report/first-revision.Rmd
Adds more pointing graphs

@@
 
+```{r}
+ggplot(sub_sop_all, aes(learning.condition, abs_error, group=factor(phase), fill=factor(phase))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
+ggplot(sub_sop_all, aes(factor(phase), abs_error, group=learning.condition, fill=learning.condition)) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
+```
+
 ### Experiment 2 phase 1
 ```{r}
-summary(aov(abs_error ~ learning.condition, data = sub_sop_all_2[sub_sop_all_2$phase == 2, ]))
+summary(aov(abs_error ~ learning.condition, data = sub_sop_all_2[sub_sop_all_2$phase == 1, ]))
 ggplot(sub_sop_all_2[sub_sop_all_2$phase == 1, ], aes(learning.condition, abs_error)) +
@@
 
+```{r}
+ggplot(sub_sop_all_2, aes(learning.condition, abs_error, group=factor(phase), fill=factor(phase))) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
+ggplot(sub_sop_all_2, aes(factor(phase), abs_error, group=learning.condition, fill=learning.condition)) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
+```
+
 ### Both experiments split by final condition
@@
 ggplot(sop_all[sop_all$phase == 2, ], aes(type, abs_error)) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
+```
+
+```{r}
+ggplot(sop_all, aes(learning.condition, abs_error, group=factor(phase), fill=factor(phase))) +
+  facet_wrap(~testing.condition)+
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
+ggplot(sop_all, aes(factor(phase), abs_error, group=learning.condition, fill=learning.condition)) +
+  facet_wrap(~testing.condition)+
   stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
@@
 ```
+
+###
+```{r}
+ez_point <- sop_all[complete.cases(sop_all),]
+ez_point$phase <- as.factor(ez_point$phase)
+ezANOVA(ez_point,
+        within = .(phase),
+        between = .(learning.condition),
+        dv=abs_error,
+        wid=id
+        )
+```

### 2019-04-06 Report/first-revision.Rmd
Adds a bit of normalised vs absolute results!

@@
 ```{r}
-summary(aov(abs_error ~ learning.condition, data = sub_sop_all[sub_sop_all$phase == 1, ]))
+av <- aov(abs_error ~ learning.condition, data = sub_sop_all[sub_sop_all$phase == 1, ])
+summary(av)
+apa_print(av)$full_result$learning_condition
 ggplot(sub_sop_all[sub_sop_all$phase == 1, ], aes(learning.condition, abs_error)) +
@@
 ```{r}
-summary(aov(abs_error ~ learning.condition, data = sub_sop_all[sub_sop_all$phase == 2, ]))
+av <- aov(abs_error ~ learning.condition, data = sub_sop_all[sub_sop_all$phase == 2, ])
+summary(av)
+apa_print(av)$full_result$learning_condition
 ggplot(sub_sop_all[sub_sop_all$phase == 2, ], aes(learning.condition, abs_error)) +
@@
 ```{r}
-summary(aov(abs_error ~ learning.condition, data = sub_sop_all_2[sub_sop_all_2$phase == 1, ]))
+av <- t.test(abs_error ~ learning.condition, data = sub_sop_all_2[sub_sop_all_2$phase == 1, ])
+summary(av)
+apa_print(av)$full_result
 ggplot(sub_sop_all_2[sub_sop_all_2$phase == 1, ], aes(learning.condition, abs_error)) +
@@
 ```{r}
-summary(aov(abs_error ~ learning.condition, data = sub_sop_all_2[sub_sop_all_2$phase == 2, ]))
+av <- t.test(abs_error ~ learning.condition, data = sub_sop_all_2[sub_sop_all_2$phase == 2, ])
+apa_print(av)$full_result
 ggplot(sub_sop_all_2[sub_sop_all_2$phase == 2, ], aes(learning.condition, abs_error)) +
@@
 ```{r}
-summary(aov(abs_error ~ type, data = sop_all[sop_all$phase == 2, ]))
+av <- aov(abs_error ~ type, data = sop_all[sop_all$phase == 2, ])
+summary(av)
+apa_print(av)$full_result$learning_condition
 ggplot(sop_all[sop_all$phase == 2, ], aes(type, abs_error)) +
@@
 ```
+
+## Absolute vs normalised values 
+```{r}
+b4 <- only_blocks(sub_walk_all, c(4))
+apa_print(aov(b4$distance ~ b4$learning.condition))$full_result$b4_learning_condition
+apa_print(aov(b4$min_norm_distance ~ b4$learning.condition))$full_result$b4_learning_condition
+b4 <- only_blocks(sub_walk_all, c(6))
+apa_print(aov(b4$distance ~ b4$learning.condition))$full_result$b4_learning_condition
+apa_print(aov(b4$min_norm_distance ~ b4$learning.condition))$full_result$b4_learning_condition
+```
+
+## Gender effects
+
+## NEw charts

### 2019-04-08 Report/final-results.Rmd
Adds gender effects

@@
 date: "9 August 2018"
-output: github_document
+output: html_document
 ---
@@
 library(ez)
+library(effsize)
 library(Hmisc)
@@
 
-settings <- load_google_sheets()
+#settings <- load_google_sheets()
+load('../settings.data')
 df_participants <- settings$participants
@@
 Data for distance walked and number of door visitation errors were each entered into a 3 (Learning Condition: real, desktop, treadmill VR) × 2 (Block: block 1, block 4) mixed-model analysis of variance (ANOVA).
-For distance walked, there was a main effect for Learning Condition, `r apa_print(ez_fit_distance$aov)$statistic$learning_condition`.
-There was a also a main effect for Block, `r apa_print(ez_fit_distance$aov)$statistic$exp_block_id`.
-The Learning Condition by Block interaction effect did not reach significance `r apa_print(ez_fit_distance$aov)$statistic$learning_condition_exp_block_id`.
+For distance walked, there was a main effect for Learning Condition, `r apa_print(ez_fit_distance$aov)$full_result$learning_condition`.
+There was a also a main effect for Block, `r apa_print(ez_fit_distance$aov)$full_result$exp_block_id`.
+The Learning Condition by Block interaction effect did not reach significance `r apa_print(ez_fit_distance$aov)$full_result$learning_condition_exp_block_id`.
 Follow-up comparisons showed that for walked distance all three modalities showed an improvement from block 1 (first exposure to the modality) to block 4 (first exposure to the real building for immersed/impoverished conditions, fourth exposure for real-world).
@@
 
-For visitation errors, there was no main effect for Learning Condition,`r apa_print(ez_fit_errors$aov)$statistic$learning_condition`.
-There was a main effect for Block, `r apa_print(ez_fit_errors$aov)$statistic$exp_block_id`.
-The Learning Condition by Block interaction effect did not reach significance, `r apa_print(ez_fit_errors$aov)$statistic$learning_condition_exp_block_id`.
+For visitation errors, there was no main effect for Learning Condition,`r apa_print(ez_fit_errors$aov)$full_result$learning_condition`.
+There was a main effect for Block, `r apa_print(ez_fit_errors$aov)$full_result$exp_block_id`.
+The Learning Condition by Block interaction effect did not reach significance, `r apa_print(ez_fit_errors$aov)$full_result$learning_condition_exp_block_id`.
 Follow-up comparisons showed that all three modalities showed an improvement in visitation errors from block 1 (first exposure to the modality) to block 4 (first exposure to the real building for immersed/impoverished conditions, fourth exposure for real-world).
@@
 
-3x2 repeated measures ANOVA revealed a main effect of block on distance performance `r apa_print(ez_fit_distance$aov)$statistic$exp_block_id`, suggesting that participants improved their paths in all three conditions.
-Although the effect of the learning modality was significant (`r apa_print(ez_fit_distance$aov)$statistic$learning_condition`), the interaction between block and modality was not significant, `r apa_print(ez_fit_distance$aov)$statistic$learning_condition_exp_block_id`, suggesting that learning rate did not differ as a function of modality.
-In contrast, for visitation errors,  we again found a main effect of block (`r apa_print(ez_fit_errors$aov)$statistic$exp_block_id`) adn modality (`r apa_print(ez_fit_errors$aov)$statistic$learning_condition`), but, importantly, a block X modality interaction effect, `r apa_print(ez_fit_errors$aov)$statistic$learning_condition_exp_block_id`.
+3x2 repeated measures ANOVA revealed a main effect of block on distance performance `r apa_print(ez_fit_distance$aov)$full_result$exp_block_id`, suggesting that participants improved their paths in all three conditions.
+Although the effect of the learning modality was significant (`r apa_print(ez_fit_distance$aov)$full_result$learning_condition`), the interaction between block and modality was not significant, `r apa_print(ez_fit_distance$aov)$full_result$learning_condition_exp_block_id`, suggesting that learning rate did not differ as a function of modality.
+In contrast, for visitation errors,  we again found a main effect of block (`r apa_print(ez_fit_errors$aov)$full_result$exp_block_id`) adn modality (`r apa_print(ez_fit_errors$aov)$full_result$learning_condition`), but, importantly, a block X modality interaction effect, `r apa_print(ez_fit_errors$aov)$full_result$learning_condition_exp_block_id`.
 This finding suggested that visitation errors improved differently as a function of modality.
@@
 To investigate further, we conducted separate one-way ANOVAs to compare learning conditions for walked distance and visitation errors for the last block of learning (block 3).
-There were main effects of learning condition for both walked distance, `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 3)))$statistic`, and visitation errors, `r apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 3)))$statistic`).
-Comparing the conditions closer, we find that there is a significant difference in visitations error performance between real world and desktop learning (`r apa_print(t.test(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Desktop'), 3)))$statistic`) and real world nad treadmill VR learning (`r apa_print(t.test(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Treadmill VR'), 3)))$statistic`), but no difference between Desktop and Treadmill learning (`r apa_print(t.test(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Desktop', 'Treadmill VR'), 3)))$statistic`).
+There were main effects of learning condition for both walked distance, `r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 3)))$full_result`, and visitation errors, `r apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 3)))$full_result`).
+Comparing the conditions closer, we find that there is a significant difference in visitations error performance between real world and desktop learning (`r t_test_effect(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Desktop'), 3))`) and real world nad treadmill VR learning (`r t_test_effect(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Treadmill VR'), 3))`), but no difference between Desktop and Treadmill learning (`r t_test_effect(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Desktop', 'Treadmill VR'), 3))`).
 The situation is the same for distance comparisons.
-real world and desktop learning (`r apa_print(t.test(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Desktop'), 3)))$statistic`) and real world nad treadmill VR learning (`r apa_print(t.test(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Treadmill VR'), 3)))$statistic`), but no difference between Desktop and Treadmill learning (`r apa_print(t.test(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Desktop', 'Treadmill VR'), 3)))$statistic`).
+real world and desktop learning (`r t_test_effect(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Desktop'), 3))`) and real world nad treadmill VR learning (`r t_test_effect(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Treadmill VR'), 3))`), but no difference between Desktop and Treadmill learning (`r t_test_effect(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Desktop', 'Treadmill VR'), 3))`).
 
@@
 
-We also observed significant differences between conditions in pointing performance at the end of the first phase, `r apa_print(aov_point_phase1)$statistic$learning_condition`.
-Specifically, participants in the real-world learning group  (`r m_sd_report(only_conditions(point_data_1, "Real")$abs_error)`) performed significantly better compared to both the immersive VR group (`r m_sd_report(only_conditions(point_data_1, "Treadmill VR")$abs_error)`), `r apa_print(t.test(abs_error ~ learning.condition, only_conditions(point_data_1, c('Real', 'Treadmill VR'))))$statistic`, and the desktop group (`r m_sd_report(only_conditions(point_data_1, "Desktop")$abs_error)`), `r apa_print(t.test(abs_error ~ learning.condition, only_conditions(point_data_1, c('Real', 'Desktop'))))$statistic`.
+We also observed significant differences between conditions in pointing performance at the end of the first phase, `r apa_print(aov_point_phase1)$full_result$learning_condition`.
+Specifically, participants in the real-world learning group  (`r m_sd_report(only_conditions(point_data_1, "Real")$abs_error)`) performed significantly better compared to both the immersive VR group (`r m_sd_report(only_conditions(point_data_1, "Treadmill VR")$abs_error)`), `r t_test_effect(abs_error ~ learning.condition, only_conditions(point_data_1, c('Real', 'Treadmill VR')))`, and the desktop group (`r m_sd_report(only_conditions(point_data_1, "Desktop")$abs_error)`), `r t_test_effect(abs_error ~ learning.condition, only_conditions(point_data_1, c('Real', 'Desktop')))`.
 Overall, these findings suggested that participants learned the building the best as a result of real-world navigation, followed by the immersive condition, with the desktop condition showing the slowest acquisition.
@@
 Data for distance walked and number of door visitation errors were each entered into a 3 (Learning Condition: real, desktop, treadmill VR) × 2 (Block: block 1, block 4) mixed-model ANOVA.
-For visitation errors, there was a main effect for Learning Condition,`r apa_print(ez_fit_errors$aov)$statistic$learning_condition`.
-There was a main effect for Block, `r apa_print(ez_fit_errors$aov)$statistic$exp_block_id`.
-The Learning Condition by Block interaction effect was significant, `r apa_print(ez_fit_errors$aov)$statistic$learning_condition_exp_block_id`.
-For walked distance, there was a main effect of Learning Condition,`r apa_print(ez_fit_distance$aov)$statistic$learning_condition`, but no effect of Block, `r apa_print(ez_fit_distance$aov)$statistic$exp_block_id`, and the interaction also failed to reach significance, `r apa_print(ez_fit_distance$aov)$statistic$learning_condition_exp_block_id`.
+For visitation errors, there was a main effect for Learning Condition,`r apa_print(ez_fit_errors$aov)$full_result$learning_condition`.
+There was a main effect for Block, `r apa_print(ez_fit_errors$aov)$full_result$exp_block_id`.
+The Learning Condition by Block interaction effect was significant, `r apa_print(ez_fit_errors$aov)$full_result$learning_condition_exp_block_id`.
+For walked distance, there was a main effect of Learning Condition,`r apa_print(ez_fit_distance$aov)$full_result$learning_condition`, but no effect of Block, `r apa_print(ez_fit_distance$aov)$full_result$exp_block_id`, and the interaction also failed to reach significance, `r apa_print(ez_fit_distance$aov)$full_result$learning_condition_exp_block_id`.
 
 To better understand these differences, we ran separate pairwise t-tests to assess the change from block 3 (last block of learning phase) to block 4 (first block of transfer phase).
-For distance, we found no significant differences from block 3 (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Real", 3)$min_norm_distance)`) to block 4 (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Real", 4)$min_norm_distance)`) for the real-world learning condition, `r apa_print(t.test(min_norm_distance ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Real'), 3:4)))$statistic`.
-Similarly, there was no difference between block 3 (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Treadmill VR", 3)$min_norm_distance)`) and block 4 (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Treadmill VR", 4)$min_norm_distance)`) for the treadmill VR condition, `r apa_print(t.test(min_norm_distance ~ exp_block_id, only_conditions_blocks(sub_walk_all, 'Treadmill VR', 3:4)))$statistic`.
+For distance, we found no significant differences from block 3 (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Real", 3)$min_norm_distance)`) to block 4 (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Real", 4)$min_norm_distance)`) for the real-world learning condition, `r t_test_effect(min_norm_distance ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Real'), 3:4))`.
+Similarly, there was no difference between block 3 (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Treadmill VR", 3)$min_norm_distance)`) and block 4 (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Treadmill VR", 4)$min_norm_distance)`) for the treadmill VR condition, `r t_test_effect(min_norm_distance ~ exp_block_id, only_conditions_blocks(sub_walk_all, 'Treadmill VR', 3:4))`.
 For visitation errors, we found an improvement in all conditions, although the effects sizes were greater for the treadmill compared to the desktop conditions.
@@
 
-Comparing the improvement in walked distance using one-way ANOVA, we see did not find any  significant difference between groups `r apa_print(aov(block34 ~ learning.condition, df_wide_distance))$statistic$learning_condition `.
+Comparing the improvement in walked distance using one-way ANOVA, we see did not find any  significant difference between groups `r apa_print(aov(block34 ~ learning.condition, df_wide_distance))$full_result$learning_condition `.
 Visual inspection of the data, however, suggested that the desktop condition again showed the least transfer from block 3 to block 4.
-For visitation errors, there were significant group differences in error rate improvement, `r apa_print(aov(block34 ~ learning.condition, df_wide_errors))$statistic$learning_condition`.
-Comparing the conditions pairwise, we can asses that there is a significant difference in improvement between real world learnign and desktop (`r apa_print(t.test(block34 ~ learning.condition, only_conditions(df_wide_errors, c('Real', 'Desktop'))))$statistic`), as well as Treadmill VR, `r apa_print(t.test(block34 ~ learning.condition, only_conditions(df_wide_errors, c('Real', 'Treadmill VR'))))$statistic`, but we find no difference in improvement betweem desktop learning and Treadmill VR learning, `r apa_print(t.test(block34 ~ learning.condition, only_conditions(df_wide_errors, c('Treadmill VR', 'Desktop'))))$statistic`.
+For visitation errors, there were significant group differences in error rate improvement, `r apa_print(aov(block34 ~ learning.condition, df_wide_errors))$full_result$learning_condition`.
+Comparing the conditions pairwise, we can asses that there is a significant difference in improvement between real world learnign and desktop (`r t_test_effect(block34 ~ learning.condition, only_conditions(df_wide_errors, c('Real', 'Desktop')))`), as well as Treadmill VR, `r t_test_effect(block34 ~ learning.condition, only_conditions(df_wide_errors, c('Real', 'Treadmill VR')))`, but we find no difference in improvement betweem desktop learning and Treadmill VR learning, `r t_test_effect(block34 ~ learning.condition, only_conditions(df_wide_errors, c('Treadmill VR', 'Desktop')))`.
 
@@
 As an additional assay of transfer, we wished to determine whether learning post modality switch was faster in the group that had been exposed to the real-world in the learning blocks compared to the VR groups.
-Comparing walked distance and visitation errors across between blocks 4 and 6 by running repeated measures ANOVA, we found a significant effect of learning condition on distance,`r apa_print(ez_fit_distance$aov)$statistic$learning_condition` and visitation errors `r apa_print(ez_fit_errors$aov)$statistic$learning_condition`.
-We also wound significant interaction effect between learning condition and block for both distance `r apa_print(ez_fit_distance$aov)$statistic$learning_condition` and visitation errors `r apa_print(ez_fit_errors$aov)$statistic$learning_condition`.
+Comparing walked distance and visitation errors across between blocks 4 and 6 by running repeated measures ANOVA, we found a significant effect of learning condition on distance,`r apa_print(ez_fit_distance$aov)$full_result$learning_condition` and visitation errors `r apa_print(ez_fit_errors$aov)$full_result$learning_condition`.
+We also wound significant interaction effect between learning condition and block for both distance `r apa_print(ez_fit_distance$aov)$full_result$learning_condition` and visitation errors `r apa_print(ez_fit_errors$aov)$full_result$learning_condition`.
 
-Looking at the improvement in visitations errors from block 4 to block 6 in separate learning conditions using t-tests, we can assess that only Desktop group (`r apa_print(t.test(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Desktop'), c(4,6))))$statistic`) and Treadmill VR group (`r apa_print(t.test(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Treadmill VR'), c(4,6))))$statistic`) have improved, whereas group learned in the real world did not improve any more in the transfer phase (`r apa_print(t.test(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Real'), c(4,6))))$statistic`).
+Looking at the improvement in visitations errors from block 4 to block 6 in separate learning conditions using t-tests, we can assess that only Desktop group (`r t_test_effect(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Desktop'), c(4,6)))`) and Treadmill VR group (`r t_test_effect(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Treadmill VR'), c(4,6)))`) have improved, whereas group learned in the real world did not improve any more in the transfer phase (`r t_test_effect(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Real'), c(4,6)))`).
 
-The story is similar with distance performance, with both Desktop (`r apa_print(t.test(min_norm_distance ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Desktop'), c(4,6))))$statistic`) and Treadmill VR (`r apa_print(t.test(min_norm_distance ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Treadmill VR'), c(4,6))))$statistic`) improve from the 4th to the 6th block, but so does the real world learning group (`r apa_print(t.test(min_norm_distance ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Real'), c(4,6))))$statistic`).
+The story is similar with distance performance, with both Desktop (`r t_test_effect(min_norm_distance ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Desktop'), c(4,6)))`) and Treadmill VR (`r t_test_effect(min_norm_distance ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Treadmill VR'), c(4,6)))`) improve from the 4th to the 6th block, but so does the real world learning group (`r t_test_effect(min_norm_distance ~ exp_block_id, only_conditions_blocks(sub_walk_all, c('Real'), c(4,6)))`).
 
-In the last testing block (block 6), we found no differences among the groups in either error rate (`r apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 6)))$statistic`), nor pointing performance (`r apa_print(aov(abs_error ~ learning.condition, data = sub_sop_all[sub_sop_all$phase == 2,]))$statistic$learning_condition`).
-However we still see small difference in distance performance (`r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 6)))$statistic`), with real group (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Real", 6)$min_norm_distance)`) performing better than Treadmill VR group (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Treadmill VR", 6)$min_norm_distance)`)(`r apa_print(t.test(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Treadmill VR', 'Real'), 6)))$statistic`) and desktop group (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Desktop", 6)$min_norm_distance)`)(`r apa_print(t.test(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Desktop'), 6)))$statistic`).
+In the last testing block (block 6), we found no differences among the groups in either error rate (`r apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 6)))$full_result`), nor pointing performance (`r apa_print(aov(abs_error ~ learning.condition, data = sub_sop_all[sub_sop_all$phase == 2,]))$full_result$learning_condition`).
+However we still see small difference in distance performance (`r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 6)))$full_result`), with real group (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Real", 6)$min_norm_distance)`) performing better than Treadmill VR group (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Treadmill VR", 6)$min_norm_distance)`)(`r t_test_effect(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Treadmill VR', 'Real'), 6))`) and desktop group (`r m_sd_report(only_conditions_blocks(sub_walk_all, "Desktop", 6)$min_norm_distance)`)(`r t_test_effect(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Desktop'), 6))`).
 
@@
 To attempt to address this issue, we compared the three different modalities, navigating in the real-world (real-world), navigating on the treadmill with the head-mounted display (immersed), and navigating with a joystick in desktop VR (impoverished) on the first block of the experiment.
-Comparing normalized distances, there were no difference between (`r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 1)))$statistic$learning_condition`), nor in visitation errors (`r apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 1)))$statistic$learning_condition`).
+Comparing normalized distances, there were no difference between (`r apa_print(aov(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all, 1)))$full_result$learning_condition`), nor in visitation errors (`r apa_print(aov(errors ~ learning.condition, only_blocks(sub_walk_all, 1)))$full_result$learning_condition`).
 This suggests that participants were readily able to navigate in the different modalities regardless of the interface.
-We did, however, find a significant difference in normalized time (`r apa_print(aov(min_norm_time ~ learning.condition, only_blocks(sub_walk_all, 1)))$statistic$learning_condition`).
+We did, however, find a significant difference in normalized time (`r apa_print(aov(min_norm_time ~ learning.condition, only_blocks(sub_walk_all, 1)))$full_result$learning_condition`).
 This effect appeared to be driven largely by the fact that treadmill VR participants took to get used to the interface and therefore having longer trial times at the start.
@@
 First we assessed difference between the two learning conditions in the last block performance usint t-tests.
-The conditions did not differ in the walking performance (`r apa_print(t.test(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all_2, c(6))))$statistic`) nor visiting errors (`r apa_print(t.test(errors ~ learning.condition, only_blocks(sub_walk_all_2, c(6))))$statistic`), therefore showing that both treadmill VR and real world learning leads to the same performance in the VR.
+The conditions did not differ in the walking performance (`r t_test_effect(min_norm_distance ~ learning.condition, only_blocks(sub_walk_all_2, c(6)))`) nor visiting errors (`r t_test_effect(errors ~ learning.condition, only_blocks(sub_walk_all_2, c(6)))`), therefore showing that both treadmill VR and real world learning leads to the same performance in the VR.
 
 ## Comparing to previous conditions
-Running a one-way ANOVA togeher together with other conditions from the first experiment (real to real, Vr - real and desktop to real), shows a significant effect of condition on final walking performance (`r apa_print(aov(min_norm_distance ~ condition, only_blocks(walk_all, c(6))))$statistic`).
+Running a one-way ANOVA togeher together with other conditions from the first experiment (real to real, Vr - real and desktop to real), shows a significant effect of condition on final walking performance (`r apa_print(aov(min_norm_distance ~ condition, only_blocks(walk_all, c(6))))$full_result`).
 
@@
 Looking at the final walking performance in the Table XXX, we can see that the differences between conditions are truly negligible.
-Interestingly, we found no significant difference between the conditions when we analysed absolute distance, rather than normalised (`r apa_print(aov(distance ~ condition, only_blocks(walk_all, c(6))))$statistic`).
+Interestingly, we found no significant difference between the conditions when we analysed absolute distance, rather than normalised (`r apa_print(aov(distance ~ condition, only_blocks(walk_all, c(6))))$full_result`).
 This means that all aprticipants walked the same distance in the end on average, regardless of them being tested on a treadmill or in the real building.
-We also found no difference between any of the conditions in their final pointing performance, `r apa_print(aov(abs_error ~ condition, sop_all[sop_all$phase==2,]))$statistic`.
+We also found no difference between any of the conditions in their final pointing performance, `r apa_print(aov(abs_error ~ condition, sop_all[sop_all$phase==2,]))$full_result`.
 
@@
 
-A different story is in comparing the visitation errors across conditions in the last block, where one-way ANOVA shows significant difference (`r apa_print(aov(errors ~ condition, only_blocks(walk_all, c(6))))$statistic`) and Tukey's post hoc tests demonstrates significant differences between all conditions that differs in transfer phase modality, e.g.
+A different story is in comparing the visitation errors across conditions in the last block, where one-way ANOVA shows significant difference (`r apa_print(aov(errors ~ condition, only_blocks(walk_all, c(6))))$full_result`) and Tukey's post hoc tests demonstrates significant differences between all conditions that differs in transfer phase modality, e.g.
 participants in the real world always perform better than participants on the treadmill VR.
@@
 Assessing the effects of trasfering from the real world to the Treadmill VR, we ran 2 (Learning condition: real vs treadmill VR) x 2 (block 3 vs blcok 4) ANOVA on data for distance walked and visitation errors.
-There was no effect of block `r apa_print(ez_fit_distance$aov)$statistic$exp_block_id` nor learning condition (`r apa_print(ez_fit_distance$aov)$statistic$learning_condition`) on distance walked and we observed no interaction effects (`r apa_print(ez_fit_distance$aov)$statistic$learning_condition_exp_block_id`).
-But we found significant interaction effect between block and learning condition in visitation errors, `r apa_print(ez_fit_errors$aov)$statistic$learning_condition_exp_block_id`.
+There was no effect of block `r apa_print(ez_fit_distance$aov)$full_result$exp_block_id` nor learning condition (`r apa_print(ez_fit_distance$aov)$full_result$learning_condition`) on distance walked and we observed no interaction effects (`r apa_print(ez_fit_distance$aov)$full_result$learning_condition_exp_block_id`).
+But we found significant interaction effect between block and learning condition in visitation errors, `r apa_print(ez_fit_errors$aov)$full_result$learning_condition_exp_block_id`.
 
-Pairwise t-tests comparing change in error rate from 3rd to 4th block show, that the group learning on the treadmill VR does show significant difference (`r apa_print(t.test(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all_2, 'Treadmill VR', 3:4)))$statistic`) as it effectively didn't change the environment, but the group switching to the VR from the real world demonstrates increase in visitations errors from block 3 (`r m_sd_report(only_conditions_blocks(sub_walk_all_2, "Real", 3)$errors)`) to block 4 (`r m_sd_report(only_conditions_blocks(sub_walk_all_2, "Real", 4)$errors)`), `r apa_print(t.test(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all_2, 'Real', 3:4)))$statistic`.
+Pairwise t-tests comparing change in error rate from 3rd to 4th block show, that the group learning on the treadmill VR does show significant difference (`r t_test_effect(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all_2, 'Treadmill VR', 3:4))`) as it effectively didn't change the environment, but the group switching to the VR from the real world demonstrates increase in visitations errors from block 3 (`r m_sd_report(only_conditions_blocks(sub_walk_all_2, "Real", 3)$errors)`) to block 4 (`r m_sd_report(only_conditions_blocks(sub_walk_all_2, "Real", 4)$errors)`), `r t_test_effect(errors ~ exp_block_id, only_conditions_blocks(sub_walk_all_2, 'Real', 3:4))`.
 

### 2019-04-08 Report/first-revision.Rmd
Adds gender effects

@@
 library(gridExtra)
+library(effsize)
 library(ez)
@@
 source("../TwoWorlds/helpers-report.R")
-
-settings <- load_google_sheets()
+#settings <- load_google_sheets()
+load('../settings.data')
 df_participants <- settings$participants
@@
 walk_all[!is.na(walk_all$norm_distance) & walk_all$norm_distance > 3, c("distance", "min_norm_distance")] <- NA
+sub_walk_all[!is.na(sub_walk_all$norm_distance) & sub_walk_all$norm_distance > 3, c("distance", "min_norm_distance")] <- NA
 
@@
 summary(av)
-apa_print(av)$full_result$learning_condition
 ggplot(sub_sop_all[sub_sop_all$phase == 1, ], aes(learning.condition, abs_error)) +
@@
 
+Anova results are `r apa_print(av)$full_result$learning_condition`
+
 ### Experiment 1 phase 2
@@
 summary(av)
-apa_print(av)$full_result$learning_condition
 ggplot(sub_sop_all[sub_sop_all$phase == 2, ], aes(learning.condition, abs_error)) +
@@
 ```
+Anova results are `r apa_print(av)$full_result$learning_condition`
 
@@
 ### Experiment 2 phase 1
+T test results is `r t_test_effect(abs_error ~ learning.condition, data = sub_sop_all_2[sub_sop_all_2$phase == 1, ])`
 ```{r}
-av <- t.test(abs_error ~ learning.condition, data = sub_sop_all_2[sub_sop_all_2$phase == 1, ])
-summary(av)
-apa_print(av)$full_result
 ggplot(sub_sop_all_2[sub_sop_all_2$phase == 1, ], aes(learning.condition, abs_error)) +
@@
 ### Experiment 2 phase 2
+T test results is `r t_test_effect(abs_error ~ learning.condition, data = sub_sop_all_2[sub_sop_all_2$phase == 2, ])`
 ```{r}
-av <- t.test(abs_error ~ learning.condition, data = sub_sop_all_2[sub_sop_all_2$phase == 2, ])
-apa_print(av)$full_result
 ggplot(sub_sop_all_2[sub_sop_all_2$phase == 2, ], aes(learning.condition, abs_error)) +
@@
 
+### Both experiments split condition
+```{r}
+av <- aov(abs_error ~ condition, data = sop_all[sop_all$phase == 2, ])
+summary(av)
+ggplot(sop_all[sop_all$phase == 2, ], aes(condition, abs_error)) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
+```
+Anova results are `r apa_print(av)$full_result$condition`
+
 ### Both experiments split by final condition
+T test results is `r t_test_effect(abs_error ~ type, data = sop_all[sop_all$phase == 2, ])`
 ```{r}
-av <- aov(abs_error ~ type, data = sop_all[sop_all$phase == 2, ])
-summary(av)
-apa_print(av)$full_result$learning_condition
 ggplot(sop_all[sop_all$phase == 2, ], aes(type, abs_error)) +
@@
 ## Gender effects
+```{r}
+walk_all <- walk_all %>% left_join(df_participants[, c('Code', 'sex')], by=c("id"="Code"))
+sub_walk_all <- sub_walk_all %>% left_join(df_participants[, c('Code', 'sex')], by=c("id"="Code"))
+sop_all <- sop_all %>% left_join(df_participants[, c('Code', 'sex')], by=c("id"="Code"))
+sub_sop_all <- sub_sop_all %>% left_join(df_participants[, c('Code', 'sex')], by=c("id"="Code"))
+sub_sop_all_2 <- sub_sop_all_2 %>% left_join(df_participants[, c('Code', 'sex')], by=c("id"="Code"))
+```
 
-## NEw charts
+To assess effects of gender, we pooled data per gender and per modality (desktop, immersive VR and real world) and analysed some of the effects separately.
+
+```{r}
+ez_data <- ez_prepare(sop_all, 'sex')
+ez_fit <- ezANOVA(
+        ez_data,
+        dv = abs_error,
+        wid = id,
+        between = .(type, sex),
+        return_aov = T,
+        type = 3)
+```
+
+Analysing complete pointing performance (averaged across both sessions) overall has not been affected by gender `r apa_print(ez_fit$aov)$full_result$sex` or interaction between gender and modality  `r apa_print(ez_fit$aov)$full_result$type_sex`.
+
+### Walkign distance
+```{r}
+ez_data <- ez_prepare(walk_all, 'sex')
+ez_fit <- ezANOVA(
+        ez_data,
+        dv = min_norm_distance,
+        wid = id,
+        between = .(type, sex),
+        return_aov = T,
+        type = 3)
+#ez_data %>% group_by(sex, type) %>% summarise(mean = mean(min_norm_distance, na.rm=T))
+ggplot(ez_data, aes(graph.type, min_norm_distance, group=sex, fill=sex)) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
+```
+Analysing complete walking distance as a product of either gender or modality, (averaged across all blocks) the walked normalised distance has not been affected by gender `r apa_print(ez_fit$aov)$full_result$sex` or interaction between gender and modality  `r apa_print(ez_fit$aov)$full_result$type_sex`.
+
+### Visitation errors
+```{r}
+ez_data <- ez_prepare(walk_all, 'sex')
+ez_fit <- ezANOVA(
+        ez_data,
+        dv = errors,
+        wid = id,
+        between = .(type, sex),
+        return_aov = T,
+        type = 3)
+#ez_data %>% group_by(sex, type) %>% summarise(mean = mean(min_norm_distance, na.rm=T))
+ggplot(ez_data, aes(graph.type, errors, group=sex, fill=sex)) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
+```
+
+Analysing complete visitations errors as a product of either gender or modality, (averaged across all blocks) number of errors has not been affected by gender `r apa_print(ez_fit$aov)$full_result$sex` or interaction between gender and modality  `r apa_print(ez_fit$aov)$full_result$type_sex`.
+
+Therefore we decided not to include gender specific analyses in our paper, as it might muddy the analyses with another predictor.
+
+## New charts
+
+## Effect sizes

### 2019-04-08 Scripts/paper-prepare.R

@@ -4,6 +4,8 @@ source_folder("../TwoWorlds")
 source('../Scripts/analysis-helpers.R')
 sop_all <- read.table("../sop.csv", sep=";", header = T, stringsAsFactors = F)
 walk_all <- read.table("../walk.csv", sep=";", header = T, stringsAsFactors = F)
+walk_all$graph.type <- dplyr::recode(walk_all$type, "real" = "Real-world", "vr" = "Immersive VR", 've'='Desktop')
+sop_all$graph.type <- dplyr::recode(sop_all$type, "real" = "Real-world", "vr" = "Immersive VR", 've'='Desktop')
 walk_all$learning.condition <- walk_all$condition
 walk_all$learning.condition <- dplyr::recode(walk_all$learning.condition, "real-real" = "Real", "ve-real" = "Desktop", "vr-real" = "Treadmill VR", "real-vr" = "Real", "vr-vr" = "Treadmill VR")
 walk_all$graph.condition <- dplyr::recode(walk_all$condition, "real-real" = "Real-world to Real-world", "ve-real" = "Desktop to Real-world", "vr-real" = "Immersive VR to Real-world", "real-vr" = "Real-world to Immersive VR", "vr-vr" = "Immersive VR to Immersive VR")

### 2019-04-08 TwoWorlds/helpers-results.R

@@ -30,3 +30,11 @@ m_sd_report <- function(vec, digits=3){
   s <- round(sd(vec, na.rm=T), digits)
   return(paste("M = ", m, ", SD = ", s, collapse=""))
 }
+
+t_test_effect <- function(f, data, paired = F){
+  t <- t.test(formula = f, data = data, paired = paired)
+  text <- apa_print(t)$statistic
+  d <- cohen.d(f, data, paired = paired)
+  text <- paste0(text, ", $d = ", round(d$estimate, 3), "$")
+  return(text)
+}

### 2019-06-19 Report/final-results.Rmd
Updates final results

@@
 df_wide_errors$block34 <- (df_wide_errors$block3-df_wide_errors$block4)/(df_wide_errors$block3+df_wide_errors$block4)
+df_wide_errors$block34[is.na(df_wide_errors$block34)] <- 0 #fixes case where person scores 0 in block 3 and 4 and so we have
 df_wide_distance$block34 <- (df_wide_distance$block3-df_wide_distance$block4)/(df_wide_distance$block3+df_wide_distance$block4)
@@
 We also observed significant differences between conditions in pointing performance at the end of the first phase, `r apa_print(aov_point_phase1)$full_result$learning_condition`.
-Specifically, participants in the real-world learning group  (`r m_sd_report(only_conditions(point_data_1, "Real")$abs_error)`) performed significantly better compared to both the immersive VR group (`r m_sd_report(only_conditions(point_data_1, "Treadmill VR")$abs_error)`), `r t_test_effect(abs_error ~ learning.condition, only_conditions(point_data_1, c('Real', 'Treadmill VR')))`, and the desktop group (`r m_sd_report(only_conditions(point_data_1, "Desktop")$abs_error)`), `r t_test_effect(abs_error ~ learning.condition, only_conditions(point_data_1, c('Real', 'Desktop')))`.
+Specifically, participants in the real-world learning group  (`r m_sd_report(only_conditions(point_data_1, "Real")$abs_error)`) performed significantly better compared to both the immersive VR group (`r m_sd_report(only_conditions(point_data_1, "Treadmill VR")$abs_error)`), `r t_test_effect(abs_error ~ learning.condition, only_conditions(point_data_1, c('Real', 'Treadmill VR')))`, and the desktop group (`r m_sd_report(only_conditions(point_data_1, "Desktop")$abs_error)`), `r t_test_effect(abs_error ~ learning.condition, only_conditions(point_data_1, c('Real', 'Desktop')))`, but no difference between VR and desktop (`r t_test_effect(abs_error ~ learning.condition, only_conditions(point_data_1, c('Treadmill VR', 'Desktop')))`).
 Overall, these findings suggested that participants learned the building the best as a result of real-world navigation, followed by the immersive condition, with the desktop condition showing the slowest acquisition.
@@
 
+```{r}
+### graphs
+ggplot(only_blocks(sub_walk_all_factor, 3:4), aes(x = graph.exp_block_id, y = min_norm_distance, fill = graph.exp_block_id)) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
+  ylab("Average normalised trial distance") + xlab("Learning condition") + 
+  scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F) +
+  facet_wrap(~graph.learning.condition, strip.position = "bottom", scales = "free_x") + 
+  theme(panel.spacing = unit(0, "lines"), 
+         strip.background = element_blank(),
+         strip.placement = "outside")
+
+ggplot(only_blocks(sub_walk_all_factor, 3:4), aes(x = graph.exp_block_id, y = errors, fill = graph.exp_block_id)) +
+  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
+  ylab("Average visitation errors") + xlab("Learning condition") + 
+  scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F) +
+  facet_wrap(~graph.learning.condition, strip.position = "bottom", scales = "free_x") + 
+  theme(panel.spacing = unit(0, "lines"), 
+         strip.background = element_blank(),
+         strip.placement = "outside")
+
+```
+
 ## Quantifying transfer 
@@
 Comparing walked distance and visitation errors across between blocks 4 and 6 by running repeated measures ANOVA, we found a significant effect of learning condition on distance,`r apa_print(ez_fit_distance$aov)$full_result$learning_condition` and visitation errors `r apa_print(ez_fit_errors$aov)$full_result$learning_condition`.
-We also wound significant interaction effect between learning condition and block for both distance `r apa_print(ez_fit_distance$aov)$full_result$learning_condition` and visitation errors `r apa_print(ez_fit_errors$aov)$full_result$learning_condition`.
+We also wound significant interaction effect between learning condition and block for both distance `r apa_print(ez_fit_distance$aov)$full_result$learning_condition_exp_block_id` and visitation errors `r apa_print(ez_fit_errors$aov)$full_result$learning_condition_exp_block_id`.
 
@@
 kable(tukey_report_table(TukeyHSD(aov(min_norm_distance ~ condition, only_blocks(walk_all, c(6))))))
+
+kable(tukey_report_table(TukeyHSD(aov(errors ~ condition, only_blocks(walk_all, c(6))))))
 ```

### 2019-10-28 Report/second-revision.Rmd
Slight changes, adds second revision results

---
title: "2nd Revision"
author: "Lukáš Hejtmánek"
date: "20 March 2019"
output: html_document
---

```{r setup, include=FALSE}
#needs lme and car in the top because of the recode
library(ggplot2)
library(effsize)
library(ez)
library(Hmisc)
library(dplyr)
library(reshape2)
library(broom)
library(knitr)
library(papaja)
library(googlesheets)

source("../Scripts/paper-prepare.R")
source("../TwoWorlds/helpers-report.R")
load('../settings.data')
df_participants <- settings$participants
df_participants <- df_participants %>% mutate(condition=paste0(First.phase,"-",Second.phase))

overview <- settings$versions
overview <- overview %>% mutate(condition=paste0(First.phase,"-",Second.phase))
overview <- overview %>% left_join(df_participants[, c('Code', 'sex')], by="Code")

### Try running it without outliers
walk_all[!is.na(walk_all$norm_distance) & walk_all$norm_distance > 3, c("distance", "min_norm_distance")] <- NA
sub_walk_all[!is.na(sub_walk_all$norm_distance) & sub_walk_all$norm_distance > 3, c("distance", "min_norm_distance")] <- NA

knitr::opts_chunk$set(echo=T, warning = F, message = F)
```

## Difference in experiment 1 block 1

```{r, result="asis"}
apa_print(aov(min_norm_distance ~ learning.condition, data = sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$full_result$learning_condition
apa_print(aov(errors ~ learning.condition, data = sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$full_result$learning_condition
```
## Difference in experiment 1 block 2

```{r}
apa_print(aov(min_norm_distance ~ learning.condition, data = sub_walk_all[sub_walk_all$exp_block_id == 2, ]))$full_result$learning_condition
apa_print(aov(errors ~ learning.condition, data = sub_walk_all[sub_walk_all$exp_block_id == 2, ]))$full_result$learning_condition
```
```{r}
only_blocks(sub_walk_all, 2) %>% group_by(learning.condition) %>% summarize(avg = mean(errors, na.rm=T))
t_test_effect(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Treadmill VR'), 2))
t_test_effect(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Desktop'), 2))
t_test_effect(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Treadmill VR', 'Desktop'), 2))
```

## Difference in experiment 2 block 1

```{r}
apa_print(aov(min_norm_distance ~ learning.condition, data = sub_walk_all_2[sub_walk_all_2$exp_block_id == 1, ]))$full_result$learning_condition
apa_print(aov(errors ~ learning.condition, data = sub_walk_all_2[sub_walk_all_2$exp_block_id == 1, ]))$full_result$learning_condition
```

## Block 4 comparisons in experiment 1

```{r}
only_blocks(sub_walk_all, 4) %>% group_by(learning.condition) %>% summarize(avg = mean(errors, na.rm=T))
t_test_effect(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Treadmill VR'), 4))
t_test_effect(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Desktop'), 4))
t_test_effect(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Treadmill VR', 'Desktop'), 4))

only_blocks(sub_walk_all, 4) %>% group_by(learning.condition) %>% summarize(avg = mean(min_norm_distance, na.rm=T))
t_test_effect(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Treadmill VR'), 4))
t_test_effect(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Desktop'), 4))
t_test_effect(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Treadmill VR', 'Desktop'), 4))
```

## Block 4 comparisons in experiment 2

```{r}
only_blocks(sub_walk_all_2, 4) %>% group_by(learning.condition) %>% summarize(avg = mean(errors, na.rm=T))
t_test_effect(errors ~ learning.condition, only_conditions_blocks(sub_walk_all_2, c('Real', 'Treadmill VR'), 4))
only_blocks(sub_walk_all_2, 4) %>% group_by(learning.condition) %>% summarize(avg = mean(min_norm_distance, na.rm=T))
t_test_effect(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all_2, c('Real', 'Treadmill VR'), 4))
```

## Absolute vs normalised values 
```{r}
b4 <- only_blocks(sub_walk_all, c(4))
apa_print(aov(b4$distance ~ b4$learning.condition))$full_result$b4_learning_condition
apa_print(aov(b4$min_norm_distance ~ b4$learning.condition))$full_result$b4_learning_condition
ggplot(b4, aes(learning.condition, min_norm_distance, fill)) + 
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
b6 <- only_blocks(sub_walk_all, c(6))
apa_print(aov(b6$distance ~ b6$learning.condition))$full_result$b6_learning_condition
apa_print(aov(b6$min_norm_distance ~ b6$learning.condition))$full_result$b6_learning_condition
ggplot(b6, aes(learning.condition, distance, fill)) + 
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```

## Dropout changes
```{r}
finished_condition <- overview %>% filter(Second.phase != "ve") %>% select(First.phase, Second.phase, finished)
finished_condition <- finished_condition %>% mutate(has_vr = First.phase == "vr" |  Second.phase == "vr")
exp_1 <- finished_condition[finished_condition$Second.phase == "vr", ]
chsqt_1 <- table(exp_1$First.phase, exp_1$finished) %>% chisq.test()
exp_2 <- finished_condition[finished_condition$Second.phase == "real", ]
chsqt_2 <- table(exp_2$First.phase, exp_2$finished) %>% chisq.test()
chsqt_1
chsqt_2
```

### 2019-10-28 TwoWorlds/helpers-report.R

@@ -1,7 +1,7 @@
 apa_p <- function(values){
   output <- c()
   for (value in values){
-    if(value < 0.01){ output <- c(output, "p < .001"); next}
+    if(value < 0.001){ output <- c(output, "p < .001"); next}
     output <- c(output, paste0("p = ", round(value, 3)))
   }
   return(output)

### 2019-12-12 Report/paper-graphs.Rmd
Adds final results graphs fro the paper modification

---
title: "paper-graphs"
author: "Lukáš 'hejtmy' Hejtmánek"
date: "12/12/2019"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(papaja)

source("../Scripts/paper-prepare.R")
source("../TwoWorlds/helpers-report.R")

#settings <- load_google_sheets()
load('../settings.data')

#green, violet, blue, yellow, pink, dark blue
estimote_pal <- c('#b6d6c1', '#54003d','#93d5f6','#d5d215','#f3aacb','#2d2556')
theme_set(theme_minimal() + theme(axis.text=element_text(size=14),
        axis.title=element_text(size=16)))

### Try running it without outliers
sub_walk_all[!is.na(sub_walk_all$norm_distance) & sub_walk_all$norm_distance > 3, c("distance", "min_norm_distance")] <- NA
walk_all[!is.na(walk_all$norm_distance) & walk_all$norm_distance > 3, c("distance", "min_norm_distance")] <- NA

sub_walk_all_factor <- sub_walk_all
sub_walk_all_factor$exp_block_id <- factor(sub_walk_all_factor$exp_block_id)
sub_walk_all_factor$id <- factor(sub_walk_all_factor$id)
sub_walk_all_factor$learning.condition <- factor(sub_walk_all_factor$learning.condition)

point_data_1 <- sub_sop_all[sub_sop_all$phase == 1,]
knitr::opts_chunk$set(echo=F, warning = F, message = F)
```

## 5a
```{r}
ggplot(only_blocks(sub_walk_all_factor, 1:3), aes(x = graph.exp_block_id, y = errors, fill = graph.exp_block_id)) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
  ylab("Average visitation errors") + xlab("Learning condition") + 
  scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F) +
  facet_wrap(~graph.learning.condition, strip.position = "bottom", scales = "free_x") + 
  theme(panel.spacing = unit(0, "lines"), 
         strip.background = element_blank(),
         strip.placement = "outside")
```

## 5b
```{r}
# Pre transfer plot of normalised distances
ggplot(only_blocks(sub_walk_all_factor, 1:3), aes(x = graph.exp_block_id, y = min_norm_distance, fill = graph.exp_block_id)) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
  ylab("Average normalised trial distance") + xlab("Learning condition") + 
  scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F) +
  facet_wrap(~graph.learning.condition, strip.position = "bottom", scales = "free_x") + 
  theme(panel.spacing = unit(0, "lines"), 
         strip.background = element_blank(),
         strip.placement = "outside")

```

## 5c

```{r}
ggplot(point_data_1, aes(x = graph.learning.condition, y = abs_error, fill = graph.learning.condition)) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
  ylab("Average pointing error after learning phase") + xlab("Learning condition") + 
  scale_fill_manual(values = estimote_pal, "Learning condition") +
  guides(fill=F) + theme(axis.title.y = element_text(size = rel(0.9)))

```

## 6a
```{r}
ggplot(only_blocks(sub_walk_all_factor, 3:4), aes(x = graph.exp_block_id, y = errors, fill = graph.exp_block_id)) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
  ylab("Average visitation errors") + xlab("Learning condition") + 
  scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F) +
  facet_wrap(~graph.learning.condition, strip.position = "bottom", scales = "free_x") + 
  theme(panel.spacing = unit(0, "lines"), 
         strip.background = element_blank(),
         strip.placement = "outside")
```

## 6b
```{r}
### graphs
ggplot(only_blocks(sub_walk_all_factor, 3:4), aes(x = graph.exp_block_id, y = min_norm_distance, fill = graph.exp_block_id)) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
  ylab("Average normalised trial distance") + xlab("Learning condition") + 
  scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F) +
  facet_wrap(~graph.learning.condition, strip.position = "bottom", scales = "free_x") + 
  theme(panel.spacing = unit(0, "lines"), 
         strip.background = element_blank(),
         strip.placement = "outside")

```

## 7a

```{r, fig.width=10, fig.height=8}
ggplot(only_blocks(walk_all, 6), aes(x = testing.condition, y = min_norm_distance, fill = graph.learning.condition)) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
  ylab("Average path distance in the last block of the transfer phase") + xlab("Transfer condition") + 
  scale_fill_manual(values = estimote_pal, "Condition")  + guides(fill=F) + 
  facet_wrap(~graph.learning.condition, strip.position = "top", scales = "free_x") + 
  ggtitle("Learning condition")+
  theme(panel.spacing = unit(0, "lines"), 
         strip.background = element_blank(),
         strip.placement = "outside",
         strip.text = element_text(size = 16),
          plot.title = element_text(hjust = 0.5))

```

##• 7b

```{r}
ggplot(only_blocks(walk_all, 6), aes(x = testing.condition, y = errors, fill = graph.learning.condition)) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
  ylab("Average number of visitation errors in the last block of the transfer phase") + xlab("Transfer condition") + 
  scale_fill_manual(values = estimote_pal, "Condition")  + guides(fill=F) + 
  facet_wrap(~graph.learning.condition, strip.position = "top", scales = "free_x") + 
  ggtitle("Learning condition")+
  theme(panel.spacing = unit(0, "lines"), 
         strip.background = element_blank(),
         strip.placement = "outside",
         strip.text = element_text(size = 16),
          plot.title = element_text(hjust = 0.5))
```

### 2020-02-12 Report/paper-graphs.Rmd
Updates graphs for the paper

@@
 estimote_pal <- c('#b6d6c1', '#54003d','#93d5f6','#d5d215','#f3aacb','#2d2556')
-theme_set(theme_minimal() + theme(axis.text=element_text(size=14),
-        axis.title=element_text(size=16)))
+theme_set(theme_minimal() + 
+          theme(axis.text=element_text(size=14),
+                axis.title=element_text(size=16),
+                axis.title.y = element_text(size=20),
+                strip.text.x = element_text(size=20),
+                plot.subtitle = element_text(hjust = 1, vjust = 0, size = 18, face = "italic"))
+          )
 
@@
 ## 5a
-```{r}
+```{r, fig.width=10, fig.height=8}
 ggplot(only_blocks(sub_walk_all_factor, 1:3), aes(x = graph.exp_block_id, y = errors, fill = graph.exp_block_id)) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
-  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
-  ylab("Average visitation errors") + xlab("Learning condition") + 
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95), geom="bar") +
+  stat_summary(fun.data=mean_se, position=position_dodge(0.95), geom="errorbar", width = 0.2, size = 1.25) + 
+  facet_wrap(~graph.learning.condition, strip.position = "bottom", scales = "free_x") + 
   scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F) +
-  facet_wrap(~graph.learning.condition, strip.position = "bottom", scales = "free_x") + 
+  labs(x = NULL, y ="Average number of visitation errors", subtitle = "5a") +
   theme(panel.spacing = unit(0, "lines"), 
@@
 ## 5b
-```{r}
+```{r, fig.width=10, fig.height=8}
 # Pre transfer plot of normalised distances
 ggplot(only_blocks(sub_walk_all_factor, 1:3), aes(x = graph.exp_block_id, y = min_norm_distance, fill = graph.exp_block_id)) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
-  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
-  ylab("Average normalised trial distance") + xlab("Learning condition") + 
-  scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F) +
+  stat_summary(fun.y=mean,position=position_dodge(width=0.95), geom="bar") +
+  stat_summary(fun.data=mean_se, position=position_dodge(0.95), geom="errorbar", width = 0.2, size = 1.25) + 
   facet_wrap(~graph.learning.condition, strip.position = "bottom", scales = "free_x") + 
+  scale_fill_manual(values=estimote_pal,"Block") + 
+  labs(x = NULL, y = "Average normalised trial distance", subtitle = "5b") +
+  guides(fill=F) +
   theme(panel.spacing = unit(0, "lines"), 
@@
          strip.placement = "outside")
-
 ```
@@
 
-```{r}
+```{r, fig.width=10, fig.height=8}
 ggplot(point_data_1, aes(x = graph.learning.condition, y = abs_error, fill = graph.learning.condition)) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
-  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
-  ylab("Average pointing error after learning phase") + xlab("Learning condition") + 
+  stat_summary(fun.y=mean, position=position_dodge(width=0.95), geom="bar") +
+  stat_summary(fun.data=mean_se, position=position_dodge(0.95), geom="errorbar", width = 0.2, size = 1.25) + 
   scale_fill_manual(values = estimote_pal, "Learning condition") +
-  guides(fill=F) + theme(axis.title.y = element_text(size = rel(0.9)))
-
+  labs(x = NULL, y = "Average pointing error after learning phase", subtitle = "5c") +
+  guides(fill=F) + 
+  theme(axis.text.x = element_text(size=18, color= "black"))
 ```
@@
 ## 6a
-```{r}
+
+```{r, fig.width=10, fig.height=8}
 ggplot(only_blocks(sub_walk_all_factor, 3:4), aes(x = graph.exp_block_id, y = errors, fill = graph.exp_block_id)) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
-  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
-  ylab("Average visitation errors") + xlab("Learning condition") + 
-  scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F) +
+  stat_summary(fun.y = mean,position=position_dodge(width = 0.95),geom="bar") +
+  stat_summary(fun.data = mean_se, position = position_dodge(0.95), geom="errorbar", width = 0.2, size = 1.25) + 
+  scale_fill_manual(values = estimote_pal,"Block") + 
   facet_wrap(~graph.learning.condition, strip.position = "bottom", scales = "free_x") + 
+  labs(x=NULL, y = "Average number of visitation errors", subtitle = "6a") +
+  guides(fill = F) +
   theme(panel.spacing = unit(0, "lines"), 
@@
 ## 6b
-```{r}
+```{r, fig.width=10, fig.height=8}
 ### graphs
 ggplot(only_blocks(sub_walk_all_factor, 3:4), aes(x = graph.exp_block_id, y = min_norm_distance, fill = graph.exp_block_id)) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
   stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
-  ylab("Average normalised trial distance") + xlab("Learning condition") + 
-  scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F) +
+  stat_summary(fun.data=mean_se, position=position_dodge(0.95), geom="errorbar", width = 0.2, size = 1.25) + 
+  scale_fill_manual(values=estimote_pal,"Block") + 
   facet_wrap(~graph.learning.condition, strip.position = "bottom", scales = "free_x") + 
+  labs(x= NULL, y = "Average normalised trial distance", subtitle = "6b") +
+  guides(fill = F) +
   theme(panel.spacing = unit(0, "lines"), 
@@
 ```{r, fig.width=10, fig.height=8}
-ggplot(only_blocks(walk_all, 6), aes(x = testing.condition, y = min_norm_distance, fill = graph.learning.condition)) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
+only_blocks(walk_all, 6) %>%
+  select(testing.condition, min_norm_distance, graph.learning.condition) %>%
+  rbind(data.frame(testing.condition = "Immersive VR", graph.learning.condition="Desktop", min_norm_distance = NA)) %>%
+  ggplot(aes(x = testing.condition, y = min_norm_distance, fill = testing.condition)) +
   stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
-  ylab("Average path distance in the last block of the transfer phase") + xlab("Transfer condition") + 
-  scale_fill_manual(values = estimote_pal, "Condition")  + guides(fill=F) + 
-  facet_wrap(~graph.learning.condition, strip.position = "top", scales = "free_x") + 
-  ggtitle("Learning condition")+
+  stat_summary(fun.data=mean_se, position=position_dodge(0.95), geom="errorbar", width = 0.2, size = 1.25) + 
+  scale_fill_manual(values = estimote_pal, "Transfer condition") +
+  facet_wrap(~graph.learning.condition, strip.position = "bottom", scales = "free_x") +
+  labs(x = NULL, y = "Average path distance in the last\n block of the transfer phase", subtitle = "7a") + 
   theme(panel.spacing = unit(0, "lines"),
@@
         strip.placement = "outside",
+        legend.position = "top",
+        legend.title = element_text(size = 16),
+        legend.text = element_text(size = 16),
+        axis.text.x = element_blank(),
         strip.text = element_text(size = 16),
         plot.title = element_text(hjust = 0.5))
-
 ```
 
-##• 7b
+## 7b
 
-```{r}
-ggplot(only_blocks(walk_all, 6), aes(x = testing.condition, y = errors, fill = graph.learning.condition)) +
-  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
-  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
-  ylab("Average number of visitation errors in the last block of the transfer phase") + xlab("Transfer condition") + 
-  scale_fill_manual(values = estimote_pal, "Condition")  + guides(fill=F) + 
-  facet_wrap(~graph.learning.condition, strip.position = "top", scales = "free_x") + 
-  ggtitle("Learning condition")+
+```{r, fig.width=10, fig.height=8}
+only_blocks(walk_all, 6) %>%
+  select(testing.condition, errors, graph.learning.condition) %>%
+  rbind(data.frame(testing.condition = "Immersive VR", graph.learning.condition="Desktop", errors = NA)) %>% 
+  ggplot(aes(x = testing.condition, y = errors, fill = testing.condition)) +
+  stat_summary(fun.y = mean, position = position_dodge(width=0.95), geom = "bar") +
+  stat_summary(fun.data = mean_se, position=position_dodge(0.95), geom="errorbar", width = 0.2, size = 1.25) +
+  facet_wrap(~graph.learning.condition, strip.position = "bottom", scales = "free_x") +
+  scale_fill_manual(values = estimote_pal, "Transfer condition") +
+  labs(x = NULL, y = "Average number of visitation errors in\n the last block of the transfer phase", subtitle = "7b") +
   theme(panel.spacing = unit(0, "lines"),
@@
         strip.placement = "outside",
+        legend.position = "top",
+        legend.title = element_text(size = 16),
+        legend.text = element_text(size = 16),
+        axis.text.x = element_blank(),
         strip.text = element_text(size = 16),

### 2020-02-12 app/global.R
Removes server app

@@ -1,38 +1,0 @@
-library(brainvr.R)
-library(ggplot2)
-
-source("../TwoWorlds/twoworld-getters.R")
-source("../TwoWorlds/twoworld-visualisation.R")
-source("../TwoWorlds/visualisation-helpers.R")
-source("../TwoWorlds/twoworld-results.R")
-dir <- "D:/GoogleDrive/Davis/Data/pilot/neo1/"
-
-
-walk <- load_experiment(dir, exp_timestamp = '17-41-52-03-12-2017')
-sop <- load_experiment(dir, exp_timestamp = '18-07-09-03-12-2017')
-
-walk$map_limits <- list(x = c(-5, 105), y = c(0, 100))
-walk <- mirror_axes(walk)
-walk <- translate_positions(walk, c(33.5, 0, 47.75))
-
-sop <- mirror_axes(sop)
-sop <- translate_positions(sop, c(33.5, 0, 47.75))
-sop$map_limits <- list(x = c(-5, 105), y = c(0, 100))
-
-n_walk_trials <- length(unique(walk$data$experiment_log$data$Index))
-n_sop_trials <- length(unique(sop$data$experiment_log$data$Index))
-
-restimoteObj <- load_restimote_log(dir)
-restimoteObj <- load_restimote_companion_log(dir, obj = restimoteObj)
-
-restimoteObj <- preprocess_companion_log(restimoteObj)
-restimoteObj <- preprocess_restimote_log(restimoteObj)
-restimoteObj <- remove_pointing(restimoteObj, c(7:8))
-restimoteObj$map_limits <- list(x = c(0, 25), y = c(0, 25))
-n_restimote_trials <- restimoteObj$n_trials
-ls <- get_n_pointings(restimoteObj)
-n_restimote_sop <- ls$companion
-
-
-dt_log <- get_position_trial(restimoteObj, 1)
-restimote_nrow <- nrow(dt_log)

### 2020-02-12 app/server.R

@@ -1,27 +1,0 @@
-server <- function(input, output, session){
-  output$plotTrialPath <- renderPlot({
-    plot_walking_trial(walk, input$sliderTrial)
-  })
-  
-  output$plotSop <- renderPlot({
-    plot_sop_point(sop, input$sliderSop)
-  })
-  
-  restimoteLog <- reactive({
-    dt_log <- get_position_trial(restimoteObj, input$sliderRestimoteTrial)
-    restimote_nrow <<- nrow(dt_log)
-    updateSliderInput(session, "sliderRestimoteTrialTime", value = restimote_nrow, max = restimote_nrow)
-    return(dt_log)
-  })
-  
-  subRestimoteLog <- reactive({
-    dt <- restimoteLog()
-    return(dt[1:input$sliderRestimoteTrialTime, ])
-  })
-  
-  output$plotRestimoteTrial <- renderPlot({
-    plt <- restimoter::create_plot(restimoteObj)
-    plt <- plot_add_restimote_path(plt, df_pos = subRestimoteLog())
-    plt + theme_bw()
-  })
-}

### 2020-02-12 app/ui.R

@@ -1,17 +1,0 @@
-ui <- fluidPage(
-  sliderInput(inputId = "sliderTrial", label = "Select trial to plot", 
-              value = 1, min = 1, step = 1, max = n_walk_trials),
-  plotOutput(outputId = "plotTrialPath"),
-  sliderInput(inputId = "sliderSop", label = "Slide to select pointing", 
-              value = 1, min = 1, max = n_sop_trials, step = 1),
-  plotOutput(outputId = "plotSop"),
-  sliderInput(inputId = "sliderRestimoteTrial", label = "Slide to select restimote trial", 
-              value = 1, min = 1, max = n_restimote_trials, step = 1),
-  sliderInput(inputId = "sliderRestimoteTrialTime", label = "Slide to select restimote trial time", 
-              value = 2, min = 2, max = restimote_nrow, step = 1,
-              animate = animationOptions(interval = 100, loop = FALSE)),
-  plotOutput(outputId = "plotRestimoteTrial"),
-  sliderInput(inputId = "sliderRestimotePoint", label = "Slide to select restimote pointing", 
-              value = 1, min = 1, max = n_restimote_sop, step = 1),
-  plotOutput(outputId = "plotRestimoteSop")
-)

### 2020-02-12 build.R
removes build.R file

@@ -1,8 +1,0 @@
-detach("package:brainvr.R", unload = TRUE)
-detach("package:restimoter", unload = TRUE)
-detach("package:navr", unload = TRUE)
-#install.packages("../brainvr-reader/",  type = "source", repos = NULL)
-#install.packages("../restimoter/", type = "source", repos = NULL )
-
-library(brainvr.R)
-library(restimoter)
